---
title: "Recipe data analysis"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
```

# Steadystate fate factors

## Make batch files for steadystate fate factors
Calculating the steadystate masses for 1000 runs per polymer takes a lot of time.
The script 01_Write_batch_files_SS.R was written to make one batch file per polymer,
so that the fate factors for the polymers can be calculated in parallel. The chunk
below shows how to make these batch files. 

```{r Steadystate batch files}
source("vignettes/CaseStudies/Recipe/01_Write_batch_files_SS.R")
```

The files are created in the folder "vignettes/CaseStudies/Recipe/BatchFilesRecipe". 
All files can be run in the background in RStudio, or on a high performance cluster 
using the commands written to the script HPC_commands.txt.

## Load output data
```{r Load output data, warning=FALSE, message=FALSE}
folderpath <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_SS"

filepaths <- list.files(folderpath)

RData_files <- filepaths[endsWith(filepaths, '.RData')]

SS_masses_continental <- tibble()

for(file in RData_files){
  load(paste0(folderpath, "/", file))
  
  SS_masses_allScale <- Output |> 
    unnest(SBoutput) |> 
    mutate(OutputType = names(SBoutput)) |> 
    rename(EmisScale = Scale) |> 
    filter(OutputType == "SteadyStateMass") |> 
    unnest(SBoutput) |> 
    filter(Species != "Unbound") |> 
    ungroup() |> 
    group_by(Polymer, EmisComp, EmisScale, RUN, Scale, SubCompart, Unit) |> 
    summarise(EqMass_SAP = sum(EqMass)) 
  
  SS_masses_continental_pol <- SS_masses_allScale |> 
  filter(EmisScale == "Continental" & (Scale == c("Regional")|Scale == c("Continental"))) |> 
  ungroup() |> 
  group_by(Polymer, EmisComp, EmisScale, RUN, SubCompart, Unit) |> 
  summarise(EqMass_SAP = sum(EqMass_SAP)) |>  # sum nested regional mass and rest of EU mass
  mutate(
    CompartmentFF = case_when(
      SubCompart == "agriculturalsoil" ~ "agriculturalSoil",
      SubCompart == "naturalsoil" ~ "naturalSoil",
      SubCompart == "othersoil" ~ "otherSoil",
      SubCompart == "cloudwater" ~ "air",
      SubCompart == "lake" ~ "freshwater",
      SubCompart == "river" ~ "freshwater",
      SubCompart == "lakesediment" ~ "freshwatersediment",
      TRUE ~ SubCompart)) |> 
    ungroup() |>
    group_by(Polymer, EmisComp, EmisScale, RUN, CompartmentFF, Unit) |>
    summarise(EqMass_SAP = sum(EqMass_SAP)) |>
    ungroup() |>
  mutate(Unit = "kg[ss]/kg[e] seconds")
  
  SS_masses_continental <- bind_rows(SS_masses_continental, SS_masses_continental_pol)
}

rm(SS_masses_allScale, SS_masses_continental_pol)
```

### Clean up data

TO DO: There are still a lot of very small values in the boxplots. Maybe there
should be more cleanup before making the plots.. 

Not all runs are successful due to the sensitivity of the solver. To remove the
values from unsuccessful runs, we will remove all values that are larger than 
the 95th percentile, or smaller than the 5th percentile.

```{r Clean up output, warning=FALSE, message=FALSE}
SS_masses_clean <- SS_masses_continental |>
  group_by(Polymer, CompartmentFF, EmisComp) |>
  mutate(
    p05 = quantile(EqMass_SAP, 0.05, na.rm = TRUE),
    p95 = quantile(EqMass_SAP, 0.95, na.rm = TRUE)
  ) |>
  filter(EqMass_SAP >= p05 & EqMass_SAP <= p95) |>
  select(-p05, -p95) |> 
  ungroup()

rm(SS_masses_continental)
```

```{r Make table, warning=FALSE, message=FALSE}
output_table_percentiles <- SS_masses_clean |>
  group_by(Polymer,CompartmentFF,EmisComp) |> 
  summarise(FF_SteadyState_avg = mean(EqMass_SAP),
            FF_SteadyState_std = sd(EqMass_SAP),
            FF_SteadyState_p95 = quantile(EqMass_SAP, 0.95),
            FF_SteadyState_p75 = quantile(EqMass_SAP, 0.75),            
            FF_SteadyState_p25 = quantile(EqMass_SAP, 0.25),
            FF_SteadyState_p05 = quantile(EqMass_SAP, 0.05))

output_table_average <- output_table_percentiles |>
  pivot_wider(
    id_cols = c(Polymer,  EmisComp),
    names_from = CompartmentFF,
    values_from = FF_Dynamic_avg
  )

knitr::kable(output_table_average)

write.xlsx(output_table_average, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_output_table_average_SS.xlsx")
write.xlsx(output_table_percentiles, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_output_table_percentiles_SS.xlsx")
```

```{r Make figures, warning=FALSE, message=FALSE}
plot_theme <-  theme(
  axis.title.x = element_text(size = 30),    
  axis.title.y = element_text(size = 30),    
  axis.text.x = element_text(size = 28, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 28),
  legend.text = element_text(size = 28),
  title = element_text(size=30),
  strip.text = element_text(size = 20))

figurefolder <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/Figures/"

for(pol in unique(SS_masses_clean$Polymer)){
  plot_data <- SS_masses_clean |>
    filter(Polymer == pol) |>
    filter(!is.na(EqMass_SAP)) 
  
  facetplot <- ggplot(plot_data, aes(y=EqMass_SAP, x=CompartmentFF, fill=CompartmentFF))+
    geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=1, notch=FALSE)+
      scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x, n = 10),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
    facet_wrap(~EmisComp) +
  theme_bw() +
  labs(title = pol,
       y = paste0("Fate factor ", unique(plot_data$Unit)),
       x = "") +
  scale_fill_discrete() +
  plot_theme +
  theme(legend.position = "none")
  
  print(facetplot)

  ggsave(paste0(figurefolder, "FF_facetplot_SS_", pol, ".png"), plot=facetplot, width = 25, height = 15, dpi = 1000)
}
```

### Clear the environment and initialize a new environment for dynamic fate factor calculations

```{r Clear environment}
rm(list=ls())
```

```{r Initialize new environment}
library(tidyverse)
library(scales)
source("basescripts/initWorld_onlyParticulate.R")
```

# Dynamic fate factors

## Make batch files for dynamic fate factors
We can also calculate fate factors from dynamic masses. The goal is to calculate
fate factors after 20 and 100 years of emissions. The script below is used to 
create batch files for dynamic fate factors. 

```{r Dynamic batch files}
source("vignettes/CaseStudies/Recipe/03_Write_batch_files_Dyn.R")
```

## Load output data
```{r Load output data, warning=FALSE, message=FALSE}
folderpath <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_Dyn"

filepaths <- list.files(folderpath)

RData_files <- filepaths[endsWith(filepaths, '.RData')]

Dyn_masses_continental <- tibble()

states <- World$states$asDataFrame

years <- c(20, 100)

for(file in RData_files){
  load(paste0(folderpath, "/", file))
  
  Dyn_masses_allScale <- Output |> 
    unnest(SBoutput) |> 
    mutate(OutputType = names(SBoutput)) |> 
    rename(EmisScale = Scale) |> 
    filter(OutputType == "DynamicMass") |> 
    unnest(SBoutput) |> 
    filter(time %in% years) |>
    select(-OutputType) |>
    pivot_longer(!c(EmisComp, Polymer, time, RUN, EmisScale), values_to = "EqMass", names_to = "Abbr") |>
    rename(Year = time) |>
    left_join(states, by="Abbr")|>
    filter(Species != "Unbound") |> 
    ungroup() |> 
    mutate(Unit ="kg") |>
    group_by(Polymer, EmisComp, EmisScale, RUN, Scale, SubCompart, Unit, Year) |> 
    summarise(EqMass_SAP = sum(EqMass)) 
  
  Dyn_masses_continental_pol <- Dyn_masses_allScale |> 
  filter(EmisScale == "Continental" & (Scale == c("Regional")|Scale == c("Continental"))) |> 
  ungroup() |> 
  group_by(Polymer, EmisComp, EmisScale, RUN, SubCompart, Unit, Year) |> 
  summarise(EqMass_SAP = sum(EqMass_SAP)) |>  # sum nested regional mass and rest of EU mass
  mutate(
    CompartmentFF = case_when(
      SubCompart == "agriculturalsoil" ~ "agriculturalSoil",
      SubCompart == "naturalsoil" ~ "naturalSoil",
      SubCompart == "othersoil" ~ "otherSoil",
      SubCompart == "cloudwater" ~ "air",
      SubCompart == "lake" ~ "freshwater",
      SubCompart == "river" ~ "freshwater",
      SubCompart == "lakesediment" ~ "freshwatersediment",
      TRUE ~ SubCompart)) |> 
    ungroup() |>
    group_by(Polymer, EmisComp, EmisScale, RUN, CompartmentFF, Unit, Year) |>
    summarise(EqMass_SAP = sum(EqMass_SAP)) |>
    ungroup() |>
  mutate(Unit = "kg[dyn]/kg[e] seconds")
  
  Dyn_masses_continental <- bind_rows(Dyn_masses_continental, Dyn_masses_continental_pol)
}

rm(Dyn_masses_allScale, Dyn_masses_continental_pol)
```

### Clean up data

Not all runs are successful due to the sensitivity of the solver. To remove the
values from unsuccessful runs, we will remove all values that are larger than  
the 95th percentile, or smaller than the 5th percentile.

```{r Clean up output, warning=FALSE, message=FALSE}
Dyn_masses_clean <- Dyn_masses_continental |>
  group_by(Polymer, CompartmentFF, EmisComp, Year) |>
  mutate(
    p05 = quantile(EqMass_SAP, 0.05, na.rm = TRUE),
    p95 = quantile(EqMass_SAP, 0.95, na.rm = TRUE)
  ) |>
  filter(EqMass_SAP >= p05 & EqMass_SAP <= p95) |>
  select(-p05, -p95) |> 
  ungroup()

rm(Dyn_masses_continental)
```

```{r Make table, warning=FALSE, message=FALSE}
output_table_percentiles <- Dyn_masses_clean |>
  group_by(Polymer,CompartmentFF,EmisComp, Year) |> 
  summarise(FF_Dynamic_avg = mean(EqMass_SAP),
            FF_Dynamic_std = sd(EqMass_SAP),
            FF_Dynamic_p95 = quantile(EqMass_SAP, 0.95),
            FF_Dynamic_p75 = quantile(EqMass_SAP, 0.75),            
            FF_Dynamic_p25 = quantile(EqMass_SAP, 0.25),
            FF_Dynamic_p05 = quantile(EqMass_SAP, 0.05))

output_table_average <- output_table_percentiles |>
  pivot_wider(
    id_cols = c(Polymer,  EmisComp, Year),
    names_from = CompartmentFF,
    values_from = FF_Dynamic_avg
  )

knitr::kable(output_table_average)

write.xlsx(output_table_average, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_output_table_average_Dyn.xlsx")
write.xlsx(output_table_percentiles, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_output_table_percentiles_Dyn.xlsx")
```

```{r Make figures, warning=FALSE, message=FALSE}
plot_theme <-  theme(
  axis.title.x = element_text(size = 30),    
  axis.title.y = element_text(size = 30),    
  axis.text.x = element_text(size = 28, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 28),
  legend.text = element_text(size = 28),
  title = element_text(size=30),
  strip.text = element_text(size = 20))

figurefolder <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/Figures/"

for(pol in unique(Dyn_masses_clean$Polymer)){
  for(year in unique(Dyn_masses_clean$Year)){
      plot_data <- Dyn_masses_clean |>
      filter(Polymer == pol) |>
      filter(!is.na(EqMass_SAP)) |>
      filter(Year == year)
    
    facetplot <- ggplot(plot_data, aes(y=EqMass_SAP, x=CompartmentFF, fill=CompartmentFF))+
      geom_boxplot(outlier.colour="black", outlier.shape=16,
         outlier.size=1, notch=FALSE)+
        scale_y_log10(
      breaks = scales::trans_breaks("log10", function(x) 10^x, n = 10),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
      facet_wrap(~EmisComp) +
    theme_bw() +
    labs(title = paste0(pol, ", year ", year),
         y = paste0("Fate factor ", unique(plot_data$Unit)),
         x = "") +
    scale_fill_discrete() +
    plot_theme +
    theme(legend.position = "none")
    
    print(facetplot)
  
    ggsave(paste0(figurefolder, "FF_facetplot_Dyn_year_", year, "_", pol, ".png"), plot=facetplot, width = 25, height = 15, dpi = 1000)
  }
}
```
