---
title: "Recipe data analysis"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
```

## Make batch files for steadystate fate factors
Calculating the steadystate masses for 1000 runs per polymer takes a lot of time.
The script 01_Write_batch_files_SS.R was written to make one batch file per polymer,
so that the fate factors for the polymers can be calculated in parallel. The chunk
below shows how to make these batch files. 

```{r Steadystate batch files}
source("vignettes/CaseStudies/Recipe/01_Write_batch_files_SS.R")
```

The files are created in the folder "vignettes/CaseStudies/Recipe/BatchFilesRecipe". 
All files can be run in the background in RStudio, or on a high performance cluster 
using the commands written to the script HPC_commands.txt.

## Make batch files for dynamic fate factors
We can also calculate fate factors from dynamic masses. The goal is to calculate
fate factors after 20 and 100 years of emissions. The script below is used to 
create batch files for dynamic fate factors. 

```{r Dynamic batch files}
source("vignettes/CaseStudies/Recipe/03_Write_batch_files_Dyn.R")
```

## Load output data
```{r Load output data, warning=FALSE, message=FALSE}
folderpath <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_other"

filepaths <- list.files(folderpath)

RData_files <- filepaths[endsWith(filepaths, '.RData')]

SS_masses_continental <- tibble()

for(file in RData_files){
  load(paste0(folderpath, "/", file))
  
  SS_masses_allScale <- Output |> 
    unnest(SBoutput) |> 
    mutate(OutputType = names(SBoutput)) |> 
    rename(EmisScale = Scale) |> 
    filter(OutputType == "SteadyStateMass") |> 
    unnest(SBoutput) |> 
    filter(Species != "Unbound") |> 
    ungroup() |> 
    group_by(Polymer, EmisComp, EmisScale, RUN, Scale, SubCompart, Unit) |> 
    summarise(EqMass_SAP = sum(EqMass)) 
  
  SS_masses_continental_pol <- SS_masses_allScale |> 
  filter(EmisScale == "Continental" & (Scale == c("Regional")|Scale == c("Continental"))) |> 
  ungroup() |> 
  group_by(Polymer, EmisComp, EmisScale, RUN, SubCompart, Unit) |> 
  summarise(EqMass_SAP = sum(EqMass_SAP)) |>  # sum nested regional mass and rest of EU mass
  mutate(
    CompartmentFF = case_when(
      SubCompart == "agriculturalsoil" ~ "agriculturalSoil",
      SubCompart == "naturalsoil" ~ "naturalSoil",
      SubCompart == "othersoil" ~ "otherSoil",
      SubCompart == "cloudwater" ~ "air",
      SubCompart == "lake" ~ "freshwater",
      SubCompart == "river" ~ "freshwater",
      SubCompart == "lakesediment" ~ "freshwatersediment",
      TRUE ~ SubCompart)) |> 
    ungroup() |>
    group_by(Polymer, EmisComp, EmisScale, RUN, CompartmentFF, Unit) |>
    summarise(EqMass_SAP = sum(EqMass_SAP)) |>
    ungroup() |>
  mutate(Unit = "kg[ss]/kg[e] seconds")
  
  SS_masses_continental <- bind_rows(SS_masses_continental, SS_masses_continental_pol)
}

```

```{r Clean up output, warning=FALSE, message=FALSE}
SS_masses_clean <- SS_masses_continental |>
  filter(EqMass_SAP > 0)
```

```{r Make table, warning=FALSE, message=FALSE}
output_table <- SS_masses_clean |>
  group_by(Polymer,CompartmentFF,EmisComp) |> 
  summarise(FF_SteadyState_avg = mean(EqMass_SAP),
            FF_SteadyState_std = sd(EqMass_SAP),
            FF_SteadyState_p95 = quantile(EqMass_SAP, 0.95),
            FF_SteadyState_p75 = quantile(EqMass_SAP, 0.75),            
            FF_SteadyState_p25 = quantile(EqMass_SAP, 0.25),
            FF_SteadyState_p05 = quantile(EqMass_SAP, 0.05))

knitr::kable(output_table)

write.xlsx(output_table, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/FF_output_table.xlsx")
```

```{r Make figures, warning=FALSE, message=FALSE}
plot_theme <-  theme(
  axis.title.x = element_text(size = 30),    
  axis.title.y = element_text(size = 30),    
  axis.text.x = element_text(size = 28, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 28),
  legend.text = element_text(size = 28),
  title = element_text(size=30),
  strip.text = element_text(size = 20))

figurefolder <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Recipe/Figures/"

for(pol in unique(SS_masses_clean$Polymer)){
  plot_data <- SS_masses_clean |>
    filter(Polymer == pol) |>
    filter(EqMass_SAP > 10e-25) |>
    filter(!is.na(EqMass_SAP)) |>
    filter(EqMass_SAP != 0)
  
  facetplot <- ggplot(plot_data, aes(y=EqMass_SAP, x=CompartmentFF, fill=CompartmentFF))+
    geom_violin()+
      scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x, n = 10),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
    facet_wrap(~EmisComp) +
  theme_bw() +
  labs(title = pol,
       y = paste0("Fate factor ", unique(plot_data$Unit))) +
  scale_fill_discrete() +
  plot_theme +
  theme(legend.position = "none")
  
  print(facetplot)

  ggsave(paste0(figurefolder, "FF_facetplot_", pol, ".png"), plot=facetplot, width = 25, height = 15, dpi = 1000)
    
}
```
