---
title: "Microplastic verification"
author: "Anne Hids & Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.width = 14, fig.height = 10)
```

This vignette demonstrates the verification process of the particulate
version of Simplebox, in this case for microplastics. First, the k's are
compared between R and excel and consequently the steady state masses
are compared.

The world needs to be initialized for a substance. In this case, that
substance is microplastic, which has a default radius of 2.5e-5 m.

We will first show the version of the model without adjustments to match the 
excel version. This will show the differences between excel and R. Consequently,
we will show the version of the model that was adjusted to match the excel version. 
This adjusted version should yield the same results as the model in excel, except 
for rounding differences. 

## Run the model without adjustments

```{r Initialize,include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
substance <- "microplastic"
source("baseScripts/initWorld_onlyPlastics.R")

World$substance <- substance
```

```{r comparison of fluxes with test FALSE, message=FALSE, include=TRUE, echo=FALSE, warning=FALSE}

library(openxlsx)
library(tidyverse)
library (ggplot2)
library(plotly)

SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4plastics_rev009-1_",substance,".xlsx") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]
      
#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

```

## Compare k's

When comparing k's between R and excel, the goal is that the relative difference
is less than 0.001 for each k. The reason is that smaller
differences often are a result of differences in rounding values between
excel and R, and not the result of mistakes in calculations or different
input values. In this vignette two types of k's are compared: diagonal
k's and from-to k's.

At the time of this verification, some improvements were already made in
the R version versus the excel version. This meant that some k's differ
between R and excel, but not because the calculations or input values
are wrong. In order to still be able to compare the two versions, the
'Test' variable was created. This variable is a boolean, that can be
used to calculate some processes in R the same way as in excel for the
verification without removing the improvements that are made. When this
test variable was used and why will be explained below.

### Diagonal k's

Diagonal k's are k's that are on the diagonal of the k matrix. They are
calculated as the sum of all the k's leaving the subcompartment plus the
sum of the removal process k's (i.e. degradation or burial).

```{r Plot diagonal differences, echo=FALSE, warning=FALSE}

custom_theme <- function() {
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)#,
    #panel.background = element_rect(fill = "white", colour = "grey"),   # White background
    #panel.grid.major = element_line(colour = "grey")#,                   # Black major grid lines
    #panel.grid.minor = element_line(colour = "black")                    # Black minor grid lines
  )
} 

# Plot of differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = diff)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 1: Absolute differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Absolute diff")

# Plot of relative differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 2: Relative differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff")
```

#### Dry deposition
As can be seen in Figure 2 above, the only relative differences between the diagonal
k's in excel and R are in the air compartment. This is caused by a difference in 
the dry deposition process. In R, dry deposition is implemented in a new manner, 
according to the Loteur v2 reference guide. See v.2.2002 <https://www.rivm.nl/lotos-euros>.

When using Test=TRUE, the implementation of dry deposition in R is temporarily 
set to the old implementation that was used in excel. 

### From-to k's

```{r Plot k differences, echo=FALSE, warning=FALSE}
# "To" and "from" in one plot
ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") +
  ggtitle(paste0("Figure 3: Absolute differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(hjust = 1, size = 8),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 4: Relative differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```

```{r filter large differences, include=FALSE, echo=FALSE}
large_differences <- mergedkaas |>
  mutate(relDif = abs(relDif)) |>
  filter(relDif > 0.001) |>
  arrange(desc(relDif))
```

#### Dry deposition
As can be seen in Figure 4, the largest relative differences in k's between R and
excel are in the k's from air to soil and water. These differences can also be 
attributed to the different implementation of the dry deposition process between 
excel and R.

#### Thermal velocity heteroagglomeration
A smaller difference that was found between the k's in R and excel is the value 
for thermal velocity in heteroagglomeration. In excel, a constant value of 285K 
was used. In R, this value was updated by calculating the value specifically 
for the temperature of the current scale. 

## Run again with Test=TRUE

Now the value for the Test variable can be changed to TRUE, and the difference
in k's between excel and R can be tested again: 

```{r include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
World$SetConst(Test = "TRUE")
World$UpdateKaas(mergeExisting = FALSE)

ProcessNanoFunctions <- c("k_Advection", "k_Burial",
                         "k_HeteroAgglomeration.a", "k_HeteroAgglomeration.wsd",
                         "k_CWscavenging", "k_Degradation",
                         "k_DryDeposition", "k_Erosion", "k_Escape",
                         "k_Leaching", "k_Resuspension", "k_Runoff", "k_Sedimentation",
                        "k_WetDeposition")



SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  
```

```{r Plots test TRUE, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Plot of relative differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 5: Relative differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  #custom_theme()+
  labs(x = "from", y = "Relative diff")

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 6: Relative differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```

As can be seen in Figures 5 and 6, there are no k's with a relative difference
large than 0.001 between excel and R when Test=TRUE.

## Compare steady state emissions

The steady state masses in R and Excel were compared by calculating the
relative differences between the masses in R and Excel (Figure 7). The
figure shows that all masses between R and Excel relatively differ less
than 0.001.

```{r comparison of steady state emissions using SB1Solve, include= TRUE, echo= FALSE, warning=FALSE}

library(stringi)

World$NewSolver("SB1Solve")

emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(365*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=8,
                             colNames=TRUE,
                             rows=c(167,169))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything())  |>
  mutate(Species = case_when(
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "G" ~ "Gas",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "D" ~ "Dissolved",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "P" ~ "Large",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "A" ~ "Small",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "U" ~ "Unbound",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "S" ~ "Solid",
    TRUE ~ "Other"
  )) |>
  filter(EqMass != 0) |>
  mutate(Abbr = str_sub(Abbr, 1, -2))

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  filter(EqMass != 0)

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by=c("Abbr", "Species"), suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R) |>
  mutate(Abbr = case_when(
    Species == "Large" ~ paste0(Abbr,"P"),
    Species == "Small" ~ paste0(Abbr,"A"),
    Species == "Solid" ~ paste0(Abbr,"S"),
    TRUE ~ "Other"
  ))

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 7: Relative differences between steady state masses in excel and R (SB1solve) - ", substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) 
```