---
title: "Verification of SimpleBox4Plastics - spreadsheet versus R implementation"
author: "Valerie de Rijk, Anne Hids, Matthias Hof and Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.width = 14, fig.height = 8)
```

This vignette demonstrates the verification process of SimpleBox
implemented in R (version 2024.8.0) and in an Excel^TM^ spreadsheet
(xl4plastic v4.0.5). To do this the 1^st^ order rate constants (k's) and
steady state masses are compared between the two model implementations.
The differences should be negligible and only based on rounding errors.
In this case we choose a relative difference of the k's or masses
between the models to not exceed 0.1%.

# Verification method

```{r InitializeSBoo, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
substance <- "microplastic"
source("baseScripts/initWorld_onlyPlastics.R")
World$substance <- substance
```

The SBoo world is initialized for a substance. In this case, that
substance is `r World$fetchData("Substance")`, which is of class: `r World$fetchData("ChemClass")` and a default
radius of `r World$fetchData("RadS")*1e6` µm.

At release already improvements or developments have been implemented in
the R version of SimpleBox (SBoo) which are not implemented in Excel
which will result in differences between the spreadsheet and R
implementation. For this reason a TEST variable has been introduced to
the changed algorithms in R in order to verify the outcome of SimpleBox
in R with the original implementation in the spreadsheet version. So,
TEST variable is a boolean, that can be used to calculate some processes
in R the same way as in excel for the verification without removing the
improvements that are made. For this reason we show the verification in
two steps:

1.  Compare k's and steady state masses of SBoo with updates to the
    spreadsheet.

2.  Compare k's and steady state masses of adapted SBoo using TEST
    variable to the spreadsheet.

When comparing k's and steady state masses between SimpleBox in R and
Excel^TM^, the goal is that the relative difference is less than 0.1
percent for each k and steady state mass. The reason is that smaller
differences are almost inevitable due to differences in rounding values
between excel and R, and not the result of mistakes in calculations or
input values.

# Step 1. Compare SBoo (incl. updates) to spreadsheet

```{r comparison_1, message=FALSE, include=TRUE, echo=FALSE, warning=FALSE}

library(openxlsx)
library(tidyverse)
library (ggplot2)
library(plotly)

SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/20240815 SimpleBox4plastics.xlsx") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes
filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

# R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)
diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 

# filter out molecular (unbound) state
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]
      
#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

```

## preliminary matrix check
In the chunk below we check if our matrix is as large as we expect it to be. 

i) Matrix 
The SimpleBox model operates over 5 scales, 12 subcompartments and 4 different speciations. As such, we would expect our matrix to have 5 * 12 * 4 = 240 rows. However, some exceptions exist within our SB world: 

- The Regional and Continental scale do now have a deepocean layer (w3), reducing the matrix by 2 * 1 * 4 = 8 rows. 
- The Global Scales (Arctic, Moderate, Tropic) only have one type of soil instead of 3, reducing the matrix by 3 * 2 * 4 = 24 rows.
- The Global Scales (Arctic, Moderate, Tropic) only have one type of sediment instead of 3, reducing the matrix by 3 * 2 * 4 = 24 rows.
- The Global Scales (Arctic, Moderate, Tropic) only have one type of water instead of 3, reducing the matrix by 3 * 2 * 4 = 24 rows.
- Cloudwater (compartment) does not have any values, reducing our matrix by 5 * 1 * 1 = 5 rows. 

With these exceptions, we expect our matrix to be 240-85 = 155 rows. This script will stop running if this is not the case. 

ii) We expect all processes to be included in Processes4SpeciesTp.csv to also be calculated. Therefore, we compare the calculated unique processes with the processes defined in this csv. 

```{r check if matrix is 155x155, include=TRUE, echo=FALSE, warning=FALSE}}
dims <- dim(diagonal_R)

# Check if the matrix is 155x155
is_155x155 <- all(dims == c(155, 2))

# If the matrix is not 155x155, stop the execution
if (!is_155x155) {
  stop("The matrix does not have 155 rows. Execution stopped.")
}

# Continue with the rest of the code if the matrix is 155x155
matrix_info <- list(dimensions = dims, is_155x155 = is_155x155)


## check if all processes included
Processes4SpeciesTp <- read.csv("data/Processes4SpeciesTp.csv")
Functions <- Processes4SpeciesTp$Process[grepl("[a-z,A-Z]", Processes4SpeciesTp$Plastic)]
Functions <- paste0("k_", Functions)
Functions[!Functions %in% unique(World$kaas$process)]
world_processes <- unique(World$kaas$process)

# Check if Functions and unique world_processes match exactly
if (!setequal(Functions, world_processes)) {
  # Find elements in world_processes not in Functions
  missing_from_functions <- world_processes[!world_processes %in% Functions]
  
  # Find elements in Functions not in world_processes
  extra_in_functions <- Functions[!Functions %in% world_processes]
  
  # Print the mismatched elements
  cat("The following process names are missing from Functions:\n")
  print(missing_from_functions)
  cat("The following process names are extra in Functions:\n")
  print(extra_in_functions)
  
  # Stop the script
  stop("Mismatch found between Functions and unique(World$kaas$process).")
}

```

## Compare first order rate constants

Two approaches are taken to comparing the 'engine' matrix of k's. First
only the diagonal is taken and compared because this consists of all the
k's relevant for that 'from' compartment including the removal
processes. Second, the separate k's are compared per 'from' and 'to'
compartment.

In summary k's are compared using:

1.  The diagonal sum of k's (from + removal)

2.  The separate from-to k's

### Diagonal sum of 'from' k's

Diagonal k's are k's that are on the diagonal of the k matrix. They are
calculated as the sum of all the k's leaving the subcompartment plus the
sum of the removal process k's (i.e. degradation or burial).

```{r PlotsDiagonalk_1,fig.cap = "Figure 1: Relative differences sum of from-k's between R and Spreadsheet implementation of SimpleBox (Test=FALSE)", echo=FALSE, warning=FALSE}
custom_theme <- function() {
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)#,
    #panel.background = element_rect(fill = "white", colour = "grey"),   # White background
    #panel.grid.major = element_line(colour = "grey")#,                   # Black major grid lines
    #panel.grid.minor = element_line(colour = "black")                    # Black minor grid lines
  )
} 

# Plot of differences of the diagonals per compartment
# ggplot(sorted_diff, aes(x = from, y = diff)) +
#   geom_boxplot() +
#   ggtitle(paste0("Figure 1: Absolute differences removal k's between R and excel for ", substance)) +
#   geom_hline(yintercept = 0.001, color = "red") +
#   geom_hline(yintercept = -0.001, color = "red") +
#   custom_theme()+
#   labs(x = "from-Subcompartment Code", y = "Absolute diff")

# Plot of relative differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from-Subcompartment Code", y = "Relative diff")


```

### From-to k's

```{r PlotFromTok_1, echo=FALSE, warning=FALSE, fig.cap="Figure 2: Relative differences from-to k's between R and Spreadsheet implementation of SimpleBox (Test=FALSE)"}
# "To" and "from" in one plot
# ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
#   geom_point() +
#   scale_color_gradient(low = "green", high = "red", trans = "log10") +
#   ggtitle(paste0("Figure 3: Absolute differences k's between R and excel for ", substance)) +
#   theme(
#     axis.text.y = element_text(hjust = 1, size = 8),
#     axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
#     plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
#   )

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0(substance)) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```

#### Dry deposition {#sec-dry-deposition}

As can be seen in Figure 1 and 2 above, the only relative differences larger than 0.1% (e.g. for
the diagonal k's in excel and R) are in the air compartment. This is
caused by a difference in the implementation of the dry deposition process. In R, dry
deposition is implemented in a new manner, according to the Loteur v2
reference guide. See v.2.2002 <https://www.rivm.nl/lotos-euros>. Whereas in the spreadsheet implementation a slightly more simplified approach was taken.

Below Test=TRUE will be used, setting the implementation of dry deposition in R temporarily to the old implementation that was used in the Spreadsheet implementation of SimpleBox.

We can filter out the exact k's that have a relative difference larger than 0.1%:

```{r TableDifferences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
mergedkaas |>
  mutate(relDif = abs(relDif)) |>
  filter(relDif > 0.001) |>
  arrange(desc(relDif)) |> knitr::kable(digits = 30)
```

### Steadystate mass
```{r PlotSteadyState_1, include= TRUE, echo= FALSE, warning=FALSE, fig.cap="Figure 3: Relative differences in steady state mass per compartment between R (SB1solve) and Spreadsheet implementation of SimpleBox (Test=FALSE)"}
library(stringi)
World$NewSolver("SB1Solve")
emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s
emissions <- emissions |>
  mutate(Emis = Emis*1000/(365*24*60*60))

SSsolve.R <- World$Solve(emissions)
SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=8,
                             colNames=TRUE,
                             rows=c(167,169))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything())  |>
  mutate(Species = case_when(
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "G" ~ "Gas",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "D" ~ "Dissolved",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "P" ~ "Large",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "A" ~ "Small",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "U" ~ "Unbound",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "S" ~ "Solid",
    TRUE ~ "Other"
  )) |>
  filter(EqMass != 0) |>
  mutate(Abbr = str_sub(Abbr, 1, -2))

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  filter(EqMass != 0)

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by=c("Abbr", "Species"), suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R) |>
  mutate(Abbr = case_when(
    Species == "Large" ~ paste0(Abbr,"P"),
    Species == "Small" ~ paste0(Abbr,"A"),
    Species == "Solid" ~ paste0(Abbr,"S"),
    TRUE ~ "Other"
  ))

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) 



```

The differences in k's drives the model output: the steady state mass. So a final check is to see how much the steady state masses differ between both implementations of SimpleBox (Figure 3). From this it is clear that there  are differences up to `r merged_SS_SB1$reldiff |> max() |> round(digits=3)*100`%. This is larger than the level we can consider negligible.

# Step 2. Compare SBoo and Spreadsheet excluding updates (Test=TRUE)
The verification's goal is to make sure no mistakes are made in porting SimpleBox from the spreadsheet implementation to R. For this reason the Test variable was included in algorithms that already implemented changes for specific variables or processes in SimpleBox. With the Test variable changed to TRUE the
difference in k's and steady state masses is shown again in relation to the intended 0.1% verification level.

```{r comparison_2, message=FALSE, warning=FALSE, include=FALSE}
World$SetConst(Test = "TRUE")
World$UpdateKaas(mergeExisting = FALSE)

ProcessNanoFunctions <- c("k_Advection", "k_Burial",
                         "k_HeteroAgglomeration.a", "k_HeteroAgglomeration.wsd",
                         "k_CWscavenging", "k_Degradation",
                         "k_DryDeposition", "k_Erosion", "k_Escape",
                         "k_Leaching", "k_Resuspension", "k_Runoff", "k_Sedimentation",
                        "k_WetDeposition")



SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  
```

## Compare first order rate constants

```{r PlotDiagonalk_2, fig.cap = "Figure 4: Relative differences sum of from-k's between R and Spreadsheet implementation of SimpleBox (Test=TRUE)", include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Plot of relative differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff")

```


```{r PlotFromTok_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Figure 5: Relative differences from-to k's between R and Spreadsheet implementation of SimpleBox (Test=TRUE)"}
ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", limits = c(0, 0.001)) +
  ggtitle(paste0(substance)) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```

As can be seen in Figures 4 and 5, there are no k's with a relative
difference large than 0.1% between excel and R when Test=TRUE. Several k's have a relative difference between 0.1 and 0.01%. These are related to:
Heteroagglomeration in porous media between particulate (S) and large heteroagglomerate/attached (P) species.
Deposition from air to water/soil
Erosion from soil to water

```{r TableDifferences_2, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
mergedkaas |>
  mutate(relDif = abs(relDif)) |>
  filter(relDif > 0.0001) |>
  arrange(desc(relDif)) |> knitr::kable(digits = 30)
```


## Steadystate mass
```{r PlotSteadyState_2, include= TRUE, echo= FALSE, warning=FALSE, fig.cap="Figure 6: Relative differences in steady state mass per compartment between R (SB1solve) and Spreadsheet implementation of SimpleBox (Test=TRUE)"}

library(stringi)

World$NewSolver("SB1Solve")

emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

# MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(365*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=8,
                             colNames=TRUE,
                             rows=c(167,169))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything())  |>
  mutate(Species = case_when(
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "G" ~ "Gas",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "D" ~ "Dissolved",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "P" ~ "Large",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "A" ~ "Small",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "U" ~ "Unbound",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "S" ~ "Solid",
    TRUE ~ "Other"
  )) |>
  filter(EqMass != 0) |>
  mutate(Abbr = str_sub(Abbr, 1, -2))

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  filter(EqMass != 0)

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by=c("Abbr", "Species"), suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R) |>
  mutate(Abbr = case_when(
    Species == "Large" ~ paste0(Abbr,"P"),
    Species == "Small" ~ paste0(Abbr,"A"),
    Species == "Solid" ~ paste0(Abbr,"S"),
    TRUE ~ "Other"
  ))

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) 

```

To test if the small differences (<0.1%) in first order rate constants is negligible (Figures 4 and 5), the steady state masses should also not differ by more than 0.1% between the R and Spreadsheet implementations of SimpleBox. This is indeed the case (Figure 6) as the max difference in now only `r merged_SS_SB1$reldiff |> abs() |> max() |> round(digits=4)*100`%. This proves that the port of SimpleBox4Plastics to R is successful in reproducing the results from the original spreadsheet implementation.
