---
title: "Verification of SimpleBox - spreadsheet versus R implementation for metals"
author: "Anne Hids, Valerie de Rijk, Matthis Hof and Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# knitr::knit_meta()
# knitr::opts_chunk$set(echo = TRUE)
# projectRoot <- paste(getwd(), "..", "..", "..", "..", sep = "/")
# knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.width = 14, fig.height = 8)
```

This vignette demonstrates the verification process of SimpleBox
implemented in R (version 2024.8.0) and in an Excel^TM^ spreadsheet
(xl v4.0.5). To do this the 1^st^ order rate constants (k's) and
steady state masses are compared between the two model implementations.
The differences should be negligible and only based on rounding errors.
In this case we choose a relative difference of the k's or masses
between the models to not exceed 0.1%.

# Verification method
```{r Initialize, message=FALSE, warning=FALSE, include=FALSE}
# Create a list with the names of substances
Potential_substances <- c("1-aminoanthraquinone", # no class
                          "1-HYDROXYANTHRAQUINONE", # acid
                          "1-Hexadecanamine, N,N-dimethyl-", # base
                          "1-Chloro-2-nitro-propane", # neutral
                          "Sb(III)" # metal
                          ) 
              
substance <- Potential_substances[5]

source("baseScripts/initWorld_onlyMolec.R")
```


The SBoo world is initialized for a substance. In this case, that
substance is `r World$fetchData("Substance")`, which is of class: `r World$fetchData("ChemClass")`.

At release already improvements or developments have been implemented in
the R version of SimpleBox (SBoo) which are not implemented in Excel
which will result in differences between the spreadsheet and R
implementation. For this reason a TEST variable has been introduced to
the changed algorithms in R in order to verify the outcome of SimpleBox
in R with the original implementation in the spreadsheet version. So,
TEST variable is a boolean, that can be used to calculate some processes
in R the same way as in excel for the verification without removing the
improvements that are made. For this reason we show the verification in
two steps:

1.  Compare k's and steady state masses of SBoo with updates to the
    spreadsheet.

2.  Compare k's and steady state masses of adapted SBoo using TEST
    variable to the spreadsheet.

When comparing k's and steady state masses between SimpleBox in R and
Excel^TM^, the goal is that the relative difference is less than 0.1
percent for each k and steady state mass. The reason is that smaller
differences are almost inevitable due to differences in rounding values
between excel and R, and not the result of mistakes in calculations or
input values.

# Step 1. Compare SBoo (incl. updates) to spreadsheet

```{r comparison_1, echo=FALSE, warning=FALSE}
#Comparing K's from R model to K's from excel model
library(openxlsx)
library(tidyverse)

ProcessMolFunctions <- c("k_Adsorption", "k_Advection", "k_Burial",
                          "k_Degradation", "k_Deposition", "k_Desorption",
                          "k_Erosion", "k_Escape", 
                         "k_Leaching", "k_Resuspension", "k_Runoff", "k_Sedimentation", 
                         "k_Volatilisation")

#read in K matrix from excel
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.05_20240816_",substance,".xlsm") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names
SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", 
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale]))

# Issue that compartments sediment and soil at global scale in excel have sd and s as acronyms instead of sd2 and s1
kaas <- 
  kaas |> 
  mutate(from = 
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") & 
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"), 
                          str_replace_all(from, c("sd2"="sd","s1"="s")), 
                          from)) |> 
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") & 
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

diagonal_R <- 
  aggregate(k ~ from, data = kaas, FUN = sum) #R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(diagonal_R, diagonal_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
merged_diagonal$reldif <- merged_diagonal$diff/merged_diagonal$k_R

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  
```

## Compare first order rate constants
When comparing k's between R and excel, the goal is that the relative difference is less
than 1 percentile for each k. The reason is that smaller differences often are 
a result of differences in rounding values between excel and R, and not the result
of mistakes in calculations or different input values. In this vignette two types of 
k's are compared: diagonal k's and from-to k's.

At the time of this verification, some improvements were already made in the R 
version versus the excel version. This meant that some k's differ between R and
excel, but not because the calculations or input values are wrong. In order to 
still be able to compare the two versions, the 'Test' variable was created. This
variable is a boolean, that can be used to calculate some processes in R the same 
way as in excel for the verification without removing the improvements that are 
made. When this test variable was used and why will be explained below. 

### Diagonal sum of 'from' k's

Diagonal k's are k's that are on the diagonal of the k matrix. They are calculated
as the sum of all the k's leaving the subcompartment plus the sum of the removal process
k's (i.e. degradation or burial). 

```{r PlotsDiagonalk_1,fig.cap = "Figure 1: Relative differences sum of from-k's between R and Spreadsheet implementation of SimpleBox (Test=FALSE)", echo=FALSE, warning=FALSE}

custom_theme <- function() {
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)#,
    #panel.background = element_rect(fill = "white", colour = "grey"),   # White background
    #panel.grid.major = element_line(colour = "grey")#,                   # Black major grid lines
    #panel.grid.minor = element_line(colour = "black")                    # Black minor grid lines
  )
} 

# Plot of differences of the diagonals per compartment
# ggplot(merged_diagonal, aes(x = from, y = diff)) +
#   geom_boxplot() +
#   ggtitle(paste0("Figure 1: Absolute differences removal k's between R and excel for ", substance)) +
#   geom_hline(yintercept = 0.001, color = "red") +
#   geom_hline(yintercept = -0.001, color = "red") +
#   custom_theme()+
#   labs(x = "from", y = "Absolute diff")

# Plot of relative differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff")
```

As can be seen in Figure 1, relative differences larger than 0.1% are in the lake and
sediment subcompartments.

```{r DiagonalDifferences, include=FALSE, echo=FALSE}
merged_diagonal |>
  mutate(reldif = abs(reldif)) |>
  filter(reldif > 0.001) |>
  arrange(desc(reldif)) |> knitr::kable(digits = 30)
```

#### Lake difference

The relative difference in lake removal rate is caused by lake sedimentation
being included in R but not in excel. To make an accurate comparison between R 
and excel, the Test variable was used to exclude flows from lake to lakesediment in 
the processes k_Sedimentation and k_Adsorption.

```{r LakeTest, include=FALSE, echo=FALSE}
lake <- kaas |>
  filter(fromSubCompart == "lake") |>
  filter(toSubCompart == "lake")

lake2 <- kaas |>
  filter(fromSubCompart == "lake") |>
  filter(fromScale == "Continental")
```

#### Settling velocity

The difference in the diagonal k's for sedimentation comes from a difference in k's for the 
sedimentation and resuspension processes. 

This is caused by the use of different formulas to calculate settling
velocity between excel and R. In excel, settling velocity is always calculated as: 

`SetVel <- 2.5/(24*3600)`

While in R, an improved version of this formula is used: 

`SetVel <- 2*(radius^2*(rhoParticle-rhoWater)*GN) / (9*DynViscWaterStandard)`

Using the Test variable, the settling velocity formula is temporarily changed to 
the formula used in excel for the resuspension and sedimentation processes (K_resuspension
and k_Sedimentation). This solves the differences in diagonal k's related to 
resuspension or sedimentation.

### From-to k's

```{r PlotFromTok_1, echo=FALSE, warning=FALSE, fig.cap="Figure 2: Relative differences from-to k's between R and Spreadsheet implementation of SimpleBox (Test=FALSE)"}
# "To" and "from" in one plot
# ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
#   geom_point() +
#   scale_color_gradient(low = "green", high = "red", trans = "log10") +
#   ggtitle(paste0("Figure 3: Absolute differences k's between R and excel for ", substance)) +
#   theme(
#     axis.text.y = element_text(size = 12),
#     axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
#     plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
#   )

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 4: Relative differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```


```{r kDifferences, include=FALSE, echo=FALSE}
mergedkaas |>
  mutate(relDif = abs(relDif)) |>
  filter(relDif > 0.001) |>
  arrange(desc(relDif)) |> knitr::kable()
```

#### Sedimentation and resuspension
As can be seen in Figure 2 and the table above, the k's that have a relative difference larger
than 0.1% go from water to sediment or from sediment to water. Changing 
the formula for calculating settling velocity when Test = TRUE for the sedimentation
and resuspension processes (explained above under 'Settling velocity') also solves these differences.

### Steadystate mass
```{r PlotSteadyState_1, include= TRUE, echo= FALSE, warning=FALSE, fig.cap="Figure 3: Relative differences in steady state mass per compartment between R (SB1solve) and Spreadsheet implementation of SimpleBox (Test=FALSE)"}
library(stringi)

World$NewSolver("SB1Solve")

emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(MW*365*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=11,
                             colNames=TRUE,
                             rows=c(44,45))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE, X2)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything()) 

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  mutate(SubCompart = str_replace(SubCompart, "lakesediment", "lake")) |>
  mutate(Abbr = str_replace(Abbr, "sd0R", "w0R")) |>
  mutate(Abbr = str_replace(Abbr, "sd0C", "w0C")) |>
  group_by(Scale, SubCompart, Species, Abbr) |>
  summarise(EqMass = sum(EqMass))

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by="Abbr", suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R)

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme()



```

The differences in k's drives the model output: the steady state mass. So a final check is to see how much the steady state masses differ between both implementations of SimpleBox (Figure 3). From this it is clear that there  are differences up to `r merged_SS_SB1$reldiff |> abs() |> max() |> round(digits=3)*100`%.

# Step 2. Compare SBoo and Spreadsheet excluding updates (Test=TRUE)

Now the value for the Test variable can be changed to TRUE, and the difference
in k's between excel and R can be tested again: 

```{r comparison_2, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
World$SetConst(Test = "TRUE")
World$UpdateKaas(mergeExisting = FALSE)

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names
SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", 
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale]))

# Issue that compartments sediment and soil at global scale in excel have sd and s as acronyms instead of sd2 and s1
kaas <- 
  kaas |> 
  mutate(from = 
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") & 
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"), 
                          str_replace_all(from, c("sd2"="sd","s1"="s")), 
                          from)) |> 
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") & 
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

diagonal_R <- 
  aggregate(k ~ from, data = kaas, FUN = sum) #R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(diagonal_R, diagonal_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
merged_diagonal$reldif <- merged_diagonal$diff/merged_diagonal$k_R

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

```

## Compare first order rate constants

```{r PlotDiagonalk_2, fig.cap = "Figure 4: Relative differences sum of from-k's between R and Spreadsheet implementation of SimpleBox (Test=TRUE)", include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Plot of relative differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff")

```


```{r PlotFromTok_2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Figure 5: Relative differences from-to k's between R and Spreadsheet implementation of SimpleBox (Test=TRUE)"}
ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0( substance)) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```


As can be seen in Figures 4 and 5, the temporary changes made using the Test variable 
solved the large differences in k's between R and excel for both the diagonal and the 
from-to k's. All relative differences are now smaller than 1 percentile. 

## Steadystate mass

```{r PlotSteadyState_2, include= TRUE, echo= FALSE, warning=FALSE, fig.cap="Figure 6: Relative differences in steady state mass per compartment between R (SB1solve) and Spreadsheet implementation of SimpleBox (Test=TRUE)"}

library(stringi)

World$NewSolver("SB1Solve")

emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(MW*365*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=11,
                             colNames=TRUE,
                             rows=c(44,45))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE, X2)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything()) 

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  mutate(SubCompart = str_replace(SubCompart, "lakesediment", "lake")) |>
  mutate(Abbr = str_replace(Abbr, "sd0R", "w0R")) |>
  mutate(Abbr = str_replace(Abbr, "sd0C", "w0C")) |>
  group_by(Scale, SubCompart, Species, Abbr) |>
  summarise(EqMass = sum(EqMass))

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by="Abbr", suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R)

#print("Difference in emissions between R and Excel")

#knitr::kable(merged_SS_SB1, format="markdown")

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0(substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme()
```

To test if the small differences (<0.1%) in first order rate constants is negligible (Figures 4 and 5), the steady state masses should also not differ by more than 0.1% between the R and Spreadsheet implementations of SimpleBox. This is indeed the case (Figure 6) as the max difference in now only `r merged_SS_SB1$reldiff |> abs() |> max() |> round(digits=5)*100`%. This proves that the port of SimpleBox to R is successful in reproducing the results from the original spreadsheet implementation for chemicals of class `r World$fetchData("ChemClass")`.