---
title: "Particulate verification"
author: "Anne Hids & Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.width = 14, fig.height = 8)
```

This vignette demonstrates the verification process of the particulate
version of Simplebox, in this case for a particle that is not a microplastic.
First, the k's are compared between R and excel,
and consequently the steady state masses are compared. This is done for
5 molecular substances; each of a different chemical class (no class,
acid, base, neutral and metal). The reason that the verification is performed for 
each of these classes is that some processes differ per class.

First, the world needs to be initialized for a substance. In this case, that substance
is nAg 10nm, which is a particle. 

We will first show the version of the R-code that has been adjusted to match the Excel Version. This means that the code is initialized with the parameter Test = TRUE. As such, this code should match (except for rounding differences) the results of the Excel version.

## Test = TRUE
```{r Initialize,include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
substance <- "nAg_10nm"

source("baseScripts/initWorld_onlyParticulate.R")
```

```{r warning=FALSE}
library(openxlsx)
library(tidyverse)
library (ggplot2)
library(plotly)

ProcessNanoFunctions <- c("k_Advection", "k_Burial",
                         "k_HeteroAgglomeration.a", "k_HeteroAgglomeration.wsd",
                         "k_CWscavenging", "k_Degradation",
                         "k_DryDeposition", "k_Erosion", "k_Escape",
                         "k_Leaching", "k_Resuspension", "k_Runoff", "k_Sedimentation",
                        "k_WetDeposition")

SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4plastics_rev009-1_",substance,".xlsx") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

length(SBexcel.K$k)

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)
unique(kaas$fromScale)
unique(kaas$fromSubCompart)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
print(diagonal_R$from)
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]
      
#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1

filtered_R$from[!filtered_R$from %in% filtered_excel$from]
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R
filtered_excel$from[!filtered_excel$from %in% filtered_R$from]

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))



SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

```

## Compare k's
When comparing k's between R and excel, the goal is that the difference is less
than 1 percentile for each k. The reason is that smaller differences often are 
a result of differences in rounding values between excel and R, and not the result
of mistakes in calculations or different input values. In this vignette two types of 
k's are compared: diagonal k's and from-to k's.

At the time of this verification, some improvements were already made in the R 
version versus the excel version. This meant that some k's differ between R and
excel, but not because the calculations or input values are wrong. In order to 
still be able to compare the two versions, the 'Test' variable was created. This
variable is a boolean, that can be used to calculate some processes in R the same 
way as in excel for the verification without removing the improvements that are 
made. When this test variable was used and why will be explained below. 

### Diagonal k's

Diagonal k's are k's that are on the diagonal of the k matrix. They are calculated
as the sum of all the k's leaving the subcompartment plus the sum of the removal process
k's (i.e. degradation or burial). 

```{r Plot diagonal differences, echo=FALSE, warning=FALSE}

custom_theme <- function() {
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
} 

# Plot of differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = diff)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 1: Absolute differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Absolute diff")

# Plot of relative differences of the diagonals per compartment
ggplot(sorted_diff, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 2: Relative differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        aspect.ratio = 1/3) 
```



```{r filter large diagonal differences, include=FALSE, echo=FALSE}
diagonal_diffs <- sorted_diff |>
  mutate(reldif = abs(reldif)) |>
  filter(reldif > 0.001) |>
  arrange(desc(reldif))
```


### From-to k's

```{r Plot k differences, echo=FALSE, warning=FALSE}
# "To" and "from" in one plot
ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") +
  ggtitle(paste0("Figure 3: Absolute differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 4: Relative differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )
```

```{r filter large differences, include=FALSE, echo=FALSE}
large_differences <- mergedkaas |>
  mutate(relDif = abs(relDif)) |>
  filter(relDif > 0.001) |>
  arrange(desc(relDif))
```

As can be seen in Figure 4, there are no k's with a relative difference large than 1 percentile between excel and R. 


## Run again with an altered copy of the excel file

An altered copy of the Excel file used for comparison was created. In this file, 
all kdeg rates were changed from 0 to 1e-20.
```{r Test TRUE, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4plastics_rev009-1_",substance,"_copy_deg.xlsx") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

length(SBexcel.K$k)

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)
unique(kaas$fromScale)
unique(kaas$fromSubCompart)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
print(diagonal_R$from)
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]
      
#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1

filtered_R$from[!filtered_R$from %in% filtered_excel$from]
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R
filtered_excel$from[!filtered_excel$from %in% filtered_R$from]

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
merged_diagonal$reldif <- merged_diagonal$diff/merged_diagonal$k_R

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

# Plot of relative differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = reldif)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 5: Relative differences removal k's between R and excel for ", substance)) +
  geom_hline(yintercept = 0.001, color = "red") +
  geom_hline(yintercept = -0.001, color = "red") +
  custom_theme()+
  labs(x = "from", y = "Relative diff")

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 6: Relative differences k's between R and excel for ", substance)) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )

```

As can be seen in Figures 5 and 6, the changes made to the copy of the excel file 
solved the large differences in k's between R and excel for the diagonal k's. All relative differences are now smaller than 1 percentile. 

## Compare steady state emissions

The steady state masses in R and Excel were compared by calculating the relative differences
between the masses in R and Excel (Figure 7). The figure shows that all masses between R and Excel relatively differ less than 1 percentile. 

```{r comparison of steady state emissions using SB1Solve, include= TRUE, echo= FALSE, warning=FALSE}

library(stringi)

World$NewSolver("SB1Solve")

emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(MW*365*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=8,
                             colNames=TRUE,
                             rows=c(167,168))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything())  |>
  mutate(Species = case_when(
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "G" ~ "Gas",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "D" ~ "Dissolved",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "P" ~ "Large",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "A" ~ "Small",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "U" ~ "Unbound",
    substr(Abbr, nchar(Abbr), nchar(Abbr)) == "S" ~ "Solid",
    TRUE ~ "Other"
  )) |>
  filter(EqMass != 0) |>
  mutate(Abbr = str_sub(Abbr, 1, -2))

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale])) |>
  filter(EqMass != 0)

merged_SS_SB1 <- merge(SSsolve.R, SSsolve.excel, by=c("Abbr", "Species"), suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R) |>
  mutate(Abbr = case_when(
    Species == "Large" ~ paste0(Abbr,"P"),
    Species == "Small" ~ paste0(Abbr,"A"),
    Species == "Solid" ~ paste0(Abbr,"S"),
    TRUE ~ "Other"
  ))

#print("Difference in emissions between R and Excel")

#knitr::kable(merged_SS_SB1, format="markdown")

#Diff per "to" compartment
ggplot(merged_SS_SB1, aes (x = Abbr, y = reldiff)) +
  geom_boxplot() +
  ggtitle(paste0("Figure 7: Relative differences between steady state masses in excel and R (SB1solve) - ", substance)) +
  geom_hline(yintercept = 0.001, color="red") +
  geom_hline(yintercept = -0.001, color="red") +
  custom_theme() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        aspect.ratio = 1/3) 
```

## Test = FALSE
Now, we will observe the differences made to the code for the R-version by returning to Test = False. This excludes any adjustments to the DragCoefficient options for the settling velocity, which are highlighted in the vignette about Sedimentation. 
The following adjustments are then initialized: 

- kdis is not included

- Dry Deposition is implemented in a new manner, according to the Loteur v2 reference guide. See v.2.2002 https://www.rivm.nl/lotos-euros

- Thermal velocity in heteroagglomeration is now calculated according to the temperature of the scale, instead of a constant value of T = 285K. 

As can be observed from Figure 8 and 9, Relative differences become large (>0.8). Most of these differences are attributed to the adjustments in the Dry Deposition implementation. 
```{r comparison of fluxes with test FALSE, include=TRUE, echo=FALSE, warning=FALSE}
World$SetConst(Test = "FALSE")
World$UpdateKaas(mergeExisting = FALSE)
SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

length(SBexcel.K$k)

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)
unique(kaas$fromScale)
unique(kaas$fromSubCompart)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0", 
                "cloudwater" = "cw")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

accronym_map3 <- c("Dissolved" = "D", 
                   "Gas" = "G", 
                   "Large" = "P", 
                   "Small" = "A",
                   "Solid" = "S", 
                   "Unbound" = "U")

# kaas |> filter(fromScale == "Tropic" & process == "k_Degradation") |> print(n=50)

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale], 
                            accronym_map3[fromSpecies]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale], 
                           accronym_map3[toSpecies]))

kaas <-
  kaas |>
  mutate(from =
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                          str_replace_all(from, c("sd2"="sd","s1"="s")),
                          from)) |>
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

#filter out dissolved and gas processes

filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]

#filter out molecular (unbound) state
#R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

diagonal_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
print(diagonal_R$from)
filtered_R <- diagonal_R[!endsWith(diagonal_R$from, "U"), ]
      
#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1

filtered_R$from[!filtered_R$from %in% filtered_excel$from]
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R
filtered_excel$from[!filtered_excel$from %in% filtered_R$from]

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
  mutate(reldif  = abs(diff/k_R))



SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)  

ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") +
  ggtitle(paste0("Figure 8: Absolute differences k's between R and excel for ", substance, "with Test = FALSE")) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )

ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") +
  ggtitle(paste0("Figure 9: Relative differences k's between R and excel for ", substance," with Test = FALSE")) +
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10), 
    plot.margin = margin(t = 10, r = 10, b = 30, l = 10)
  )


```
