---
title: "Setup Simplebox PFAS Tjebbe"
output: html_document
date: "2025-01-06"
author: "Tjebbe Janson"
editor_options: 
  chunk_output_type: console
---

## Setup libraries and working directory
```{r setup, include=FALSE, warning=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")

#conflict_prefer("select", "dplyr")
#conflict_prefer("filter", "ggdag")
```

```{R Packages laden}
print(getwd())
check_and_install <- function(package) {
  tryCatch({
    # Load the package
    library(package, character.only = TRUE)
    message(paste("Package", package, "is already installed and loaded."))
  }, error = function(e) {
    # If an error occurs, install the package
    message(paste("Package", package, "is not installed. Installing now..."))
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
    message(paste("Package", package, "has been successfully installed and loaded."))
  })
}

# Install the required packages
check_and_install("patchwork")
check_and_install("ggplot2")
check_and_install("tidyverse")
check_and_install("constants")
check_and_install("deSolve")
check_and_install("knitr")
check_and_install("dplyr")
check_and_install("zyp")
check_and_install("slider")
check_and_install("readxl")
check_and_install("Kendall")
check_and_install("ggpubr")
check_and_install("ncdf4")
check_and_install("raster")
check_and_install("sf")
check_and_install("exactextractr")
check_and_install("here")
check_and_install("reshape2")
check_and_install("scales")
library(dplyr)
library(here)
library(lhs)
library(tidyverse)
library(treemapify)
library(gridExtra)
```

##Stof kiezen, Wereld aanpassen
```{r Substance list, warning=FALSE, include=TRUE}
#Stof instellingen, handmatig 
substances <- read.csv(here("data", "Substances.csv"))
gekozen_stoffen = c("Perfluorobutane_sulfonic", "Perfluorobutanoic", 
                    "Perfluoropentanoic", "Perfluoroheptanoic", 
                    "Perfluorooctanoic", "Trifluoroacetic")
#substances <- filter(substances, Substance %in% gekozen_stoffen)

substance <- "Trifluoroacetic"

chemclass <- substances |>
  filter(Substance == substance) |>
  dplyr::select(ChemClass)
chemclass <- chemclass$ChemClass

#World Module
if(substance == "microplastic"){
  source("baseScripts/initWorld_onlyPlastics.R")
} else if (chemclass == "particle") {
  source("baseScripts/initWorld_onlyParticulate.R")
} else {
  source("baseScripts/initWorld_onlyMolec.R")
}
World$substance <- substance

#Model aanpassen met AdjustModel{catchment}.R 
catchment <- 'Rhine'
if(catchment == "Rhine"){
  source("vignettes/CaseStudies/PFAS Tjebbe/Script/AdjustModelRhine.R")
} else if (catchment == "Meuse") {
  source("vignettes/CaseStudies/PFAS Tjebbe/Script/AdjustModelMeuse.R")
}

#Tijds instelling
runtime <- 140 #years
tmax <- runtime*365.25*24*60*60 #in seconds
nTIMES <- runtime+1 #Number of timesteps, e.g. yearly, 2-year



```

##Emissie dataframe maken met EmissionModel.R
```{r}
#Emissies
Wereldwijd = TRUE
catchment = 'Rhine'
scenario = 1
source("vignettes/CaseStudies/PFAS Tjebbe/Script/EmissionModel.R")

#Variabelen:RAINrate, degradation, Runoff??, Infiltratie?? tdry/twet??


```

##Informatie over afkortingen, scales en oppervlaktes
```{r Scales}
scales <- data.frame(Scale = c("Arctic", "Continental", "Moderate", "Regional", "Tropic"), Abbreviation = c("A", "C", "M", "R", "T"))

subcompartments <- read.csv("data/SubCompartSheet.csv") |>
  dplyr::select(SubCompartName, AbbrC) |>
  rename(Abbreviation = AbbrC) |>
  rename(SubCompartment = SubCompartName)

species <- read.csv("data/SpeciesSheet.csv") |>
  dplyr::select(Species, AbbrP) |>
  rename(Abbreviation = AbbrP)

substancesubcompartments <- read.csv("data/SubstanceCompartments.csv")

kable(scales)
kable(subcompartments)
kable(species)

# Area simplebox is ongeveer 51% van aarde
Area_simplebox <- World$fetchData("TotalArea") %>%
  dplyr::select(TotalArea) %>%
  summarize(sum = sum(TotalArea, na.rm=TRUE)) %>%
  pull(sum)
print(Area_simplebox * 100 /5.1e+14)

# Area continental
print(World$fetchData("Area"))

# Area regional

```

##Eventueel stof eigenschappen aanpassen
```{r}
# pas_stofeigenschap_aan <- function(value, variable){
#   #print var
#   print(paste("adjusting var: ", variable, "...."))
#   print(paste("Old var:", World$fetchData(variable)))
#   #Make df
#   df = data.frame(
#   Waarde = value,
#   varName = (variable)
#   )
#   #Mutating
#   World$mutateVars(df)
#   #print new var and updatedirty
#   print(paste("Updated var: ", World$fetchData(variable)))
#   World$UpdateDirty(variable) 
# }
# 
# # Stof eigenschappen uit 'Data pilot study stock assessments_QSAR estimates' - Joris Meesters?
# #MolecularWeight <- pas_stofeigenschap_aan()
# MeltingPoint <- pas_stofeigenschap_aan(136.16, "Tm")
# WaterSolubility <- pas_stofeigenschap_aan(253.8, "Sol25")  #0.4813
# DissociationConstant <- pas_stofeigenschap_aan(-4.2, "pKa") #-4.2
# VapourPressure <- pas_stofeigenschap_aan(0.0000475, "Pvap25")    #19.3
# Kaw <- pas_stofeigenschap_aan(0.57, "Kaw25") #0.57                
# Kow <- pas_stofeigenschap_aan(1, "Kow") #4.81                  
# Ksw <- pas_stofeigenschap_aan(1.03, "Ksw") #1.03 obv kow          


# World$fetchData("kdeg")
# degradation = data.frame(
#   SubCompart = c("air", "cloudwater", "agriculturalsoil", 
#                  "naturalsoil", "othersoil", "deepocean", 
#                  "river", "sea", "lake"),
#   Waarde = c(2.2e-08, 2.2e-08, 1.1e-10, 1.1e-10, 1.1e-10, 2.2e-10, 2.2e-10, 2.2e-10, 2.2e-10),
#   varName = ("kdeg")
# )
# 
# degradation_test = data.frame(
#   SubCompart = c("air", "cloudwater", "agriculturalsoil", 
#                  "naturalsoil", "othersoil", "deepocean", 
#                  "river", "sea", "lake"),
#   Waarde = c(2.2e-4, 2.2e-4, 1.1e-5, 1.1e-5, 1.1e-5, 2.2e-5, 2.2e-5, 2.2e-5, 2.2e-5),
#   varName = ("kdeg")
# )
# 
# World$mutateVars(degradation)
# World$fetchData("kdeg")
# World$UpdateDirty("kdeg")

#kdeg lager = snellere degradatie
```

##Solven
``` {r Solve}
#steadystate
#T/y
#Simpel scenario met de continentale emissies relatief tot oppervlakte verschillen
emissions_simple <- data.frame(Abbr = c("aRU", "s2RU", "s1RU", "s3RU", "w0RU", "w1RU", "w2RU",
                                        "aCU", "s2CU", "s1CU", "s3CU", "w0CU", "w1CU", "w2CU"), 
                               Emis = c(2, 0.5, 0.5, 0.5, 0.1, 0.1, 0.10, 
                                        124.5, 15.38271, 15.38271, 15.38271, 95, 95, 95))
emissions_simple <- emissions_simple |>
  mutate(Emis = Emis*1000/(365*24*60*60))

World$NewSolver("SteadyStateSolver")
World$Solve(emissions = emissions_simple)

#New solver
#World$NewSolver("DynamicSolver")
#World$Solve(emissions = emissions, tmax = tmax, nTIMES = nTIMES)


solution_tfa <- World$Masses()
emission_tfa <- World$Emissions()
concentration_tfa <- World$Concentration()
```

``` {r Plot staafdiagram}
##Staaf diagram maken 

#SteadyStates omzetten naar percentages totale massa
#regionaal
solutions_pfoa <- solution_pfoa %>%
  filter(grepl("R", Abbr)) %>%
  mutate(percentage_pfoa = 100 * Mass_kg /sum(Mass_kg)) 
solutions_tfa <- solution_tfa %>%
  filter(grepl("R", Abbr)) %>%
  mutate(percentage_tfa = 100 * Mass_kg /sum(Mass_kg))
solutions_reg <- left_join(solutions_pfoa, solutions_tfa, by = "Abbr", suffix = c("_pfoa", "_tfa"))
#Continentaal
solutions_pfoa <- solution_pfoa %>%
  filter(grepl("C", Abbr)) %>%
  mutate(percentage_pfoa = 100 * Mass_kg /sum(Mass_kg)) 
solutions_tfa <- solution_tfa %>%
  filter(grepl("C", Abbr)) %>%
  mutate(percentage_tfa = 100 * Mass_kg /sum(Mass_kg))
solutions_con <- left_join(solutions_pfoa, solutions_tfa, by = "Abbr", suffix = c("_pfoa", "_tfa"))

#Pivot
solutions_regional <- solutions_reg %>%
  dplyr::select(Abbr, percentage_pfoa, percentage_tfa) %>%
  pivot_longer(
    cols = c(percentage_pfoa, percentage_tfa),
    names_to = "stof",
    values_to = "waarde"
  )
solutions_continental <- solutions_con %>%
  dplyr::select(Abbr, percentage_pfoa, percentage_tfa) %>%
  pivot_longer(
    cols = c(percentage_pfoa, percentage_tfa),
    names_to = "stof",
    values_to = "waarde"
  )

#concentratie dataframes 
concentrations_reg <- concentration_pfoa %>%
  filter(grepl("R", Abbr)) %>%
  filter(grepl("s", Abbr)) %>%
  left_join(concentration_tfa, by='Abbr', suffix = c('_PFOA', '_TFA')) %>%
  pivot_longer(
    cols = c(Concentration_PFOA, Concentration_TFA),
    names_to = "stof",
    values_to = "waarde"
  ) %>%
  mutate(log_waarde = log10(waarde)) %>%
  mutate(large_waarde = 1 * waarde)
  

concentrations_con <- concentration_pfoa %>%
  filter(grepl("C", Abbr)) %>%
  filter(grepl("s", Abbr)) %>%
  left_join(concentration_tfa, by='Abbr', suffix = c('_PFOA', '_TFA')) %>%
  pivot_longer(
    cols = c(Concentration_PFOA, Concentration_TFA),
    names_to = "stof",
    values_to = "waarde"
  ) %>%
  mutate(log_waarde = log10(waarde)) %>%
  mutate(large_waarde = 1 * waarde)


kable(subcompartments)
#staaf diagram maken van verdeling per compartiment en Stof
Kolom = 'waarde'

p1 <- ggplot(concentrations_reg, aes(x = Abbr, y = large_waarde, fill = stof)) +
                geom_bar(stat = "identity", position = "dodge") +
                scale_y_log10() +
                labs(
                  title = "Concentration PFOA en TFA (steady state - regional)",
                  x = "Compartment",
                  y = "LOG(Concentration * 1e9)"
                ) +
                theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(concentrations_con, aes(x = Abbr, y = large_waarde, fill = stof)) +
                geom_bar(stat = "identity", position = "dodge") +
                scale_y_log10() + 
                labs(
                  title = "Concentration PFOA en TFA (steady state - continental)",
                  x = "",
                  y = "LOG(Concentration * 1e9)"
                ) +
                theme(axis.text.x = element_text(angle = 45, hjust = 1))
p1 / p2

kable(subcompartments)
```

``` {r Plotten nieuwe solver}
# You can specify a scale and a subcompartment
World$PlotMasses(scale = "Regional", subcompart = "agriculturalsoil")

# Or just a scale and all subcompartments are plotted:
World$PlotMasses(scale = "Continental", subcompart=c("agriculturalsoil", "naturalsoil", "othersoil"))

# Plot all concentrations at regional scale
#Waters
World$PlotMasses(scale = "Regional", subcompart=c("sea", "river", "lake", "deepocean"))
World$PlotConcentration(scale = "Regional", subcompart=c("sea", "river", "lake", "deepocean"))
#Sediments 
World$PlotMasses(scale = "Regional", subcompart=c("freshwatersediment", "lakesediment", "marinesediment"))
World$PlotConcentration(scale = "Regional", subcompart=c("freshwatersediment", "lakesediment", "marinesediment"))
#Soils
World$PlotMasses(scale = "Regional", subcompart=c("agriculturalsoil", "naturalsoil", "othersoil"))
World$PlotConcentration(scale = "Regional", subcompart=c("agriculturalsoil", "naturalsoil", "othersoil"))

World$PlotConcentration(scale = "Regional")
World$PlotMasses(scale = "Regional")

# Continental
#Sediments 
World$PlotMasses(scale = "Continental", subcompart=c("freshwatersediment", "lakesediment", "marinesediment"))
World$PlotConcentration(scale = "Continental", subcompart=c("freshwatersediment", "lakesediment", "marinesediment"))
#Soils
World$PlotMasses(scale = "Continental", subcompart=c("agriculturalsoil", "naturalsoil", "othersoil"))
World$PlotConcentration(scale = "Continental", subcompart=c("agriculturalsoil", "naturalsoil", "othersoil"))
#Waters
World$PlotMasses(scale = "Continental", subcompart=c("sea", "river", "lake", "deepocean"))
World$PlotConcentration(scale = "Continental", subcompart=c("sea", "river", "lake", "deepocean"))

World$PlotConcentration(scale = "Continental")
World$PlotMasses(scale = "Continental")


#All 
World$PlotConcentration(scale = "Continental")
World$PlotConcentration(scale = "Continental", subcompart=c("sea"))
World$PlotConcentration(scale = "Arctic")
World$PlotConcentration(scale = "Tropic", subcompart=c("sea", "deepocean", "air", "naturalsoil"))
World$PlotConcentration(scale = "Tropic")
World$PlotConcentration(scale = "Moderate")
World$PlotConcentration(scale = "Moderate", subcompart=c("air", "deepocean", "marinesediment", "sea"))
World$PlotMasses(scale = "Tropic")
World$PlotMasses(scale = "Moderate")
World$PlotMasses(scale = "Arctic")

World$PlotMassDistribution(scale = 'Continental')

```




############################################
###----------Overige Code----------------###
############################################


#Plotten
``` {r Plotten Resultaat:}
solved_long <- solved |> as_tibble() |> 
  dplyr::select(!starts_with("emis")) |>
  pivot_longer(!time, names_to = "Abbr", values_to = "Mass") |>
  mutate(Year = time/(365.25*24*60*60)) |>
  left_join(World$states$asDataFrame, by="Abbr") # Join the abbreviations to more understandable Scale, SubCompart and Species

"Functie voor plot"
plot_scale <- function(mass_dataframe, scale, SubCompartments){
  # Filter masses for the specific scale
  if (is.null(SubCompartments)) {
    print("Plotting all SubCompartments")
    mass_data_scale <- mass_dataframe %>%
      filter(Scale == scale)
  }
  else {
    print("Plotting selected SubCompartments")
    mass_data_scale <- mass_dataframe %>%
      filter(Scale == scale) %>%
      filter(SubCompart %in% SubCompartments) 
  }
  
  # Plot masses on the specific scale
  mass_plot <- ggplot(mass_data_scale, aes(x=Year, y=Mass, col=SubCompart)) + 
    geom_line() +
    ylab("Mass (kg)") +
    theme_bw() +
    ggtitle(paste0("Mass of ", substance, " in the ", catchment, " catchment")) +
    scale_y_continuous() +
    scale_x_continuous()
  
  return(mass_plot)
}

plot <- plot_scale(solved_long, "Regional", SubCompartments = NULL)
regional_plot <- plot_scale(solved_long, "Regional", SubCompartments = list("river", "naturalsoil", "agriculturalsoil", "othersoil", "air"))

#Plot data
print(regional_plot)
```


##Wat plots 
```{r debiet Rijn/Maas plotten en correleren met neerslag}
#Dataframe met beide rivieren maken
debiet <- left_join(maas, rijn, by="WAARNEMINGDATUM") %>%
  rename("Maas" = AVERAGE_VALUE.x)%>%
  rename("Rijn" = AVERAGE_VALUE.y)

#Functie om te plotten
plot_debiet <- function(data) {
  # Ensure WAARNEMINGDATUM is in Date format (though it seems to already be)
  data$WAARNEMINGDATUM <- as.Date(data$WAARNEMINGDATUM)

  plot <- ggplot(data) +
    # Plotting Maas and Rijn as separate lines
    geom_line(aes(x = WAARNEMINGDATUM, y = Maas, color = "Maas")) + 
    geom_line(aes(x = WAARNEMINGDATUM, y = Rijn, color = "Rijn")) +
    ylab("Q [m3/s]") +
    ggtitle("Debiet over tijd") +
    scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "2 year") +  # Date formatting on x-axis
    scale_y_continuous() +
    theme_bw() +
    scale_color_manual(values = c("Maas" = "blue", "Rijn" = "red"))  # Customize colors
  return(plot)
}

#Functie aanroepen om te plotten
debiet_plot <- plot_debiet(debiet)
print(debiet_plot)
```

```{r Neerslag}
'''
#Verschillende plots 
ggplot(data = yearly_summary[-c(1:60), ], aes(x = year_number, y = TotalPrecipitationMean)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(
    title = "Yearly Average Precipitation",
    x = "Year",
    y = "Average Precipitation (mm)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#histogram of cell values
raster_values <- values(mask)  # mask_raster is your masked raster object
raster_values <- raster_values[!is.na(raster_values)]  # Remove NA values
mean_value <- mean(raster_values * 1000 * 30.43, na.rm = TRUE)
median_value <- median(raster_values * 1000 * 30.43, na.rm = TRUE)

# Create histogram
hist(
  raster_values * 1000 * 30.43,
  breaks = 30,
  main = "Distribution of Precipitation Values",
  xlab = "Precipitation (mm)",
  ylab = "Frequency",
  col = "skyblue",
  border = "white"
)

# Add mean and median lines
abline(v = mean_value, col = "red", lwd = 2, lty = 2)  # Red dashed line for mean
abline(v = median_value, col = "blue", lwd = 2, lty = 2)  # Blue dashed line for median

# Add legend
legend(
  "topright",
  legend = c("Mean", "Median"),
  col = c("red", "blue"),
  lty = 2,
  lwd = 2,
  bty = "n"  # No border around the legend
)
#mask on raster
precipitation.slide <- precipitation.array[, , 500]
r <- raster(t(precipitation.slide), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
plot(r * 1000 *30, main = "Raster with Shapefile Overlay", col = terrain.colors(100))
transparent_red <- rgb(1, 0, 0, alpha = 1)  # 50% transparent red
plot(shape, add = TRUE, border = transparent_red, col= NA, lwd = 2)
'''

```

```{r correlatie}
#Correleren van data na 1989, op jaarlijkse basis
cor_data <- data.frame(x = precipitation_1989$TotalPrecipitation, y = debiet_1989$Rijn)
result <- cor(cor_data$x, cor_data$y, method = "spearman")

ggscatter(cor_data, x = "x", y= "y",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Precipitation (mm/yr)", ylab = "Flow (m3/yr)") +
  ggtitle(paste("Annual Precipitation vs Flow (n =", nrow(cor_data), ")"))
          

```

```{r Partitioning}
# Solid-Water K - f_Ksw
#Ocatnol-water coefficient
World$SetConst(Kow = 1995262)
Kow <- World$fetchData("Kow")
World$UpdateDirty("Kow")

#Corg organic constant
Corg <- World$fetchData("Corg")
print(Corg)
CorgStandard <- World$fetchData("CorgStandard")
print(CorgStandard)

#PKA 
pKa <- World$fetchData("pKa")
if(is.na(pKa)) {
  World$SetConst(pKa = 7) 
  warning("pKa not given in input data. Substance assumed to be neutral (pKa = 7).")
}

#Calculating Ksw
Substance_ChemClass <- World$fetchData("ChemClass")
QSARtable <- World$fetchData("QSARtable")
print(QSARtable)
QSARrecord <- QSARtable[QSARtable$QSAR.ChemClass == Substance_ChemClass,]
RhoTable <- World$fetchData("rhoMatrix")
RHOsolid <- RhoTable$rhoMatrix[RhoTable$SubCompart == "othersoil"]
KswModelled <- f_Ksw(Kow=Kow, pKa=pKa, CorgStandard=CorgStandard, 
                     a = QSARrecord$a, b = QSARrecord$b, 
                     ChemClass=Substance_ChemClass,
                     RHOsolid=RHOsolid,
                     alt_form = F)
#print(KswModelled)

World$SetConst(Ksw = KswModelled)
World$fetchData("Ksw")

#Ksw in alternative form
Ksw.alt
test <- World$NewCalcVariable("Ksw.alt")
World$CalcVar("Ksw.alt")
World$fetchData("Ksw.alt")
Kswalt = World$fetchData("Ksw.alt")

```

```{r Flows}
"Berekent afvoer, komt enigzins overeen met werkelijkheid voor Rijn &Maas"
"Rijn 1714 m3/s <- gemiddeld is dit zo'n 2025 m3/s"
"Maas 320 m3/s <- gemiddeld is dit zo'n 220 m3/s"

flow2 <- World$NewFlow("x_RiverDischarge")

flow2$FromAndTo
flow2$execute()
World$CalcVar("x_RiverDischarge")

x_RiverDischarge_flow <- World$fetchData("x_RiverDischarge")
x_ContRiver2Reg_flow <- World$fetchData("x_ContRiver2Reg")
LakeFracRiver_val <- World$fetchData("LakeFracRiver")
Scalename <- World$fetchData("ScaleName")

flow3 <- World$NewFlow("x_LakeOutflow")

flow3$FromAndTo
flow3$execute()
World$CalcVar("x_LakeOutflow")

flows = World$fetchData("Flows") %>%
  filter(toSubCompart == "sea" | toSubCompart == "river") %>%
  filter(toScale == "Regional" | toScale == "Continental")
  


















