---
title: "PFAS_prob"
output: html_document
---

### Use the dynamic solver probabilistically

For the dynamic deterministic solver, the emissions need to be given as
a dataframe with three columns:

-   'Abbr', which contains the abbreviations for the compartments

-   'Emis', which contains the emissions to the compartments, in kg/s

-   'Time', which contains the time of the emission in seconds,

-   'RUN', which contains the run number of the emission.

```{r Prepare uncertain dynamic emissions, warning=FALSE, message=FALSE}

substance <- "Trifluoroacetic"
source('baseScripts/initWorld.R')

source("AdjustModelRhine.R")

load("data/Examples/example_uncertain_data.RData")

example_data <- 
  example_data |>
  dplyr::select(To_Compartment, `2020`, `2021`,`2022`, `2023`, RUN) |>
  pivot_longer(!c(To_Compartment, RUN), names_to = "year", values_to = "Emis") |>
  mutate(Abbr = case_when(
    To_Compartment == "Agricultural soil (micro)" ~ "s2RU",
    To_Compartment == "Residential soil (micro)" ~ "s3RU",
    To_Compartment == "Surface water (micro)" ~ "w1RU"
  )) |>
  dplyr::select(-To_Compartment) |>
  mutate(Time = ((as.numeric(year))*365.25*24*3600)) |>
  mutate(Emis = (Emis*1000000)/(365.25*24*3600)) |> # Convert kt/year to kg/s
  dplyr::select(-year)

# The raw emission data from 2020 to 2023 will not solve due to instability in the intital state starting at 0. This can be solved by starting at t=0 and mass=0 see below:

example_data <-
  example_data |> 
  mutate(Time = Time - 2019*365.25*24*3600) |> # converting to 0 being one year before 2020
  full_join(expand.grid(Abbr = unique(example_data$Abbr), 
                        RUN = c(1:20)) |> 
              mutate(Time = 0,
                     Emis = 0))

```

Prepare the variables the same way as used for the variables in the
steady state probabilistic solver.

```{r Prepare variable df and functions dyn, warning=FALSE, message=FALSE}

# Load the Excel file containing example distributions for variablese
Example_vars <- readxl::read_xlsx("data/Examples/Example_uncertain_variables.xlsx", sheet = "Variable_data")

varFuns <- World$makeInvFuns(Example_vars)

```

`World$Solve` needs the following variables to solve steady state
probabilistically:

-   emissions: the emissions dataframe

-   var_box_df: a dataframe with the example variables

-   var_invFun: the functions created from the var_box_df in a list

-   nRUNs: the number of runs (should match the number of RUNs in the
    emissions dataframe provided)

-   tmax: the end time for running the solver, in seconds

-   nTIMES: the number of calculation time steps.

Now we can solve and get the solution, variable values, concentration
and emissions. NOTE: These calculations take a bit longer than the
previous calculations.

```{r Solve dyn probabilistic, warning=FALSE, message=FALSE}
World$NewSolver("DynamicSolver")


tmax <- max(unique(example_data$Time))
tmin <- min(unique(example_data$Time))
nTIMES <- length(seq(0, tmax, length.out = 10))

World$Solve(emissions = example_data, 
            var_box_df = Example_vars, 
            var_invFun = varFuns, 
            nRUNs = length(unique(example_data$RUN)), 
            tmin = tmin,
            tmax = tmax, 
            nTIMES = nTIMES)

solution = World$Masses()
emission = World$Emissions()
variable_values = World$VariableValues()
concentrations = World$Concentration()
```

Finally, we can plot the outcome using the predefined plot functions. If
there is no scale specified, the Regional scale will be selected. If
there is no subcompart specified, "agriculturalsoil" will be selected.

```{r Plot outcome with uncertainty, warning=FALSE, message=FALSE}
World$PlotConcentration(scale = "Regional", subcompart = "agriculturalsoil")

World$PlotMasses(scale = "Regional", subcompart = "agriculturalsoil")
```