---
title: "DPMFA LEON-T"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

# *Initializing World and Substance*

Depending on the substance you want to analyze, a selection needs to be
made from the three different"worlds" : Molecular, Paticulate and
Plastics. Here, we initialize the world for plastics.

```{r Initialize World}
library(lhs)
library(readxl)
library(viridis)

source("baseScripts/initWorld_onlyPlastics.R")

World$substance <- "microplastic"
```

# Load the DPMFA data and make a nested emission dataframe


```{r Load DPMFA data}
abspath <- "R:/Projecten/E121554 LEON-T/03 - uitvoering WP3/DPMFA_output/Baseline_PMFA_EU.RData"

load(abspath)

# Check if the loaded data is DPMFA or PMFA data
if(exists("DPMFA_stocks")) {
  type <- "DPMFA"
} else {
  type <- "PMFA"
}

# Convert to long format
data_long <- 
  DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
  pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
               names_to = "Year",
               values_to = "Mass_Polymer_kt")

# Choose a source of interest if you want, else write NA
source_of_interest <- "Tyre wear"

# Choose material of interest 
mat <- "RUBBER"

if(!is.na(source_of_interest)){
  if(!source_of_interest %in% unique(data_long$Source)){
    print("Selected source not in dataframe")
  } else {
    filtersource <- source_of_interest
  }
} else if (is.na(source_of_interest)){
  filtersource <- unique(data_long$Source)
}

# Calculate kg/s from kt/y 
data_summed <- data_long |>
  filter(Source %in% filtersource) |>
  #filter(between(RUN, 1, 100)) |>
  mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |>
  filter(Material_Type == "micro") 

if(unique(data_long$Scale) == "EU"){
  SBscale <- "C"
} else if (unique(data_long$Scale) == "NL") {
  SBscale <- "R"
}

# Assign SB compartments to DPMFA compartments
data_filtered <- data_summed |>
  select(To_Compartment, Mass_Polymer_kg_s, Year, RUN, Polymer) |>
  mutate(Scale = SBscale) |>
  mutate(Compartment = case_when(
    str_detect(To_Compartment, "soil") ~ "s",
    str_detect(To_Compartment, "water") ~ "w",
    str_detect(To_Compartment, "air") ~ "a"
  )) |>
  mutate(Subcompartment = case_when(
    str_detect(To_Compartment, "Agricultural") ~ "2",
    str_detect(To_Compartment, "Natural") ~ "1",
    str_detect(To_Compartment, "Sub-surface") ~ "3",
    str_detect(To_Compartment, "Road side") ~ "3",
    str_detect(To_Compartment, "Residential") ~ "3",
    str_detect(To_Compartment, "Sea") ~ "2",
    str_detect(To_Compartment, "Surface") ~ "1",
    str_detect(To_Compartment, "Outdoor") ~ ""
  )) |>
  mutate(Species = "S") |>
  mutate(Abbr = paste0(Compartment, Subcompartment, Scale, Species)) |>
  group_by(Abbr, Year, RUN, Polymer) |>
  summarise(Mass_Polymer_kg_s = sum(Mass_Polymer_kg_s)) |>
  ungroup() |>
  rename(value = Mass_Polymer_kg_s) |>
  select(Abbr, Year, Polymer, value, RUN)

```

```{r Make a separate emission df for every polymer}
polymers <- unique(data_filtered$Polymer)


filtered <- data_filtered |>
 filter(Polymer == mat)
 
df_names <- c()
 
if(type == "DPMFA"){
 # Make an emission dataframe for the dynamic uncertain solver
 emis_df_dyn <- filtered |>
   group_by(Abbr, Year) |>
   rename(Timed = Year) |>
   mutate(Timed = as.double(Timed)*(365.25*24*3600)) |>
   nest(Emis = c(RUN, value)) 
 
 df_name <- paste0("emis_dyn_", i)

 assign(df_name, emis_df_dyn)

 df_names <- c(df_names, df_name)
   
 ymin <- min(emis_df_dyn$Year)
 ymax <- max(emis_df_dyn$Year)
   
} else if(type == "PMFA"){
  
 # Select data for 2019
 y <- 2019
  
 # Make an emission dataframe for the steady uncertain solver
 emis_df_ss <- filtered |>
   filter(Year == y) |>
   nest(Emis = c(RUN, value)) 
 
 df_name <- paste0("emis_ss_", i)
   
 assign(df_name, emis_df_ss)
   
 df_names <- c(df_names, df_name)
}

```

```{r Functions for distributions}
# Define triangular distribution function
Triangular <- function(u, a, b, c) {               # u = samples, a = , b = , c = 
  ifelse(u < (c-a)/(b-a),
         a + sqrt(u * (b-a) * (c-a)),
         b - sqrt((1-u) * (b-a) * (b-c)))
}

# Define uniform distribution function
Uniform <- function(u, a, b) {                     # u = samples, a = min, b = max
  transformed_samples <- a + (b - a) * lhs_samples
  return(transformed_samples)
}

# Define power law distribution function



```

# Create dummy dataset with uncertain variables

```{r}
# Path to excel file with distribution values
path_dist <- "N:/Documents/Simplebox/Microplastic_variables.xlsx"

# Generate the correct number of samples
n_samples <- nrow(emis_df_ss$Emis[[1]]) # Number of emission runs 
sample_df_names <- c()

input_vars <- read_excel(path_dist, sheet = mat) |>
  filter(!is.na(Distribution)) |>
  select(-c(Source, Comments))
  
varnames <- unique(input_vars$VarName)
  
n_vars <- nrow(input_vars)
# Generate LHS
lhs_samples <- randomLHS(n_samples, n_vars)
var_df_names <- c()
  
for(j in 1:nrow(input_vars)){
  df <- input_vars[j, ]

  if(is.na(df$Scale) && is.na(df$SubCompart) && is.na(df$Species)){
    varname <- df$VarName
      
    var <- World$fetchData(varname)
    var <- data.frame(setNames(var, varname))
    
    name <- paste0("var", j)
    var_df_names <- c(var_df_names, name)
      
    var_df <- var |>
      mutate(Scale = df$Scale) |>
      mutate(SubCompart = df$SubCompart)

    var_df$a <- df$Min
    var_df$b <- df$Max
    var_df$c <- df$Peak
      
    assign(name, var_df)
  } 
}
  
params <- tibble(
  varName = varnames,
  Scale = sapply(var_df_names, function(v) get(v)$Scale),
  SubCompart = sapply(var_df_names, function(v) get(v)$SubCompart),
  data = lapply(var_df_names, function(v) {
    df <- get(v)
    tibble(id = c("a", "b", "c"), value = c(df$a, df$b, df$c))
  }))

sample_df <- params
  # Transform each LHS sample column to the corresponding triangular distribution

for (k in 1:n_vars) {
  a <- filter(params$data[[k]], id == "a") %>% pull(value)
  b <- filter(params$data[[k]], id == "b") %>% pull(value)
  c <- filter(params$data[[k]], id == "c") %>% pull(value)
    samples <- Triangular(lhs_samples[, k], a, b, c)
    # Create a new tibble for 'data' with samples replacing original values
  new_data <- tibble(value = samples)
    # Update the data column in the sample_df
  sample_df$data[[k]] <- new_data
}
  
sample_df_name <- paste0("sample_df_",i)
sample_df_names <- paste0(sample_df_names, sample_df_name)
assign(sample_df_name, sample_df)

```

# Solve the matrix 

Solve the matrix depending on the type of input emissions given (PMFA = Steady state, DPMFA = dynamic)

```{r Solve steady state}

if(type == "PMFA"){
  World$NewSolver("UncertainSolver")
  
  for(i in polymers){
    emis_df_name <- paste0("emis_ss_", i)
    sample_df_name <- paste0("sample_df_", i)
    
    solved <- World$Solve(get(emis_df_name), needdebug = FALSE, get(sample_df_name))
    assign(paste0("solved_ss_", i), solved)
  }
  
} else if(type == "DPMFA"){
  World$NewSolver("UncertainDynamicSolver")
  
  for(i in polymers){
    emis_df_name <- paste0("emis_dyn_", i)
    sample_df_name <- pasteo("sample_df_", i)
    
    solved <- World$Solve(World$Solve(get(i)), sample_df, tmax = tmax, needdebug = F)
    assign(paste0("solved_dyn_", i), solved)
  }
}

```

# Calculate concentrations
```{r Calculate concentrations}

dfname <- paste0("solved_ss_", mat)
  
sol <- get(dfname)
  
Solution <- sol$SteadyStateMass |>
  mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  mutate(Abbr = str_replace_all(Abbr, "cw", "a")) |>
  group_by(RUN, Scale, SubCompart, Species, Unit, Abbr) |>
  summarise(EqMass = sum(EqMass)) |>
  ungroup() 
  
# Fetch necessary data
  
# Sum cloudwater and air under air
Volume <- World$fetchData("Volume") |>
  mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  group_by(Scale, SubCompart) |>
  summarise(Volume = sum(Volume)) |>
  ungroup()
  
Area <- World$fetchData("Area")
FRACw <- World$fetchData("FRACw")
FRACa <- World$fetchData("FRACa")
Fractrial <- FRACa$FRACa[FRACa$SubCompart =="air" & FRACa$Scale =="Arctic" ]
Rho <- World$fetchData("rhoMatrix")
Concentration_eq <- merge(Solution, Volume, by = c("SubCompart", "Scale"))
Concentration_eq$Concentration <- Concentration_eq$EqMass / Concentration_eq$Volume
RhoWater_value <- Rho$rhoMatrix[Rho$SubCompart == "river"]

f_adjust_concentration <- function(Concentration, FRACw, FRACa, SubCompart, Scale, Rho, RhoWater_value) {
  # Fetch Fracw based on SubCompart and Scale
  Fracw <- FRACw$FRACw[FRACw$SubCompart == SubCompart & FRACw$Scale == Scale]
  
  # Fetch Fraca based on SubCompart and Scale
  Fraca <- FRACa$FRACa[FRACa$SubCompart == SubCompart & FRACa$Scale == Scale]
  
  # Fetch RHOsolid based on SubCompart
  RHOsolid <- Rho$rhoMatrix[Rho$SubCompart == SubCompart]
    
  # Check if any of Fracw, Fraca, or RHOsolid are N
    
  Concentration * 1000 / (Fracw * RhoWater_value + (1 - Fracw - Fraca) * RHOsolid)
}
  
subcomparts <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment")
  
filtered_data <- Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, ]
  
# Apply the adjustment function to the filtered data
adjusted_concentrations <- apply(filtered_data, 1, function(row) {
  f_adjust_concentration(
    Concentration = as.numeric(row["Concentration"]),
    FRACw = subset(FRACw, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
    FRACa = subset(FRACa, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
    Rho = subset(Rho, SubCompart == row["SubCompart"]),
    RhoWater_value = 998,  # Replace with your actual RhoWater_value
    SubCompart = row["SubCompart"],
    Scale = row["Scale"]
  )
})
# Update Concentration_eq with adjusted concentrations
Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, "Concentration"] <- adjusted_concentrations

#Define the units
subcompart <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment",  "lakesediment", "air", "deepocean", "lake" , "river", "sea")
units <- c("g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw",
           "kg/m^3", "kg/L", "kg/L", "kg/L", "kg/L", "kg/L")
  
# Combine into a named list
subcompart_units <- setNames(units, subcompart)
Concentration_eq <- Concentration_eq |>
  mutate(Units_per_SubCompart = subcompart_units[SubCompart])
  
convert_units <- function(concentration, unit) {
  if (unit == "g/kg dw") {
    return(concentration * 1e9)
  } else if (unit == "kg/m^3") {
    return(concentration * 1e12)
  } else if (unit == "kg/L") {
    return(concentration * 1e12)
  } else {
    return(concentration)
  }
}

# Convert concentrations to ng/kg or ng/L and update units
Concentration_eq <- Concentration_eq |>
  mutate(
         Concentration = mapply(convert_units, Concentration, Units_per_SubCompart),
         Units_per_SubCompart = ifelse(Units_per_SubCompart == "g/kg dw", "ng/kg dw", 
                                       ifelse(Units_per_SubCompart == "kg/m^3", "ng/m^3", "ng/L")))
Concentration_eq <- Concentration_eq |>
  mutate(across(where(is.numeric), ~ format(., scientific = TRUE)))

solname <- paste0("Solution_", mat)
assign(solname, Concentration_eq)
  
```

# Prep data and make plots
```{r Solve steady state}

# Create a plot theme
plot_theme <-  theme(
  axis.title.x = element_text(size = 14),    
  axis.title.y = element_text(size = 14),    
  axis.text.x = element_text(size = 12, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 12),
  title = element_text(size=20),
  panel.background = element_rect(fill = "white"),  # White background
  panel.grid.major = element_line(color = "lightgrey", size = 0.5),  # Major grid lines in light grey
  panel.grid.minor = element_line(color = "lightgrey", size = 0.25)
)

# Plot the data for each polymer separately, for all selected compartments (different plots depending on PMFA or DPMFA)
if(type == "PMFA"){
  solution_name <- paste0("Solution_", mat)
  df_name <- paste0("solved_ss_", mat)
  
  Solution <- get(solution_name) |>
    mutate(RUN = as.integer(RUN)) |>
    mutate(EqMass = as.double(EqMass)) |>
    mutate(Concentration = as.double(Concentration)) 
  
  solved <- get(df_name)
  
  # Prepare data for plots
  Input_Variables <- 
    solved$Input_Variables |> unnest(data) 
  Input_Emission <- 
    solved$Input_Emission |> unnest(Emis) 
  
  Plot_data <- 
    Input_Variables |> 
    pivot_wider(names_from = c(varName,Scale,SubCompart,Unit), values_from = value) |> 
    full_join(Solution) |>
    full_join(Input_Emission, by = c("Abbr", "RUN"))
  
  # Aggregate plot data for all species
  Plot_data_agg <- Plot_data |>
    group_by(Scale, SubCompart, RUN, Unit.x, Units_per_SubCompart) |>
    summarise(EqMass = sum(EqMass), Concentration = sum(Concentration))
  
  scales <- unique(Plot_data_agg$Scale)
  
  # Make plots for each scale, one with mass and one with concentrations
  for (scale in scales) {
    df <- Plot_data_agg |>
      filter(Scale == scale) |>
      mutate(concname = paste0(SubCompart, " (", Units_per_SubCompart, ")"))
    
    # Filter out the compartments without mass
    masscomps <- df |>
      filter(RUN == 1) |>
      filter(EqMass != 0)
    
    masscomps <- masscomps$SubCompart
    
    df <- df |>
      filter(SubCompart %in% masscomps)
    
    # Mass plot
    mass_p <- ggplot(df, mapping = aes(x = SubCompart, y = EqMass, fill = SubCompart)) +  
      geom_violin() + 
      labs(title = paste0("Masses at ", scale, " scale"),
           subtitle = subt,
           x = "Compartment",
           y = "Mass (kg)") +
      plot_theme +
      scale_y_continuous(trans = 'log10') +
      scale_fill_viridis_d() + 
      # Customize the legend for this plot as well
      theme(axis.text.x = element_blank(),  # Remove x-axis labels
            legend.position = "bottom") +   # Move legend to the bottom
      guides(fill = guide_legend(title = NULL))  # Remove legend title
    
    print(mass_p)
    
    ggsave(paste0("N:/Documents/Simplebox/LEON-T_plots/Mass_plot_", scale, ".png"), plot=mass_p)
    
    # Concentration plot
    conc_p <- ggplot(df, mapping = aes(x = concname, y = Concentration, fill = concname)) +  
      geom_violin() +
      labs(title = paste0("Concentrations at ", scale, " scale"),
           subtitle = subt,
           x = "Compartment",
           y = "Concentration") +
      plot_theme +
      scale_y_continuous(trans = 'log10') +
      scale_fill_viridis_d() +
      theme(axis.text.x = element_blank(),  # Remove x-axis labels
            legend.position = "bottom") +   # Move legend to the bottom
      guides(fill = guide_legend(title = NULL))  # Remove legend title
    
    print(conc_p)
    
    ggsave(paste0("N:/Documents/Simplebox/LEON-T_plots/Concentration_plot_", scale, ".png"), plot=conc_p)
  }

} else if(type == "DPMFA"){

  # Prepare the plot data
  df_name <- paste0("solved_dyn_",i)
  solved <- get(df_name)
  
  ymin <- as_character(ymin)
  ymax <- as_character(ymax)
  
  Input_Variables <- 
    solved$Input_Variables |> unnest(data) 
  
  inputcomps <- unique(emissions$Abbr)
  
  Input_Emission <- solved$DynamicMass |>
    select(starts_with("emis2"), "time", "RUN") |>
    rename_with(
      .fn = ~ str_remove(.x, "emis2"),              # Function to add "_new" to column names
      .cols = !c("time", "RUN")  # Select columns to modify
    ) |>
    pivot_longer(cols = -c("time", "RUN"),
                 names_to = "Abbr",
                 values_to = "Emis") 
  
  masses_long <- solved$DynamicMass |>
    select(!starts_with("emis2")) |>
    rename_with(
      .fn = ~ paste0("mass_", .x),              # Function to add "_new" to column names
      .cols = !c("time", "RUN", "Unit") & !starts_with("emis")  # Select columns to modify
    ) |>
    pivot_longer(cols = starts_with("mass_"),
                 names_to = "Abbr",
                 values_to = "Mass") |>
    mutate(Abbr =  str_remove(Abbr, "mass_")) |>
    left_join(Input_Emission, by = c("Abbr", "time", "RUN")) |>
    left_join(solved$States, by = "Abbr")
  
  Plot_data <- 
    Input_Variables |> 
    pivot_wider(names_from = c(varName,Scale,SubCompart,Unit), values_from = value) |> 
    full_join(masses_long) |>
    mutate(time = time/(365.25*24*60*60))
  
  datatmax <- Plot_data |>
    filter(time == tmax/(365.25*24*60*60)) |>
    left_join(solved$States)
  
  maxyear <- tmax/(365.25*24*60*60)
  
  subt_range <- paste0(ymin, "-", ymax, ", ", filtersource , ", ", mat)
  subt_tmax <- paste0( as.character(y), ", ", filtersource , ", ", mat)
  
  # Plot the outcome
  varnames <- colnames(Plot_data)[2:(1+n_vars)]
  
  var_data <- Plot_data |> 
    filter(Abbr %in% emiscomps)
  
  var_tmax <- datatmax |>
    filter(Abbr %in% emiscomps)
  
  for(i in varnames) {
    p1 <- ggplot(var_data, mapping = aes(x = time, y = Mass, group = RUN, color = .data[[i]])) +  
      geom_line() + 
      facet_wrap(vars(Abbr)) +
      labs(subtitle = subt_range,
           x = "Time (years)", 
           y = "Mass (kg)", 
           color = i) +  
      ggtitle("Mass in compartment over time (default substance)") +
      plot_theme
    print(p1)  
  }
  
  for(i in varnames) {
    p2 <- ggplot(var_tmax, mapping = aes(x = .data[[i]], y = Mass)) +
      geom_point() + 
      facet_wrap(vars(Abbr)) + 
      ggtitle(paste0("Mass in compartment at year ", as.character(maxyear))) + 
      labs(subtitle = subt_tmax,
           x = i,  
           y = "Mass (kg)") +
      plot_theme                 
    print(p2)
  }
  
  # Filter the plot data for the compartments that received emissions
  EM_data <- Plot_data |>
    filter(Abbr %in% emiscomps)
  
  EM_tmax <- datatmax |>
    filter(Abbr %in% emiscomps)
  
  # Plot the emissions against the masses for these compartments
  EM_p1 <- ggplot(EM_tmax, mapping = aes(x = Emis, y = Mass)) +
    geom_point() + 
    facet_wrap(vars(Abbr)) +
    ggtitle("Relation between emissions and mass") + 
    labs(subtitle = subt_tmax,
         x = "Emissions (kg/s)",
         y = "Mass (kg)") + 
    plot_theme
  print(EM_p1)
}

```

