---
title: "DPMFA LEON-T"
author: "Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

# *Initializing World and Substance*

Depending on the substance you want to analyze, a selection needs to be
made from the three different"worlds" : Molecular, Particulate and
Plastics. Here, we initialize the world for plastics.

```{r Initialize World}
library(lhs)
library(readxl)
library(viridis)
library(trapezoid)

source("baseScripts/initWorld_onlyPlastics.R")
```

# Load the DPMFA data and make a nested emission dataframe

```{r Load DPMFA data}
#source_of_interest <- NA # NA for all except TWP (needs to be run separately)
source_of_interest <- "Tyre wear"

# Choose substance based on chosen sources
if(!is.na(source_of_interest) && length(source_of_interest) == 1 && source_of_interest == "Tyre wear") {
  World$substance <- "TRWP"
} else {
  World$substance <- "microplastic"
}

# Select PMFA data for 2019
PMFA_year <- 2019

MFAtype <- "DPMFA" # other option is DPMFA

if(MFAtype == "PMFA"){
  # Load EU data
  abspath <- "vignettes/Case Studies/CaseData/PMFA_EU.RData"
  load(abspath)
  
  # Convert to long format
  data_long_EU <- 
    DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
    pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
                 names_to = "Year",
                 values_to = "Mass_Polymer_kt") |>
    mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |> # Convert kt/year to kg/s
    filter(Material_Type == "micro") |> # Select microplastics only
    mutate(SBscale = ifelse(Scale == "EU", "C", "R"))
    
  # Load NL data
  abspath <- "vignettes/Case Studies/CaseData/PMFA_NL.RData"
  load(abspath)
  
  # Convert to long format
  data_long_NL <- 
    DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
    pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
                 names_to = "Year",
                 values_to = "Mass_Polymer_kt") |>
    mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |> # Convert kt/year to kg/s
    filter(Material_Type == "micro") |> # Select microplastics only
    mutate(SBscale = ifelse(Scale == "EU", "C", "R"))
  
  data_long <- rbind(data_long_EU, data_long_NL)
  
  if (all(!is.na(source_of_interest))) {
    # Check if all the elements of source_of_interest are present in data_long$Source
    if (!all(source_of_interest %in% unique(data_long$Source))) {
      print("Selected source(s) not in dataframe")
    } else {
      sources <- source_of_interest
    }
  } else if (all(is.na(source_of_interest))) {
    # If all elements are NA, return all unique sources
    sources <- unique(data_long$Source)
    sources <- sources[sources != "Tyre wear"]
  }
  
  # Assign SB compartments to DPMFA compartments
  DPMFA_sink_micro <- data_long |>
    filter(Year == PMFA_year) |>
    filter(Source %in% sources) |>
    select(Source, To_Compartment, Mass_Polymer_kg_s, Year, RUN, Polymer, SBscale) |>
    rename(Scale = SBscale) |>
    filter(To_Compartment != "Sub-surface soil (micro)") |> # Exclude sub-surface soil because this is currently outside the scope of SimpleBox
    mutate(Compartment = case_when(
      str_detect(To_Compartment, "soil") ~ "s",
      str_detect(To_Compartment, "water") ~ "w",
      str_detect(To_Compartment, "air") ~ "a"
    )) |>
    mutate(Subcompartment = case_when(
      str_detect(To_Compartment, "Agricultural") ~ "2",
      str_detect(To_Compartment, "Natural") ~ "1",
      str_detect(To_Compartment, "Road side") ~ "3",
      str_detect(To_Compartment, "Residential") ~ "3",
      str_detect(To_Compartment, "Sea") ~ "2",
      str_detect(To_Compartment, "Surface") ~ "1",
      str_detect(To_Compartment, "Outdoor") ~ ""
    )) |>
    mutate(Species = case_when(
      Source == "Tyre wear" ~ "P",
      TRUE ~ "S")) |>
    mutate(Abbr = paste0(Compartment, Subcompartment, Scale, Species)) |>
    mutate(Subcompartment = paste0(Compartment, Subcompartment)) |>
    #group_by(Source, Abbr, Year, RUN, Polymer) |>
    group_by(Abbr, Year, Polymer, RUN, Subcompartment) |>
    summarise(Mass_Polymer_kg_s = sum(Mass_Polymer_kg_s)) |>
    ungroup() |>
    rename(value = Mass_Polymer_kg_s)
} else if(MFAtype == "DPMFA"){
   # Load EU data
  abspath <- "vignettes/Case Studies/CaseData/DPMFA_EU.RData"
  load(abspath)
  
  # Convert to long format
  data_long_EU <- 
    DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
    pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
                 names_to = "Year",
                 values_to = "Mass_Polymer_kt") |>
    mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |> # Convert kt/year to kg/s
    filter(Material_Type == "micro") |> # Select microplastics only
    mutate(SBscale = ifelse(Scale == "EU", "C", "R")) 
  
  # Load NL data
  abspath <- "vignettes/Case Studies/CaseData/DPMFA_NL.RData"
  load(abspath)
  
  # Convert to long format
  data_long_NL <- 
    DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
    pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
                 names_to = "Year",
                 values_to = "Mass_Polymer_kt") |>
    mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |> # Convert kt/year to kg/s
    filter(Material_Type == "micro") |> # Select microplastics only
    mutate(SBscale = ifelse(Scale == "EU", "C", "R")) 
  
  data_long <- rbind(data_long_EU, data_long_NL)
  
  if (all(!is.na(source_of_interest))) {
    # Check if all the elements of source_of_interest are present in data_long$Source
    if (!all(source_of_interest %in% unique(data_long$Source))) {
      print("Selected source(s) not in dataframe")
    } else {
      sources <- source_of_interest
    }
  } else if (all(is.na(source_of_interest))) {
    # If all elements are NA, return all unique sources
    sources <- unique(data_long$Source)
    sources <- sources[sources != "Tyre wear"]
  }
  
  # Assign SB compartments to DPMFA compartments
  DPMFA_sink_micro <- data_long |>
    filter(Source %in% sources) |>
    select(Source, To_Compartment, Mass_Polymer_kg_s, Year, RUN, Polymer, SBscale) |>
    rename(Scale = SBscale) |>
    filter(To_Compartment != "Sub-surface soil (micro)") |> # Exclude sub-surface soil because this is currently outside the scope of SimpleBox
    mutate(Compartment = case_when(
      str_detect(To_Compartment, "soil") ~ "s",
      str_detect(To_Compartment, "water") ~ "w",
      str_detect(To_Compartment, "air") ~ "a"
    )) |>
    mutate(Subcompartment = case_when(
      str_detect(To_Compartment, "Agricultural") ~ "2",
      str_detect(To_Compartment, "Natural") ~ "1",
      str_detect(To_Compartment, "Road side") ~ "3",
      str_detect(To_Compartment, "Residential") ~ "3",
      str_detect(To_Compartment, "Sea") ~ "2",
      str_detect(To_Compartment, "Surface") ~ "1",
      str_detect(To_Compartment, "Outdoor") ~ ""
    )) |>
    mutate(Species = case_when(
      Source == "Tyre wear" ~ "P",
      TRUE ~ "S")) |>
    mutate(Abbr = paste0(Compartment, Subcompartment, Scale, Species)) |>
    mutate(Subcompartment = paste0(Compartment, Subcompartment)) |>
    group_by(Abbr, Year, RUN, Polymer, Subcompartment) |>
    summarise(Mass_Polymer_kg_s = sum(Mass_Polymer_kg_s)) |>
    ungroup() |>
    rename(value = Mass_Polymer_kg_s) |>
    select(Abbr, Year, Polymer, value, RUN, Subcompartment)
}
```

```{r Make a df with the emissions}
TESTING = T
if(TESTING==TRUE) DPMFA_sink_micro <- DPMFA_sink_micro |> filter(RUN<11)

DPMFA_sink_micro <- DPMFA_sink_micro |>
  filter(Year %in% 2019:2025)

# Filter out emission subcompartments for which SimpleBox does not have a compartment (yet)
DPMFA_sink_micro <-
  DPMFA_sink_micro |> filter(Subcompartment %in% World$fetchData("AbbrC")$AbbrC)

if(MFAtype == "PMFA"){
  DPMFA_sink_micro <- DPMFA_sink_micro |>
    nest(Emis = c(RUN, value))
} else if(MFAtype == "DPMFA"){ # TODO when using DPMFA filtering should not occur for 1 year. All years should be used.
  DPMFA_sink_micro <- DPMFA_sink_micro |>
    mutate(Timed = as.double(Year)*(365.25*24*3600)) |>
    nest(Emis = c(RUN, value))
}
```

```{r Functions for distributions}
# Define triangular distribution function
triangular <- function(u, a, b, c) {               # u = samples, a = min, b = max, c = peak
  ifelse(u < (c-a)/(b-a),
         a + sqrt(u * (b-a) * (c-a)),
         b - sqrt((1-u) * (b-a) * (b-c)))
}

# Define uniform distribution function
uniform <- function(u, a, b) {                     # u = samples, a = min, b = max
  transformed_samples <- a + (b - a) * u
  return(transformed_samples)
}

# Define power law distribution function
power_law <- function(u, a, b, c){                 # u = samples, a = min, b = max, c = alpha
  
  # Ensure that samples are within [0, 1]
  samples <- pmin(pmax(u, 0), 1)
  
  # Transform samples to the power-law distribution
  scaled_samples <- a * ((b / a) ^ samples) ^ (1 / (1 - c))
  
  return(scaled_samples)
}

trapezoidal <- function(u, a, b, c, d) {
  # Ensure u is in the range [0, 1]
  u <- pmin(pmax(u, 0), 1)  # Clip u to [0, 1]
  
  # Total width of the trapezoid
  width_total <- d - a
  base1 <- b - a    # Width of the left base
  base2 <- d - c    # Width of the right base
  
  # Calculate the CDF segments
  CDF_left <- base1 / width_total         # Area under the left triangle
  CDF_flat <- 1 - base2 / width_total     # Area under the flat top
  
  result <- ifelse(u < (base1 / width_total), 
                   a + sqrt(u * (base1) * width_total),  # Left triangle
                   ifelse(u <= (CDF_flat + base1 / width_total), 
                          b + (u - base1 / width_total) * (d - b),  # Flat top
                          d - sqrt((1 - u) * (base2) * width_total)  # Right triangle
                   )
  )
  
  return(result)
}

```

# Create dataset with uncertain variables

## Prepare excel data

```{r Prepare excel data}
# Path to excel file with distribution values
#path_dist <- "N:/Documents/Simplebox/Microplastic_variables_v2.xlsx"
path_parameters_file <- "vignettes/Case studies/CaseData/Microplastic_variables_v1.xlsx"

Material_Parameters <- read_excel(path_parameters_file, sheet = "Polymer_data") |> 
  
  # change um to nm unit coversion
  mutate(across(c(a, b, c, d), as.numeric)) |>
  mutate(across(c(a, b, c, d), ~ case_when(
    str_detect(Unit, "um") ~ . * 1000,
    TRUE ~ .
  ))) |>
  mutate(Unit = case_when(
    str_detect(Unit, "um") ~ "nm",
    TRUE ~ Unit
  ))

materials <- unique(Material_Parameters$Polymer)
materials <- unique(DPMFA_sink_micro$Polymer) # materials in selected sources



# TODO: get these from SB objects and not set in text input:
# water <- c("lake", "sea", "deepocean", "river")
# soil <- c("naturalsoil", "agriculturalsoil", "othersoil")
# sediment <- c("marinesediment", "freshwatersediment", "lakesediment")
# soil_sediment <- c(soil, sediment)
# air <- c("air", "cloudwater")
# all_SC <- c(water, soil, sediment, air)
# species <- c("Large", "Small", "Solid")
# small_large <- c("Small", "Large")
# scales <- c("Arctic", "Tropic", "Moderate", "Regional", "Continental")

# Only the ones with data in xlsx or with emission data are needed no? Why this vector?
# materials <- c("HDPE", "LDPE", "PP", "PS", "PVC", "Acryl", "PA", "PET", "ABS", "EPS", "PC", "PMMA", "PUR", "RUBBER", "OTHER")
explodeF <- function(df, target_col, explode_value, new_values) {
  df %>%
    # Use mutate to create a new column if the target column equals explode_value
    mutate(!!sym(target_col) := ifelse(!!sym(target_col) == explode_value, list(new_values), !!sym(target_col))) %>%
    # Unnest the target column to duplicate rows
    unnest(!!sym(target_col))
}

DefinedVariables <- lapply(unique(Material_Parameters$VarName),World$fetchData)
names(DefinedVariables) = unique(Material_Parameters$VarName)


# how to cope with any. For now this, but materials should be only for material being calculated for.
suppressWarnings({
  Material_Parameters <- explodeF(Material_Parameters, target_col = "Polymer", explode_value = "any", new_values = materials)
})

```


```{r Make sample df}
# Generate the correct number of samples
n_samples <- nrow(DPMFA_sink_micro$Emis[[1]]) # Number of emission runs 

polymers <- unique(DPMFA_sink_micro$Polymer)

Material_Parameters_n <- data.frame()

for(pol in polymers){

  input_vars <- 
    Material_Parameters |>
    filter(if_all(c(Distribution, a, b), ~ !is.na(.) )) |> # filter out non defined distributions
    filter(Polymer == pol) |> 
    # Not sure what the 3 lines below do.
    group_by(VarName, Scale, SubCompart, Species, Polymer) |>
    filter(if (n() > 1) MP_source == source_of_interest | (MP_source == "" & !any(MP_source == source_of_interest)) else TRUE) |>
    ungroup()
  
  n_vars <- nrow(input_vars)
  
  # Generate LHS
  lhs_samples <- randomLHS(n_samples, n_vars)
  # var_df_names <- c()
  
  # somthing like this should work to get the nested tibble with data of samples
  sample_df_var <-  
    input_vars |> 
    mutate(nvar = c(1:n_vars)) |> 
    rowwise() |> 
    mutate(
      data = 
        case_match(Distribution,
                   "Triangular" ~ list(tibble(value=triangular(lhs_samples[, nvar], a, b, c))),
                   "Uniform" ~  list(tibble(value=uniform(lhs_samples[, nvar], a, b))),
                   "Powerlaw" ~  list(tibble( value=power_law(lhs_samples[, nvar], a, b, c))),
                   "Trapezoidal" ~  list(tibble( value=trapezoidal(lhs_samples[, nvar], a, b, c, d))),
                   .default = NA
        )
    ) 
  
  # sample_df_var <- sample_df_var |>
  #   mutate(Polymer = pol) |>
  #   mutate(Source = source)
  
  Material_Parameters_n <- rbind(Material_Parameters_n, sample_df_var)
}

#TODO before explosion export table with some statistics of applied variable ranges for reporting in relation to input variables of each distribution. Aim at: min, p5, p50, mean, p95, max.

scales <- union((World$FromDataAndTo()$fromScale),(World$FromDataAndTo()$toScale))
subCompartments <-  union((World$FromDataAndTo()$fromSubCompart),(World$FromDataAndTo()$toSubCompart))
species <- union((World$FromDataAndTo()$fromSpecies),(World$FromDataAndTo()$toSpecies))

Material_Parameters_n <- 
  Material_Parameters_n %>% 
  separate_rows(Species, sep = "_") |> 
  separate_rows(SubCompart, sep = "_") |>
  mutate( Scale = str_replace_all(Scale, "any",
                                  paste(scales,collapse="__"))) |> 
  separate_rows(Species, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Water",
                                       paste(c("lake", "sea", "deepocean", "river"),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Soil",
                                       paste(subCompartments |> str_subset(c("soil")),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Sediment",
                                       paste(subCompartments |> str_subset(c("sediment")),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( Species = str_replace_all(Species, "any",
                                    paste(species,collapse="__"))) |> 
  separate_rows(Species, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "any",
                                       paste(subCompartments,collapse="__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  rename(Source = MP_source)

```

```{r Make some figures to show the values chosen from the distributions}

#Example of plot:
#TODO: include units

PlotVariable = "alpha"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(PlotVariable) + scale_x_log10() 

PlotVariable = "RadS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) + scale_x_log10() 

PlotVariable = "RhoS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kdeg"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kfrag"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable)) 
```

# Solve the matrix and calculate concentrations

Solve the matrix depending on the type of input emissions given (PMFA =
Steady state, DPMFA = dynamic)

```{r Solve steady state}
start_time <- Sys.time()

Output <- tibble(Polymer = polymers, 
                 SBoutput = NA)

if(MFAtype == "PMFA"){ 
  World$NewSolver("UncertainSolver")
  
  for(pol in polymers){
    emis_source <- DPMFA_sink_micro |>
      filter(Polymer == pol) |>
      select(Abbr, Year, Polymer, Emis)
    
    sample_source <- Material_Parameters_n |>
      filter(Polymer == pol) |>
      select(VarName, Scale, SubCompart, Species, data) |> 
      rename(varName = VarName)
    
    solved <- World$Solve((emis_source), needdebug = F, sample_source)
    solved$SteadyStateConc <- World$GetConcentration()
    
    Output$SBoutput[Output$Polymer == pol] <- list(solved)
  }
  
} else if(MFAtype == "DPMFA"){
  World$NewSolver("UncertainDynamicSolver")
  tmax <- max(DPMFA_sink_micro$Timed)
  
  for(pol in polymers){
    emis_source <- DPMFA_sink_micro |>
      filter(Polymer == pol) |>
      select(Abbr, Timed, Emis)
    
    sample_source <- Material_Parameters_n |>
      filter(Polymer == pol) |>
      select(VarName, Scale, SubCompart, Species, data) |> 
      rename(varName = VarName)
    
    solved <- World$Solve((emis_source), sample_source, tmax = tmax, needdebug = F)
    solved$DynamicConc <- World$GetConcentration()
    
    Output$SBoutput[Output$Polymer == pol] <- list(solved)
    
  }
}
end_time <- Sys.time()

elapsed_time <- end_time - start_time

print(paste0("Elapsed time is ", elapsed_time))

# Save the ouput as RData file
if(is.na(source_of_interest)){
  save(Output,
     file = paste0("vignettes/Case studies/Casedata/LEON-T_output_OTHER_", MFAtype, "_", as.character(n_samples) , "_RUNS_", format(Sys.Date(),"%Y_%m_%d"),".RData"),
     compress = "xz",
     compression_level = 9)   
} else if(source_of_interest == "Tyre wear"){
  save(Output,
     file = paste0("vignettes/Case studies/Casedata/LEON-T_output_TYRE_WEAR_", MFAtype, "_", as.character(n_samples) , "_RUNS_", format(Sys.Date(),"%Y_%m_%d"),".RData"),
     compress = "xz",
     compression_level = 9)  
}
```

# paralel version

```{r}
# library(foreach)
# registerDoParallel(2)  # use multicore, set to the number of our cores
# 
# out.data <- 
#   foreach(i=c(1:nrow(unique_combinations)), .inorder=FALSE, .errorhandling = "pass") %dopar%  {
#     
#     source <- unique_combinations$Source[i]
#     pol <- unique_combinations$Polymer[i]
#     
#     emis_source <- DPMFA_sink_micro |>
#       filter(Source == source) |>
#       filter(Polymer == pol) |>
#       select(Source, Abbr, Year, Polymer,Emis)
#     
#     sample_source <- Material_Parameters_n |>
#       filter(Source == source) |>
#       filter(Polymer == pol) |>
#       select(VarName, Scale, SubCompart, Species, data) |> 
#       rename(varName = VarName)
#     
#     solved <- World$Solve((emis_source), needdebug = F, sample_source)
#     solved$SteadyStateConc <- World$GetConcentration()
#     solved
#   }
# 
# stopImplicitCluster()
# 
# # equivalent to colMeans(m)
# m <- matrix(rnorm(9), 3, 3)
# foreach(i=1:ncol(m), .combine=c) %do%
#   mean(m[,i])
# foreach(i=1:nrow(m), .combine=rbind) %dopar%
#   (m[i,] / mean(m[i,]))

```

# Prep data and make plots

```{r Load output data}
otherdataname <- "LEON-T_output_OTHER_DPMFA_10_RUNS_2024_11_04.RData"
twdataname <- "LEON-T_output_TYRE_WEAR_DPMFA_10_RUNS_2024_11_01.RData"

data_folder <-  "vignettes/Case Studies/CaseData/"

Solution <- data.frame()
Concentrations <- data.frame()

load(paste0(data_folder, otherdataname))
for(i in 1:nrow(Output)){
  
  
  
}

concentration_other <- concentration_df |>
  mutate(Source = "Other sources")
input_vars_other <- input_var_df
solution_other <- solved_df |>
  mutate(Source = "Other sources")

load(paste0(data_folder, twdataname))
concentration_tw <- concentration_df |>
  mutate(Source = "Tyre wear")
input_vars_tw <- input_var_df
solution_tw <- solved_df |>
  mutate(Source = "Tyre wear")

concentration_df <- rbind(concentration_other, concentration_tw)
solution_df <- rbind(solution_other, solution_tw)

```

```{r}

# Create a plot theme
plot_theme <-  theme(
  axis.title.x = element_text(size = 14),    
  axis.title.y = element_text(size = 14),    
  axis.text.x = element_text(size = 12, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 12),
  title = element_text(size=20),
  panel.background = element_rect(fill = "white"),  # White background
  panel.grid.major = element_line(color = "lightgrey", size = 0.5),  # Major grid lines in light grey
  panel.grid.minor = element_line(color = "lightgrey", size = 0.25)
)

# Plot the data for each polymer separately, for all selected compartments (different plots depending on PMFA or DPMFA)
if(type == "PMFA"){
  for(mat in polymers){
    solution_name <- paste0("solution_", mat)
    df_name <- paste0("solved_ss_", mat)
    
    Solution <- get(solution_name) |>
      mutate(RUN = as.integer(RUN)) |>
      # mutate(EqMass = as.double(EqMass)) |>
      mutate(Concentration = as.double(Concentration)) 
    
    solved <- get(df_name)
    
    subt <- paste0(mat, ", ", y)
    
    # Prepare data for plots
    Input_Variables <- 
      solved$Input_Variables |> unnest(data) 
    Input_Emission <- 
      solved$Input_Emission |> unnest(Emis) 
    
    Plot_data <- 
      Input_Variables |> 
      pivot_wider(names_from = c(varName,Scale,SubCompart,Unit), values_from = value) |> 
      full_join(Solution) |>
      full_join(Input_Emission, by = c("Abbr", "RUN"))
    
    # Aggregate plot data for all species
    Plot_data_agg <- Plot_data |>
      group_by(Scale, SubCompart, RUN, Unit.x, Units_per_SubCompart) |>
      summarise(EqMass = sum(EqMass), Concentration = sum(Concentration))
    
    scales <- unique(Plot_data_agg$Scale)
    
    # Make plots for each scale, one with mass and one with concentrations
    for (scale in scales) {
      df <- Plot_data_agg |>
        filter(Scale == scale) |>
        mutate(concname = paste0(SubCompart, " (", Units_per_SubCompart, ")"))
      
      # Filter out the compartments without mass
      masscomps <- df |>
        filter(RUN == 1) |>
        filter(EqMass != 0)
      
      masscomps <- masscomps$SubCompart
      
      df <- df |>
        filter(SubCompart %in% masscomps)
      
      # Mass plot
      mass_p <- ggplot(df, mapping = aes(x = SubCompart, y = EqMass, fill = SubCompart)) +  
        geom_violin() + 
        labs(title = paste0("Masses at ", scale, " scale"),
             subtitle = subt,
             x = "Compartment",
             y = "Mass (kg)") +
        plot_theme +
        scale_y_continuous(trans = 'log10') +
        scale_fill_viridis_d() + 
        # Customize the legend for this plot as well
        theme(axis.text.x = element_blank(),  # Remove x-axis labels
              legend.position = "bottom") +   # Move legend to the bottom
        guides(fill = guide_legend(title = NULL))  # Remove legend title
      
      print(mass_p)
      
      ggsave(paste0("vignettes/Case Studies/CaseData/Mass_plot_", scale, ".png"), plot=mass_p)
      
      # Concentration plot
      conc_p <- ggplot(df, mapping = aes(x = concname, y = Concentration, fill = concname)) +  
        geom_violin() +
        labs(title = paste0("Concentrations at ", scale, " scale"),
             subtitle = subt,
             x = "Compartment",
             y = "Concentration") +
        plot_theme +
        scale_y_continuous(trans = 'log10') +
        scale_fill_viridis_d() +
        theme(axis.text.x = element_blank(),  # Remove x-axis labels
              legend.position = "bottom") +   # Move legend to the bottom
        guides(fill = guide_legend(title = NULL))  # Remove legend title
      
      print(conc_p)
      
      ggsave(paste0("vignettes/Case Studies/CaseData/Concentration_plot_", scale, ".png"), plot=conc_p)
    }
  }
  
} else if(type == "DPMFA"){        
  
  # TO DO: Make plots for DPMFA output
  
}

```
