---
title: "DPMFA LEON-T"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

# *Initializing World and Substance*

Depending on the substance you want to analyze, a selection needs to be
made from the three different"worlds" : Molecular, Particulate and
Plastics. Here, we initialize the world for plastics.

```{r Initialize World}
library(lhs)
library(readxl)
library(viridis)
library(doParallel)

source("baseScripts/initWorld_onlyPlastics.R")

World$substance <- "microplastic"
```

# Load the DPMFA data and make a nested emission dataframe

```{r Load DPMFA data}
abspath <- "R:/Projecten/E121554 LEON-T/03 - uitvoering WP3/DPMFA_output/Baseline_PMFA_EU.RData"

load(abspath)

# Check if the loaded data is DPMFA or PMFA data
if(exists("DPMFA_stocks")) {
  type <- "DPMFA"
} else {
  type <- "PMFA"
}

# Convert to long format
data_long <- 
  DPMFA_sink |> unnest(Mass_Polymer_kt, keep_empty = TRUE) |> 
  pivot_longer(cols=-c(Type, Scale, Source, Polymer, To_Compartment, Material_Type, iD_source, RUN),
               names_to = "Year",
               values_to = "Mass_Polymer_kt")

# Choose a source of interest if you want, else write NA
source_of_interest <- "Tyre wear"

if(!is.na(source_of_interest)){
  if(!source_of_interest %in% unique(data_long$Source)){
    print("Selected source not in dataframe")
  } else {
    filtersource <- source_of_interest
  }
} else if (is.na(source_of_interest)){
  filtersource <- unique(data_long$Source)
}

# Calculate kg/s from kt/y 
data_summed <- data_long |>
  filter(Source %in% filtersource) |>
  filter(between(RUN, 1, 10)) |>
  mutate(Mass_Polymer_kg_s = Mass_Polymer_kt*1000000/(365.25*24*3600)) |>
  filter(Material_Type == "micro") 

if(unique(data_long$Scale) == "EU"){
  SBscale <- "C"
} else if (unique(data_long$Scale) == "NL") {
  SBscale <- "R"
}

# Assign SB compartments to DPMFA compartments
data_filtered <- data_summed |>
  select(To_Compartment, Mass_Polymer_kg_s, Year, RUN, Polymer) |>
  mutate(Scale = SBscale) |>
  mutate(Compartment = case_when(
    str_detect(To_Compartment, "soil") ~ "s",
    str_detect(To_Compartment, "water") ~ "w",
    str_detect(To_Compartment, "air") ~ "a"
  )) |>
  mutate(Subcompartment = case_when(
    str_detect(To_Compartment, "Agricultural") ~ "2",
    str_detect(To_Compartment, "Natural") ~ "1",
    str_detect(To_Compartment, "Sub-surface") ~ "3",
    str_detect(To_Compartment, "Road side") ~ "3",
    str_detect(To_Compartment, "Residential") ~ "3",
    str_detect(To_Compartment, "Sea") ~ "2",
    str_detect(To_Compartment, "Surface") ~ "1",
    str_detect(To_Compartment, "Outdoor") ~ ""
  )) |>
  mutate(Species = "S") |>
  mutate(Abbr = paste0(Compartment, Subcompartment, Scale, Species)) |>
  group_by(Abbr, Year, RUN, Polymer) |>
  summarise(Mass_Polymer_kg_s = sum(Mass_Polymer_kg_s)) |>
  ungroup() |>
  rename(value = Mass_Polymer_kg_s) |>
  select(Abbr, Year, Polymer, value, RUN)

```

```{r Make a separate emission df for every polymer}
polymers <- unique(data_filtered$Polymer)

for(i in polymers){
 filtered <- data_filtered |>
   filter(Polymer == i)
 
 df_names <- c()
 
 if(type == "DPMFA"){
   # Make an emission dataframe for the dynamic uncertain solver
   emis_df_dyn <- filtered |>
     group_by(Abbr, Year) |>
     rename(Timed = Year) |>
     mutate(Timed = as.double(Timed)*(365.25*24*3600)) |>
     nest(Emis = c(RUN, value)) 
   
   df_name <- paste0("emis_dyn_", i)
   
   assign(df_name, emis_df_dyn)

   df_names <- c(df_names, df_name)
   
   ymin <- min(emis_df_dyn$Year)
   ymax <- max(emis_df_dyn$Year)
   
 } else if(type == "PMFA"){
  
   # Select data for 2019
   y <- 2019
  
   # Make an emission dataframe for the steady uncertain solver
   emis_df_ss <- filtered |>
     filter(Year == y) |>
     nest(Emis = c(RUN, value)) 
   
   df_name <- paste0("emis_ss_", i)
   
   assign(df_name, emis_df_ss)
   
   df_names <- c(df_names, df_name)
 }
}

```

```{r Functions for distributions}
# Define triangular distribution function
triangular <- function(u, a, b, c) {               # u = samples, a = min, b = max, c = peak
  ifelse(u < (c-a)/(b-a),
         a + sqrt(u * (b-a) * (c-a)),
         b - sqrt((1-u) * (b-a) * (b-c)))
}

# Define uniform distribution function
uniform <- function(u, a, b) {                     # u = samples, a = min, b = max
  transformed_samples <- a + (b - a) * u
  return(transformed_samples)
}

# Define power law distribution function
power_law <- function(u, a, b, c){                 # u = samples, a = min, b = max, c = alpha

  # Ensure that samples are within [0, 1]
  samples <- pmin(pmax(u, 0), 1)
  
  # Transform samples to the power-law distribution
  scaled_samples <- a * ((b / a) ^ samples) ^ (1 / (1 - c))
  
  return(scaled_samples)
}

```

# Create dataset with uncertain variables

## Prepare excel data

```{r}
# Path to excel file with distribution values
path_dist <- "N:/Documents/Simplebox/Microplastic_variables_v1.xlsx"

excel_df <- read_excel(path_dist, sheet = "Polymer_data")

materials <- unique(excel_df$Polymer)

water <- c("lake", "sea", "deepsea", "river")
soil <- c("naturalsoil", "agriculturalsoil", "othersoil")
sediment <- c("marinesediment", "freshwatersediment", "lakesediment")
soil_sediment <- c(soil, sediment)
air <- c("air", "cloudwater")
all_SC <- c(water, soil, sediment, air)
materials <- c("HDPE", "LDPE", "PP", "PS", "PVC", "Acryl", "PA", "PET", "ABS", "EPS", "PC", "PMMA", "PUR", "RUBBER", "OTHER")

scales <- c("Arctic", "Tropic", "Moderate", "Regional", "Continental")

explode <- function(df, target_col, explode_value, new_values) {
  df %>%
    # Use mutate to create a new column if the target column equals explode_value
    mutate(!!sym(target_col) := ifelse(!!sym(target_col) == explode_value, list(new_values), !!sym(target_col))) %>%
    # Unnest the target column to duplicate rows
    unnest(!!sym(target_col))
}

var_df <- explode(excel_df, target_col = "Polymer", explode_value = "any", new_values = materials) |>
  mutate(a = as.numeric(a)) |>
  mutate(b = as.numeric(b)) |>
  mutate(c = as.numeric(c))
```

```{r}
# Generate the correct number of samples
n_samples <- nrow(emis_df_ss$Emis[[1]]) # Number of emission runs 
sample_df_names <- c()

for(i in polymers){
  
  input_vars <- var_df |>
    filter(if_all(c(Distribution, a, b), ~ !is.na(.))) |>
    filter(Polymer == i) 
  
  varnames <- input_vars$VarName
  
  n_vars <- nrow(input_vars)
  
  # Generate LHS
  lhs_samples <- randomLHS(n_samples, n_vars)
  var_df_names <- c()
  
  for(j in 1:nrow(input_vars)){
    df <- input_vars[j, ]
    
    varname <- df$VarName
      
    var <- df |>
      select(VarName, Scale, SubCompart, a, b, c)
      
    name <- paste0("var", j)
    var_df_names <- c(var_df_names, name)
      
    assign(name, var)
  }  
  
  params <- tibble(
    varName = sapply(var_df_names, function(v) get(v)$VarName),
    Scale = sapply(var_df_names, function(v) get(v)$Scale),
    SubCompart = sapply(var_df_names, function(v) get(v)$SubCompart),
    data = lapply(var_df_names, function(v) {
      df <- get(v)
      tibble(id = c("a", "b", "c"), value = c(df$a, df$b, df$c))
    }))
  
  sample_df <- params

  # Transform each LHS sample column to the corresponding triangular distribution
  for (k in 1:n_vars) {
    #print(params$varName[[k]])
    a <- filter(params$data[[k]], id == "a") %>% pull(value)
    b <- filter(params$data[[k]], id == "b") %>% pull(value)
    c <- filter(params$data[[k]], id == "c") %>% pull(value)
    
    if(input_vars$Distribution[k] == "Triangular"){
      samples <- triangular(lhs_samples[, k], a, b, c)
    } else if(input_vars$Distribution[k] == "Uniform"){
      samples <- uniform(lhs_samples[, k], a, b)
    } else if(input_vars$Distribution[k] == "Powerlaw"){
      samples <- power_law(lhs_samples[, k], a, b, c)
    }

    # Create a new tibble for 'data' with samples replacing original values
    new_data <- tibble(value = samples)

    # Update the data column in the sample_df
    sample_df$data[[k]] <- new_data
  }
  
  # Save a separate dataframe for each material
  sample_df_name <- paste0("sample_df_",i)
  sample_df_names <- paste0(sample_df_names, sample_df_name)
  
  assign(sample_df_name, sample_df)
}

# Remove illegal Scale-SubCompart combinations from the dataframes
allowd_combs <- World$fetchData("Area") |>
    select(Scale, SubCompart)


for(sample_df_name in sample_df_names) {
  sample_df <- get(sample_df_name)

  # Step 1: Explode for Scale
  exploded_scales <- explode(sample_df, target_col = "Scale", explode_value = "any", new_values = scales)
  print(paste("After exploding scales:", nrow(exploded_scales)))  # Check size

  # Step 2: Explode for SubCompart: water
  exploded_water <- explode(exploded_scales, target_col = "SubCompart", explode_value = "water", new_values = water)
  print(paste("After exploding water:", nrow(exploded_water)))  # Check size

  # Step 3: Explode for SubCompart: soil_sediment
  exploded_soil_sediment <- explode(exploded_water, target_col = "SubCompart", explode_value = "soil_sediment", new_values = soil_sediment)
  print(paste("After exploding soil_sediment:", nrow(exploded_soil_sediment)))  # Check size

  # Step 4: Explode for SubCompart: any
  exploded_subcomparts <- explode(exploded_soil_sediment, target_col = "SubCompart", explode_value = "any", new_values = all_SC)
  print(paste("After exploding any:", nrow(exploded_subcomparts)))  # Check size

  # Step 5: Perform the semi-join
  sample_df <- exploded_subcomparts %>%
    semi_join(allowd_combs, by = c("Scale", "SubCompart"))
  print(paste("After semi-join:", nrow(sample_df)))  # Check size

  assign(sample_df_name, sample_df)  # Save it back to the environment
}

```

# Solve the matrix

Solve the matrix depending on the type of input emissions given (PMFA =
Steady state, DPMFA = dynamic)

```{r Solve steady state}

if(type == "PMFA"){
  World$NewSolver("UncertainSolver")
  
  for(i in polymers){
    emis_df_name <- paste0("emis_ss_", i)
    sample_df_name <- paste0("sample_df_", i)
    
    solved <- World$Solve(get(emis_df_name), needdebug = FALSE, get(sample_df_name))
    assign(paste0("solved_ss_", i), solved)
  }
  
} else if(type == "DPMFA"){
  World$NewSolver("UncertainDynamicSolver")
  
  for(i in polymers){
    emis_df_name <- paste0("emis_dyn_", i)
    sample_df_name <- pasteo("sample_df_", i)
    
    solved <- World$Solve(World$Solve(get(i)), sample_df, tmax = tmax, needdebug = F)
    assign(paste0("solved_dyn_", i), solved)
  }
}

```

# Calculate concentrations

```{r Calculate concentrations}
for(mat in polymers){ 
  dfname <- paste0("solved_ss_", mat)
    
  sol <- get(dfname)
    
  Solution <- sol$SteadyStateMass |>
    mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
    mutate(Abbr = str_replace_all(Abbr, "cw", "a")) |>
    group_by(RUN, Scale, SubCompart, Species, Unit, Abbr) |>
    summarise(EqMass = sum(EqMass)) |>
    ungroup() 
    
  # Fetch necessary data
    
  # Sum cloudwater and air under air
  Volume <- World$fetchData("Volume") |>
    mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
    group_by(Scale, SubCompart) |>
    summarise(Volume = sum(Volume)) |>
    ungroup()
    
  Area <- World$fetchData("Area")
  FRACw <- World$fetchData("FRACw")
  FRACa <- World$fetchData("FRACa")
  Fractrial <- FRACa$FRACa[FRACa$SubCompart =="air" & FRACa$Scale =="Arctic" ]
  Rho <- World$fetchData("rhoMatrix")
  Concentration_eq <- merge(Solution, Volume, by = c("SubCompart", "Scale"))
  Concentration_eq$Concentration <- Concentration_eq$EqMass / Concentration_eq$Volume
  RhoWater_value <- Rho$rhoMatrix[Rho$SubCompart == "river"]
  
  f_adjust_concentration <- function(Concentration, FRACw, FRACa, SubCompart, Scale, Rho, RhoWater_value) {
    # Fetch Fracw based on SubCompart and Scale
    Fracw <- FRACw$FRACw[FRACw$SubCompart == SubCompart & FRACw$Scale == Scale]
    
    # Fetch Fraca based on SubCompart and Scale
    Fraca <- FRACa$FRACa[FRACa$SubCompart == SubCompart & FRACa$Scale == Scale]
    
    # Fetch RHOsolid based on SubCompart
    RHOsolid <- Rho$rhoMatrix[Rho$SubCompart == SubCompart]
      
    # Check if any of Fracw, Fraca, or RHOsolid are N
      
    Concentration * 1000 / (Fracw * RhoWater_value + (1 - Fracw - Fraca) * RHOsolid)
  }
    
  subcomparts <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment")
    
  filtered_data <- Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, ]
    
  # Apply the adjustment function to the filtered data
  adjusted_concentrations <- apply(filtered_data, 1, function(row) {
    f_adjust_concentration(
      Concentration = as.numeric(row["Concentration"]),
      FRACw = subset(FRACw, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
      FRACa = subset(FRACa, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
      Rho = subset(Rho, SubCompart == row["SubCompart"]),
      RhoWater_value = 998,  # Replace with your actual RhoWater_value
      SubCompart = row["SubCompart"],
      Scale = row["Scale"]
    )
  })
  # Update Concentration_eq with adjusted concentrations
  Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, "Concentration"] <- adjusted_concentrations
  
  #Define the units
  subcompart <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment",  "lakesediment", "air", "deepocean", "lake" , "river", "sea")
  units <- c("g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw", "g/kg dw",
             "kg/m^3", "kg/L", "kg/L", "kg/L", "kg/L", "kg/L")
    
  # Combine into a named list
  subcompart_units <- setNames(units, subcompart)
  Concentration_eq <- Concentration_eq |>
    mutate(Units_per_SubCompart = subcompart_units[SubCompart])
    
  convert_units <- function(concentration, unit) {
    if (unit == "g/kg dw") {
      return(concentration * 1e9)
    } else if (unit == "kg/m^3") {
      return(concentration * 1e12)
    } else if (unit == "kg/L") {
      return(concentration * 1e12)
    } else {
      return(concentration)
    }
  }
  
  # Convert concentrations to ng/kg or ng/L and update units
  Concentration_eq <- Concentration_eq |>
    mutate(
           Concentration = mapply(convert_units, Concentration, Units_per_SubCompart),
           Units_per_SubCompart = ifelse(Units_per_SubCompart == "g/kg dw", "ng/kg dw", 
                                         ifelse(Units_per_SubCompart == "kg/m^3", "ng/m^3", "ng/L")))
  Concentration_eq <- Concentration_eq |>
    mutate(across(where(is.numeric), ~ format(., scientific = TRUE)))
  
  solname <- paste0("Solution_", mat)
  assign(solname, Concentration_eq)
}
  
```

# Prep data and make plots

```{r Solve steady state}

# Create a plot theme
plot_theme <-  theme(
  axis.title.x = element_text(size = 14),    
  axis.title.y = element_text(size = 14),    
  axis.text.x = element_text(size = 12, angle = 45, hjust = 1),     
  axis.text.y = element_text(size = 12),
  title = element_text(size=20),
  panel.background = element_rect(fill = "white"),  # White background
  panel.grid.major = element_line(color = "lightgrey", size = 0.5),  # Major grid lines in light grey
  panel.grid.minor = element_line(color = "lightgrey", size = 0.25)
)

# Plot the data for each polymer separately, for all selected compartments (different plots depending on PMFA or DPMFA)
if(type == "PMFA"){
  for(mat in polymers){
    solution_name <- paste0("Solution_", mat)
    df_name <- paste0("solved_ss_", mat)
    
    Solution <- get(solution_name) |>
      mutate(RUN = as.integer(RUN)) |>
      mutate(EqMass = as.double(EqMass)) |>
      mutate(Concentration = as.double(Concentration)) 
    
    solved <- get(df_name)
    
    subt <- paste0(mat, ", ", y)
    
    # Prepare data for plots
    Input_Variables <- 
      solved$Input_Variables |> unnest(data) 
    Input_Emission <- 
      solved$Input_Emission |> unnest(Emis) 
    
    Plot_data <- 
      Input_Variables |> 
      pivot_wider(names_from = c(varName,Scale,SubCompart,Unit), values_from = value) |> 
      full_join(Solution) |>
      full_join(Input_Emission, by = c("Abbr", "RUN"))
    
    # Aggregate plot data for all species
    Plot_data_agg <- Plot_data |>
      group_by(Scale, SubCompart, RUN, Unit.x, Units_per_SubCompart) |>
      summarise(EqMass = sum(EqMass), Concentration = sum(Concentration))
    
    scales <- unique(Plot_data_agg$Scale)
    
    # Make plots for each scale, one with mass and one with concentrations
    for (scale in scales) {
      df <- Plot_data_agg |>
        filter(Scale == scale) |>
        mutate(concname = paste0(SubCompart, " (", Units_per_SubCompart, ")"))
      
      # Filter out the compartments without mass
      masscomps <- df |>
        filter(RUN == 1) |>
        filter(EqMass != 0)
      
      masscomps <- masscomps$SubCompart
      
      df <- df |>
        filter(SubCompart %in% masscomps)
      
      # Mass plot
      mass_p <- ggplot(df, mapping = aes(x = SubCompart, y = EqMass, fill = SubCompart)) +  
        geom_violin() + 
        labs(title = paste0("Masses at ", scale, " scale"),
             subtitle = subt,
             x = "Compartment",
             y = "Mass (kg)") +
        plot_theme +
        scale_y_continuous(trans = 'log10') +
        scale_fill_viridis_d() + 
        # Customize the legend for this plot as well
        theme(axis.text.x = element_blank(),  # Remove x-axis labels
              legend.position = "bottom") +   # Move legend to the bottom
        guides(fill = guide_legend(title = NULL))  # Remove legend title
      
      print(mass_p)
      
      ggsave(paste0("N:/Documents/Simplebox/LEON-T_plots/Mass_plot_", scale, ".png"), plot=mass_p)
      
      # Concentration plot
      conc_p <- ggplot(df, mapping = aes(x = concname, y = Concentration, fill = concname)) +  
        geom_violin() +
        labs(title = paste0("Concentrations at ", scale, " scale"),
             subtitle = subt,
             x = "Compartment",
             y = "Concentration") +
        plot_theme +
        scale_y_continuous(trans = 'log10') +
        scale_fill_viridis_d() +
        theme(axis.text.x = element_blank(),  # Remove x-axis labels
              legend.position = "bottom") +   # Move legend to the bottom
        guides(fill = guide_legend(title = NULL))  # Remove legend title
      
      print(conc_p)
      
      ggsave(paste0("N:/Documents/Simplebox/LEON-T_plots/Concentration_plot_", scale, ".png"), plot=conc_p)
    }
  }

} else if(type == "DPMFA"){        
  
  # TO DO: Make plots for DPMFA output

}

```
