---
title: "ODE solver set up"
author: " Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

# *Initializing World and Substance*

Depending on the substance you want to analyze, a selection needs to be
made from the three different"worlds" : Molecular, Paticulate and
Plastics. Here, we initialize the world for particulates.

```{r Initialize World}
source("baseScripts/initWorld_onlyParticulate.R")


```

Next, we initialize the solver. SimpleBox in R gives the opportunity to
choose from multiple solvers, but here we choose the Base-R solve.

```{r SBsteady}
World$NewSolver("SB1Solve")
```

What solving means is that using matrix algebra a set of differential
equations is solved:

`K %*% m + e`

Where:

K is the matrix of rate constants for each process describing the mass
transfers to and from and out of a state (e.g. substance in freshwater
(w1U) or small heteroagglomerate in natural soil (s1A)).

m is the mass in each compartment, e.g. 0 at t=0.

e is the emission to each compartment per unit of time, e.g. 1 t/y.

# *Testing the Solver with real- life emissions*

To test this solver, we use data from a case study on GO-Chitosan. The
emission data is based on Monte Carlo simulations of a DMPFA model. To
test, we first select emissions from one scenario and towards one
compartment.

```{r emission data}
file_paths <- 
  list.files("data/emissions",recursive = TRUE)
Emissions <-
  read_csv(paste0("data/emissions/",file_paths), id="file_name", col_names = c("RUN",0:24),skip = 1)

Emissions <- 
  Emissions |>
  pivot_longer(
    cols = !c(file_name,RUN),
    names_to = "year",
    values_to = "emission_t" ) |> mutate_at('year',as.numeric) |> 
  ungroup() |> 
  group_by(file_name,year) |> 
  summarise(Emission_p50_kg = quantile(emission_t,probs = 0.5),
            Emission_mean = mean(emission_t)) |> ungroup()

Emissions <- 
  Emissions |> 
  mutate(compartment = 
           case_when(str_detect(file_name, "(?i)Air") ~ "Air",
                     str_detect(file_name, "(?i)Soil") ~ "SludgeTreatedSoil",
                     str_detect(file_name, "(?i)Water") ~ "SurfaceWater",
                     str_detect(file_name, "(?i)Subsurface") ~ "Subsurface",
                     TRUE ~ "Other"),
         scale = 
           case_when(str_detect(file_name, "(?i)EU") ~ "EU_average",
                     str_detect(file_name, "(?i)Ireland") ~ "EU_STsoil",
                     str_detect(file_name, "(?i)Switzerland") ~ "EU_noSTsoil",
                     TRUE ~ "Other"),
         Substance = "GO-Chitosan",
  )
Emissions <- 
  Emissions |> 
  mutate(compartment = 
           case_when(str_detect(file_name, "(?i)Air") ~ "Air",
                     str_detect(file_name, "(?i)Soil") ~ "SludgeTreatedSoil",
                     str_detect(file_name, "(?i)Water") ~ "SurfaceWater",
                     str_detect(file_name, "(?i)Subsurface") ~ "Subsurface",
                     TRUE ~ "Other"),
         scale = 
           case_when(str_detect(file_name, "(?i)EU") ~ "EU_average",
                     str_detect(file_name, "(?i)Ireland") ~ "EU_STsoil",
                     str_detect(file_name, "(?i)Switzerland") ~ "EU_noSTsoil",
                     TRUE ~ "Other"),
         Substance = "GO-Chitosan",
  )

Emissions <- Emissions |>
  filter(compartment == "SurfaceWater" & scale == "EU_average") |> 
  select(everything())


plot(Emissions$year, Emissions$Emission_mean, 
     type = "l",  # "l" for line plot, or use "p" for points
     xlab = "Time", ylab = "Emission_mean_tonnes",
     main = "Emission Mean over Time")

```

# *Solving for Steady State*

We are first going to solve for a steady state output, meaning that we
will read in a dataframe with constant emissions in kg/s. For this, we
take the mean of the yearly averages (which is the mean of the 1000 MC
runs)

```{r Prep Emission Steady}
Emissions <- Emissions |>
  mutate(Abr = "w1CS")
Emissions_avg <- Emissions[-nrow(Emissions), ]

average_emission <- mean(Emissions_avg$Emission_mean[7:23], na.rm = TRUE)
print(average_emission)
emissions <- data.frame(Abbr = "w1CS", 
                        Emis = average_emission)
emissions$Emis<- emissions$Emis * 1000 / (365.25 * 24 * 60 * 60) # convert 1 t/y to si units: kg/s


```

Now we are ready to run the solver, which results in the mass in each
compartment. We filter for unbound species, which are 0 since they are only relevant when running the model for molecules. 

```{r Solution Steady}
SolutionSteady <- World$Solve(emissions)
SolutionSteady <- SolutionSteady |>
  filter(Species != "Unbound")
```

# *Dynamically Solving with average Emissions*

With these average emissions, we are now going to solve dynamically. As
such, we divide the average emission again to be kg/s and use the
deSolve::ODE to run the model.

```{r ODE Steady}
SimpleBoxODE = function(t, m, parms) {

  with(as.list(c(parms, m)), {
    e <- c(rep(0, length(SBNames)))
    for (name in names(emis)) {
      e[grep(name, SBNames)] <- emis[[name]] #* (t[2] -t [1])
    }
    dm <- K%*% m + e
    res <- c(dm)
    list(res, signal = e)
  })
}

SBsolve4 <- function( tmax = 1e10, nTIMES = 100, Engine) {

  SB.K = Engine
  SBNames = colnames(Engine)
  SB.m0 <- rep(0, length(SBNames))
  SBtime <- seq(0,tmax,length.out = nTIMES)
  emis <- data.frame(w1CS = (average_emission*1000)/(365.25*24*3600))


  out <- deSolve::ode(
    y = as.numeric(SB.m0),
    times = SBtime ,
    func = SimpleBoxODE,
    parms = list(K = SB.K, SBNames=SBNames, emis= emis),
    rtol = 1e-10, atol = 1e-2)
  #if(as.character(class(deS)[1])!="data.frame") return (list(errorstate="error", deS))
  colnames(out)[1:length(SBNames)+1] <- SBNames
  colnames(out)[grep("signal",colnames(out))] <- paste("emis",SBNames,sep = "2")
  as.data.frame(out) }

SolutionConstantEmissions <- SBsolve4(tmax = 24*(365.25*24*3600),
  nTIMES = 25,
  Engine = World$exportEngineR())

```

```{r Emission Check}
ggplot(SolutionConstantEmissions, aes(x = time, y = emis2w1CS)) +
  geom_line() +  
  #geom_hline(yintercept = 43*1000, linetype = "dashed", color = "red") +# Add a line for w1Cs2emis over time
  labs(title = "Plot of w1Cs2emis against Time",
       x = "Time",
       y = "w1Cs2emis") +   # Add labels for axes and title
  theme_minimal()      # Use a minimal theme
```

## Comparing Steady and constant
In these plots we compare the output of two compartments for both deterministic Steady State calculation and a dynamic calculation with average emissions. 

```{r plots Steady vs Constant Emissions}
library (ggplot2)
w1CS_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "w1CS"]
ggplot(SolutionConstantEmissions, aes(x = time, y = w1CS)) +
  geom_line(aes(color = "Average Emissions")) +
  geom_hline(aes(yintercept = w1CS_value, color = "Steady"), linetype = "dashed") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue")) +
  labs(title = "Plot of w1CS against Time",
       x = "Time [s]",
       y = "w1CS [kg]",
       color = "Legend") +  # Adding legend title
  theme_minimal()

w1CA_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "w1CA"]  
ggplot(SolutionConstantEmissions, aes(x = time, y = w1CA)) +
  geom_line(aes(color = "Average Emissions")) +
  geom_hline(aes(yintercept = w1CA_value, color = "Steady"), linetype = "dashed") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue")) +
  labs(title = "Plot of w1CA against Time",
       x = "Time [s]",
       y = "w1CA [kg]",
       color = "Legend") +  # Adding legend title
  theme_minimal()
```

# *Testing ODE's ApproxFun*

In this part, we are testing the approxfun from base-R. We will later
use this to interpolate between the emission data points. However, first
we test it on the average emission to ensure it generates the same
output.

```{r ODE approxfun}
tmax <- 25 * (365.25 * 24 * 3600)  # Total number of seconds in 25 years
#generating approxfun for constant emissions
# Calculate emissions (constant rate)
emissions <- (average_emission * 1000) / ( 365.25 * 24 * 3600)  # Emissions rate in kg/second
times <- seq(0, tmax, length.out = 1000)
# Generate time points (assuming you want to interpolate over some interval)
emis_values <- rep(emissions, length(times))
# Create approxfun
emislist <- list(approxfun(times, emis_values))
emis_fun <- emislist[["w1CS"]]
names(emislist) <- "w1CS"




SimpleBoxODE = function(t, m, parms) {

  with(as.list(c(parms, m)), {
    e <- c(rep(0, length(SBNames)))
    for (name in names(emislist)) {
      e[grep(name, SBNames)] <- emislist[[name]](t) 
    }
    dm <- K%*% m + e
    res <- c(dm)
    list(res, signal = e)
  })
}

SBsolve4 <- function( tmax = 1e10, nTIMES = 100, Engine, emislist) {

  SB.K = Engine
  SBNames = colnames(Engine)
  SB.m0 <- rep(0, length(SBNames))
  SBtime <- seq(0,tmax,length.out = nTIMES)


  out <- deSolve::ode(
    y = as.numeric(SB.m0),
    times = SBtime ,
    func = SimpleBoxODE,
    parms = list(K = SB.K, SBNames=SBNames, emislist= emislist),
    rtol = 1e-10, atol = 1e-2)
  #if(as.character(class(deS)[1])!="data.frame") return (list(errorstate="error", deS))
  colnames(out)[1:length(SBNames)+1] <- SBNames
  colnames(out)[grep("signal",colnames(out))] <- paste("emis",SBNames,sep = "2")
  as.data.frame(out) }

SolutionConstantApproxFun <- SBsolve4(tmax = 24*(365.25*24*3600),
  nTIMES = 25,
  Engine = World$exportEngineR(), 
  emislist = emislist)

```

## Comparison of the three results so far

```{r Comparison of 3 results}
w1CA_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "w1CA"]
w1CS_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "w1CS"]
w2CA_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "w2CA"]
s1CA_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "s1CA"]


ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w1CS, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w1CS, color = "Approximation Function avg Emissions"), linetype = "dashed") +
  geom_hline(aes(yintercept = w1CS_value, color = "Steady"), linetype = "dashed") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Approximation Function avg Emissions" = "green")) +
  labs(title = "Plot of w1CS against Time",
       x = "Time [s]",
       y = "w1CS [kg]",
       color = "Legend") +  # Adding legend title
  theme_minimal()

ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w1CA, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w1CA, color = "Approximation Function avg Emissions"), linetype = "dashed") +
  geom_hline(aes(yintercept = w1CA_value, color = "Steady"), linetype = "dashed") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Approximation Function avg Emissions" = "green")) +
  labs(title = "Plot of w1CA against Time",
       x = "Time [s]",
       y = "w1CA [kg]",
       color = "Legend") +  # Adding legend title
  theme_minimal()

ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w2CA, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w2CA, color = "Approximation Function avg Emissions"), linetype = "dashed") +
  geom_hline(aes(yintercept = w2CA_value, color = "Steady"), linetype = "dashed") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Approximation Function avg Emissions" = "green")) +
  labs(title = "Plot of w2CA against Time",
       x = "Time [s]",
       y = "w2CA [kg]",
       color = "Legend") +  # Adding legend title
  theme_minimal()


```

# *Solving for yearly emission data with approxFun*

Here, we will start solving for the actual emission data (one value per
year). We first create the approxfun which we solve in a list to use as
input for the ODE Solver. Importantly, the emissions should be converted
into mass/seconds to ensure proper working of the solver. This is due to
the fact that the matrix solves per second.

## prepping ApproxFun

```{r emission prep}

Emissions <- Emissions |>
  mutate(time_s = year*(365.25*24*60*60)+(365.25*24*60*60))

Emissions <- Emissions |> 
  mutate(Emission_mean_kg = (Emission_mean* 1000)/(365.25*24*60*60) )

SBEmissions <- 
  Emissions |> 
  group_by(Abr) |> 
  summarise(n=n(),
            EmisFun = list(
              approxfun(
                data.frame(time_s = c(0,time_s), 
                           emis_kg=c(0,Emission_mean_kg)),
                rule = 1)
            )
  )

funlist <- SBEmissions$EmisFun
names(funlist) <- "w1CS"
w1CS <- funlist[["w1CS"]]


times <- seq(0, 25*365.25*24*3600, by = 1000)


time <- Emissions$year
time_s <-  c(0, time*(365.25*24*60*60)+(365.25*24*60*60))
emis_kg = c(0,(Emissions |> pull(Emission_mean_kg)))

#after 1
timesemisvalue <- seq(31557600, 31557600*2, by = 1000000)
# Calculate the area under the curve using numerical integration
emission_values <- w1CS(timesemisvalue)




```

## Solving for yearly emission data

```{r original ode with approxfun}

##ODE

SimpleBoxODE = function(t, m, parms) {

  with(as.list(c(parms, m)), {
    e <- c(rep(0, length(SBNames)))

    for (name in names(funlist)) {
      e[grep(name, SBNames)] <- funlist[[name]](t)
    }

    dm <- K%*% m + e
    res <- c(dm)
    list(res, signal = e)
  })
}


#Function to Solve
SBsolve4 <- function( tmax = 1e10, nTIMES = 100, Engine, funlist) {

  SB.K = Engine
  SBNames = colnames(Engine)
  SB.m0 <- rep(0, length(SBNames))
  SBtime <- seq(0,tmax,length.out = nTIMES)


  out <- deSolve::ode(
    y = as.numeric(SB.m0),
    times = SBtime ,
    func = SimpleBoxODE,
    parms = list(K = SB.K, SBNames=SBNames, funlist=funlist),
    rtol = 1e-10, atol = 1e-2)
  #if(as.character(class(deS)[1])!="data.frame") return (list(errorstate="error", deS))
  colnames(out)[1:length(SBNames)+1] <- SBNames
  colnames(out)[grep("signal",colnames(out))] <- paste("emis",SBNames,sep = "2")
  as.data.frame(out)

}

# Can also test with xlsx engine
# read in K matrix from xlsx:
# K.Matrix <- readxl::read_xlsx("vignettes/Development/Quality control/SBExcel/SimpleBox4plastics 4.03_GO-Chitosan.xlsx",
#                               sheet = "engine",
#                               range = "D8:FB163") |> as.data.frame()
#
# colnames(K.Matrix) <- readxl::read_xlsx("vignettes/Development/Quality control/SBExcel/SimpleBox4plastics 4.03_GO-Chitosan.xlsx",
#                               sheet = "engine",
#                               range = "D3:FB3",
#                               col_names = FALSE) |> slice(1) |> as.character()
#
# Solving
Solution <- SBsolve4(tmax = 24*(365.25*24*3600),
nTIMES = 25,
Engine = World$exportEngineR(),
funlist = funlist)

# Solution <- SBsolve4(tmax = 25*(365.25*24*3600),
#                      nTIMES = 130,
#                      Engine = as.matrix(K.Matrix),
#                      funlist = funlist)

SolutionEmissionData <- as.data.frame(Solution)

# Same plot of emission as above
plot(Solution$time,Solution$emis2w1CS)



```

# *Comparison of all mass output together*

```{r comparison to steady}


# Plot for w2CA
plot_w2CA <- ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w2CA, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w2CA, color = "Constant ApproxFun")) +
  geom_line(data = SolutionEmissionData, aes(x = time, y = w2CA, color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = w2CA_value, color = "Steady"), linetype = "dashed") +
  labs(title = "Plot of w2CA against Time",
       x = "Time (s)",
       y = "w2CA [kg]",
       color = "Data Source") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Constant ApproxFun" = "green", "Approxfun Emission Data" = "pink")) +
  theme_minimal()

# Display the plot
print(plot_w2CA)

# Plot for w1CS
plot_w1CS <- ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w1CS, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w1CS,  color = "Constant ApproxFun")) +
  geom_line(data = SolutionEmissionData, aes(x = time, y = w1CS, color = "Approxfun Emission Data")) +
 geom_hline(aes(yintercept = w1CS_value, color = "Steady"), linetype = "dashed") +
  labs(title = "Plot of w1CS against Time",
       x = "Time (s)",
       y = "w1CS [kg]",
       color = "Data Source") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Constant ApproxFun" = "green", "Approxfun Emission Data" = "pink")) +
  theme_minimal()

# Display the plot for w1CS
print(plot_w1CS)

# Plot for w1CA
plot_w1CA <- ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = w1CA, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = w1CA,  color = "Constant ApproxFun")) +
  geom_line(data = SolutionEmissionData, aes(x = time, y = w1CA, color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = w1CA_value, color = "Steady"), linetype = "dashed") + 
  labs(title = "Plot of w1CA against Time",
       x = "Time (s)",
       y = "w2CA [kg]",
       color = "Data Source") +
   scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Constant ApproxFun" = "green", "Approxfun Emission Data" = "pink")) +
  theme_minimal()
print(plot_w1CA)

s1CA_value <- SolutionSteady$EqMass[SolutionSteady$Abbr == "s1CA"]

plot_s1CA <- ggplot() +
  geom_line(data = SolutionConstantEmissions, aes(x = time, y = s1CA, color = "Average Emissions")) +
  geom_line(data = SolutionConstantApproxFun, aes(x = time, y = s1CA,  color = "Constant ApproxFun")) +
  geom_line(data = SolutionEmissionData, aes(x = time, y = s1CA, color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = s1CA_value, color = "Steady"), linetype = "dashed") + 
  labs(title = "Plot of s1CA against Time",
       x = "Time (s)",
       y = "s1CA [kg]",
       color = "Data Source") +
  scale_color_manual(values = c("Steady" = "red", "Average Emissions" = "blue", "Constant ApproxFun" = "green", "Approxfun Emission Data" = "pink")) +
  theme_minimal()

print(plot_s1CA)


```

# *Computation of Concentrations*

To have a better understanding of the orders of magnitude of the output,
we compute the concentrations per compartment. First, we will do so for
the Steady State output. Since this output is relatively
straightforward, the computation is simple.

## Steady State concentrations

For soil and sediment, we decide to compute the output /kg wet weight
(w.w.), as is standard in concentration reporting. This requires some
extra computational steps.

```{r Concentrations Steady}
library(knitr)
Volume <- World$fetchData("Volume")
Area <- World$fetchData("Area")
FRACw <- World$fetchData("FRACw")
FRACa <- World$fetchData("FRACa")
Fractrial <- FRACa$FRACa[FRACa$SubCompart =="air" & FRACa$Scale =="Arctic" ]
Rho <- World$fetchData("rhoMatrix")
Concentration_eq <- merge(SolutionSteady, Volume, by = c("SubCompart", "Scale"))
Concentration_eq$Concentration <- Concentration_eq$EqMass / Concentration_eq$Volume
RhoWater_value <- Rho$rhoMatrix[Rho$SubCompart == "river"]

f_adjust_concentration <- function(Concentration, FRACw, FRACa, SubCompart, Scale, Rho, RhoWater_value) {
  # Fetch Fracw based on SubCompart and Scale
  Fracw <- FRACw$FRACw[FRACw$SubCompart == SubCompart & FRACw$Scale == Scale]
  
  # Fetch Fraca based on SubCompart and Scale
  Fraca <- FRACa$FRACa[FRACa$SubCompart == SubCompart & FRACa$Scale == Scale]
  
  # Fetch RHOsolid based on SubCompart
  RHOsolid <- Rho$rhoMatrix[Rho$SubCompart == SubCompart]
  
  # Check if any of Fracw, Fraca, or RHOsolid are N
  
  Concentration * 1000 / (Fracw * RhoWater_value + (1 - Fracw - Fraca) * RHOsolid)
}

subcomparts <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment")

filtered_data <- Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, ]

# Apply the adjustment function to the filtered data
adjusted_concentrations <- apply(filtered_data, 1, function(row) {
  f_adjust_concentration(
    Concentration = as.numeric(row["Concentration"]),
    FRACw = subset(FRACw, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
    FRACa = subset(FRACa, SubCompart == row["SubCompart"] & Scale == row["Scale"]),
    Rho = subset(Rho, SubCompart == row["SubCompart"]),
    RhoWater_value = 998,  # Replace with your actual RhoWater_value
    SubCompart = row["SubCompart"],
    Scale = row["Scale"]
  )
})
# Update Concentration_eq with adjusted concentrations
Concentration_eq[Concentration_eq$SubCompart %in% subcomparts, "Concentration"] <- adjusted_concentrations

#Define the units
subcompart <- c("agriculturalsoil", "naturalsoil", "othersoil", "freshwatersediment", "marinesediment",  "lakesediment", "air", "deepocean", "lake" , "river", "sea", "cloudwater")
units <- c("g/kg w", "g/kg w", "g/kg w", "g/kg w", "g/kg w", "g/kg w",
           "kg/kg", "kg/L", "kg/L", "kg/L", "kg/L", "kg/L", "kg/L")

# Combine into a named list
subcompart_units <- setNames(units, subcompart)


Concentration_eq <- Concentration_eq |>
  mutate(Units_per_SubCompart = subcompart_units[SubCompart])

convert_units <- function(concentration, unit) {
  if (unit == "g/kg w") {
    return(concentration * 1e9)
  } else if (unit == "kg/kg") {
    return(concentration * 1e12)
  } else if (unit == "kg/L") {
    return(concentration * 1e12)
  } else {
    return(concentration)
  }
}

# Convert concentrations to ng/kg or ng/L and update units
Concentration_eq <- Concentration_eq |>
  mutate(
         Concentration = mapply(convert_units, Concentration, Units_per_SubCompart),
         Units_per_SubCompart = ifelse(Units_per_SubCompart == "g/kg w", "ng/kg w", 
                                       ifelse(Units_per_SubCompart == "kg/kg", "ng/kg", "ng/L")))
Concentration_eq <- Concentration_eq |>
  mutate(across(where(is.numeric), ~ format(., scientific = TRUE)))

#Concentration_eq <- subset(Concentration_eq, select = -old_EqMass)

kable(Concentration_eq)
```

## Dynamic concentrations

The output of the ODE solver currently only returns output per
compartment abbreviation. As such, we need to ensure that we match the
right constants with the right compartments. This requires some
additional steps. The principles of calculation are the same as above.
We create some visual output to look at the first results for the
dynamic concentration.

```{r concentrations dynamic approxfun output}

library(dplyr)
nTIMES <- 25
#seperate signals and matrix
#compartments
compartments <- SolutionEmissionData[, 1:156]
compartments <- compartments |> select(-matches("U$"))
signals <- SolutionEmissionData[, 157:311]
# Fetch necessary data
Volume <- World$fetchData("Volume")
Area <- World$fetchData("Area")
FRACw <- World$fetchData("FRACw")
FRACa <- World$fetchData("FRACa")
Rho <- World$fetchData("rhoMatrix")


# Define acronyms maps
acronym_map <- c(
  "marinesediment" = "sd2", "freshwatersediment" = "sd1", "lakesediment" = "sd0",
  "agriculturalsoil" = "s2", "naturalsoil" = "s1", "othersoil" = "s3",
  "air" = "a", "deepocean" = "w3", "sea" = "w2", "river" = "w1", "lake" = "w0", "cloudwater" = "cw"
)
acronym_map2 <- c("Arctic" = "A", "Moderate" = "M", "Tropic" = "T", "Continental" = "C", "Regional" = "R")
acronym_map3 <- c("Dissolved" = "D", "Gas" = "G", "Large" = "P", "Small" = "A", "Solid" = "S", "Unbound" = "U")

# Add compartment column to data
Volume <- Volume |> mutate(compartment = paste0(acronym_map[SubCompart], acronym_map2[Scale]))
FRACw <- FRACw |> mutate(compartment = paste0(acronym_map[SubCompart], acronym_map2[Scale]))
FRACa <- FRACa |> mutate(compartment = paste0(acronym_map[SubCompart], acronym_map2[Scale]))
Rho <- Rho |> mutate(compartment = paste0(acronym_map[SubCompart]))

CompartmentsConc <- compartments
# List of columns in CompartmentsConc that need transformation
columns_to_transform <- setdiff(names(CompartmentsConc), "time")

# Calculate concentrations by dividing by the corresponding volume
for (col in columns_to_transform) {
  compartment <- substr(col, 1, nchar(col) - 1)
  volume <- Volume$Volume[Volume$compartment == compartment]
  if (length(volume) == 1) {
    CompartmentsConc[[col]] <- CompartmentsConc[[col]] / volume
  } else {
    warning(paste("Volume not found for compartment", compartment))
  }
}

## Extract soil and sediment columns
CompartmentsConc_soil_names <- grep("^s[123]", names(CompartmentsConc), value = TRUE)
CompartmentsConc_sediment_names <- grep("^sd[0123]", names(CompartmentsConc), value = TRUE)
CompartmentsConc_soil <- CompartmentsConc[, CompartmentsConc_soil_names]
CompartmentsConc_sediment <- CompartmentsConc[, CompartmentsConc_sediment_names]


# Combine soil and sediment columns
CompartmentsConc_combined <- cbind(CompartmentsConc_soil, CompartmentsConc_sediment)


# Identify the columns that are not in the excluded list
included_names <- setdiff(names(CompartmentsConc), names(CompartmentsConc_combined))
included_names <- included_names[!grepl("time", included_names, ignore.case = TRUE)]
print(included_names)

# Multiply the included columns by 1000
CompartmentsConc[included_names] <- CompartmentsConc[included_names] * 1000

# Get RhoWater value for river compartment
RhoWater_value <- Rho$rhoMatrix[Rho$SubCompart == "river"]

# Define function to adjust concentrations
f_adjust_concentration <- function(CompConc, Fracw, Fraca, RHOsolid) {
  CompConc * 1000 / (Fracw * RhoWater_value + (1 - Fracw - Fraca) * RHOsolid)
}

# Get compartment prefixes for combined soil and sediment
compartment_prefixes_scale_sd <- substr(names(CompartmentsConc_sediment), 1, 4)
compartment_prefixes_scale_s <- substr(names(CompartmentsConc_soil), 1, 3)
compartment_prefixes_combined_scale <- cbind(compartment_prefixes_scale_s, compartment_prefixes_scale_sd)
compartment_prefixes_sd <- substr(names(CompartmentsConc_sediment), 1, 3)
compartment_prefixes_s <- substr(names(CompartmentsConc_soil),1, 2)
compartment_prefixes_combined <- cbind(compartment_prefixes_s, compartment_prefixes_sd)

# Get corresponding FRACw, FRACa, and Rho values for combined soil and sediment
FRACw_values_combined <- FRACw$FRACw[match(compartment_prefixes_combined_scale, FRACw$compartment)]
FRACa_values_combined <- FRACa$FRACa[match(compartment_prefixes_combined_scale, FRACa$compartment)]
Rho_values_combined <- Rho$rhoMatrix[match(compartment_prefixes_combined, Rho$compartment)]

# Apply the function to adjust combined soil and sediment concentrations
CompartmentsConc_combined_adjusted <- mapply(
  f_adjust_concentration,
  CompartmentsConc_combined,
  Fracw = FRACw_values_combined,
  Fraca = FRACa_values_combined,
  RHOsolid = Rho_values_combined,
  SIMPLIFY = FALSE
) |> as.data.frame()

# Assign column names to adjusted data
colnames(CompartmentsConc_combined_adjusted) <- colnames(CompartmentsConc_combined)


CompartmentsConc_corrected <- CompartmentsConc

# Overwrite the original soil and sediment columns with the adjusted values
CompartmentsConc_corrected[, names(CompartmentsConc_combined_adjusted)] <- CompartmentsConc_combined_adjusted


plot(CompartmentsConc_corrected$time, CompartmentsConc_corrected$w1CA, type = "l",
     xlab = "Time", ylab = "Concentration [g/L]",
     main = "Freshwater concentrations over time",
     col = "blue") # Set color for w1CA

# Add the line for w1CP to the same plot
lines(CompartmentsConc_corrected$time, CompartmentsConc_corrected$w1CP, col = "red") 
lines(CompartmentsConc_corrected$time, CompartmentsConc_corrected$w1CS, col = "green")

# Add a legend to distinguish the lines
legend("topright", legend = c("W1CA", "W1CP", "W1CS"), col = c("blue", "red", "green"), lty = 1)


plot(CompartmentsConc_corrected$time, CompartmentsConc_corrected$sd1CA, type = "l",
     xlab = "Time", ylab = "Concentration [g/kg w.w.]",
     main = "Soil concentrations over time",
     col = "blue") # Set color for w1CA

# Add the line for w1CP to the same plot
lines(CompartmentsConc_corrected$time, CompartmentsConc_corrected$s2CP, col = "red") 
lines(CompartmentsConc_corrected$time, CompartmentsConc_corrected$s2CS, col = "green")

# Add a legend to distinguish the lines
legend("topright", legend = c("s2CA", "s2CP", "s2CS"), col = c("blue", "red", "green"), lty = 1)

CompartmentsConc_corrected[is.na(CompartmentsConc_corrected)] <- 0
CompartmentsConc_corrected[, -which(names(CompartmentsConc_corrected) == "time")] <- 
  CompartmentsConc_corrected[, -which(names(CompartmentsConc_corrected) == "time")] * 1e9

# CompartmentsConc_corrected <- CompartmentsConc_corrected |>
#   mutate(across(where(is.numeric), ~ format(., scientific = TRUE)))
CompartmentsConc_corrected  <- CompartmentsConc_corrected |>
  mutate(row_number = row_number())
FinalConcentration <- CompartmentsConc_corrected |>
  filter(row_number == (nTIMES - 1)) |>
  select(-row_number)
FinalConcentration <- FinalConcentration %>%
  pivot_longer(
    cols = -time,
    names_to = "Compartment",
    values_to = "Concentration"
  )

FinalConcentration <- FinalConcentration |> 
  mutate(Concentration = replace_na(Concentration, 0))
kable(FinalConcentration)

```

## Comparison with Steady State


Here, we look at the differences in concentrations by creating both
plots and tables.

```{r Concentration Plots}
# Convert all columns in CompartmentsConc to numeric
CompartmentsConc[] <- lapply(CompartmentsConc, as.numeric)

# Convert the Concentration column to numeric
Concentration_eq$Concentration <- as.numeric(Concentration_eq$Concentration)
w1CS_conc <- Concentration_eq$Concentration[Concentration_eq$Abbr == "w1CS"]
ggplot(CompartmentsConc_corrected, aes(x = time, y = w1CS)) +
  geom_line(aes(color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = w1CS_conc, color = "steady"), linetype = "dashed") +
  labs(title = "Plot of w1CS against Time",
       x = "Time",
       y = "w1CS [ng/L]") +
  theme_minimal() +
  scale_color_manual(name = "Legend", 
                     values = c("Approxfun Emission Data" = "pink", "steady" = "red"))

# Extract w1CA concentration
w1CA_conc <- Concentration_eq$Concentration[Concentration_eq$Abbr == "w1CA"]


# Plot w1CA against time
ggplot(CompartmentsConc_corrected, aes(x = time, y = w1CA)) +
  geom_line(aes(color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = w1CA_conc, color = "steady"), linetype = "dashed") +
  labs(title = "Plot of w1CA against Time",
       x = "Time",
       y = "w1CA [ng/L]") +
  theme_minimal() +
  scale_color_manual(name = "Legend", values = c("Approxfun Emission Data" = "pink", "steady" = "red"))

w2CA_conc <- Concentration_eq$Concentration[Concentration_eq$Abbr == "w2CA"]
# Plot w2CA against time
ggplot(CompartmentsConc_corrected, aes(x = time, y = w2CA)) +
  geom_line(aes(color ="Approxfun Emission Data")) +
  geom_hline(aes(yintercept = w2CA_conc, color = "steady"), linetype = "dashed") +
  labs(title = "Plot of w2CA against Time",
       x = "Time [s]",
       y = "w2CA [ng/L]") +
  theme_minimal() +
  scale_color_manual(name = "Legend", values = c("Approxfun Emission Data" = "pink", "steady" = "red"))

s1CA_conc <- Concentration_eq$Concentration[Concentration_eq$Abbr == "s1CA"]
# Plot s1CA against time
ggplot(CompartmentsConc_corrected, aes(x = time, y = s1CA)) +
  geom_line(aes(color = "Approxfun Emission Data")) +
  geom_hline(aes(yintercept = s1CA_conc, color = "steady"), linetype = "dashed")  +
  labs(title = "Plot of s1CA against Time",
       x = "Time [s]",
       y = "s1CA [ng/L]") +
  theme_minimal() +
  scale_color_manual(name = "Legend", values = c("Approxfun Emission Data" = "pink", "steady" = "red"))

```

The concentration in certain compartments is less for dynamic calculations due to the fact that a steady state is not yet reached in some compartments, as is examplified in the table below. 
```{r Comparison Steady Concentration}

Concentration_eq <- Concentration_eq |>
  mutate(Concentration = as.numeric(format(Concentration, scientific = FALSE)))
FinalConcentration <- FinalConcentration |>
  mutate(Concentration = as.numeric(format(Concentration, scientific = FALSE)))


DiffSteadyDynamic <- FinalConcentration |>
  left_join(Concentration_eq, by = c("Compartment" = "Abbr")) |>
  # Rename concentration.x and concentration.y
  mutate(ConcDynamic = Concentration.x,
         ConcSteady = Concentration.y) |>
  # Calculate concentration difference
  mutate(concentration_difference = ConcDynamic - ConcSteady) |>
  # Select relevant columns
  select(Compartment, ConcDynamic, ConcSteady, concentration_difference)
kable(DiffSteadyDynamic)


```
