---
title: "Comparison shape update"
author: "Nadim Saadi, Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}



```

# Explanation of update

In this update, we introduce the PlasticFADE (Plastic Fragmentation And
Degradation in the Environment) degradation model, based on the work of
Tang & Boulay (under review). PlasticFADE was developed as an empirical
model to predict polymer-specific rate constants in various
environmental compartments.

For degradation, the model considers UV intensity and microbial
concentration in the various environmental compartments, as well as the
polymer specific empirical constants, to compute k_deg. The empirical
constants for each polymer can be found in the original publication.

with plasticFADE, the equation is:

kdeg = deg_x \* SAV\^deg_tau \* (deg_y \* UVintensity\^deg_theta + deg_z
\* MICROBconc\^deg_eta) / (24*60*60)

to derive kdeg in s-1. UVintensity and MICROBconc are defined in
ScaleSubCompartData.csv

-   v_KdegDorC is changed to include the new plasticFADE input
    parameters.

There is now a choice of three types of degradation input to be used:
Kssdr, plasticFADE and the Default by setting kdeg directly for all the
relevant compartments based on seperately estimated half lifes.

This is done using `World$SetConst(DegApproach = "Default")`.

-   Degrading_enzyme: Further, the update returns a 0 kdeg in
    compartments without UV intensity, only for polymers that do not
    have naturally occurring degrading enzymes. For example, the
    degradation of polyolefins (PE, HDPE, LDPE, EPS, ...) cannot be
    initiated in the absence of UV, but it can for polymers, such as
    PET, PA, PU, etc. (Chow et al., 2023,
    doi.org/10.1111/1751-7915.14135). This represented with the new
    boolean called "Degrading_ezyme" (TRUE or FALSE). TRUE means that
    degrading enzymes exist in the environment, and therefore SBoo will
    compute kdeg (based on default, Kssdr, or plasticFADE). If FALSE,
    the degradation function will test for the presence of UV before
    returning 0 or computing kdeg.

-   DegApproach = "Default", "Kssdr", or "PlasticFADE"

The following code test the swithing between the different DegApproach
variables for PET_sphere_example (defined in the Substances.csv file).
We use a PET spherical particles of 100um diameter as an example. For
the PET example, the plasticFADE empirical constants are defined:
deg_x=5.50E-03 deg_tau=1.37E-02 deg_y=1.72E-02 deg_theta=7.05E-01
deg_z=1.62E-05 deg_eta=4.42E-01

The SSDR is also defined.

```{r}

substance = "PET_sphere_example"
source("baseScripts/initWorld.R")

kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(DegApproach = World$fetchData("DegApproach"))

 World$SetConst(DegApproach = "Kssdr")
    # PET_ssdr_df <- data.frame(varName = c("Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr"), 
    #                       SubCompart = c("deepocean","river","sea","lake","marinesediment","freshwatersediment"
    #                                      ,"lakesediment","naturalsoil","othersoil","agriculturalsoil"),
    #                       Waarde = c(9.51E-15, 9.51E-15,9.51E-15,9.51E-15,9.51E-15,9.51E-15,9.51E-15,  1.88E-14 , 1.88E-14,1.88E-14)) 
    # World$mutateVars(PET_ssdr_df)
 World$UpdateDirty(unique(PET_ssdr_df$varName))
    
kaas <-
  kaas |> 
  rbind (World$kaas |> mutate(DegApproach = World$fetchData("DegApproach")) )

World$SetConst(DegApproach = "PlasticFADE")
World$UpdateDirty("DegApproach")

kaas <-
  kaas |> 
  rbind (World$kaas |> mutate(DegApproach = World$fetchData("DegApproach")) )

kaas |> filter(process == "k_Degradation")

kaas |> distinct(process)

kaasw <- kaas |> 
  pivot_wider(names_from = DegApproach, values_from = k) |> 
  filter(process == "k_Degradation")


```

```{R Specify directories, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# InstallSBoo (Release_SBoo = NA, # tag for release use NA for branch
#              Release_SBooScripts = NA, # tag for release use NA for branch
#              devBranch_SBoo = "Runoff_update_clean",
#              devBranch_SBooScripts = "Runoff_MPinterception",
#              Temp_Folder = "C:/Temp" # an existing folder where SimpleBox is to be installed
# )

InstallSBoo(Release_SBoo = NA, # tag for release use NA for branch
            Release_SBooScripts = NA, # tag for release use NA for branch
            devBranch_SBoo = "development",
            devBranch_SBooScripts = "development",
            Temp_Folder = NULL # an existing folder where SimpleBox is to be installed
)

```

As this update was only implemented for microplastics and tyre road wear
particles, we will test for these substances. To be sure nothing changed
for the other substances, we will also test one other substance.

```{R Specify substance, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "PET_sphere_example", "1-aminoanthraquinone","nAg_10nm") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```

```{R, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas <- data.frame()
# this_variables <- data.frame()
#substance <- "microplastic"
for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
  print(substance)
  source("baseScripts/initWorld.R")
  
  if (substance == "PET_sphere_example") { #Use SSDR for PET
    World$SetConst(DegApproach = "Kssdr")
    # PET_ssdr_df <- data.frame(varName = c("Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr","Kssdr"), 
    #                       SubCompart = c("deepocean","river","sea","lake","marinesediment","freshwatersediment"
    #                                      ,"lakesediment","naturalsoil","othersoil","agriculturalsoil"),
    #                       Waarde = c(9.51E-15, 9.51E-15,9.51E-15,9.51E-15,9.51E-15,9.51E-15,9.51E-15,  1.88E-14 , 1.88E-14,1.88E-14)) 
    # World$mutateVars(PET_ssdr_df)
    World$UpdateDirty(unique(PET_ssdr_df$varName))
  }

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas <- rbind(this_kaas, kaas)
}

```

Do the same for the other implementation.

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

```{R Run SB and get degradation rates using PlasticFADE, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()
  
for(substanceK in substances$Substance){
  Temp_Folder=NULL
  substance = substanceK
  
  source("baseScripts/initWorld.R")
  if (substance == "PET_sphere_example") { #Use plasticFADE for PET
    World$SetConst(DegApproach = "PlasticFADE")
    #World$UpdateKaas()
    World$UpdateDirty("DegApproach")
  }
  
  
  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

## Verification of previous and new implementation

```{R Calculate the differences between ks, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))
changed_kaas
```

As can be seen in the figures below, the new degradation rate constants
for `r paste(substances$Substance)[2]`, have a very large difference,
max `r max(abs(unique(changed_kaas$rel_diff)))*(-100)` % larger compared
to the previous default kdeg value's. The figures below also show this
and and that the code still works as expected for the other compounds
for which no Kssdr is defined. The larger degradation rates for
microplastics is as expected due to the Kssdr implementation. Further
examples should illustrate the validity of this modification, e.g.
comparing to measuresments etc.

```{R Plot the differences between ks, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs <- changed_kaas |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  absdif_plot
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  reldif_plot
}
```
