---
title: "Comparison shape update"
author: "Nadim Saadi, Valerie de Rijk, Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

source("baseScripts/initWorld.R")


```

# Explanation of update

This update introduces the Surface Specific Degradation Rate (SSDR)
approach of Maga et al. (2022)
(<https://doi.org/10.1007/s11367-022-02040-1>) and adapted by
Corella-Puertas et al. (2023)
(<https://doi.org/10.1016/j.jclepro.2023.138197>) to account for
degradation of microplastics in the environment. The approach considers
multiple parameters to derive a degradation rate (kdeg [yr-1]).

Further, this update introduces MinSettVel, as a place holder settling
velocity for low and medium density microplastics to account for
biofouling. In fact, SBoo used to compute a settling velocity of 0 m/s
for particles with density smaller than that of water. However, this
approach ignores the change in density of particles due to the formation
of a biofilm on the surface of microplastics, which can increase the
density and lead to the (slow) sedimentation of low and medium density
microplastics. MinSettVel is implemented using sedimentation rate values
from Corella-Puertas et al. (2023) to account for this.

-   v_KdegDorC is changed to include the new Kssdr input parameter
    instead of just kdeg

    -   The Kssdr approach uses

        -   ShapeFac: A factor to consider the shape 2 for films/boxes,
            3 for fibers/cylinders, 4 for spheres\
            (Corella-Puertas et al., 2023)

        -   CorFacSSA: A surface area correction factor, which accounts
            for polymer surface roughness and porosity.

        -   Shortest_side: The characteristic dimension through which
            degradation occurs. The degradation is assumed to mainly
            occur perpendicular to the largest surface, along the
            shortest_side axis.

        -   KSSDR: The specific surface degradation rate (SSDR) in
            [m/year]. kSSDR can be found in the literature for different
            polymers (Maga et al., 2022)

        -   kdeg = ShapeFac \* KSSDR \* CorFacSSA / Shortest_side

-   v_SettlingVelocity is changed

    -   MinSettVel added when the density of the particle is smaller
        than that of water, to avoid a null settling velocity.

```{R Specify directories, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# InstallSBoo (Release_SBoo = NA, # tag for release use NA for branch
#              Release_SBooScripts = NA, # tag for release use NA for branch
#              devBranch_SBoo = "Runoff_update_clean",
#              devBranch_SBooScripts = "Runoff_MPinterception",
#              Temp_Folder = "C:/Temp" # an existing folder where SimpleBox is to be installed
# )

InstallSBoo(Release_SBoo = NA, # tag for release use NA for branch
            Release_SBooScripts = NA, # tag for release use NA for branch
            devBranch_SBoo = "development",
            devBranch_SBooScripts = "development",
            Temp_Folder = NULL # an existing folder where SimpleBox is to be installed
)

```

As this update was only implemented for microplastics and tyre road wear
particles, we will test for these substances. To be sure nothing changed
for the other substances, we will also test one other substance.

```{R Specify substance, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "TRWP", "1-aminoanthraquinone","nAg_10nm") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```

```{R, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas <- data.frame()
this_variables <- data.frame()
for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas <- rbind(this_kaas, kaas)
}

```

Do the same for the other implementation.

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

```{R Run SB and get advection ks for FLux_work branch, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()
  
for(substanceK in substances$Substance){
  Temp_Folder=NULL
  substance = substanceK
  
  source("baseScripts/initWorldOther.R")
  
  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

## Verification of previous and new implementation

```{R Calculate the differences between ks, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))
changed_kaas
```

As can be seen in the figures below, the new rate constants for the
substances, `r paste(substances$Substance)`, have a negligible
difference of max `r max(abs(unique(changed_kaas$rel_diff)))*100` %
difference to the previous version. The figures below also show this and
therefor this verification of replicating the previous version with the
new code is complete.

```{R Plot the differences between ks, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs <- kaas_comparison |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}
```

## Quantify and illustrate changes specific to kssdr

[TODO] script showing the effects of different drag methods and shapes
on particle settling velicities and rate constants.

```{r}

rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
source("baseScripts/initWorld.R")

# to be added, script showing for the effects of different drag methods and shapes on particle settling velicities and rate constants.

```
