---
title: "Comparison shape update"
author: "Nadim Saadi, Valerie de Rijk, Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

source("baseScripts/initWorld.R")
```

# Explanation of update

This explains the way we include drag of particles with shapes different
to a sphere.

Several processes which depend on the settling velocity now are
dependant on the Settling Velocity solver based on a specific Drag
method. The supported particle shapes can be found in section [3.2
Particle Properties]{.underline} of the documentation.

The updated functions are:

-   f_SetVelWater and v_SettlingVelocity

    -   For which several functions are introduced or changed to cope
        with shape:

    -   R/f_DragCoefficient.R

    -   R/f_SetVelSolver.R

    -   R/f_PerimeterParticle.R

    -   R/f_SurfaceAreaParticle.R

    -   R/f_Vol.R

    -   R/v_rad_species.R

    -   R/v_rho_species.R

-   f_Grav.R and R/k_HeteroAgglomeration.wsd.R

    -   In future shape could also be included for SPM, now still sphere
        assumed representative

-   k_Resuspension due to the update of f_SetVelWater

-   R/k_Sedimentation.R due to the update of f_SetVelWater

```{R Specify directories, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

InstallSBoo(Release_SBoo = NA, # tag for release use NA for branch
            Release_SBooScripts = NA, # tag for release use NA for branch
            devBranch_SBoo = "development",
            devBranch_SBooScripts = "development",
            Temp_Folder = NULL # an existing folder where SimpleBox is to be installed
)


```

As this update was only implemented for microplastics and tyre road wear
particles, we will test for these substances. To be sure nothing changed
for the other substances, we will also test one other substance.

```{R Specify substance, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "TRWP", "1-aminoanthraquinone","nAg_10nm") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```

```{R Run SB and get ks for this branch (new code), include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas <- data.frame()
this_variables <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas <- rbind(this_kaas, kaas)
}

```

Do the same for the other implementation.

```{R Clean environment, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

```{R Run SB and get ks for other (old code), include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()

for(substanceK in substances$Substance){
Temp_Folder=NULL
substance = substanceK

source("baseScripts/initWorldOther.R")

  kaas <- World$kaas |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

## Verification of previous and new implementation (Original drag method)

```{R Calculate the differences between ks, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))

# changed_kaas

paste(c(changed_kaas |> distinct(process) |> pull()), sep = )

```

As can be seen in the figures below, the new rate constants for the
substances, `r paste(substances$Substance)`, have a negligible
difference of max `r max(abs(unique(changed_kaas$rel_diff)))*100` %
difference to the previous version, related to processes:
`r c(changed_kaas |> distinct(process) |> pull())`. The figures below
also show this and therefore this verification of replicating the
previous version with the new code is complete.

```{R Plot the differences between ks, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs <- kaas_comparison |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}
```

## Quantify changes for new drag method options

The Original and Bagheri drag method will be compared.

```{R include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "TRWP", "1-aminoanthraquinone","nAg_10nm") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```

```{R include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
this_kaas <- data.frame()
this_variables <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")
  DragMethod_df <- data.frame(
  varName = "DragMethod",
  Waarde = "Original"
  )
  World$mutateVars(DragMethod_df)
  World$UpdateDirty("DragMethod")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas <- rbind(this_kaas, kaas)
}

```

Change drag method and recalculate k values

```{R Run SB and get k values for Bagheri drag method, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()

for(substanceK in substances$Substance){
  
  substance = substanceK
  
  source("baseScripts/initWorld.R")
  
  DragMethod_df <- data.frame(
    varName = "DragMethod",
    Waarde = "Bagheri"
  )
  World$mutateVars(DragMethod_df)
  World$UpdateDirty("DragMethod")
  
  kaas <- World$kaas |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

```{R include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))

changed_kaas |> distinct(process) |> pull()

```

As can be seen in the figures below, the new rate constants for the
substances, `r paste(substances$Substance)`, have a difference of max
`r max(abs(unique(changed_kaas$rel_diff)))*100` % between the original
and Bagheri drag methods. This affects the
`r changed_kaas |> distinct(process) |> pull()` processes.

```{R include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs <- kaas_comparison |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between k's for originial and Bagheri's drag method",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative  between k's for originial and Bagheri's drag method",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}
```

## Example for differently shaped particles

Change shapes based on substance data

Run SBoo for a sphere Using Bagheri Dragmethod.
<!--# TODO: Compare the new shapes to spheres with same volume to show the implications of using shapes. And doing this for different sizes. -->

```{R Run SB and get k values for different shapes, echo=TRUE, message=FALSE, warning=FALSE}

substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(str_detect(Substance, "^microplastic"))

this_kaas_shape <- data.frame()

for(substanceK in substances$Substance){
  
  substance <- substanceK
      source("baseScripts/initWorld.R")

  kaas <- World$kaas |>
    mutate(Substance = substance)
  
  this_kaas_shape <- rbind(this_kaas_shape, kaas)
}

```

All are calculated, but let's check if the new microplastics_sphere (set
as Shortest_side) provides the same k's as the default microplastic (set
as RadS).

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

substance <- substances$Substance[1]
      source("baseScripts/initWorld.R")
this_kaas <- World$kaas

# set the substance to the test substance using World$substance
# after initiatlisation using the default microplastic substance
# otherwise alpha, kdis and kdeg are not set properly

World$substance <- substances$Substance[2]

# World$fetchData("Shortest_side")

other_kaas <- World$kaas

common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    filter(process == "k_HeteroAgglomeration.wsd") |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))

if(nrow(changed_kaas) == 0) print("test passed, no changes in k's found") else stop("different k's found, plese check!")

```
