---
title: "Comparison shape update"
author: "Nadim Saadi, Valerie de Rijk, Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

source("baseScripts/initWorld.R")
```

# Explanation of update

This explains the way we include drag of particles with shapes different
to a sphere.

Several processes which depend on the settling velocity now are
dependant on the Setlling Velocity solver based on a specific Drag
method.

The updated functions are:

-   f_SetVelWater and v_SettlingVelocity

    -   For which several functions are introduced or changed to cope
        with shape:

    -   R/f_DragCoefficient.R

    -   R/f_SetVelSolver.R

    -   R/f_PerimeterParticle.R

    -   R/f_SurfaceAreaParticle.R

    -   R/f_Vol.R

    -   R/v_rad_species.R

    -   R/v_rho_species.R

-   f_Grav.R and R/k_HeteroAgglomeration.wsd.R

    -   In future shape could also be included for SPM, now still sphere
        assumed representative

-   k_Resuspension due to update of f_SetVelWater

-   R/k_Sedimentation.R due to update of f_SetVelWater

```{R Specify directories, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# InstallSBoo (Release_SBoo = NA, # tag for release use NA for branch
#              Release_SBooScripts = NA, # tag for release use NA for branch
#              devBranch_SBoo = "Runoff_update_clean",
#              devBranch_SBooScripts = "Runoff_MPinterception",
#              Temp_Folder = "C:/Temp" # an existing folder where SimpleBox is to be installed
# )

InstallSBoo(Release_SBoo = NA, # tag for release use NA for branch
            Release_SBooScripts = NA, # tag for release use NA for branch
            devBranch_SBoo = "Shape_setvel",
            devBranch_SBooScripts = "Shape_setvel",
            Temp_Folder = NULL # an existing folder where SimpleBox is to be installed
)

# Define branches to test
Comparison <- c("Previous", "ThisOne")

# Specify file paths for saving results
result_files <- paste0("results_", Comparison, ".rds")

# Directory of your Git repository (modify this path to the correct repository location)
script_dir <-getwd()
repo_dir <- file.path(getwd(), "..", "SBoo")

```

As this update was only implemented for microplastics and tyre road wear
particles, we will test for these substances. To be sure nothing changed
for the other substances, we will also test one other substance.

```{R Specify substance, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "TRWP", "1-aminoanthraquinone","nAg_10nm") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```


```{R Run SB and get advection ks for original dragmethod, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas <- data.frame()
this_variables <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")
  DragMethod_df <- data.frame(
  varName = "DragMethod",
  Waarde = "Original"
  )
  World$mutateVars(DragMethod_df)
  World$UpdateDirty("DragMethod")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas <- rbind(this_kaas, kaas)
}

```

Do the same for the other implementation.

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

## Quantify changes based drag method chosen

Change drag method and recalculate k values

```{R Run SB and get k values for Bagheri drag method, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()

for(substanceK in substances$Substance){
Temp_Folder=NULL
substance = substanceK

source("baseScripts/initWorldOther.R")
DragMethod_df <- data.frame(
  varName = "DragMethod",
  Waarde = "Bagheri"
  )
World$mutateVars(DragMethod_df)
World$UpdateDirty("DragMethod")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

## Verification of previous and new implementation

```{R Calculate the differences between ks, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas <- kaas_comparison |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))
changed_kaas
```

As can be seen in the figures below, the new rate constants for the
substances, `r paste(substances$Substance)`, have a negligible
difference of max `r max(abs(unique(changed_kaas$rel_diff)))*100` %
difference to the previous version. The figures below also show this and
therefor this verification of replicating the previous version with the
new code is complete.

```{R Plot the differences between ks, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs <- kaas_comparison |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}
```

## Quantify changes based on shape 

Change shape and observe the difference (using Bagheri)

```{r}

rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```


##Run SBoo for a sphere
Using Bagheri Dragmethod

```{R Run SB and get k values for Bagheri drag method for a sphere, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas_shape <- data.frame()
#this_variables <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")
  DragMethod_df <- data.frame(
  varName = "DragMethod",
  Waarde = "Original"
  )
  World$mutateVars(DragMethod_df)
  World$UpdateDirty("DragMethod")
  
  shape_df <- data.frame(
  varName = "Shape",
  Waarde = "Sphere"
  )
  World$mutateVars(shape_df)
  World$UpdateDirty("Shape")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  # SetVel <- World$fetchData("SettlingVelocity") |> 
    # mutate(Substance = substance)
  # this_variables <- rbind(this_variables,SetVel)
  # this_variables <- cross_join(SetVel)
  this_kaas_shape <- rbind(this_kaas_shape, kaas)
}

```

##Run SBoo for a fiber
Using Bagheri drag method


```{R Run SB and get k values for Bagheri drag method for a fiber, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas_shape <- data.frame()

for(substanceK in substances$Substance){
Temp_Folder=NULL
substance = substanceK

source("baseScripts/initWorldOther.R")
DragMethod_df <- data.frame(
  varName = "DragMethod",
  Waarde = "Bagheri"
  )
World$mutateVars(DragMethod_df)
World$UpdateDirty("DragMethod")

shape_df <- data.frame(
  varName = "Shape",
  Waarde = "Fiber"
  )
World$mutateVars(shape_df)
World$UpdateDirty("Shape")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas_shape <- rbind(other_kaas_shape, kaas)
}

```


## Change between shape (sphere and fiber)

```{R Calculate the differences between ks (shape), include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols_shape <- setdiff(intersect(colnames(this_kaas_shape), colnames(other_kaas_shape)), "k")

kaas_comparison_shape <- merge(this_kaas_shape, other_kaas_shape, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison_shape <- kaas_comparison_shape |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas_shape <- kaas_comparison_shape |>
 filter(diff != 0) |>
    # filter(abs(rel_diff) > 0.001) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))
changed_kaas_shape
```

As can be seen in the figures below, the new rate constants for the
substances, `r paste(substances$Substance)`, have a negligible
difference of max `r max(abs(unique(changed_kaas_shape$rel_diff)))*100` %
difference to the previous version. The figures below also show this and
therefor this verification of replicating the previous version with the
new code is complete.

```{R Plot the differences between ks (shape), include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
all_diffs_shape <- kaas_comparison_shape |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs_shape$Substance)){
  diffs_substance_shape <- all_diffs_shape |>
    filter(Substance == i)
  
  absdif_plot_shape <- ggplot(diffs_substance_shape, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between sphere and fiber k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(linewidth = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot_shape)
  
  reldif_plot_shape <- ggplot(diffs_substance_shape, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between sphere and fiber k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot_shape)
}
```
