---
title: "Comparison runoff update"
author: "Anne Hids, Nadim Saadi, Joris Quik"
date: "2025-05-27"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

substance <- "microplastic"

source("baseScripts/initWorld.R")
```

# Explanation of update


=======
In the previous itteration Runoff was only reduced for the largest
species of heteroagglomerates (P/attached species). Runoff for the solid
and aggregate species would be calculated using the runoff rate for
microplastics as used in the orignial SimpleBox4Plastics model (Quik
etal. 2024). As microplastics themselves can already be relatively
large, runoff is overestimated for those larger microplastics because
larger particles are likely to be trapped or slowed down by vegetation
present on the soil.

To avoid overestimation of the runoff rate, an interception fractiobn
was added. This fraction represents the fraction of microplastics
intercepted by the vegetation on soil. The default value of interception
fraction for microplastics larger than about 100 micrometers is 0.9715,
which is the average between the interception at high, medium and low
density vegetation as found by Han et al. (2022). The option to use this
interception fraction is now added in SimpleBox based on the previous
implementation by Louvet et al. (in prep).


Old function:

$(Runoff * f_CORRsoil(VertDistance, relevant_depth_s, penetration_depth_s)) / Volume * to.FracROWatComp$

New function:

$(Runoff * f_CORRsoil(VertDistance, relevant_depth_s, penetration_depth_s)) / Volume * to.FracROWatComp * (1-IntrcptFrac)$


Where IntrcptFrac is calculated based on rad_species being larger than
SizeRunoff and being set to VegInterceptFrac. Default VegInterceptFrac =
1, which means no Runoff for those larger particles, but can be adjusted
to 0.9715 using MutateVars, see also general documentation in [7.3
OtherIntermedia.md](/vignettes/7.3-OtherIntermedia.md).


# Change in k values for runoff flows

<!-- TO DO: fetch development at a certain date instead of most recent version. This would ensure consequent outcomes of this comparison, no matter when the script is run. -->

```{R Specify directories, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)

# InstallSBoo (Release_SBoo = NA, # tag for release use NA for branch
#              Release_SBooScripts = NA, # tag for release use NA for branch
#              devBranch_SBoo = "Runoff_update_clean",
#              devBranch_SBooScripts = "Runoff_MPinterception",
#              Temp_Folder = "C:/Temp" # an existing folder where SimpleBox is to be installed
# )

InstallSBoo(Release_SBoo = NA, # tag for release use NA for branch
            Release_SBooScripts = NA, # tag for release use NA for branch
            devBranch_SBoo = "development",
            devBranch_SBooScripts = "development",
            Temp_Folder = NULL # an existing folder where SimpleBox is to be installed
)

# Define branches to test
Comparison <- c("development", "ThisOne")

# Specify file paths for saving results
result_files <- paste0("results_", Comparison, ".rds")

# Directory of your Git repository (modify this path to the correct repository location)
#script_dir <-getwd()
#repo_dir <- file.path(getwd(), "..", "SBoo")

```

As this update was only implemented for microplastics and tyre road wear
particles, we will test for these substances. To be sure nothing changed
for the other substances, we will also test one other substance.

```{R Specify substance, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
substance_names <- c("microplastic", "TRWP", "1-aminoanthraquinone") 
substances <- read.csv("data/Substances.csv")
substances <- substances |>
  filter(Substance %in% substance_names)
```

```{R Switch to SBoo Development branch, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# Switch to the correct Git directory
# setwd(repo_dir)

# Checkout the branch
# branch <- branches[1]
# message("Switching to branch: ", branch)
# system(paste("git checkout -f ", branch))
# system("git pull", intern = TRUE)  # Ensure the branch is up-to-date
  
# setwd(script_dir)
```

```{R Run SB and get advection ks for development, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
this_kaas <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  this_kaas <- rbind(this_kaas, kaas)
}

```

Do the same for the other implementation.

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

```{R Switch to SBoo Flux_work branch, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
# Switch to the correct Git directory
# setwd(repo_dir)

# Checkout the branch
# branch <- branches[2]
# message("Switching to branch: ", branch)
# system(paste("git checkout -f ", branch))
# system("git pull", intern = TRUE)  # Ensure the branch is up-to-date
  
# setwd(script_dir)


```

```{R Run SB and get advection ks for FLux_work branch, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}

other_kaas <- data.frame()
  
for(substanceK in substances$Substance){
Temp_Folder=NULL
substance = substanceK

source("baseScripts/initWorldOther.R")

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

```

## Compare the k values for each of the substances

```{R Calculate the differences between ks, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")


kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_New", "_Old"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_Old)


changed_kaas <- kaas_comparison |>
  filter(diff != 0) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))
```

As can be seen in the figures below, the new runoff rates for
microplastics and TRWP are about
`r round(abs(min(unique(changed_kaas$rel_diff))), 0)` times lower than
in the previous version. For other substances k_Runoff remains
unchanged.


```{R Plot the differences between ks, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}

all_diffs <- kaas_comparison |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new runoff k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new runoff k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}
```

Further analysis

Use of Runoff with 200 micrometer sized microplastics

```{r message=FALSE, warning=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])

this_kaas <- data.frame()

for(substanceK in substances$Substance){
  # Get the substance type
  substance <- substanceK
      source("baseScripts/initWorld.R")

  # World$substance
  # World$fetchData("RadS")
  if(World$fetchData("ChemClass") == "particle") {
    attr(World$fetchDataUnits("RadS"),"unit") # unit of input!
    World$SetConst(RadS = 250000)
    World$UpdateDirty("RadS")
    # World$fetchData("SizeRunoff")
    # unique(World$fetchDataUnits("SizeRunoff")$Unit) # unit of input!
    # World$UpdateKaas()
  }
  
  #World$fetchData("IntrcptFrac")
  #World$fetchData("rad_species")
      
  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  this_kaas <- rbind(this_kaas, kaas)
}

rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])

other_kaas <- data.frame()
  
for(substanceK in substances$Substance){
Temp_Folder=NULL
substance = substanceK

source("baseScripts/initWorldOther.R")

  if(World$fetchData("ChemClass") == "particle") {
    attr(World$fetchDataUnits("RadS"),"unit") # unit of input!
    World$SetConst(RadS = 250000)
    # World$fetchData("SizeRunoff")
    # unique(World$fetchDataUnits("SizeRunoff")$Unit) # unit of input!
    World$UpdateDirty("RadS")
  }

  kaas <- World$kaas |>
    # filter(process == "k_Runoff") |>
    mutate(Substance = substance)
  
  other_kaas <- rbind(other_kaas, kaas)
}

common_cols <- setdiff(intersect(colnames(this_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(this_kaas, other_kaas, by=common_cols, suffixes = c("_New", "_Old"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_Old)

changed_kaas <- kaas_comparison |>
  filter(diff != 0) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale))

all_diffs <- changed_kaas |>
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

for(i in unique(all_diffs$Substance)){
  diffs_substance <- all_diffs |>
    filter(Substance == i)
  
  absdif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = diff)) + 
    geom_point() + 
    labs(
      title = "Difference between old and new runoff k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Difference"  
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),  
      panel.background = element_blank() 
    )
  
  print(absdif_plot)
  
  reldif_plot <- ggplot(diffs_substance, mapping = aes(x = toname, y = fromname, color = abs(rel_diff))) + 
    geom_point() + 
    labs(
      title = "Relative difference between old and new runoff k's",
      subtitle = i,
      x = "To",  
      y = "From",  
      color = "Relative difference"  
    ) +
    # scale_color_gradient(low = "black", high = "blue", limits = c(0, 1)) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1), 
      panel.grid.major = element_line(size = 0.2, color = "gray90"),
      panel.background = element_blank()  
    )
  print(reldif_plot)
}

```

In the above plots for large microplastics and TWRP (250 micrometer),
you can see that now Runoff in the new situation is switched off, due to
the interception being set to 1 (100%). The relative differences between
old and new is thus also 1 ((Old-New)/Old)

