---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
source("baseScripts/initWorld_onlyMolec.R")
World$fetchData("TotalArea")
World$fetchData("FRACsea")
```

SHowing the workflow and debugging for adding a local scale

added a scale nest grouping

1.  Added Local scale row in ScaleSheet.csv

    -   `TotalArea` set to 25 by default

    -   Addded ScaleNestGroup <!--#  not clear why yet -->

    -   `FRACsea = 0` by default, but sea can be part of Local Scale

2.  Make `v_Area` more generic to accommodate Local scale area calculation

    -   Basic assumptions on air, cloudwater, sea, marinesediment and deepocean remain.

    -   Other compartments are all calculated based on specification of `landFRAC` (`landFRAC * AreaLand`).

3.  Added Local to FlowIO.csv

    -   Local to Regional

        -   For air, river and sea

    -   Regional to local

        -   For air, river and sea

4.  Add Local data to

    -   ScaleSubCompartData.csv

        -   VertDistance data needed for air, cloudwater, river, marine sediment, natural soil and sea.

        -   landFRAC

    -   

5.  Local air flow

    -   Adjust x_Advection_Air.R

```{r}

CalCSea <-  World$NewCalcVariable("AreaSea")
CalCSea$execute(debugAt = list(ScaleName = "Regional"))

oCalCSea <- CalCSea$execute()

World$fetchData("AreaSea")
World$fetchData("AreaLand")

# fist debug v_Area
Calc_Area <- World$NewCalcVariable("Area")
area_1 <- Calc_Area$execute()
length(area_1$Area)
# seems fine now
# Calc_Area$execute(debugAt = list(ScaleName = "Local"))

World$fetchData("Area")

FRACcldw
VertDistance

World$fetchData("Volume")
World$fetchData("AirFlow")

World$fetchData("WINDspeed")
World$NewCalcVariable("AirFlow")
World$CalcVar("AirFlow")

x_Advection_Air

```
