---
title: "Concentrations and mass flux"
author: "Jaap Slootweg, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

## Concentrations and mass fluxes

This vignette demonstrates how to obtain concentrations and mass fluxes for a steady-state solution. We first initialize the world for molecules and add arbitrary emissions to the regional air and river compartments. 

```{r, warning = FALSE}
#We need to initialize, by default a molecular substance is selected
source("baseScripts/initWorld_onlyMolec.R")

Solvr <- World$NewSolver("SB1Solve")
Solution <- World$Solve(emissions = data.frame(
  Abbr = c( "aRU", "w1RU"),
  Emis = 10e6)
  )


```

## Calculating concentrations.

After computing the steady state solution we end up with equilibrium masses for the compartments. To compute the concentration, we simply divide by the Volume of the respective compartment. 

```{r}
Concentrations <- function(EqMass, Volume) {
  EqMass / Volume
}
World$NewCalcVariable("Concentrations")
ConcPM <- World$CalcVar("Concentrations")
pivot_wider(ConcPM[, c("SubCompart", "Scale", "Concentrations")],
            values_from = "Concentrations",
            values_fill = NULL,
            names_from = "Scale")

```

## Mass fluxes
Now, if we're interested in the mass fluxes between the compartments, we compute the mass fluxes. This is simply the value of the first order rate constant k (or kaas) and the equilibrium mass of the from compartment. As such, we end up with mass fluxes in weight per second.
```{r}
World$fetchData("EqMass")

MsFlux <- left_join(World$kaas, World$fetchData("EqMass"), 
                    join_by(fromScale == Scale, fromSubCompart == SubCompart, fromSpecies == Species))
MsFlux$mFlux <- MsFlux$k * MsFlux$EqMass
print(MsFlux)
```

