---
title: "Dependencies in R"
author: "Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```
## Dependencies of R functions 
Just like in Excel, we can trace the dependencies of processes and variables. For this, we can use the function World$nodelist. This returns all calculated variables in SBoo and they're required parameters. In this vignette, we demonstrate both the dependencies for a process module (k_Adsorption and AreaSea). You can use this technique to generate plots like these for all calculated parameters.

### Generating nodes

```{r Initializing World, warning = FALSE}
source("baseScripts/initWorld_onlyMolec.R")
allnodes <- World$nodelist
head(allnodes)
filtered_nodelist <- World$nodelist[World$nodelist$Calc == "k_Adsorption", ]
filtered_nodelist_sea <- World$nodelist[World$nodelist$Calc == "AreaSea",]
print(filtered_nodelist)
```

### k_Adsorption dependents
k_Adsorption is a great example of a process module that is dependent on a significant amount of parameters. 
```{r World nodes}
NodeAsText <- paste(filtered_nodelist$Params, "->", filtered_nodelist$Calc)
AllNodesAsText <- do.call(paste, c(as.list(NodeAsText), list(sep = ";")))

# Create the DAG string
dag_string <- paste("dag{", AllNodesAsText, "}")

# Convert the string to a DAG object
dag <- dagitty::dagitty(dag_string)
tidy_dag <- tidy_dagitty(dag)



# Plot with customized node colors
ggdag(tidy_dag, text = FALSE) +
  theme_dag_blank() +  # Apply the default DAG theme
  geom_dag_node(aes(color = "black"), fill = "black", size = 25) +  # Customize node colors and size
  geom_dag_text(aes(label = name), size = 2, color = "white") +  # Customize node text size and color
  scale_fill_manual(values = "green") +  # Apply custom fill colors
  theme(
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none"  # Remove the legend
  ) +
  labs(
    title = "Dependents of k_Adsorption"
  )
```

### v_AreaSea dependents 
As we can see below, v_AreaSea is dependent on way less parameters. 
```{r v_AreaSea dependents}
# Assuming filtered_nodelist_sea is a data frame and Params and Calc are columns in it
NodeAsText <- paste(filtered_nodelist_sea$Params, "->", filtered_nodelist_sea$Calc)

# Combine all nodes into a single string, separating them with ";"
AllNodesAsText <- do.call(paste, c(as.list(NodeAsText), list(sep = ";")))

# Create the DAG string
dag_string <- paste("dag{", AllNodesAsText, "}")

# Convert the string to a DAG object
dag <- dagitty::dagitty(dag_string)

# Convert the DAG object to a tidy format for plotting
tidy_dag <- tidy_dagitty(dag)

# Plot the DAG with customized aesthetics
ggdag(tidy_dag, text = FALSE) +
  theme_dag_blank() +  # Apply the default DAG theme
  geom_dag_node(aes(color = "black"), fill = "black", size = 25) +  # Customize node colors and size
  geom_dag_text(aes(label = name), size = 2, color = "white") +  # Customize node text size and color
  scale_fill_manual(values = "green") +  # Apply custom fill colors
  theme(
    plot.title = element_text(hjust = 0.5), 
    legend.position = "none"  # Remove the legend
  ) +
  labs(
    title = "Dependents of v_AreaSea"
  )
```

