---
title: "classic solve"
author: "JS"
date: "12 Apr 2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
## impossible to generalise? projectRoot <- paste(getwd(), "..", sep = "/")
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

## A matrix of first order differential equations
This vignettes demonstrates the mechanism behind SB.
#  We need to initialize, by default a molecular substance is selected.
#  initWorld.R does just that, and calculates SBvars, flows and k's. But we want to start with a single state (box): Unbound (Molecular) Regional air.

```{r}
#We need to initialize, by default a molecular substance is selected
source("baseScripts/initWorld.R")

Solvr <- World$NewSolver("SBsolve")
Solution <- World$Solve(emissions = data.frame(
  Abbr = c("aRP", "aRU", "w1RU"),
  Emis = 10e6),
  needdebug = F,
  nTIMES = 200 #,
  #EmisAsPulse = T
  )

library(htmlTable)
World$kaas <- World$kaas[World$kaas$fromAbbr == "aRU" & World$kaas$toAbbr == "aRU",]
```
Mind the warning... The format of kaas should conform the expected layout.
This particular k has the same from- as to state; It is a sink, like degradation.
We add an emission to the same state, and see what happens in time.

```{r}

Solvr <- World$NewSolver("SBsolve")
#kaasSquare <- Solvr$PrepKaasM()
Solution <- World$Solve(emissions = data.frame(
  Abbr = c("aRP", "aRU", "w1RU"),
  Emis = 10e6),
  needdebug = F,
  nTIMES = 200 #,
  #EmisAsPulse = T
  )

emissions <- ClassicExcel$ExcelEmissions("current.settings")[1,]
World$NewSolver("SBsolve") #
steadSolve <- World$Solve(emissions = emissions)
#World$SolutionAsRelational()
```

## The constants package
Some global constants are imported from this package. For convenience two functions are availeable, demonstrated below.
```{r}
Concentrations <- function(EqMass, Volume) {
  EqMass / Volume
}
World$NewCalcVariable("Concentrations")
ConcPM <- World$CalcVar("Concentrations")
pivot_wider(ConcPM[, c("SubCompart", "Scale", "Concentrations")],
            values_from = "Concentrations",
            values_fill = NULL,
            names_from = "Scale")
# tidyHtmlTable(
#   x = ConcPM[, c("SubCompart", "Scale", "Concentrations")],
#   value = "Concentrations",
#   header = "Scale",
#   rnames = "SubCompart",
#   #col.header = "#FEFEFE",
#   col.rgroup = c("#FEFEFE", "#F7F7F7")
# )
```
## Mass fluxes
```{r}
MsFlux <- left_join(World$kaas, World$fetchData("EqMass"), 
                    join_by(fromScale == Scale, fromSubCompart == SubCompart, fromSpecies == Species))
MsFlux$mFlux <- MsFlux$k * MsFlux$EqMass
aggregate(mFlux~toAbbr, data = MsFlux, FUN = sum)
``` 
accuracy? Roundings?? tol doesn't make a difference?!

Happy calculations!
