---
title: "Introduction to SB-variables"
author: "Jaap Slootweg, Valerie de Rijk"
date: "11/23/2022"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
## impossible to generalise? projectRoot <- paste(getwd(), "..", sep = "/")
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

# Variables in SimpleBox

are read initially, but you can add variables just by defining the
function how to calculate the variable. This vignette will demonstrate
how the variable "volume" is defined, how it can be retrieved, and how
to verify it against the values in the excel-version. First step is to
initiate testing objects by running a standard script
baseScripts/initTestWorld.R. We will use two objects that are created:
"World" and "ClassicExcel". Note the class properties and the
inheritance of ClassicExcel.

```{r initTestWorld, warning = FALSE}
source("baseScripts/initWorld_onlyMolec.R")
class(World)
```

SimpleBox Variables (SBVs) are defined by a function, and many of the
defining functions are present in the package. One of the variable
defining functions is "Volume". The method NewCalcVariable adds the
SimpleBox variable; it needs the name (as character) of the function.

```{r volfunction}
Volume
World$NewCalcVariable("Volume")
```

R "knows" the parameters of a function, see the function formals(). The
names of these parameters are also SBVs, partly already read from data
in the initialisation. You can lookup which variables are known, and
their values by method fetchData(varname) of the class SBcore. World, as
initiated in the baseScript initWorld.R is of this class. An empty
varname (the parameter for this method) returns all known parameters.
Note that variable are usually data.frames (tables), with keyfields
indicating the domain of the variable.

## Atomic functions and tables

Note that variables are usually tables, including the key fields
defining their domain (except constants) but the parameters in the
functions are atomic. This means you can treat it as a simple number or
character; its not the table, not even a vector! This makes writing a
SBV defining function a lot easier, think of the "normal" if()
situations that need the ifelse() functions. The variable defining
functions are called by SBOO for each relevant combination of the
variables. Results, when not NA, are stored in the in-memory database of
SBOO.

```{r volparameters}
formals(Volume)
World$fetchData()
World$fetchData("VertDistance")
World$fetchData("FRACcldw")
```

The variable Area is missing. Missing variables can be found by the
method whichUnresolved. Area is also a function, which needs other
variables. To complete the calculation of Volume we have four steps,
three in preparation. Area can (and will) be used by other functions. No
need to recalculate it.

```{r neededparameters}
#World$whichUnresolved()

lapply(c("AreaSea", "AreaLand", "Area"), function(FuName){
  World$NewCalcVariable(FuName)
  World$CalcVar(FuName)
})

World$CalcVar("Volume")
```

The calculation differs slightly from the excel version. This is
because:
+ the automation of the calculation demands a much more formal
approach 
+ a stricter differentiation between data and calculations 
+ simplifications are applied like renaming both depth and height into VertDistance

The variables as functions/tables and the automated dependencies can be
confusing initially, but the concept is powerful and helps documenting
the model as a whole. We can even create a graph showcasing all relationships between the variables.

```{r theDAG}
library(ggdag)
NodeAsText <- paste(World$nodelist$Params, "->" ,World$nodelist$Calc)
AllNodesAsText <- do.call(paste, c(as.list(NodeAsText), list(sep = ";")))
dag <- dagitty::dagitty(paste("dag{", AllNodesAsText, "}"))
plot(dagitty::graphLayout(dag))
```
