---
title: "SBoo Solver output"
author: "Anne Hids, Jaap Slootweg, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

The SteadyState solver and DynamicSolver have initial output relative to
the emission input which is by default in mass (kilograms) in each
compartment. Several transformations of this output are possible some of
which are already implemented for ease of use.

These are calculation of:

1.  Concentrations

OVERALL WORK IN PROGRESS.

```{r Initialize World, message=FALSE, warning=FALSE, include=FALSE}
source("baseScripts/initWorld_onlyMolec.R")

# Create the steady state emission dataframe
emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), Emis = c(10000, 10000, 10000)) 

# convert 1 t/y to si units: kg/s
emissions <- emissions |>
  mutate(Emis = Emis*1000/(365*24*60*60)) 
# Define the solver function to use. For steady state calculations, this is always "SteadyODE"
World$NewSolver("SteadyStateSolver")

# Solve with the emissions we defined in the previous chunk
World$Solve(emissions = emissions)
```

After initializing and solving the processing functions are available.

## World\$Cocentration()

This function is based on the `v_Mass2Conc` variable and the hard coded
default conversions in the `R6SolverModule` in the function
`ConcentrationToGrams`. No other units are supported at the moment, this
is up to the user for further analysis and conversion.

```{r}

World$Concentration()
```
