---
author: "Joris Quik, Valerie de Rijk, Nadim Saadi"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

### 3.2 Particle Properties

```{r initiate, echo=TRUE, message=FALSE, warning=FALSE}
#we initialize the test environment with a nano substance
substance <- "nAg_10nm"
substance <- "microplastic"
substance <- "microplastic_sphere"
substance <- "microplastic_cube"
#script to initialize test environment, including faking a future 'library(sboo)'
source("baseScripts/initWorld.R")
# debugonce("k_Sedimentation")



```

#### Size and Shape

The particle properties of the particle to be assessed can be defined by either:

-   `RadS` (radius of solid sphere) or

-   `Shortest_side`, `Intermediate_side` or `Longest_side`.

These can be accompanied by the `Shape` for any particle that is not a sphere and needs to be described by additional size characteristics: the Intermediate and/or longest side.

Shapes implemented are:

-   `Sphere` or `Default`

    -   Defined by `RadS` (radius) or `Shortest_side` (diameter)

-   `Ellipsoid`

    -   Defined by `Shortest_side`, `Intermediate_side` and `Longest_side`

-   `Cube` or `Box` or `Film`

    -   Defined by `Shortest_side`, `Intermediate_side` and `Longest_side`

-   `Cylindric - circular` or `Fiber`

    -   Defined by `Shortest_side (diameter)` and `Longest_side`

-   `Cylindric - elliptic`

    -   Defined by `Shortest_side` (minor diameter), `Intermediate_side` (major diameter) and `Longest_side` (height).

These shapes and sizes are used for calculating:

-   Volume (`fVol`)

-   Surface Area (`f_SurfaceAreaParticle`)

-   Perimeter (`f_PerimeterParticle`)

#### Species size and density

Three different particle species exist: The nanoparticle or microplastic particle (Soild/S), Aggregated to natural colloids (Small/A) and Attached to coarse natural particulates (Large/P).
The size and density of these heteroagglomerates (Small/A or Large/P) is dependant on both the size and density of the Solid/S species as well as those of the other particle.

The density and radius of the nanoparticle/microplastic itself is given by the substance properties RhoS and RadS/sides (see above).
For the aggregated and the properties of the "natural part" of the species; *CoarseParticulate* for the Attached species and *NaturalColloid* for the Aggregated species.
Those properties are RhoCP and RhoCOL for density and RadCP and RadCOL for their radius.
An overview of excel origins:

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
read_csv("data/Substances.csv") |> filter(ChemClass == "particle") # data cleanup needed - not all correct!
World$fetchDataUnits("RhoCP")
World$fetchDataUnits("RhoCOL")
World$fetchDataUnits("RadCP") |> mutate(RadCP = RadCP*1e6) # input unit (see units.csv)
World$fetchDataUnits("RadCOL") |> mutate(RadCOL = RadCOL*1e9) # input unit (see units.csv)
```

From these, we can derive the values for rho_species:

```{r echo=FALSE, paged.print=TRUE}
World$fetchData("rho_species")
# rp <- World$NewCalcVariable("rho_species")
# rho_species
# World$CalcVar("rho_species")
```

And the radii:

```{r echo=FALSE, paged.print=TRUE}
World$fetchData("rad_species")
# rs <- World$NewCalcVariable("rad_species")
# World$CalcVar("rad_species")

```

#### Settling velocity

An example of a process dependent on these characteristics is the Settling Velocity.
The Settling Velocity is calculated through the choice of different drag methods of which the original does not include drag.

[PM: more details on the drag methods and approach to calculate the Settling Velocity.
Functions to cover:

-   f_DragCoefficient

-   f_SetVelSolver

-   v_SettlingVelocity

```{r}
World$fetchDataUnits("SettlingVelocity")
# World$NewCalcVariable("SettlingVelocity")
# World$CalcVar("SettlingVelocity")
```
