---
author: "Joris Quik, Valerie de Rijk, Nadim Saadi"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

### 3.2 Particle Properties

```{r initiate, echo=TRUE, message=FALSE, warning=FALSE}
#we initialize the test environment with a nano substance
substance <- "nAg_10nm"
substance <- "microplastic"
substance <- "microplastic_sphere"
substance <- "microplastic_cube"
#script to initialize test environment, including faking a future 'library(sboo)'
source("baseScripts/initWorld.R")
# debugonce("k_Sedimentation")



```

#### Size and Shape

The particle properties of the particle to be assessed can be defined by either:

-   `RadS` (radius of solid sphere) or

-   `Shortest_side`, `Intermediate_side` or `Longest_side`.

These can be accompanied by the `Shape` for any particle that is not a sphere and needs to be described by additional size characteristics: the Intermediate and/or longest side.

Shapes implemented are:

-   `Sphere` or `Default`

    -   Defined by `RadS` (radius) or `Shortest_side` (diameter)

-   `Ellipsoid`

    -   Defined by `Shortest_side`, `Intermediate_side` and `Longest_side`

-   `Cube` or `Box` or `Film`

    -   Defined by `Shortest_side`, `Intermediate_side` and `Longest_side`

-   `Cylindric - circular` or `Fiber`

    -   Defined by `Shortest_side (diameter)` and `Longest_side`

-   `Cylindric - elliptic`

    -   Defined by `Shortest_side` (minor diameter), `Intermediate_side` (major diameter) and `Longest_side` (height).

If short, medium or longest side is not defined, it is defined as RadS\*2 by default (as for a sphere) in 'v_SettlingVelocity.R'.

These shapes and sizes are used for calculating:

-   Volume (`fVol`)

-   Surface Area (`f_SurfaceAreaParticle`)

-   Perimeter (`f_PerimeterParticle`)

-   Various shape factors for different drag methods (circularity, sphericity, elongation, etc.) in v_SettlingVelocity.R

#### Species size and density

Three different particle species exist: The nanoparticle or microplastic particle (Soild/S), Aggregated to natural colloids (Small/A) and Attached to coarse natural particulates (Large/P).
The size and density of these heteroagglomerates (Small/A or Large/P) is dependant on both the size and density of the Solid/S species as well as those of the other particle.

The density and radius of the nanoparticle/microplastic itself is given by the substance properties RhoS and RadS/sides (see above).
For the aggregated and the properties of the "natural part" of the species; *CoarseParticulate* for the Attached species and *NaturalColloid* for the Aggregated species.
Those properties are RhoCP and RhoCOL for density and RadCP and RadCOL for their radius.
An overview of excel origins:

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
read_csv("data/Substances.csv") |> filter(ChemClass == "particle") # data cleanup needed - not all correct!
World$fetchDataUnits("RhoCP")
World$fetchDataUnits("RhoCOL")
World$fetchDataUnits("RadCP") |> mutate(RadCP = RadCP*1e6) # input unit (see units.csv)
World$fetchDataUnits("RadCOL") |> mutate(RadCOL = RadCOL*1e9) # input unit (see units.csv)
```

From these, we can derive the values for rho_species:

```{r echo=FALSE, paged.print=TRUE}
World$fetchData("rho_species")
# rp <- World$NewCalcVariable("rho_species")
# rho_species
# World$CalcVar("rho_species")
```

And the radii:

```{r echo=FALSE, paged.print=TRUE}
World$fetchData("rad_species")
# rs <- World$NewCalcVariable("rad_species")
# World$CalcVar("rad_species")

```

#### Settling velocity

An example of a process dependent on these characteristics is the Settling Velocity.
The Settling Velocity is calculated through the choice of different drag methods (Dioguardi, Swamee, Stokes, Bagheri) of which the original does not include drag.

Dittmar et al. (2024) (<https://doi.org/10.1021/acs.est.3c09602>) reviewed multiple drag models (Waldschläger, Khatmullina, Dioguardi, Bagheri, etc.).
Bagheri & Bonadonna (2016) (<https://www.sciencedirect.com/science/article/pii/S0032591016303539>) is generally recommended, as Coyle et al. (2023) (<https://doi.org/10.1021/acsestwater.2c00466>) have shown that the model was applicable to large microplastics, while Dittmar et al. (2024) have shown its applicability to smaller MPs.
Both articles show that the model performs better than other available models for MPs, although the model was developed for small particles in general, not microplastics.

The settling velocity is computed though the following functions:

-   f_DragCoefficient: This function computes the drag coefficient (CD) according to the selected drag method (e.g.: DragMethod == "Dioguardi").
    The equations and inputs for CD depends on the drag method chosen.
    The inputs (Sphericity, etc.) are calculated in v_SettlingVelocity.R

-   f_SetVelSolver: The settling velocity (vs) depends on the reynolds number (Re) and the drag coefficient.
    However, Re itself depends on the settling veloicity, therefore this solver function is needed to solve for the settling velocity, which solves for vs until it converges.

-   v_SettlingVelocity: Computes the inputs needed (described above) to solve for the settling velocity in air and water.

```{r}
World$fetchDataUnits("SettlingVelocity")
# World$NewCalcVariable("SettlingVelocity")
# World$CalcVar("SettlingVelocity")
```
