---
title: "3.3 Characteristics of the landscape"
author: "Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Landscape characteristics

In its default settings, SimpleBox represents the Earth’s northern 
hemisphere: continents surrounded by oceans in three climatic zones, 
and continents composed of river catchments, one of which is detailed 
as regional environment. 

#### Initialisation 

```{r preparation, message=FALSE, warning=FALSE}
#First, we initialize the world object
source("baseScripts/initWorld_onlyMolec.R")
```


#### Examples of parameters
Several examples of landscape characteristics are given below. Area characteristics are specifically discussed in vignette 5.1-Area. All units are SI units.

```{r parameters}

# Each scale has its own temperature
World$fetchData("Temp")

##windspeed is constant
World$fetchData("WINDspeed")

## and so is the rain rate
World$fetchData("RAINrate")


```