---
title: "4 Intermedia Partitioning"
author: "Jaap Slootweg, Joris Quik, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Intermedia processing

#### Description of the process

This vignette describes chapter 3.4 of the Schoorl et al. (2015) paper.There are multiple transfers for substances from one matrix to the other (matrix in the sense of medium, like air or water). The speed of some of these transfers are calculated as processes; you will find them in SubCompartProcesses.csv. Other transfers are so quick and small that a continuous equilibrium is assumed. In this vignette will will focus on this intermedia partitioning. See the paper for additional description.

#### Initializing World

Before any calculations can be done, we need to initialize the World object again. We give opportuntiy to choose from 4 different molecule types.

```{r initiate, message=FALSE, warning=FALSE, include=FALSE}
#if no substance is defined, a default substance will run.
substance <- "organic acid" #example of acid
#substance <- "organic base" #example of base
#substance <- "Ag(I)" #example of metal

source("baseScripts/initWorld_onlyMolec.R")
```

#### Partitioning coefficients of substances

If "World" has found data for the substance you initiated with, this is perfect. But these properties are not always known. Therefore, SimpleBox has functions that approximate these parameters. In this part we will discuss the solid-water partioning coefficient (Ksw).Generally, all these functions are based on the paper (Schoorl et al., 2015).

The function is given below, the parameters are: Kow, pKa, Corg, a, b, ChemClass and RHOsolid; The code is not nicely formatted by Rmd.

```{r}
f_Ksw
```

If the substance does not have all constants as needed for this function, you can prescripe them through the function World\$SetConst.

```{r}
Kow <- World$fetchData("Kow")
if(is.na(World$fetchData("Kow"))) {
  World$SetConst(Kow = 2750)
}
```

For this process, we need the standard organic carbon content (Corgstandard), instead of the Corg per compartment. The differences are examplified below.

```{r}
Corg <- World$fetchData("Corg")
print(Corg)
CorgStandard <- World$fetchData("CorgStandard")
print(CorgStandard)
```

pKa could be missing (e.g. in the case of "default substance"). If the substance is neutral, you can apply a value of 7. Variables a and b come from the QSARtable. RHOsolid can be taken from the matrix that is considered for the process.

```{r ksw}
pKa <- World$fetchData("pKa")
if(is.na(pKa)) {
  World$SetConst(pKa = 7) 
  warning("pKa not given in input data. Substance assumed to be neutral (pKa = 7).")
}

Substance_ChemClass <- World$fetchData("ChemClass")
QSARtable <- World$fetchData("QSARtable")
print(QSARtable)
QSARrecord <- QSARtable[QSARtable$QSAR.ChemClass == Substance_ChemClass,]
RhoTable <- World$fetchData("rhoMatrix")
RHOsolid <- RhoTable$rhoMatrix[RhoTable$SubCompart == "othersoil"]
KswModelled <- f_Ksw(Kow=Kow, pKa=pKa, CorgStandard=CorgStandard, 
                     a = QSARrecord$a, b = QSARrecord$b, 
                     ChemClass=Substance_ChemClass,
                     RHOsolid=RHOsolid,
                     alt_form = F)
print(KswModelled)
```

We now have all the parameters and can set the value for Ksw and know that the system will use our modelled Ksw.

```{r}
World$SetConst(Ksw = KswModelled)
World$fetchData("Ksw")

```

Not in the data is Ksw for the alternative form. The same function f_Ksw is applied by the defining function Ksw.alt, creating the SB variable:

```{r ksw.alt}
Ksw.alt
test <- World$NewCalcVariable("Ksw.alt")
World$CalcVar("Ksw.alt")
World$fetchData("Ksw.alt")

```

#### Fraction molecular species in original form (based on pKa)

We will now walk through some other calculated variables.

This is the fraction of a substance that is in the original form (non-dissociated), which relates to the pH of the compartment.

The FRorig for Matrix "air" is the FRorig of aerosols in air. The pH of "air" is set to 3 in the SubCompartSheet.csv, corresponding to "pH.aerw" in the excel version.

```{r FRorig}
World$NewCalcVariable("FRorig")
World$CalcVar("FRorig")
World$fetchData("FRorig")

World$NewCalcVariable("FRorig_spw")
World$CalcVar("FRorig_spw")
```

#### Partitioning coeficient Kp

The partitioning of a subcompartment / water is Kp, which is also an object-oriented state variable. This is dependent on many parameters, of which Ksw, Ksw_alt, FRorig and Corg are examples.See Schoorl et al. (2015) or the functions for the specific formulas.

```{r Kp}

World$NewCalcVariable("Kp")
World$CalcVar("Kp")

World$NewCalcVariable("D")
World$CalcVar("D")
World$NewCalcVariable("KpCOL")
World$CalcVar("KpCOL")
```

#### Dimensionless partition coefficients per compartment/scale

Similar to the solid-water partitioning coefficient, partition coefficients for other compartments also need to be computed. 

_Air_
The substance specific air/water partition coefficient at 25 degrees Celsius (Kaw25) is required for the calculation of scale-specific partition coefficients of air/water, aerosol water/air, and aerosol solids/air. When not provided as input, it is calculated within the functions for the scale specific partition coefficients (and not first as a separate variable with CalcVar).

```{r Kaw}
World$fetchData("Kaw25")

#Dimensionless air/water partition coefficient
World$NewCalcVariable("Kacompw")
World$CalcVar("Kacompw")

#Dimensionless aerosol water/air partition coefficient
World$NewCalcVariable("Kaerw") 
World$CalcVar("Kaerw")

#Dimensionless aerosol solids/air partition coefficient
World$NewCalcVariable("Kaers")
World$CalcVar("Kaers")

```

_Soil/Sediment_

Calculating the partitioning between soil- or sediment and water requires the fractions of water and air in soil, in addition to some of the previously calculated variables such as Kp, Kacompw, and FRorig_spw. The fractions of water and air in soil are provided as input ("FRACw" and "FRACa") in "ScaleSubCompartData.csv" and can differ between both (combinations of) scale and compartment. The fraction of solids in soil and sediment is calculated based on FRACw and FRACa, except for the fraction of solids in air, where it is also provided as input. Input data is provided as subFRAC, FRACx is calculated. 


```{r}
#fractions
World$fetchData("subFRACw")
World$fetchData("subFRACa")
World$NewCalcVariable("FRACa")
World$CalcVar("FRACa")
World$NewCalcVariable("FRACw")
World$CalcVar("FRACw")
World$NewCalcVariable("FRACs")
World$CalcVar("FRACs")

#partition coefficients
World$NewCalcVariable("Ksdcompw") #sediment/water
World$CalcVar("Ksdcompw")

World$NewCalcVariable("Kscompw") #soil/water
World$CalcVar("Kscompw")

```

#### Fraction chemical in gas, water or solid phase

The fractions of the chemical in the gas, water and solid phase of a compartment are calculated as variables with the following functions:

```{r FRchem}
World$NewCalcVariable("FRingas")
World$CalcVar("FRingas")

World$NewCalcVariable("FRinaers")
World$CalcVar("FRinaers")

World$NewCalcVariable("FRinaerw")
World$CalcVar("FRinaerw")

World$NewCalcVariable("FRinw")
World$CalcVar("FRinw")

World$NewCalcVariable("FRins")
World$CalcVar("FRins")

```

These variables described the partitioning constants. The speed in which the equilibrium are reached is modelled by diffusion processes, like volatilisation and absorption. This is described in a separate vignette, namely DiffusionProcesses.
