---
title: "5.1 Area"
author: "Jaap Slootweg, Joris Quik, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Calculation of area of SubCompartment within Scales

In the R version the calculation of the scales is based on data input of
TotalArea and FRACsea, from which AreaLand and AreaSea are calculated.
Scales can be nested: e.g. by default the Regional and Continental scale
are included in the Moderate scale.

The calculation of area slightly differs from the excel version. In
theExcel version calculation of scales is the other way around for the
regional and continental scale: AreaLand and AreaSea are data inputs
from which the total system area is calculated. The way of calculating
scales in the R version is more consistent across scales, and also makes
adding additional scales (e.g. a local scale within regional) or
changing the nesting of scales (e.g continental in tropic instead of
moderate) more straightforward.

```{r start.data, message=FALSE, warning=FALSE}
source("baseScripts/initWorld_onlyMolec.R")
World$fetchData("TotalArea")
World$fetchData("FRACsea")

```

#### Nesting and splitting land and sea

Scales can be nested in other scales. TotalArea of a scale represent the
total area of that scale, *including* the TotalArea of any scales nested
within them. However, AreaLand and AreaSea represent the area of
respectively land and sea of that scale *excluding* any nested scales.

Therefore, the TotalArea of any nested scales is substracted before
calculating AreaLand and AreaSea. For example, The TotalArea of
Continental contains Regional. Therefore,the TotalArea of Continental is
reduced by Regional, before the sea and land fraction are applied.
Similarly, currently by default Continental is included in the Moderate
scale, and its TotalArea is substracted from the TotalArea of the
moderate scale before calculation of AreaLand and AreaSea. Note that the
TotalArea of the Continental scale already includes the TotalArea of the
Regional scale, so only TotalArea of the continental scale is
substracted (and not also TotalArea of regional separately, as that
would be double substracting the area of the regional scale).

Regional and Arctic currently have no nested scale, and are
straightforward calculation of AreaSea and AreaLand as respectively.
TotalArea\*FRACsea and TotalArea\*(1-FRACsea)

#### 2025.x.x: update:
Instead of hardcoding the nesting and structure, a ScaleNestGroup and the exising ScaleOrder are applied to calculate the 

```{r}
v_AreaSea <- World$NewCalcVariable("AreaSea")
tAreaSea <- World$CalcVar("AreaSea")
v_AreaLand <- World$NewCalcVariable("AreaLand")
tAreaLand <- World$CalcVar("AreaLand")
merge(tAreaLand, tAreaSea)
```

#### Variable references behind the scenes

*The part below gives an insight into the specific workings of the SBOO
repository. This is not strictly necessary for understanding the
workings of the model.*

Variables are often tables, as can be seen above for the AreaLand
variable. When a variable is applied in a “defining function” the SBoo
system looks for the proper dimension, e.g. when it finds Arealand and
it is calculating for Arctic it will use the AreaLand of Arctic. But
when calculating for instance AreaSea for Continental we need TotalArea
for both Continental AND Regional! 

In such a case we can use the “all.” preposition. The function will
receive the whole table for the variable at hand. As an example look at
the defining function for AreaSea:

```{r}
AreaSea
```

#### The areas to be used in further calculations

AreaLand and AreaSea are used in the calculation of the areas of all
Scale/SubCompart combinations. The calculation of the area differs per
subcompartment. We therefore include SubCompartName as parameter in the
defining function and use it to distinguish which formula applies (see
the Area function below).

landFRAC is the fraction that belongs to a specificic subcompartment (e.g. agricultural soil) of the total AREALand.

All Scale/SubCompart combinations (present in states) have an area,
including the subcompart “cloudwater”! It receives the same area as air,
namely the sum of AreaSea and AreaLand for the scale at hand. For sea
and for oceans, if present, the area equal that of AreaSea. Otherwise
(different Subcompartments that add up to “AreaLand”) the calculation is
straightforward (landFRAC \* AreaLand).

Below we show the Area function.


```{r}
World$fetchData("landFRAC")
Area
v_Area <- World$NewCalcVariable("Area")
tArea <- World$CalcVar("Area")
```

#### Verification of size compartments
Just taking the sum of areas grouped by scale would include double
counting the subcompartments which are on top of each other, or mixed:
air / cloudwater / sea / deepocean! But we can test if the TotalArea of
Arctic + Moderate + Tropic equals the total Area of air.

```{r}
TotAreaGlobalScales <- World$fetchData("TotalArea")
TotAreaGlobalScales <- TotAreaGlobalScales$TotalArea[TotAreaGlobalScales$Scale %in% c(
  "Arctic","Moderate","Tropic")]
print(TotAreaGlobalScales)
TotAir <- World$fetchData("Area")
TotAir <- TotAir$Area[TotAir$SubCompart == "air"]
print(TotAir)
sum(TotAreaGlobalScales) == sum(TotAir)
```
