---
title: "5.4 Advection"
author: "Jaap Slootweg, Joris Quik, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

### 5.4 Advection

Here the implementation of the advective flows of air and water in SBoo
are illustrated. An advective flow in the context of SB is the massflow
of a matrix (air, water ) from one box to another, where either the
subcompartment or scale is different.

```{r initialize, include=FALSE}
source("baseScripts/initWorld_onlyMolec.R")
```

For calculating the rate constants for advection there is data needed on
several parameters such as wind speeds, discharges, residence times etc.
All the different fluxes or flows of air or water are calculated after
which these flows can be converted to rate constants. All default data
are stored in csv files and the values can be read using
'World\$fetchData("...")'. They can be adjusted using
'World\$mutateVars(...)' and then the World can be updated for
downstream variable updates using: 'WorldUpdateDirty(unique(...))'.
Examples can be found in the case studies.

Here we illustrate all the flow processes defined.

#### Air flow

The air flow rate is calculated in two steps.

First the AirFlow is calculated based on the residence time of air:

```{r}
World$fetchData("Volume")
World$fetchData("Area")
World$fetchData("WINDspeed")
World$fetchData("SubCompartName")

f_TAU

AirFlow

World$CalcVar("AirFlow")

```

The mean residence time in the region is calculated according the f_TAU
function.

The AirFlow function calculates the amount that would enter/leave the
region; namely the volume divided by the residence time. Notice that
f_TAU is a normal function, but AirFlow will be used as a variable
defining function. This provides the overall airflow within a
compartment which affects the flow from this compartment or scale to the
next. Currently air only flows between scales.

Second, from the AirFlow the flows are actually connected to each other.

```{r}
x_Advection_Air
# AdvAir <- World$NewFlow("x_Advection_Air")
# AdvAir$execute(debugAt = list(toScale = "Continental"))
World$CalcVar("x_Advection_Air")
```

To ensure mass balance in the model the amount from one scale to the
next would need to be equal to the amount flowing the opposite way! To
honor the mass balance we give priority to the amount calculated for the
smallest volume to calculate the actual flows. This means that for
nested scales we need to know their 'place' in the world. This brings a
new challenge; we need to know properties of the scale the flow is going
to, or possibly the scale where it is coming from. Because of the
nesting, we need to know even properties of scale not directly related
to the to- or from- the scale! To accommodate this use of the "to.", the
"from." and the "all." preposition is made available. The to- and from
scaleName are atomic (a single string) but the "all." prepositions
provides a full table of the variable and its dimensions (in this case
only across Scales). If you want to see what is really going on in the
function calls you can use the debug parameter.

This provides the air flows. To test if they add up and give the same
amount flowing to and from each scale we do the following calculation:

```{r}
AirFlows <- World$fetchData("x_Advection_Air")
ToFlow <- aggregate(flow~toScale, data = AirFlows, FUN = sum)
FromFlow <- aggregate(flow~fromScale, data = AirFlows, FUN = sum)
merge(FromFlow, ToFlow, by.x = "fromScale", by.y = "toScale", suffixes = c("_from", "_to")) 
merge(FromFlow, ToFlow, by.x = "fromScale", by.y = "toScale", suffixes = c("_from", "_to")) |> mutate(from_to_Diff = flow_from-flow_to)

```

#### Water advection

All flows dealing with advective processes are denoted by x_files in
SBOO. Different dimensions are relevant for computing the water flows:
subcompartments and scale. This relates to water flowing from lake,
through the river compartment to the sea, after which the sea flows
between scales. This is defined in the following x_flow. The order of
calculation matters (for regional and continental scale function x 1-5)
!

*1. x_ContRiver2Reg*

x_ContRiver2Reg: based on Runoff calculates the flow from contintental
river to regional river subcompartment (not usually relevant, thus often
0).

```{r}
x_ContRiver2Reg

World$NewCalcVariable("Runoff")
World$CalcVar("Runoff")

World$NewCalcVariable("RainOnFreshwater")
World$CalcVar("RainOnFreshwater")

World$fetchData("ScaleName")
World$fetchData("SubCompartName")

World$fetchData("dischargeFRAC")
World$fetchData("LakeFracRiver")
World$fetchData("Runoff")
World$fetchData("Area")

World$fetchData("LakeFracRiver")

flow1 <- World$NewFlow("x_ContRiver2Reg")

flow1$FromAndTo
flow1$execute()
World$CalcVar("x_ContRiver2Reg")
```

*2. x_RiverDischarge*

```{r}

World$fetchData("Runoff")

World$fetchData("RainOnFreshwater")

World$fetchData("dischargeFRAC")
World$fetchData("x_ContRiver2Reg")
World$fetchData("ScaleName")

flow2 <- World$NewFlow("x_RiverDischarge")

flow2$FromAndTo
flow2$execute()
World$CalcVar("x_RiverDischarge")
```

*3. x_LakeOutflow*

```{r}
World$fetchData("x_RiverDischarge")
World$fetchData("x_ContRiver2Reg")
World$fetchData("LakeFracRiver")
World$fetchData("ScaleName")

flow3 <- World$NewFlow("x_LakeOutflow")

flow3$FromAndTo
flow3$execute()
World$CalcVar("x_LakeOutflow")

```

*4. x_ContSea2Reg*

```{r}
World$fetchData("LakeFracRiver")
World$fetchData("x_RiverDischarge")

flow4 <- World$NewFlow("x_ContSea2Reg")

flow4$FromAndTo
flow4$execute()
World$CalcVar("x_ContSea2Reg")
World$fetchData("x_ContSea2Reg")
World$fetchDataUnits("x_ContSea2Reg")

```

*5. x_RegSea2Cont*

```{r}

World$fetchData("LakeFracRiver")
World$fetchData("x_RiverDischarge")

flow5 <- World$NewFlow("x_RegSea2Cont")

flow5$FromAndTo
flow5$execute()
World$CalcVar("x_RegSea2Cont")

```

*6. x_ToModerateWater*

Advection of continental sea, Tropic and Arctic sea compartments to sea
and deepocean at Moderate scale.

```{r}

World$fetchData("Volume")
World$fetchData("TAUsea")
World$fetchData("x_RegSea2Cont")


flow6 <- World$NewFlow("x_ToModerateWater")

flow6$FromAndTo
flow6$execute()
World$CalcVar("x_ToModerateWater")

```

### Water flow from Moderate Scale

*7. x_FromModerate2TropWater*

Separated in different flows based on the to Scale

```{r}

World$fetchData("Volume")
World$fetchData("TAUsea")
World$fetchData("OceanCurrent")
World$fetchData("x_RegSea2Cont")
World$fetchData("ScaleName")

flow <- World$NewFlow("x_FromModerate2TropWater")
# flow$execute(debugAt = list())
flow$FromAndTo
flow$execute()
World$CalcVar("x_FromModerate2TropWater")
```

*8. x_FromModerate2ArctWater*

```{r}
World$fetchData("OceanCurrent")
World$fetchData("ScaleName")

flow <- World$NewFlow("x_FromModerate2ArctWater")
# flow$execute(debugAt = list())
flow$FromAndTo
flow$execute()
World$CalcVar("x_FromModerate2ArctWater")
```

*9. x_FromModerate2ContWater*

```{r}
# required data:
World$fetchData("Volume")
World$fetchData("TAUsea")
World$fetchData("OceanCurrent")
World$fetchData("x_RegSea2Cont")
World$fetchData("ScaleName")

flow <- World$NewFlow("x_FromModerate2ContWater")
# flow$execute(debugAt = list())
flow$FromAndTo
flow$execute()
World$CalcVar("x_FromModerate2ContWater")
```

*10. x_OceanMixing2Deep*

```{r}
x_OceanMixing2Deep

World$fetchData("Volume")
World$fetchData("TAUsea")
World$fetchData("OceanCurrent")
World$fetchData("SubCompartName")

flow <- World$NewFlow("x_OceanMixing2Deep")

flow$FromAndTo
flow$execute()
World$CalcVar("x_OceanMixing2Deep")

```

*11. x_OceanMixing2Sea*

```{r}
x_OceanMixing2Sea

World$fetchData("Volume")
World$fetchData("TAUsea")
World$fetchData("OceanCurrent")
World$fetchData("SubCompartName")

flow <- World$NewFlow("x_OceanMixing2Sea")

flow$FromAndTo
flow$execute()
World$CalcVar("x_OceanMixing2Sea")
```

#### Overview of water flows

Overview of flows included above (excludes surface - deep ocean mixing
and ocean currents at global scales)

```{r}
flows = World$moduleList[["k_Advection"]]$WithFlow

flux2Test <- World$fetchData("Flows")
flux2Test <- flux2Test[flux2Test$FlowName %in% flows & flux2Test$fromSubCompart %in% c("river","sea","lake","deepocean"),]
flux2Test$fromScale <- factor(flux2Test$fromScale, levels = c("Regional", "Continental", "Moderate","Arctic","Tropic")) #force order
flux2Test$toScale <- factor(flux2Test$toScale, levels = c("Regional", "Continental", "Moderate","Arctic","Tropic")) #force order
flux2Test$fromSubCompart <- factor(flux2Test$fromSubCompart, levels = c("lake", "river", "sea","deepocean")) #force order
flux2Test$toSubCompart <- factor(flux2Test$toSubCompart, levels = c("lake", "river", "sea","deepocean")) #force order
f2Torder <- order(flux2Test$fromScale, flux2Test$fromSubCompart, flux2Test$toScale, flux2Test$toSubCompart)

flux2Test$fcolumns <- do.call(paste, c(flux2Test[c("fromScale", "fromSubCompart")], sep="."))
flux2Test$frows <- do.call(paste, c(flux2Test[c("toScale", "toSubCompart")], sep = "."))
#flux2Test$flux = 10e9 * flux2Test$flux # tune unit 

table1 <- pivot_wider(flux2Test[f2Torder,c("frows","fcolumns","flow")],
            id_cols = frows, names_from = fcolumns, values_from = flow)

print(table1, n=40)


```

#### Advection rate constant

This process is only one step beyond the calculation of the relevant
flows, see the function f_Advection. All flows are by default associated
with this process. "flow" is used as special keyword, there is no SB
variable with the name flow. You can find the values by fetching
"Flows". This also means the script to get the k_Advection is a little
different from other k's. See below:

```{r}
World$fetchData("Flows")
 k_Advection
# Don't create a new process like: World$FromDataAndTo("k_Advection")
# find the existing one like this:
testProc <- World$moduleList[["k_Advection"]]
# FromDataAndTo is from and to from the data, ultimately the property is used
# which can derived from the flows for Advection
FAT <- testProc$FromAndTo
# the data frmae to be calculated:
left_join(rename(FAT, "Scale" = "fromScale", "SubCompart" = "fromSubCompart"), World$fetchData("Volume"))
#for testing use:
testProc$execute()

```
