---
title: "Advection"
author: "Jaap Slootweg, Joris Quik"
date: "17-11-2023"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

# Advection

This vignette demonstrates the implementation of the advective flows of
air and water in SimpleBox (SB). An advective flow in the context of SB
is the massflow of a matrix (air, water, ) from one box to another,
where either the subcompartment or scale is different.

```{r initialize}
source("baseScripts/initTestWorld.R")
World
```

## Data and variables

For calculating the rate constants for advection there is data needed
on: - Windspeed Together with the variables: -

```{r}
SBvars <- c("AreaLand",
            "AreaSea",
            "Area",
            "Volume"
            )

for (x in SBvars) {
  World$NewCalcVariable(x)
  World$CalcVar(x)
}
```

## Air flow

We calculate AirFlow in two steps. The first is the application of a
simple model of air circulation depending on windspeed; The mean
residence time in the region is according the f_TAU function. `r f_TAU`
The AirFlow function calculates the amount that would enter/leave the
region namely the volume divided by the residence time. Notice that
f_TAU is a normal function, but AirFlow will be used as a variable
defining function. After the call to World\$CalcVar("AirFlow") this
property is stored in the data within World. `r AirFlow`. This provides
the overal airflow within a compartment which affects the flow from this
compartment or scale to the next. Currently air only flows between
scales.

```{r}
World$fetchData("Volume")
World$fetchData("Area")
World$fetchData("WINDspeed")
World$fetchData("SubCompartName")
# f_TAU
# AirFlow
World$NewCalcVariable("AirFlow")
World$CalcVar("AirFlow")
```

### flow mass balance

To keep the mass balance in the model the amount from one scale to the
next would need to be equal to the amount flowing the opposite way! To
honor the mass balance we give priority to the amount calculated for the
smallest volume to calculate the actual flows. This means that for
nested scales we need to know how the nesting is. This brings a new
challange; we need to know properties of the scale the flow is going to,
or possibly the scale where it is coming from. Because of the nesting,
we need to know even properties of scale not directly related to the to-
or from- the scale! To accomodate this use of the "to.", the "from." and
the "all." preposition is made available. The to- and from scaleName are
atomic (a single string) but the "all." prepositions provides a full
table of the variable and its dimensions (in this case only accross
Scales). If you want to see what is really going on in the function
calls you can use the debug parameter. See the Debugging vignette for
further information.

```{r}
x_Advection_Air
AdvAir <- World$NewFlow("x_Advection_Air")
#AdvAir$execute(debugAt = list(toScale = "Continental"))
World$CalcVar("x_Advection_Air")

```

This provides the air flows. To test if they add up and give the same
amount flowing to and from each scale we do the following calculation:

```{r}
AirFlows <- World$fetchData("x_Advection_Air")
ToFlow <- aggregate(flow~toScale, data = AirFlows, FUN = sum)
FromFlow <- aggregate(flow~fromScale, data = AirFlows, FUN = sum)
merge(FromFlow, ToFlow, by.x = "fromScale", by.y = "toScale", suffixes = c("_from", "_to"))
```

## Water flow

For water the flows are relevant for different dimensions:
subcompartments and scale. This relates to water flowing from lake,
through the river compartment to the sea. And the sea flows between
scales. This is defined in the following x_flow algorithms for water
advection in the following sections. The order of calculation matters!

1.  x_RiverSeaScales

2.  x_ContSea2Moder

3.  x_RiverSeaScales

### 1. xContRiver2Reg

x_ContRiver2Reg: based on Runoff calculates the flow from contintental
river to regional river subcompartment (not usually relevant, thus ofter
0).

```{r}
x_ContRiver2Reg

World$NewCalcVariable("Runoff")
World$CalcVar("Runoff")

World$NewCalcVariable("RainOnFreshwater")
World$CalcVar("RainOnFreshwater")

World$fetchData("ScaleName")
World$fetchData("SubCompartName")

World$fetchData("dischargeFRAC")
World$fetchData("LakeFracRiver")
World$fetchData("Runoff")
World$fetchData("Area")

World$fetchData("LakeFracRiver")

flow1 <- World$NewFlow("x_ContRiver2Reg")

flow1$FromAndTo
flow1$execute()
World$CalcVar("x_ContRiver2Reg")
```

### 2. x_RiverDischarge

```{r}

World$fetchData("Runoff")

World$fetchData("RainOnFreshwater")

World$fetchData("dischargeFRAC")
World$fetchData("x_ContRiver2Reg")
World$fetchData("ScaleName")

flow2 <- World$NewFlow("x_RiverDischarge")

flow2$FromAndTo
flow2$execute()
World$CalcVar("x_RiverDischarge")
```

### 3. x_LakeOutflow

```{r}
World$fetchData("x_RiverDischarge")
World$fetchData("x_ContRiver2Reg")
World$fetchData("LakeFracRiver")
World$fetchData("ScaleName")

flow3 <- World$NewFlow("x_LakeOutflow")

flow3$FromAndTo
flow3$execute()
World$CalcVar("x_LakeOutflow")

```

### 4. x_ContSea2Reg

```{r}
# these cannot be found:
World$fetchData("LakeFracRiver")
World$fetchData("x_RiverDischarge")

flow4 <- World$NewFlow("x_ContSea2Reg")

flow4$FromAndTo
flow4$execute()
World$CalcVar("x_ContSea2Reg")
```

### 5. x_RegSea2Cont

```{r}

World$fetchData("LakeFracRiver")
World$fetchData("x_RiverDischarge")

flow5 <- World$NewFlow("x_RegSea2Cont")

flow4$FromAndTo
flow4$execute()
World$CalcVar("x_RegSea2Cont")

```

### x_RiverSeaScales

```{r RiverSeaScales}
x_RiverSeaScales

# varaiables:
World$fetchData("SubCompartName")
World$fetchData("ScaleName")
World$fetchData("OceanCurrent")
World$fetchData("TAUsea")
World$fetchData("ContinentalInModerate")

# these cannot be found:
World$fetchData("x_ContSea2Reg")
World$fetchData("ContRiver2Reg")
World$fetchData("RegSea2Cont")

AdvAir <- World$NewFlow("x_RiverSeaScales")
#AdvAir$execute(debugAt = list(toScale = "Continental"))
World$CalcVar("x_RiverSeaScales")

```

This provides the air flows. To test if they add up and give the same
amount flowing to and from each scale we do the following calculation:

```{r}
AirFlows <- World$fetchData("x_Advection_Air")
ToFlow <- aggregate(flow~toScale, data = AirFlows, FUN = sum)
FromFlow <- aggregate(flow~fromScale, data = AirFlows, FUN = sum)
merge(FromFlow, ToFlow, by.x = "fromScale", by.y = "toScale", suffixes = c("_from", "_to"))
```

## Advection rate constant
