---
title: "7.1.4 Resuspension from water to sediment"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

## Resuspension
Resuspension is one of two processes transporting molecules/particles from sediment to water (the other process is desorption). Besides resuspension and desorption, the net exchange between water and sediment also depends on sedimentation (see 7.1.3).

```{r, warning=FALSE}
substance <- "nAg_10nm"
source("baseScripts/initWorld_onlyParticulate.R")
```

One of the variables needed to calculate the resuspension rate is the settling velocity. This variable is calculated using the helper function f_SetVelWater. 

```{r, warning=FALSE}
World$fetchData("VertDistance")
World$fetchData("DynViscWaterStandard")
World$fetchData("rhoMatrix")          # For the to compartment
World$fetchData("NETsedrate")         # For the to compartment
World$fetchData("RadCP")              # For the to compartment
World$fetchData("RhoCP")              # For the to and from compartment
World$fetchData("FRACs")              
World$fetchData("SUSP")               # For the to compartment
World$fetchData("SpeciesName")

World$fetchData("SettlVelocitywater")
```

First, the gross sedimentation rate is calculated. Then the resuspension flow is calculated by subtracting the net sedimentation rate from the gross sedimentation rate (which cannot exceed zero, i.e. the net sedimentation cannot be larger than the gross sedimentation). Finally, the resuspension flow is calculated by the depth of the sediment subcompartment 

```{r Calculate k_Resuspension, warning=FALSE}
resuspension <- World$NewProcess("k_Resuspension")
resuspension$execute()
```
