---
title: "7.1.2 Particulate Deposition air to soil/water"
author: "Jaap Slootweg, Joris Quik, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Deposition processes for particulates
The deposition process for particulates, contrary to molecules, is split into two separate process for wet and dry. First, we have to initialize our particulate world again. 
```{r initiate, message=FALSE, warning=FALSE}

source("baseScripts/initWorld_onlyParticulate.R")

```

#### Variable check

See if the following variable are available. They are required for the DryDeposition function. 

```{r paged.print=TRUE}

World$fetchData("Area")
World$fetchData("Volume")
World$fetchData("AEROresist")
World$fetchData("Matrix")
#Diffusivity is calculated based on f_Diffusivity

World$fetchData("DynVisc")
World$fetchData("rhoMatrix")
World$fetchData("ColRad")

World$NewCalcVariable("rad_species")
World$CalcVar("rad_species")
World$NewCalcVariable("rho_species")
World$CalcVar("rho_species")

World$fetchData("Temp")
World$fetchData("alpha.surf")
# Cunningham calculated using f_Cunningham

# World$fetchData("SettlingVelocity")
World$NewCalcVariable("SettlingVelocity")
World$CalcVar("SettlingVelocity")

World$fetchData("gamma.surf")
World$fetchData("FricVel")
World$fetchData("SubCompartName")


```


#### Calculation of Dry Deposition rate constant for particulates

The implementation follows the LOTOS-EUROS v2.0 2016 guidance for dry deposition of particles. This does mean there might be small deviations from the current implementation in SimpleBox4nano xlsx. However, a Test function is provided for initializing the world that can reproduce the exact k's from Excel.

```{r Initializing test function, warning=FALSE }
#only if you want exact same output as Excel version
World$SetConst(Test = "TRUE")
#Updating of matrix because of the Test function 
World$UpdateKaas(mergeExisting = F)
```

```{r k_DryDeposition, warning=FALSE, paged.print=TRUE}

World$FromDataAndTo("k_DryDeposition")
testProc <- World$NewProcess("k_DryDeposition")
testProc$execute()

```

#### Calculation of Wet Deposition rate constant for particulates
For wet deposition there are two steps where the particulates are first scavenged into an intermediate cloud water compartment after which they deposit to soil or surface water through the process of k_WetDeposition. 

```{r k_CWscavenging}

World$fetchData("NaturalRho")
World$fetchData("SettlingVelocity") 
World$fetchData("rhoMatrix")
World$fetchData("RAINrate")
World$fetchData("FRACtwet")
World$fetchData("tdry")
World$fetchData("twet")
World$fetchData("COLLECTeff")
testProc <- World$NewProcess("k_CWscavenging")
testProc$execute()


```

```{r k_WetDeposition}

testProc <- World$NewProcess("k_WetDeposition")
testProc$execute()

```

