---
title: "7.1.3 Sedimentation from water to sediment"
output: github_document
editor_options: 
  chunk_output_type: console
---

Sedimentation rate depends on the settling velocity of the particle, which is calculated according the function given below; after the standard initialisation. Note that for Molecular species - all classic substances! - the settling velocity in excel is set at 2.5 m/day. In SB OO this is calculated using Stokes for SPM particle, assuming the not truly dissolved fraction settles at the rate of the SPM particle. TODO Joris?

```{r initiate, echo=TRUE, message=FALSE, warning=TRUE}

source("baseScripts/initTestWorld.R")

```
## v_functions and f_functions
SB OO uses functions in different ways. Some functions define the calculation of a state-variable aka landscape-variable. These functions will be passed (by name) to the method NewCalcVariable of a SBcore object, like World. Once a state-variable is calculated is is stored within the SBcore object and can be fetched, and therefore automatically used for process calculation, or for other state-variables. This function for the settlingvelocity is more generic, and will be used more generic than just for the state variable of nano species. Therefor every function that uses f_SettlingVelocity() must provide it with the correct variables.
A feature of defining functions is that parameters should be atomic. This enables easy use of if-then-else constructions. This construction is not possible if the if condition argument contains a vector. (You would have to use the ifelse function.)
(See AAAreadme.R in the SB project for more on file naming conventions)

## The process
For the Molecular species we use the settling rate of the SPM particle. This settling rate ia also needed to assess the build-up of the sediment layer, resuspention and burial. Therefore it is convenient to define it as a variable.
For Nano materials, the process rate is settling velocity divided by the depth of the layer. Because SettlingVelocity is a regular function ("f_SettlingVelocity") all the parameters for the settling are also needed. See the defining function (sources from newAlgorithmScripts at the moment):
```{r}
k_Sedimentation

World$fetchData("FRinw")
World$fetchData("SettlingVelocity")
World$fetchData("DynViscWaterStandard")
World$fetchData("VertDistance")
World$fetchData("RhoCP")
World$fetchData("RadCP")


source("newAlgorithmScripts/k_Sedimentation.R")

```
Sedimentation is directly to the sediment below.



