---
title: "Other K air"
author: "Jaap Slootweg, Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Other K Air

Here we describe the fate processess that are part of the "other" k Air,
as described in 7.1.1.

1.  \- Ab/adsorption

2.  \- Degradation

3.  \- Advection 

#### Initialization

We initialise a "World" and calculate the needed parameters.

```{r initiate, message=FALSE, warning=FALSE}
source("baseScripts/initWorld_onlyMolec.R")
```

#### Gas absorption

The overall mass transfer coefficient for gas absorption can be
estimated using the classical two - film resistance model (Schwarzenbach
et al., 1993). In the case of transport across the air/water interface,
the overall transfer coefficients follow from summation of the
resistances at the water and air sides of the interface. In the case of
transport across the air/soil interface, the soil-side of the interface
is treated as a pair of parallel resistances (air phase and water phase
of the soil).

Below, we start with calculating the fraction of speciations of the
molecules (e.g. how much is free in particular matrices)

```{r absorption}

# This needs to be calculated first:
World$NewCalcVariable("FRorig")
World$CalcVar("FRorig")
World$NewCalcVariable("FRorig_spw")
World$CalcVar("FRorig_spw")

# World$NewCalcVariable("FRACa")
World$fetchData("FRACa")
# World$NewCalcVariable("FRACw")
World$fetchData("FRACw")

World$NewCalcVariable("FRACs")
World$CalcVar("FRACs")
```

*MassTransferCoefficients* Next, we compute partial mass transfer
coefficients between compartments. These are based on wind matrices and
other parameters. We introduce all of them here:

Partial Mass Transfer Coefficients operate like resistors; to calculate
the total MTC we use 1/K = 1/k1 + 1/(kp.k2), with kp the equilibrium
constant. The soil side MTCa,s is related to the degradation of the
substance, for water/air the wind plays a role. The defining functions
are organised by the medium the film is bordering towards, but use the
properties of the medium () of the box) they are defined for, the
departure of the diffusion.

Following is the implementation of 4 MTC's.

1.  MTC_2a -\> for water side of the water-air interface (from water to
    air) and the soil side of the soil-air interface (from soil to air).

2.  MTC_2w -\> for air side of the air-water interface (from air to
    water) and the sediment side of the sediment-water interface (from
    sediment to water).

3.  MTC_2s -\> for the air side of the air-soil interface (from air to
    soil).

4.  MTC_2sd -\> for the water side of the water-sediment interface (from
    water to sediment).

```{r MTC}

World$NewCalcVariable("MTC_2a")
World$CalcVar("MTC_2a")
World$NewCalcVariable("MTC_2w")
World$CalcVar("MTC_2w")
World$NewCalcVariable("MTC_2s")
World$CalcVar("MTC_2s")
World$NewCalcVariable("MTC_2sd")
World$CalcVar("MTC_2sd")

World$NewCalcVariable("Kscompw")
World$CalcVar("Kscompw")
```

We calculate the absorption based on these parameters. Next, we also
have to correct for the relative area and vertical depth of the
compartment, e.g if it goes to water we correct for how much of the
total to area is water: AreaFrac = to.Area/(AreaLand+AreaSea)
return(GASABS/VertDistance\*AreaFrac)

```{r}

World$fetchData("VertDistance")
World$fetchData("landFRAC")
World$fetchData("Area")
World$fetchData("AreaLand")
World$fetchData("AreaSea")
World$fetchData("SpeciesName")

testProc <- World$NewProcess("k_Adsorption")
testProc$execute()


```

#### Degradation
Degradation of the molecule in the air also is part of other processes happening in the air. For a description of this process, see vignette 6.0 Removal. 

### Advection 
Advective processes (movement of air from one scale to another scale) are also part of the kaas in air. For a detailed description of this process, see vignette 5.4 Advection. 
