---
title: "7.3 Physical inter media transport processes"
author: "Jaap Slootweg, Joris Quik , Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

### Other intermedia transfer processes

Intermedia processes discussed here are:

1.  \- Erosion

2.  \- Run-off

3.  \- Volatilization

4.  \- Desorption

#### Initialization

We initialise a "World" and calculate the needed parameters.

```{r initiate, message=FALSE, warning=FALSE}
source("baseScripts/initWorld_onlyMolec.R")
```

#### Erosion

The processes of erosion and runoff are combined into one flow in the Excel version.
n the R version, if multiple processes are defined for the same transfer, the k's are automatically added.
This makes it easier to define and possibly re-define each process.
In this case the factor CORRrunoff / DEPTH will be executed twice.
We believe the loss in calculation speed is negligible.
And this approach is in favor because of the clearly defined process functions.

The factor CORRrunoff (Correction factor depth dependent soil concentration) however is calculated as EXP((-1/0.1) \* 0) \* (1/0.1) \* DEPTH / (1-EXP((-1/0.1) \* DEPTH)) for the assumed pentration depth of 0.1 m.
To make sure this remains the same for both Runoff and erosion, we define a variable to calculate it.
The only parameter, so far, is DEPTH, which has been renamed to vertDistance in the R version (combined with height).

```{r message=FALSE, warning=FALSE}
# data needed for k_Erosion:
World$fetchData("relevant_depth_s")
World$fetchData("penetration_depth_s")
World$fetchData("EROSIONsoil")
World$fetchData("VertDistance")
World$fetchData("ScaleName")

World$fetchData("SubCompartName")
World$fetchData("ScaleName")

test1 <- World$NewProcess("k_Erosion")
test1$execute()



```

#### Runoff

Runoff is implemented in a similar way.
Here the formula is: RunoffFlow / Kscompw \* CORRrunoff / Volume \* FracROWatComp.
For organic compounds we need Kscompw (Ksw), which is described in section [4.0 partitioning](#0).
The correction factor is explained above.
The RunoffFlow from soil is needed, which is dependant on the Area, FRACrun and the RAINrate.
Then a fraction of Runoff (FracROWatComp) can go to either the upstream (w0) or downstream (w1) freshwater compartment.

```{r message=FALSE, warning=FALSE}

RunoffFlow
World$fetchData("RunoffFlow")
FracROWatComp
World$fetchData("FracROWatComp")
Kscompw
World$fetchData("Kscompw")

f_CORRsoil
# with:
World$fetchData("relevant_depth_s")
World$fetchData("penetration_depth_s")
World$fetchData("VertDistance")

World$FromDataAndTo("k_Runoff")

test1 <- World$NewProcess("k_Runoff")

test1$execute()

```

#### Volatilisation

Volatilsation uses roughly the same two-film resistance model as gas absorption, as described in other K air.
However, volatilisation processes are corrected also based on the depth correction, as is described for the processes above.

```{r Volatilisation, warning= FALSE }


World$NewCalcVariable("MTC_2a")
World$CalcVar("MTC_2a")
World$NewCalcVariable("MTC_2w")
World$CalcVar("MTC_2w")
World$NewCalcVariable("MTC_2s")
World$CalcVar("MTC_2s")
World$NewCalcVariable("MTC_2sd")
World$CalcVar("MTC_2sd")

testvar <- World$NewCalcVariable("Kacompw")
testvar$execute(debugAt = list())
World$CalcVar("Kacompw")

CalcVariable$execute(debugAt = list(ScaleName = "Arctic", SubCompartName = "deepocean"))

debugonce(k_Volatilisation)


testProc <- World$NewProcess("k_Volatilisation")
testProc$execute(debugAt = list())
```

#### Desorption

Once again, Values for the overall mass transfer coefficients for direct desorption across the sediment-water interface can be obtained using the two-film resistance model.
However, for this we need the sediment water partition coefficient.
We also correct for the vertical distance of the compartment.

```{r Desorption, warning = FALSE}
World$NewCalcVariable("Ksdcompw")
World$CalcVar("Ksdcompw")

testProc <- World$NewProcess("k_Desorption")
testProc$execute(debugAt = list())


```
