---
title: "7.3 Physical inter media transport processes"
author: "Jaap Slootweg, Joris Quik , Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

### Other intermedia transfer processes

Intermedia processes discussed here are:

1.  \- Erosion

2.  \- Run-off

3.  \- Volatilization

4.  \- Desorption

#### Initialization

We initialise a "World" and calculate the needed parameters.

```{r initiate, message=FALSE, warning=FALSE}
source("baseScripts/initWorld_onlyMolec.R")
```

#### Erosion

In R processes across the same dimension, such as Runoff and Erosion are defined seperately.
The engine adds the rate constants (k's) automatically for the K matrix.
This approach is in favor because of the clearly defined process functions.

The factor CORRrunoff (Correction factor depth dependent soil concentration) however is calculated as EXP((-1/0.1) \* 0) \* (1/0.1) \* DEPTH / (1-EXP((-1/0.1) \* DEPTH)) for the assumed pentration depth of 0.1 m.
To make sure this remains the same for both Runoff and erosion, we define a variable to calculate it.
The only parameter, so far, is DEPTH, which has been renamed to vertDistance in the R version (combined with height).

At Regional and Continental scale Erossion and Runoff all goes to freshwater (w0 an w1), whereas at global scales (Tropic, Moderate and Arctic) all Erosion and Runoff goes to sea (w2).

At Regional and Continental scale the fraction of Erossion and Runoff going to either w0 or w1 is calculated using a variable: FracROWatComp.

```{r message=FALSE, warning=FALSE}
# data needed for k_Erosion:
FracROWatComp



World$fetchData("FracROWatComp")


k_Erosion

World$fetchData("relevant_depth_s")
World$fetchData("penetration_depth_s")
World$fetchData("EROSIONsoil")

World$fetchData("FracROWatComp")

World$fetchData("VertDistance")
World$fetchData("ScaleName")

World$fetchData("SubCompartName")
World$fetchData("ScaleName")

test1 <- World$NewProcess("k_Erosion")
test1$execute()



```

#### Runoff Molecular

Runoff is implemented in a similar way.
Here the formula is: RunoffFlow / Kscompw \* CORRrunoff / Volume \* FracROWatComp.
For organic compounds we need Kscompw (Ksw), which is described in section [4.0 partitioning](#0), as only the fraction in water can move with Runoff.
The correction factor is explained above.
The RunoffFlow from soil is needed, which is dependant on the Area, FRACrun and the RAINrate.
And at last the fraction of Runoff can go to either the upstream (w0) or downstream (w1) freshwater compartment as calculated by v_FracROWatComp.
FracROWatComp is now calculated based on the surface fraction (landFRAC) of w0 versus w1.
This can or should be adjusted when more specific data is available on the land surface area contributing to the upstream (w0) versus downstream (w1) Runoff.

```{r message=FALSE, warning=FALSE}



RunoffFlow
World$fetchData("RunoffFlow")
FracROWatComp
  World$fetchData("FracROWatComp")
Kscompw
World$fetchData("Kscompw")

f_CORRsoil
# with:
World$fetchData("relevant_depth_s")
World$fetchData("penetration_depth_s")
World$fetchData("VertDistance")

World$FromDataAndTo("k_Runoff")

test1 <- World$NewProcess("k_Runoff")

test1$execute()

```

#### Runoff Particulate species

For particulate species Runoff is not appliable for very large particles.
These behave more like other particulates that make up the soil or ar at least intercepted by vegetation reducing the Runoff rate.
For this reason a SizeRunoff is introduced above which Runoff is either set to 0 or to a value reduced by the VegInterceptFrac as defined in the IntrcptFrac variable.
See example code below.
SizeRunoff is set at the scale level, where currently the SizeRunoff is set at 100 micrometer for all scales.

```{r}
rm(list = ls())
source("baseScripts/initWorld_onlyParticulate.R")

RunoffFlow
World$fetchData("RunoffFlow")
FracROWatComp
World$fetchData("FracROWatComp")

World$fetchData("IntrcptFrac")
World$fetchData("VegInterceptFrac")
World$fetchData("SizeRunoff")
World$fetchData("rad_species")


World$FromDataAndTo("k_Runoff")

test1 <- World$NewProcess("k_Runoff")

test1$execute()

# if you want an Interceptation fraction  of 0.9 to be applied to all species above 100 nm
World$fetchData("VegInterceptFrac")
f_MutateHelper("VegInterceptFrac",
               c(0.9,0.9,0.9))

World$fetchData("SizeRunoff")
f_MutateHelper("SizeRunoff",
               c(0.1,0.1,0.1,0.1,0.1))

World$fetchData("IntrcptFrac")

test1 <- World$NewProcess("k_Runoff")

test1$execute()
```

#### Volatilisation

Volatilsation uses roughly the same two-film resistance model as gas absorption, as described in other K air.
However, volatilisation processes are corrected also based on the depth correction, as is described for the processes above.

```{r Volatilisation, warning= FALSE }


World$NewCalcVariable("MTC_2a")
World$CalcVar("MTC_2a")
World$NewCalcVariable("MTC_2w")
World$CalcVar("MTC_2w")
World$NewCalcVariable("MTC_2s")
World$CalcVar("MTC_2s")
World$NewCalcVariable("MTC_2sd")
World$CalcVar("MTC_2sd")

testvar <- World$NewCalcVariable("Kacompw")
# testvar$execute(debugAt = list())
World$CalcVar("Kacompw")

# CalcVariable$execute(debugAt = list(ScaleName = "Arctic", SubCompartName = "deepocean"))

# debugonce(k_Volatilisation)


testProc <- World$NewProcess("k_Volatilisation")
# testProc$execute(debugAt = list())
testProc$execute()

```

#### Desorption

Once again, Values for the overall mass transfer coefficients for direct desorption across the sediment-water interface can be obtained using the two-film resistance model.
However, for this we need the sediment water partition coefficient.
We also correct for the vertical distance of the compartment.

```{r Desorption, warning = FALSE}
World$NewCalcVariable("Ksdcompw")
World$CalcVar("Ksdcompw")

testProc <- World$NewProcess("k_Desorption")
testProc$execute()


```
