---
title: "9 Heteroagglomerationd and Fragmentation"
author: "Jaap Slootweg, Joris Quik, Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

```{r message=FALSE, warning=FALSE}
substance <- "Microplastic"
source("baseScripts/initWorld_onlyParticulate.R")

```

## Heteroagglomeration

This is a particulate process which only applies to Solid species and the transformation to the Aggregated or Attached Particulate Species.

There are three different approaches applied here depending on the matrix where this process takes place and the species affected.

The three approaches are based on: - Smoluchowski's colloid theory (water, soil/sediment pore water) - Filtration theory for porous matrices (soil and sediment) - Aerosol agglomeration approaches (air)

### Heteroagglomeration in water

This plays are role for the transformation from the primary particulate species and their agglomeration with natural colloids \< 450 nm in water, soil and sediment pore water. Also agglomeration with larger SPM (\>450 nm) in water.

The data required for calculating for instance the collision frequency is related to the particle number concentration of natural colloids present in the compartment. For this we use the per subcompartment defined characteristics of the natural particles: 'RadCOL', 'RadCP', 'RhoCOL' and 'RhoCP'. For aerosols the Accumulation mode corresponds to COL and Coarse Mode corresponds to CP characteristics.

The heteroaggregation rate constant is calculated using (k_HeteroAgglomeration.wsd.R). This depends on the following variables:

1.  rad_species: this calculates the radius of the heteroagglomerate consisting of the solid particle and the Small or Large natural particle.

2.  rho_species: this calculates the density of the he heteroagglomerate consisting of the solid particle and the Small or Large natural particle.

Furthermore there are several helper functions used to avoid repeating code and too complex variables:

-   f_Inter (Interception or peri-kinetic collision frequency coefficient)
-   f_Brown (Brownian motion or kinetic collision frequency coefficient) 
-   f_Grav (Gravitational impaction collision frequency coefficient)
-   f_NumConc (Calculate number concentration from mass concentration) 
-   f_SetVelWater (Function for calculating the Settling Velocity of a particle through water)

```{r Variables w.Heterogglomeration, include=TRUE, warning=FALSE}
# Below are all variables used to calculate heteroagglomeration in water
World$fetchData("alpha") # here to.alpha needs to be used
World$fetchData("COL")
World$fetchData("SUSP")
World$fetchData("Shear")
World$fetchData("RadS")
World$fetchData("RhoS")
World$fetchData("Temp")
World$fetchData("RhoCOL")
World$fetchData("RhoCP")
World$fetchData("RadCOL")
World$fetchData("RadCP")
World$fetchData("RadNuc")
World$fetchData("NumConcNuc")
World$fetchData("NumConcAcc")
World$fetchData("NumConcCP")
World$fetchData("DynViscWaterStandard")
World$fetchData("DynViscAirStandard")
World$fetchData("Df")
World$fetchData("SubCompartName")
World$fetchData("rhoMatrix")
World$fetchData("Matrix")
World$fetchData("SpeciesName")

# Calculate the radius of the heteroagglomerate
CalcVariable <- World$NewCalcVariable("rad_species")
World$CalcVar("rad_species")

# Calculate the density of the heteroagglomerate
World$NewCalcVariable("rho_species")
World$CalcVar("rho_species")

# for sediment and soil and het to attached species:
CalcVariable <- World$NewCalcVariable("FRACs")
World$CalcVar("FRACs")

```

```{r Calculate w.Heteroagglomeration, include=TRUE, warning=FALSE}
World$FromDataAndTo("k_HeteroAgglomeration.wsd")
testProc <- World$NewProcess("k_HeteroAgglomeration.wsd")
testProc$execute()
```

### Heteroagglomeration in air

Heteroagglomeration for attached species in soil and sediment is also included in k_HeteroAgglomeration.wsd. However, heteroagglomeration in air is calculated seperately using k_HeteroAgglomeration.a. In k_HeteroAgglomeration.a, two helper functions are defined:

-   ThermVel (calculates the thermal velocity)
-   f_Fuchs (calculates Fuch's correction factor). 

In addition, the helper function f_Diffusivity is used to calculate the diffusivitiy of the heteroagglomerate. 

```{r Calculate heteroagglomeration in air, include=TRUE, warning=FALSE}
# Below are all variables used to calculate heteroagglomeration in air
World$fetchData("RadNuc") 
World$fetchData("RadCOL") 
World$fetchData("RadCP") 

World$fetchData("NumConcNuc")
World$fetchData("NumConcAcc")
World$fetchData("NumConcCP")

World$fetchData("RhoNuc")
World$fetchData("RhoCOL")
World$fetchData("RhoCP")

World$fetchData("Temp")
World$fetchData("DynViscAirStandard")
World$fetchData("Matrix")

World$fetchData("Substance") 

# Calculate heteroagglomeration in air: 
World$FromDataAndTo("k_HeteroAgglomeration.a")
testProc <- World$NewProcess("k_HeteroAgglomeration.a")
testProc$execute()
```


## Fragmentation

This is the start of the fragmentation proces being included. This is now done for the fragmentation of heteroagglomerates to the originial species, S. This also allows for emission to be as a agglomerate, e.g. with TRWP, when this could fall appart in TWP.

The fragmentation process is only implemented for the existing species in SimpleBox. When Solid species would fall appart then this results in degradation. This could be implemented differently in the future, e.g. with classes of Solid species, but otherwise flows should be captured from Solid Fragmentation and used as emission in a new run of SimpleBox  seperately.

```{r}

World$fetchData("kdeg") 
World$fetchData("kfrag") 

World$FromDataAndTo("k_Fragmentation")
testProc <- World$NewProcess("k_Fragmentation")
testProc$execute()




```



```{r To Do, include=FALSE}
# [TO DO\* rework input of Aerosol concentrations into mass concentrations instead of number concentrations as they are now for COL and SUSP, and consider Nucc concentrations in CONSTANTS just like their Rad and Rho properties]
# 
# Potentially we can integrate air also in one algorithm, alhough it will be a big one...
```