---
title: "Getting started"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, warning=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

## Initialize

This vignette demonstrates how to use SimpleBox Object-Oriented (SBOO).

### Choose a substance

The first step is to initialize the model. Before initialization a
substance needs to be selected, otherwise the "default substance" is
used. To select another substance than "default substance", a substance
can be chosen from the "Substance" column of the data frame created
below.

```{r Substance list, include=TRUE, warning=FALSE}
#substances <- read.csv("data/Substances.csv")
substances <- read.csv("~/GitHub/SimpleBox/SBooScripts/data/Substances.csv")

# Assign a substance name from the Substance column to the variable "substance":
substance <- "1-aminoanthraquinone"
```

### Initialize the World object

The World object contains all variables and first order rate constants
(k's) for the chosen substance. This object is needed later to calculate
the masses in each compartment on each scale.

To initialize the World object, one of three script is called:

-   initWorld_onlyMolec.R (used for molecules)
-   initWorld_onlyParticulate.R (used for particulates)
-   initWorld_onlyPlastics.R (used for microplastics)

Choosing which script to use is dependent on the ChemClass of the chosen
substance (see "ChemClass" column in the "substances" data frame). Each
of these three classes require different initWorld scripts because
different processes and variables are used to calculate the k's. With
the chunk below, the correct initWorld script is automatically chosen
and run based on the chosen substance:

```{r Initialize World, include=TRUE, warning=FALSE}
library(tidyverse)

chemclass <- substances |>
  filter(Substance == substance) |>
  select(ChemClass)

chemclass <- chemclass$ChemClass

if(substance == "microplastic"){
  source("baseScripts/initWorld_onlyPlastics.R")
} else if (chemclass == "particle") {
  source("baseScripts/initWorld_onlyParticulate.R")
} else {
  source("baseScripts/initWorld_onlyMolec.R")
}

```

## Accessing variables and k's

Now that the World is initialized, its variables and calculated flows
can be accessed. To access these variables and k's, first the names of
the variables are needed:

```{r Get variable names, include=TRUE, warning=FALSE}
varnames <- World$fetchData()

print(varnames)
```

### Access variables

A specific variable (in this case AreaSea) can be accessed as follows:

```{r Access variables, include=TRUE, warning=FALSE}
World$fetchData("AreaSea")
```

### Access k's

The "kaas" variable contains a data frame with the k's, proccess name,
to-subcompartment name from-subcompartment name, to-scale name,
from-scale name, to-species name and from-species name. It can be
accessed in the same way other variables are accessed:

```{r Access kaas, include=TRUE, warning=FALSE}
df_ks <- World$fetchData("kaas")
```

## Calculate steady state output

To calculate steady state masses, emissions and a solver are needed. The
have to be given to the solver in a particular format.

### Create emissions data frame

To be able to calculate steady state masses, an emission data frame is
needed. The emissions data frame consists of one column with the
abbreviation of the scale-subcompartment-species combination, and
another column containing the emission to that compartment.

The abbreviations are as follows:

```{r Abbreviations, include=TRUE, warning=FALSE}
library(knitr)
scales <- data.frame(Scale = c("Arctic", "Continental", "Moderate", "Regional", "Tropic"), Abbreviation = c("A", "C", "M", "R", "T"))

subcompartments <- read.csv("data/SubCompartSheet.csv") |>
  select(SubCompartName, AbbrC) |>
  rename(Abbreviation = AbbrC) |>
  rename(SubCompartment = SubCompartName)

species <- read.csv("data/SpeciesSheet.csv") |>
  select(Species, AbbrP) |>
  rename(Abbreviation = AbbrP)

print(scales)
print(subcompartments)
print(species)

```

*Notes:*

-   *Compartment deepocean only exists on Global (Arctic, Moderate,
    Tropic) scale.*
-   *Compartments lake, river, agriculturalsoil and othersoil only exist
    on regional and continental scale.*

The abbreviations used in the emissions data frame are built up as
follows:

1.  Abbreviation of the subcompartment
2.  Abbreviation of the scale
3.  Abbreviation of the species.

Now the emissions data frame can be created. The column named "Abbr"
contains the abbreviations, and the column "Emis" contains the
emissions. In the example below, emissions of 10000 t/y go into regional
air, regional agricultural soil and regional river water. These
emissions in tonnes per year are then converted to mol/s.

**Notice that because this script is using a molecular substance, the
abbreviation "U" for "Unbound" is used here to specify emissions. If the
substance is a particle, use "S" for "Solid"!**

```{r Create emissions data frame, include=TRUE, warning=FALSE}
emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), Emis = c(10000, 10000, 10000) ) 

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(MW*365*24*60*60)) # convert 1 t/y to mol/s

print(emissions)

```

### Solve the matrix

To solve the matrix, a solver first needs to be specified. To solve for
a steady state we can use "SB1Solve". In this case the resulting steady
state masses are reported in the data frame "masses".

```{r Solve the matrix, include=TRUE, warning=FALSE}
# Specify which solver to use to the World object
World$NewSolver("SB1Solve")

# Use the emissions data frame and solve the matrix
masses <- World$Solve(emissions)

print(masses)

```

For more information on the different types of solver and how to use
them, please see SBooScripts/vignettes/x. Solver use.md.
