---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{R load packages}
library(openxlsx)
library(tidyverse)
library(readxl)
```

# Leslie et al. (2017) <!--# What is the source paper of this dataset? This should be referenced as well -->

The data used for validation of the SimpleBox4Plastics model is read in.
The data is filtered for samples taken in the Netherlands, and data measured in \# particles per km2 was removed from the dataset.
This leaves the measurements by Leslie et al. (2017).

Detection method: light microscopy was used to determine the count, size and shape.
Detection limit 10 - 5000 um.

Sampling method: Samples were collected in 1 or 2 liter glass jars.

```{R Leslie et al. data}
SB4P_raw <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/Concentrations_SB4N_validation.xlsx", sheet = "Plastics")

European_countries <- c("France", "Switzerland", "Italy", "England", "Germany", "Ireland", "Slovakia", "Finland", "Netherlands", "Sweden", "Portugal", "Denmark", "Chech Republic")

# Select needed data and clean up the dataframe
SB4P <- SB4P_raw |>
  filter(country %in% European_countries) |>
  mutate(Scale = case_when(
    str_detect(country, "Netherlands") ~ "Regional",
    TRUE ~ "Continental"
  )) |>
  mutate(type = tolower(type)) |>
  filter(type2 == "Surf") |>
  filter(unit == "#/m3") |>
  select(-location) |>
  mutate(min_diameter_um = as.double(map_chr(str_split(size.um, " - "), 1))) |>
  mutate(max_diameter_um = as.double(map_chr(str_split(size.um, " - "), 2))) |>
  mutate(source = paste0(ref, " (", pub.year, ")")) |>
  rename(SubCompart = type,
         value = mean) |>
  select(SubCompart, Scale, min_diameter_um, max_diameter_um, value, unit, source, method, country) |>
  rename(sampling_method = method) |>
  #mutate(detection_method = "Light microscopy") |>
  mutate(detection_method = NA) |>
  mutate(lower_detection_limit_um = 10) |>
  mutate(upper_detection_limit_um = 5000)
```

# KWR data

KWR measured microplastic concentrations in surface water at different locations in the Netherlands.

Sampling method: Two samples were taken per sampling event: one for LDIR and one for OM <!--# What is OM? -->.
Sampling depth 15 cm below the water surface.
For both detection methods, about 500L of water was run through several filters with different sizes (see Table 1 in the article by Bäuerlein et al. (2023)).

Detection method: Two detection methods were used: laser direct infrared and optical microscopy (combined detection limit of 10 to 5000 um).

<!--# Good de describe some basic statistics, e.g. number of observations, number of locations -->

```{R KWR data}
#### KWR data
KWR_data_concentrations <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/KWR/Data KWR/readme.xlsx", sheet = "Meta")

KWR_path <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/KWR/Data KWR/"
KWR_data_dfs <- data.frame()

for(i in 1:nrow(KWR_data_concentrations)){
  # Load the file
  filename <- KWR_data_concentrations$Filename[i]
  filepath <- paste0(KWR_path, filename, ".xlsx")
  file_df <- read.xlsx(filepath)
  
  MP_m3 <- KWR_data_concentrations$MP.per.m3[i]
  
  file_df <- file_df |>
    mutate(MP_m3 = MP_m3) |>
    mutate(location = filename)
  
  if(nrow(KWR_data_dfs) == 0){
    KWR_data_dfs <- file_df
  } else {
    KWR_data_dfs <- bind_rows(KWR_data_dfs, file_df)
  }
}

KWR_location_concentrations_data <- KWR_data_dfs |>
  group_by(location) |>
  summarise(min_diameter_um = min(`Diameter.(µm)`),
            max_diameter_um = max(`Diameter.(µm)`),
            mean_diameter_um = mean(`Diameter.(µm)`),
            median_diameter_um = median(`Diameter.(µm)`),
            value = mean(MP_m3)) |>
  mutate(SubCompart = "river") |>
  mutate(unit = "#/m3") |>
  mutate(institute = "KWR") 
```

<!--# Make a bit clearer what is done in each chunk, Descibe seperately the filtration out of some of the data, e.g. related to effluent. Why is downstream filtered out? -->

```{r}
#### KWR effluent data
KWR_effluent_concentrations <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/KWR/Effluent/readme.xlsx", sheet = "Data")

KWR_path <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/KWR/Effluent/"
KWR_effluent_dfs <- data.frame()

for(i in 1:nrow(KWR_effluent_concentrations)){
  # Load the file
  filename <- KWR_effluent_concentrations$File[i]
  filepath <- paste0(KWR_path, filename)
  file_df <- read.xlsx(filepath)
  
  MP_m3 <- KWR_effluent_concentrations$`MP./m3`[i]
  
  file_df <- file_df |>
    mutate(MP_m3 = MP_m3) |>
    mutate(location = filename)
  
  if(nrow(KWR_effluent_dfs) == 0){
    KWR_effluent_dfs <- file_df
  } else {
    KWR_effluent_dfs <- bind_rows(KWR_effluent_dfs, file_df)
  }
}

KWR_location_concentrations_effluent <- KWR_effluent_dfs |>
  group_by(location) |>
  summarise(min_diameter_um = min(Diameter),
            max_diameter_um = max(Diameter),
            mean_diameter_um = mean(Diameter),
            median_diameter_um = median(Diameter),
            value = mean(MP_m3)) |>
  mutate(SubCompart = "river") |>
  mutate(unit = "#/m3") |>
  filter(str_detect(location, "front|upstream")) |>
  mutate(location = str_remove(location, ".xlsx"))

KWR_data <- bind_rows(KWR_location_concentrations_data, KWR_location_concentrations_effluent) |>
  mutate(source = "Bäuerlein et al. (2023)") |>
  mutate(sampling_method = "500L water run through differently sized filters") |>
  mutate(detection_method = "LDIR + OM") |>
  mutate(lower_detection_limit_um = 10) |>
  mutate(upper_detection_limit_um = 5000) |>
  mutate(institute = "KWR") |>
  mutate(Scale = "Regional") |> 
  mutate(country = "Netherlands")
```

# TNO LEON-T measurements tyre wear

Read in and save data collected by TNO on NR and SBR particles.
(Data in mass, not particles)

```{R}
source("vignettes/CaseStudies/LEON-T/f_plot_functions.R")
abs_path_Measurements <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/LEONT_TNO_measurements_clean.xlsx" # Define path to LEON-T measurement data
LEONT_TWP_data <- prep_LEONT_data(abs_path_Measurements) |>
  select(-RUN) |>
  rename(location = Locatie,
         value = Concentration,
         unit = Unit) |>
  mutate(Scale = "NL") |>
  select(-Polymer) |>
  select(-SubCompartName) |>
  mutate(lower_detection_limit_um = 2,
         upper_detection_limit_um = 1000,
         min_diameter_um = 2,
         max_diameter_um = 1000,
         country = "Netherlands",
         Scale = "Regional") 
```

# Redondo-Hasselerharm et al., 2024

Read in the data on soil microplastic concentrations (Redondo-Hasselerharm et al., 2024)

NOTE: only mean concentrations are currently kept.

```{R}
soil_concentrations <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/Redondo_2024_soil.xlsx", sheet = "Measurements")
references <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/Redondo_2024_soil.xlsx", sheet = "References")

# Change "N.R." (not reported) to NA
soil_concentrations[soil_concentrations == "N.R."] <- NA
soil_concentrations$Mean_concentration <- gsub("[^0-9.]", "", soil_concentrations$Mean_concentration)
soil_concentrations$Min_concentration <- gsub("[^0-9.]", "", soil_concentrations$Min_concentration)
soil_concentrations$Max_concentration <- gsub("[^0-9.]", "", soil_concentrations$Max_concentration)
soil_concentrations$REF <- gsub("[^0-9.]", "", soil_concentrations$REF)

# Select background concentrations only
soil_concentrations <- soil_concentrations |>
  mutate(Mean_concentration = as.integer(Mean_concentration),
         Min_concentration = as.integer(Min_concentration),
         Max_concentration = as.integer(Max_concentration)) |>
  filter(str_detect(MP.source, "Background"))

# Filter for concentrations in Europe 
soil_concentrations <- soil_concentrations |>
  filter(str_detect(`Country,.Continent`, "Europe")) |>
  select(Land.type, `Country,.Continent`, detection_method, lower_detection_limit_um, upper_detection_limit_um, Mean_concentration, unit, REF) |>
  mutate(SubCompart = case_when(
    Land.type == "Farmland" ~ "agriculturalsoil",
    Land.type == "Grassland" ~ "naturalsoil",
    Land.type == "Forest" ~ "naturalsoil"
  )) |>
  rename(value = Mean_concentration,
         country = `Country,.Continent`) |>
  select(-Land.type) |>
  mutate(REF= as.numeric(REF)) |>
  left_join(references, by = "REF") |>
  select(-REF) |>
  mutate(Scale = case_when(
    str_detect(country, "Netherlands") ~ "Regional",
    TRUE ~ "Continental"
  )) |>
  mutate(unit = "#/kg dw") |>
  mutate(country = str_remove(country, ", Europe"))
```

# Redondo-Hasselerharm et al., 2022

Read in the data on freshwater sediment microplastic concentrations (Redondo-Hasselerharm et al., 2022)

NOTE: only mean concentrations are currently kept.

```{R}
sediment_concentrations <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/Redondo_2022_sediment.xlsx", sheet = "Measurements")
references <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/Redondo_2022_sediment.xlsx", sheet = "References")

# Change "N.R." (not reported) to NA
sediment_concentrations[sediment_concentrations == "N.R."] <- NA
sediment_concentrations$Mean_concentration <- gsub("[^0-9.]", "", sediment_concentrations$Mean_concentration)
sediment_concentrations$Min_concentration <- gsub("[^0-9.]", "", sediment_concentrations$Min_concentration)
sediment_concentrations$Max_concentration <- gsub("[^0-9.]", "", sediment_concentrations$Max_concentration)
sediment_concentrations$REF <- gsub("[^0-9.]", "", sediment_concentrations$REF)

# Select background concentrations only
sediment_concentrations <- sediment_concentrations |>
  mutate(Mean_concentration = as.integer(Mean_concentration),
         Min_concentration = as.integer(Min_concentration),
         Max_concentration = as.integer(Max_concentration)) 

# Filter for concentrations in Europe 
sediment_concentrations <- sediment_concentrations |>
  filter(str_detect(`Country,.Continent`, "Europe")) |>
  select(Ecosystem.type, `Country,.Continent`, lower_detection_limit_um, upper_detection_limit_um, Mean_concentration, Unit, REF) |>
  mutate(SubCompart = case_when(
    Ecosystem.type == "River" ~ "freshwatersediment",
    Ecosystem.type == "Lake" ~ "lakesediment"
  )) |>
  rename(value = Mean_concentration,
         country = `Country,.Continent`) |>
  select(-Ecosystem.type) |>
  filter(!is.na(SubCompart)) |>
  filter(value != 0) |>
  mutate(REF= as.numeric(REF)) |>
  left_join(references, by = "REF") |>
  select(-REF) |>
  mutate(Scale = case_when(
    str_detect(country, "Netherlands") ~ "Regional",
    TRUE ~ "Continental"
  )) |>
  mutate(country = str_remove(country, ", Europe")) |>
  filter(raw_data == "N") |>
  select(-raw_data) |>
  mutate(lower_detection_limit_um = as.numeric(lower_detection_limit_um),
         upper_detection_limit_um = as.numeric(upper_detection_limit_um)) |>
  rename(unit = Unit) |>
  filter(country != "Russia")

```

# MoreMomentum

Data from Mintenig et al. (2020)

```{R}
Mintenig <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/MoreMomentum_data_clean.xlsx", sheet = "Mintenig_2020") |>
  rename(value = concentration) |>
  mutate(Scale = "Regional") |>
  select(-polymer)

Mintenig_particles <- Mintenig |>
  filter(unit == "#/m3")

Mintenig_mass <- Mintenig |>
  filter(unit != "#/m3") |>
  mutate(unit = "µg/m^3")

```

Data from Moses et al. (2023)

```{R}
Moses <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/MoreMomentum_data_clean.xlsx", sheet = "Moses_2023") 
Moses[is.na(Moses)] <- 0

Moses <- Moses |>
  mutate(value = PP+PE+PS+PVC+Sil+PET+EVAc+PMMA+PC+ABS+PBT+PU+PA+`PLA-PBAT`)|>
  mutate(Scale = "Continental") |>
  select(value, unit, SubCompart, Scale, country, lower_detection_limit_um, upper_detection_limit_um, source)

```

Data from Liu et al. (2022)

```{R}
Liu <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/MoreMomentum_data_clean.xlsx", sheet = "Liu_2022") |>
  rename(location = areatype) |>
  select(location, Total, unit, SubCompart, country, lower_detection_limit_um, upper_detection_limit_um, source) |>
  rename(value = Total) |>
  mutate(Scale = "Continental")
```

Data from Mani et al. (2020)

```{R}
Mani <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/MoreMomentum_data_clean.xlsx", sheet = "Mani_2020") |>
  rename(value = concentration) |>
  mutate(Scale = "Continental")  |>
  select(value, unit, SubCompart, Scale, country, lower_detection_limit_um, upper_detection_limit_um, location, source)
```

Data from Semmouri et al. (2023)

```{R}
Semmouri <- read.xlsx("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/Measurements/MoreMomentum_data_clean.xlsx", sheet = "Semmouri_2023") |>
  group_by(Waterway, location, Date.of.sampling, unit, SubCompart, country, lower_detection_limit_um, upper_detection_limit_um, source) |>
  summarise(value = sum(concentration)) |>
  ungroup() |>
  mutate(Scale = "Continental")
```

# Save

Save the data.
#/m3 and weight concentrations on different tabs.

```{R Save data}
particle_data <- bind_rows(SB4P, KWR_data, soil_concentrations, Mintenig_particles, Moses, Liu, Mani, Semmouri, sediment_concentrations) |>
  select(value, unit, SubCompart, Scale, location, min_diameter_um, max_diameter_um, mean_diameter_um, median_diameter_um, sampling_method, detection_method, lower_detection_limit_um, upper_detection_limit_um, source, institute, country) |>
  rename(concentration = value,
         concentration_unit = unit)

mass_data <- bind_rows(LEONT_TWP_data, Mintenig_mass)|>
  select(value, unit, SubCompart, Scale, location, min_diameter_um, max_diameter_um, lower_detection_limit_um, upper_detection_limit_um, source, country) |>
  rename(concentration = value,
         concentration_unit = unit)

# Create a new workbook
wb <- createWorkbook()

# Add `particle_data` to the first sheet
addWorksheet(wb, "Particle Data")
writeData(wb, "Particle Data", particle_data)

# Add `mass_data` to the second sheet
addWorksheet(wb, "Mass Data")
writeData(wb, "Mass Data", mass_data)

# Save the workbook
saveWorkbook(wb, "vignettes/CaseData/MOMENTUM2_measurements.xlsx", overwrite = TRUE)
# saveWorkbook(wb, "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/MOMENTUM2/MOMENTUM2_measurements.xlsx", overwrite = TRUE)
```

```{r}

mass_data |> 
  filter(Scale == "Regional") |>
  summarise(length = n())
particle_data |> 
  filter(Scale == "Regional") |>
  summarise(length = n())
particle_data |> 
  group_by(SubCompart,Scale) |> 
  summarise(length = n())


ggplot(mass_data, mapping = aes(concentration)) + 
  geom_histogram()+
  scale_x_log10()+
  facet_wrap( ~ concentration_unit)


ggplot(particle_data, mapping = aes(concentration)) + 
  geom_histogram()+
  scale_x_log10()+
  facet_wrap( ~ concentration_unit)
```
