---
title: "DPMFA LEON-T"
author: "Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

# *Initializing World and Substance*

Depending on the substance you want to analyze, a selection needs to be
made from the three different"worlds" : Molecular, Particulate and
Plastics. Here, we initialize the world for plastics.

```{r Initialize World}
library(lhs)
library(readxl)
library(viridis)
library(trapezoid)

source("baseScripts/initWorld_onlyPlastics.R")

World$substance <- "TRWP"
```

# Fate Factor calculation

- Emission can be to at regional scale (Neglecting Cont. and Global):
Other soil - 1 kg/s
Fresh water - 1 kg/s (fraction surface area corrected to lake/fresh water)
Air - 1 kg/s

FF for (At Regional Scale): 
Other soil
Natural + Aggriculatural soil
Sea water
Fresh water + Lake Water
Sea water
Freshwater sediment (+kw Sediment)
Marine Sediment



```{r Load DPMFA data}

source_of_interest <- c("Tyre wear") # NA for all

Area_w0R <- World$fetchData("Area") |> filter(Scale == "Regional" & SubCompart == "lake") |>  pull(Area)
Area_w1R <- World$fetchData("Area") |> filter(Scale == "Regional" & SubCompart == "river") |>  pull(Area)

FracArea_w0R = Area_w0R/(Area_w0R+Area_w1R)
FracArea_w1R = Area_w1R/(Area_w0R+Area_w1R)

EmisSourceFF <- list()
EmisSourceFF$Air <- tibble(Abbr = c("aRS",
                                    "s3RP",
                                    "w1RP",
                                    "w0RP"),
                           Emis = c(1,0,0,0))
EmisSourceFF$Soil <- tibble(Abbr = c("aRS",
                                     "s3RP",
                                     "w1RP",
                                     "w0RP"),
                            Emis = c(0,1,0,0))
EmisSourceFF$Water <- tibble(Abbr = c("aRS",
                                      "s3RP",
                                      "w1RP",
                                      "w0RP"),
                             Emis = c(0,0,
                                      1*FracArea_w1R,
                                      1*FracArea_w0R))

```

```{r Functions for distributions}
# Define triangular distribution function
triangular <- function(u, a, b, c) {               # u = samples, a = min, b = max, c = peak
  ifelse(u < (c-a)/(b-a),
         a + sqrt(u * (b-a) * (c-a)),
         b - sqrt((1-u) * (b-a) * (b-c)))
}

# Define uniform distribution function
uniform <- function(u, a, b) {                     # u = samples, a = min, b = max
  transformed_samples <- a + (b - a) * u
  return(transformed_samples)
}

# Define power law distribution function
power_law <- function(u, a, b, c){                 # u = samples, a = min, b = max, c = alpha
  
  # Ensure that samples are within [0, 1]
  samples <- pmin(pmax(u, 0), 1)
  
  # Transform samples to the power-law distribution
  scaled_samples <- a * ((b / a) ^ samples) ^ (1 / (1 - c))
  
  return(scaled_samples)
}

trapezoidal <- function(u, a, b, c, d) {
  # Ensure u is in the range [0, 1]
  u <- pmin(pmax(u, 0), 1)  # Clip u to [0, 1]
  
  # Total width of the trapezoid
  width_total <- d - a
  base1 <- b - a    # Width of the left base
  base2 <- d - c    # Width of the right base
  
  # Calculate the CDF segments
  CDF_left <- base1 / width_total         # Area under the left triangle
  CDF_flat <- 1 - base2 / width_total     # Area under the flat top
  
  result <- ifelse(u < (base1 / width_total), 
                   a + sqrt(u * (base1) * width_total),  # Left triangle
                   ifelse(u <= (CDF_flat + base1 / width_total), 
                          b + (u - base1 / width_total) * (d - b),  # Flat top
                          d - sqrt((1 - u) * (base2) * width_total)  # Right triangle
                   )
  )
  
  return(result)
}

```

# Create dataset with uncertain variables

## Prepare excel data

```{r Prepare excel data}
# Path to excel file with distribution values
#path_dist <- "N:/Documents/Simplebox/Microplastic_variables_v2.xlsx"
path_parameters_file <- "vignettes/CaseStudies/CaseData/Microplastic_variables_v1.xlsx"

Material_Parameters <- read_excel(path_parameters_file, sheet = "Polymer_data") |> 
  # change um to nm unit coversion
  mutate(across(c(a, b, c, d), as.numeric)) |>
  mutate(across(c(a, b, c, d), ~ case_when(
    str_detect(Unit, "um") ~ . * 1000,
    TRUE ~ .
  ))) |>
  mutate(Unit = case_when(
    str_detect(Unit, "um") ~ "nm",
    TRUE ~ Unit
  ))

material <-  "RUBBER"



# Only the ones with data in xlsx or with emission data are needed no? Why this vector?
# materials <- c("HDPE", "LDPE", "PP", "PS", "PVC", "Acryl", "PA", "PET", "ABS", "EPS", "PC", "PMMA", "PUR", "RUBBER", "OTHER")
explodeF <- function(df, target_col, explode_value, new_values) {
  df %>%
    # Use mutate to create a new column if the target column equals explode_value
    mutate(!!sym(target_col) := ifelse(!!sym(target_col) == explode_value, list(new_values), !!sym(target_col))) %>%
    # Unnest the target column to duplicate rows
    unnest(!!sym(target_col))
}

DefinedVariables <- lapply(unique(Material_Parameters$VarName),World$fetchData)
names(DefinedVariables) = unique(Material_Parameters$VarName)


# how to cope with any. For now this, but materials should be only for material being calculated for.
suppressWarnings({
  Material_Parameters <- explodeF(Material_Parameters, target_col = "Polymer", explode_value = "any", new_values = material)
})

```

## take input samples
```{r Make sample df}
# Generate the correct number of samples
n_samples <- 10 # Number of emission runs 

Material_Parameters_n <- data.frame()

# EmisSourceFF

input_vars <- 
  Material_Parameters |>
  filter(if_all(c(Distribution, a, b), ~ !is.na(.) )) |> # filter out non defined distributions
  filter(Polymer == material) |> 
  # Not sure what the 3 lines below do.
  group_by(VarName, Scale, SubCompart, Species, Polymer) |>
  filter(if (n() > 1) MP_source == source_of_interest | (MP_source == "" & !any(MP_source == source_of_interest)) else TRUE) |>
  mutate(MP_source = replace_na(MP_source, source_of_interest)) |> 
  ungroup()

n_vars <- nrow(input_vars)

# Generate LHS
lhs_samples <- randomLHS(n_samples, n_vars)
# var_df_names <- c()

# somthing like this should work to get the nested tibble with data of samples
sample_df_var <-  
  input_vars |> 
  mutate(nvar = c(1:n_vars)) |> 
  rowwise() |> 
  mutate(
    data = 
      case_match(Distribution,
                 "Triangular" ~ list(tibble(value=triangular(lhs_samples[, nvar], a, b, c))),
                 "Uniform" ~  list(tibble(value=uniform(lhs_samples[, nvar], a, b))),
                 "Powerlaw" ~  list(tibble( value=power_law(lhs_samples[, nvar], a, b, c))),
                 "Trapezoidal" ~  list(tibble( value=trapezoidal(lhs_samples[, nvar], a, b, c, d))),
                 .default = NA
      )
  ) 

Material_Parameters_n <- rbind(Material_Parameters_n, sample_df_var)


#TODO before explosion export table with some statistics of applied variable ranges for reporting in relation to input variables of each distribution. Aim at: min, p5, p50, mean, p95, max.

scales <- union((World$FromDataAndTo()$fromScale),(World$FromDataAndTo()$toScale))
subCompartments <-  union((World$FromDataAndTo()$fromSubCompart),(World$FromDataAndTo()$toSubCompart))
species <- union((World$FromDataAndTo()$fromSpecies),(World$FromDataAndTo()$toSpecies))

Material_Parameters_n <- 
  Material_Parameters_n %>% 
  separate_rows(Species, sep = "_") |> 
  separate_rows(SubCompart, sep = "_") |>
  mutate( Scale = str_replace_all(Scale, "any",
                                  paste(scales,collapse="__"))) |> 
  separate_rows(Species, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Water",
                                       paste(c("lake", "sea", "deepocean", "river"),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Soil",
                                       paste(subCompartments |> str_subset(c("soil")),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "Sediment",
                                       paste(subCompartments |> str_subset(c("sediment")),
                                             collapse = "__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  mutate( Species = str_replace_all(Species, "any",
                                    paste(species,collapse="__"))) |> 
  separate_rows(Species, sep = "__") |> 
  mutate( SubCompart = str_replace_all(SubCompart, "any",
                                       paste(subCompartments,collapse="__"))) |> 
  separate_rows(SubCompart, sep = "__") |> 
  rename(Source = MP_source)





```
## plots inputs
```{r Make some figures to show the values chosen from the distributions}

#Example of plot:
#TODO: include units

PlotVariable = "alpha"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(PlotVariable) + scale_x_log10() 

PlotVariable = "RadS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) + scale_x_log10() 

PlotVariable = "RhoS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kdeg"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kfrag"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable)) 
```

# Solve the matrix and calculate concentrations

Solve the matrix depending on the type of input emissions given (PMFA =
Steady state, DPMFA = dynamic)

```{r Solve steady state}
start_time <- Sys.time()

Output <- tibble(EmisComp = names(EmisSourceFF),
                 SBoutput=NA) 

World$NewSolver("UncertainSolver")

for(i in 1:length(EmisSourceFF)){
  emis_source <- EmisSourceFF[[i]]
  sample_source <- Material_Parameters_n |>
    filter(Source == "Tyre wear") |>
    filter(Polymer == "RUBBER") |>
    select(VarName, Scale, SubCompart, Species, data) |> 
    rename(varName = VarName)
  
  solved <- World$Solve((emis_source), needdebug = F, sample_source)
  
  # Output$EmisComp[i] = names(EmisSourceFF) [i]
  Output$SBoutput[i] <- list(solved$SteadyStateMass)
}

end_time <- Sys.time()

elapsed_time <- end_time - start_time

print(paste0("Elapsed time is ", elapsed_time/60, "minutes"))

# 10000 runs will take 4 hours and 10 minutes per polymer-source combination.
```

# Fate Factor summary

```{r}

FF_R <-
  Output |> unnest(SBoutput) |> filter(Scale == "Regional" & Species != "Unbound") |> ungroup() |> group_by(EmisComp,RUN,SubCompart,Unit) |> 
  summarise(EqMass_SAP = sum(EqMass)) |> 
  mutate(
    CompartmentFF = case_when(
      SubCompart == "agriculturalsoil" ~ "otherSoil",
      SubCompart == "naturalsoil" ~ "otherSoil",
      SubCompart == "othersoil" ~ "RoadSoil",
      SubCompart == "cloudwater" ~ "air",
      SubCompart == "lake" ~ "freshwater",
      SubCompart == "river" ~ "freshwater",
      SubCompart == "lakesediment" ~ "freshwatersediment",
      TRUE ~ SubCompart)
  ) |> ungroup() |> 
  group_by(CompartmentFF,EmisComp) |> 
  summarise(FF_SS = mean(EqMass_SAP))

write_csv(FF_R, file = "vignettes/CaseStudies/CaseData/FateFactors.csv")

```
