---
title: "DPMFA LEON-T"
author: "Anne Hids, Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

```{r Load packages}
library(tidyverse)
library(ggplot2)
```

# Tyre Wear fate modelling using SimpleBox4Plastics

The model is run for Tyre Wear Particles and for other microplastics sources separately.

## 01: Make emission dataframes for tyre wear and other sources and save them as RData files 

The sourced script reads in the DPMFA output data, and links the emissions to the relevant SimpleBox compartments. The 
emission data in kt/y is converted to kg/s. 

```{r 01 Make emission RData files}

# Make emission RData file for tyre wear
source("vignettes/CaseStudies/LEON-T/01_get_Emission.R")
# check within the script that the DPMFA source data is available via abspath_EU and abspath_NL.

# Overview figure
load(paste0("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/DPMFAoutput_LEON-T_D3.5_TWP_20241126.RData"))

```


## 02: Make parameter dataframes for tyre wear and other sources and save them as RData files
```{r 02 Make parameter RData files}
# Make parameters file for tyre wear
source_of_interest <- "Tyre wear"
source("vignettes/CaseStudies/LEON-T/02_get_Parameters.R")

# Make emission RData file for other sources
source_of_interest <- NA
source("vignettes/CaseStudies/LEON-T/02_get_Parameters.R")
```

### Visualize the variable distributions for tyre wear

```{r Make some figures to show the values chosen from the distributions}

# load Parameters object from file if not already in memory
load(paste0("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/Parameters_LEON-T_D3.5_TWP_20241130.RData"))

Material_Parameters_n <- Parameters$Material_Parameters_n
#Example of plot:
#TODO: include units

PlotVariable = "alpha"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(PlotVariable)+ scale_x_log10() 

PlotVariable = "RadS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) + scale_x_log10() 

PlotVariable = "RhoS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kdeg"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |>
         filter(Species != "Unbound") |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable))  + scale_x_log10() 

PlotVariable = "kfrag"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> 
         filter(Species != "Solid") |>unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable))  + scale_x_log10() 

# load Parameters object from file if not already in memory
load(paste0("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/Parameters_LEON-T_D3.5_Other_20241130.RData"))
Material_Parameters_n <- Parameters$Material_Parameters_n
PlotVariable = "alpha"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable & Polymer == "ABS") |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(PlotVariable)+ scale_x_log10() 
# above alpha's the same for all polymers

PlotVariable = "RadS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer))+
  xlab(paste(PlotVariable)) + scale_x_log10() 

PlotVariable = "RhoS"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer))+
  xlab(paste(PlotVariable)) 

PlotVariable = "kdeg"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |>
         filter(Species != "Unbound") |> unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable))  + scale_x_log10() 
# aboev kdeg same for all polymers

PlotVariable = "kfrag"
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> 
         filter(Species != "Solid") |>unnest(data), aes(x=value,y=SubCompart)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Species))+
  xlab(paste(PlotVariable))  + scale_x_log10() 

```

### Fate Factor calculations
Based on the above parameters and with a generic emission of 1 kg to air, freshwater or roadside soil (othersoil/s3 in SBoo) fate factors are calculation for emission at NL and EU scale for TRWP.
Seperately following the SBR and NR fractions.
```{r}
# run the script:
source("vignettes/CaseStudies/LEON-T/03ff_get_solution_FateFactors.R")
# analysis is done sepeperately using xls.

```


# Make batch files for the High Performance Cluster
Calculating dynamic masses with uncertainty for 1950 to 2050 with 1000 runs takes too
much computing power to calculate in RStudio. Therefore, a high performance cluster (HPC)
was used to calculate the results. The script sourced below divides the 1000 runs for 
tyre wear and other sources between different 'batch scripts', which can then
be sent separately to the HPC. This significantly lowers the computing time, 
as the results are calculated in parallel and a large number of nodes is 
available. 

The following packages should be installed in the environment of your linux terminal:
- tidyverse

## Make batch scripts for tyre wear
The chunk below creates batch scripts for tyre wear, and a txt files with LSF commands. 
Copy the LSF commands into your linux terminal to run the batch scripts. 
```{r Create batch scripts for tyre wear}
# Define batch parameters for tyre wear
batch_n <- 2
batch_max <- 1000  # Should be a multiple of batch_n for the loop logic to work
Source <- NA

source("vignettes/CaseStudies/LEON-T/HPC_batch_files.R")
```

## Make batch scripts for other sources
The chunk below creates batch scripts for other sources, and a txt files with LSF commands. 
Copy the LSF commands into your linux terminal to run the batch scripts. 
```{r Create batch scripts for other sources}
# Define batch parameters for other sources
batch_n <- 2
batch_max <- 1000  # Should be a multiple of batch_n for the loop logic to work
Source <- '"Tyre wear"'

source("vignettes/CaseStudies/LEON-T/HPC_batch_files.R")
```

# SimpleBox4Plastics

## TRWP
Example run of SimpleBox4Plastics and analysis of P species as TRWP, instead of S as TWP.
```{r SimpleBox4Plastics}
load(paste0("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/Parameters_LEON-T_D3.5_TWP_20241130.RData"))
# test Rho and RadCP for Rad and Rho of SAP species.
source("baseScripts/initWorld_onlyPlastics.R")
World$substance <- "TRWP"

# Parameters$Parameter_summary
# Parameters$Material_Parameters_n$data[[1]]
# Parameters$Material_Parameters_n |> distinct(VarName)
# Calculate distribution of RadP and RhoP for air and water
SpeciesP <- 
  Parameters$Material_Parameters_n |> 
  filter((VarName == "RadS" |VarName == "RhoS") & Polymer == "SBR") |> 
  unnest(data, keep_empty = TRUE) |> 
  group_by(VarName) |> 
  mutate(RUN = 1:1000) |> 
  ungroup() |> 
  select(VarName, Source, Polymer, RUN, value) |>
  pivot_wider(names_from = c(VarName),
              values_from = value) |> 
  cross_join(World$fetchData("RhoCP")) |> 
  full_join(World$fetchData("RadCP")) |> 
  rowwise() |>
  mutate(RadS = RadS*1e-9) |> # change if/when data is in nm instead of micrometer.
  mutate(RhoP = 
           ((RhoS*4/3*pi*RadS^3+
               RhoCP*4/3*pi*RadCP^3)/(4/3*pi*(RadS^3+RadCP^3))),
         RadP = (RadCP^3+RadS^3)^(1/3)
  )

SpeciesP |> group_by(SubCompart) |> 
  filter(SubCompart %in% c("air", "river", "othersoil")) |> 
  summarise(RadS_avg = mean(RadS),
            RadP_avg = mean(RadP),
            RadS_p50 = quantile(RadS,probs = 0.5),
            RadP_p50 = quantile(RadP,probs = (0.5)),
            RhoS_avg = mean(RhoS),
            RhoP_avg = mean(RhoP),
            RhoS_p50 = quantile(RhoS,probs = 0.5),
            RhoP_p50 = quantile(RhoP,probs = 0.5),
            n = n())


World$fetchData("Df")
World$fetchData("RadS")
# goal, get graphs for RadP and RhoP
# ggplot(SpeciesP |> filter(SubCompart == "othersoil") ,aes (x=RadP) )+
  # geom_histogram()

SpeciesP_long <- 
  SpeciesP |> 
  pivot_longer(cols = -c(Source,Polymer,RUN,SubCompart),
               names_to = "VarName",
               values_to = "value") |> 
  filter(SubCompart %in% c("air", "river", "othersoil"))

PlotVariables = c("RadS", "RadP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables & SubCompart!="othersoil") ,aes (x=value, y=VarName) )+
  geom_violin() + 
facet_wrap(vars(SubCompart,Source))+xlab( "Radius (um)")+
 scale_x_log10()

PlotVariables = c("RhoS", "RhoP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables) ,aes (x=value, y=VarName) )+
  geom_violin() + 
facet_wrap(vars(SubCompart,Source))+xlab( "density (kg/m3)")
 # scale_x_log10()

# RhoP vs RadP
PlotVariables = c("RadP", "RhoP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables) |> pivot_wider(names_from = c(VarName),
              values_from = value) ,aes (y=RhoP, x=RadP*1e6) )+
  geom_point() + 
facet_wrap(vars(SubCompart,Source))+scale_x_log10()+
  xlab( "Radius TRWP (um)")+
  ylab( "Density TRWP (kg/m3)")

# RhoP vs RadP
PlotVariables = c("RadS", "RadP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables) |> pivot_wider(names_from = c(VarName),
              values_from = value) ,aes (x=RadP, y=RadS) )+
  geom_point() + 
facet_wrap(vars(SubCompart,Source))+scale_y_log10()

PlotVariables = c("RhoS", "RhoP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables) |> pivot_wider(names_from = c(VarName),
              values_from = value) ,aes (x=RhoP, y=RhoS) )+
  geom_point() + 
facet_wrap(vars(SubCompart,Source))+scale_y_log10()

PlotVariables = c("RadP")
ggplot(SpeciesP_long |> filter(VarName %in% PlotVariables) ,aes (x=value*1e6, y=VarName) )+
  geom_violin() + 
facet_wrap(vars(SubCompart))+xlab( "Radius (um)")+
 scale_x_log10()

PlotVariable = "RadP"
ggplot(SpeciesP_long |> filter(SubCompart == "othersoil") ,aes (x=value,y = SubCompart) )+
  geom_violin() +
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) + scale_x_log10() 




PlotVariables = "RadS"
ggplot(Parameters$Material_Parameters_n |> filter(VarName %in% PlotVariables) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer))+
  xlab(paste(PlotVariable)) + scale_x_log10() 

PlotVariable = "RhoP" 
ggplot(Material_Parameters_n |> filter(VarName == PlotVariable) |> unnest(data), aes(x=value,y=VarName)) +
  geom_violin()+
  facet_wrap(vars(Polymer,Source))+
  xlab(paste(PlotVariable)) 

```

## other microplastics
```{r run batchfiles}

BatchFiles <- list.files("vignettes/CaseStudies/LEON-T/BatchFiles")

for(Bfile in BatchFiles){
  source(paste0("vignettes/CaseStudies/LEON-T/BatchFiles/",Bfile))
}

```

...

# Bind the data
Once the batch scripts have been run, the data can be bound together. This is done
using the a script that binds data in parallel, as this saves time. 

Needed packages: 
- tidyverse
- stringr
- doParallel

```{r 04: Bind data in parallel}
source("vignettes/CaseStudies/LEON-T/04_bind_data_parallel.R")

```

# Make figures from SimpleBox output
The script below was used to create the SimpleBox output figures as in LEON-T
Deliverable 3.5. Before running, change the paths to the RDatafiles created in 
script 04 to where the files are located on your computer. 

Needed packages: 
- tidyverse
- readxl
- scales

```{r 05: Make SimpleBox output figures}
source("vignettes/CaseStudies/LEON-T/05_SimpleBox_figures_D3.5.R")

```

# Prepare the DPMFA emission output for plotting
Before being able to make figures from the DPMFA emission output data, the data 
is converted to smaller datasets that RStudio is better able to handle. 

```{r 06: Prepare emission data}
source("vignettes/CaseStudies/LEON-T/06_save_emission_figure_data.R")

```

# Make emission output figures
This script was used to make the DPMFA output figures. Please change the filepaths 
at the beginning of the script to the 

```{r 08: Make DPMFA output figures}
source("vignettes/CaseStudies/LEON-T/08_emission_figures_D3.5.R")

```

