---
title: "LEON-T_emission_figures"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r load data}
library(ggplot2)
library(tidyverse)
library(scales)
library(openxlsx)
library(networkD3)
library(webshot)
library(jsonlite)
library(viridis)

Figure_folder <- "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/Figures/Emission_plots/"
MFAtype <- "DPMFA"
```

# Prepare data for plotting
```{r}
# Load emission data 
load(file = "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/DPMFA_NL_EU_long.RData")
load(file = "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/DPMFA_NL_long.RData")
load(file = "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/DPMFA_EU_long.RData")
load(file = "/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/DPMFA_emission_NL_EU_TW_time.RData")

Years_available <- unique(data_long$Year)

############################### Prepare NL + EU data ###########################
data_Year <- data_long |>
  mutate(Mass_Polymer_t = Mass_Polymer_kt * 1000, .keep = "unused") |> # convert kt to ton
  mutate(`Environmental_compartment` = # make column indicating of this an air, water or soil sink.
           case_when(
             str_detect(To_Compartment, 'water') ~ "water",
             str_detect(To_Compartment, 'air') ~ "air",
             str_detect(To_Compartment, 'soil') ~ "soil",
             .default = "none"
           )) |> filter(`Environmental_compartment` != "none") |> # remove sinks that are not environmental
  mutate(IenW_Source = # aggregate inputs to the Major sources classes the study focusses on
           case_when(
             str_detect(Source, 'Import of primary plastics') ~ "Pre-production pellets",
             str_detect(Source, 'Domestic primary plastic production') ~ "Pre-production pellets",
             str_detect(Source, 'Clothing') ~ "Clothing",
             str_detect(Source, 'Household textiles') ~ "Household textiles",
             .default = Source
           )) |>
  filter(!is.na(Mass_Polymer_t)) |>
  mutate(Scale = "EU_with_NL")
  
data_Year_sep_macro <- 
  data_Year |> filter(Material_Type == "micro") |> full_join(
    data_Year |> 
      filter(Material_Type == "macro") |> 
      ungroup() |> 
      group_by(Type, Scale, Polymer,To_Compartment, Material_Type, RUN, Year, Environmental_compartment) |> 
      summarise(Mass_Polymer_t = sum(Mass_Polymer_t),
                IenW_Source = "Macroplastic",
                Source = "Macroplastic")  )

data_micro_summed_over_polymers <- data_Year |>
  filter(Material_Type == "micro") |>
  group_by(Type, Scale, Source, To_Compartment, Material_Type, RUN, Year, Environmental_compartment, IenW_Source) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

############################### Prepare NL data ################################
# Create a dataframe for just selectyear
data_Year_NL <- data_long_NL |>
  mutate(Mass_Polymer_t = Mass_Polymer_kt * 1000, .keep = "unused") |> # convert kt to ton
  mutate(`Environmental_compartment` = # make column indicating of this an air, water or soil sink.
           case_when(
             str_detect(To_Compartment, 'water') ~ "water",
             str_detect(To_Compartment, 'air') ~ "air",
             str_detect(To_Compartment, 'soil') ~ "soil",
             .default = "none"
           )) |> filter(`Environmental_compartment` != "none") |> # remove sinks that are not environmental
  mutate(IenW_Source = # aggregate inputs to the Major sources classes the study focusses on
           case_when(
             str_detect(Source, 'Import of primary plastics') ~ "Pre-production pellets",
             str_detect(Source, 'Domestic primary plastic production') ~ "Pre-production pellets",
             str_detect(Source, 'Clothing') ~ "Clothing",
             str_detect(Source, 'Household textiles') ~ "Household textiles",
             .default = Source
           )) |>
  filter(!is.na(Mass_Polymer_t)) |>
  mutate(Scale = "NL")
  
data_Year_sep_macro_NL <- 
  data_Year_NL |> filter(Material_Type == "micro") |> full_join(
    data_Year |> 
      filter(Material_Type == "macro") |> 
      ungroup() |> 
      group_by(Type, Polymer, To_Compartment, Scale, Material_Type, RUN, Year,Environmental_compartment) |> 
      summarise(Mass_Polymer_t = sum(Mass_Polymer_t),
                IenW_Source = "Macroplastic",
                Source = "Macroplastic"))

data_micro_summed_over_polymers_NL <- data_Year_NL |>
  filter(Material_Type == "micro") |>
  group_by(Type, Source, To_Compartment, Scale, Material_Type, RUN, Year, Environmental_compartment, IenW_Source) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

############################### Prepare EU data ################################
# Create a dataframe for just selectyear
data_Year_EU <- data_long_EU |>
  mutate(Mass_Polymer_t = Mass_Polymer_kt * 1000, .keep = "unused") |> # convert kt to ton
  mutate(`Environmental_compartment` = # make column indicating of this an air, water or soil sink.
           case_when(
             str_detect(To_Compartment, 'water') ~ "water",
             str_detect(To_Compartment, 'air') ~ "air",
             str_detect(To_Compartment, 'soil') ~ "soil",
             .default = "none"
           )) |> filter(`Environmental_compartment` != "none") |> # remove sinks that are not environmental
  mutate(IenW_Source = # aggregate inputs to the Major sources classes the study focusses on
           case_when(
             str_detect(Source, 'Import of primary plastics') ~ "Pre-production pellets",
             str_detect(Source, 'Domestic primary plastic production') ~ "Pre-production pellets",
             str_detect(Source, 'Clothing') ~ "Clothing",
             str_detect(Source, 'Household textiles') ~ "Household textiles",
             .default = Source
           )) |>
  filter(!is.na(Mass_Polymer_t)) |>
  mutate(Scale = "EU")
  
data_Year_sep_macro_EU <- 
  data_Year_NL |> filter(Material_Type == "micro") |> full_join(
    data_Year |> 
      filter(Material_Type == "macro") |> 
      ungroup() |> 
      group_by(Type, Polymer, To_Compartment, Scale, Material_Type, RUN, Year,Environmental_compartment) |> 
      summarise(Mass_Polymer_t = sum(Mass_Polymer_t),
                IenW_Source = "Macroplastic",
                Source = "Macroplastic"))

data_micro_summed_over_polymers_EU <- data_Year_EU |>
  filter(Material_Type == "micro") |>
  group_by(Type, Source, To_Compartment, Material_Type, Scale, RUN, Year, Environmental_compartment, IenW_Source) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

```

```{r}
# Create a theme for the plots
plot_theme = theme(
  axis.title.x = element_text(size = 32),
  legend.text = element_text(size = 26),
  axis.text = element_text(size = 26), 
  axis.title.y = element_text(size = 32),
  title = element_text(size = 32),
  plot.background = element_rect(fill = 'white'),
  panel.background = element_rect(fill = 'white'),
  axis.line = element_line(color='black'),
  plot.margin = margin(2, 4, 2, 2, "cm"),
  legend.title = element_blank(),  
  panel.grid.major = element_line(color = "lightgrey", size = 0.5),  # Major grid lines in light grey
  panel.grid.minor = element_line(color = "lightgrey", size = 0.25)
)

# Define colors for groups
source_colors <- c(
  "Agriculture" = "#F8766D",
  "Clothing" = "#E58700",
  "Household textiles" = "#E76BF3",
  "Intentionally produced microparticles" = "#C99800",
  "Macroplastic" = "#A3A500",
  "Packaging" = "#619cff",
  "Paint" = "#00BA38",
  "Pre-production pellets" = "#00C0AF",
  "Technical textiles" = "#00B0F6",
  "Textile" = "#B983FF",
  "Tyre wear" = "#ff67a4"
)

TW_other_colors <- c(
  "Tyre wear" = "#ff67a4",
  "Other sources" = "#619cff"
)

micro_macro_colors <- c(
  "Microplastic" = "#A3A500",
  "Macroplastic" = "#B983FF"
)

sink_colors <- c(
  "Sea water" = "dodgerblue4",
  "Air" = "slategray3",
  "Other soil" = "darkgoldenrod",
  "Fresh water" = "cyan3",
  "Roadside soil" = "azure4"
)

to_comp_colors <- c(
  "Agricultural soil" = "darkolivegreen3",
  "Outdoor air" = "deepskyblue",
  "Residential soil" = "chartreuse4",
  "Road side soil" = "azure4",
  "Surface water" = "cyan3",
  "Sub-surface soil" = "chocolate4"
)
```

Figure comparing all micro emissions per source for NL + EU
```{r}
source_comparison_micro_data <- data_Year_sep_macro

micro_source_plot <- ggplot(source_comparison_micro_data, aes(x = IenW_Source, y = Mass_Polymer_t, fill=IenW_Source)) +
  geom_violin() +
  scale_fill_manual(values = source_colors) +
  theme(legend.position="none")+
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "", y = "Total plastic emissions (ton)") +                   # Adjust labels
  coord_flip() +
  plot_theme 

micro_source_plot

ggsave(paste0(Figure_folder, "micro_source_EU_with_NL_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Figure comparing micro with macro emissions for NL + EU
```{r Micro macro}
micro_macro <- data_Year_sep_macro |>
  group_by(Material_Type, RUN)|>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t)) |>
  mutate(Material_Type =case_when(
    Material_Type == "micro" ~ "Microplastic",
    Material_Type == "macro" ~ "Macroplastic"
  ))

micro_macro_plot <- ggplot(micro_macro, aes(x = Material_Type, y = Mass_Polymer_t, fill=Material_Type)) +
  geom_violin() +
  scale_fill_manual(values = micro_macro_colors) +
  theme(legend.position="none")+
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "", y = "Total plastic emissions (ton)") +                   # Adjust labels
  coord_flip() +
  plot_theme 

micro_macro_plot

ggsave(paste0(Figure_folder, "Micro_macro_EU_with_NL_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Figure comparing microplastic emissions tyre wear with other microplastic emissions for NL + EU
```{r}
Data_TW_separated <- data_micro_summed_over_polymers |>
  mutate(TW_class = ifelse(Source == "Tyre wear", "Tyre wear", "Other sources")) |>
  group_by(TW_class, RUN, Scale) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

TW_Other_micro_plot <- ggplot(Data_TW_separated, aes(x = TW_class, y = Mass_Polymer_t, fill=TW_class)) +
  geom_violin() +
  scale_fill_manual(values = TW_other_colors) +
  theme(legend.position="none")+
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "", y = "Microplastic emissions (ton)") +                   # Adjust labels
  coord_flip() +
  plot_theme 

TW_Other_micro_plot

ggsave(paste0(Figure_folder, "TW_other_EU_with_NL_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Figure comparing microplastic emissions tyre wear with other microplastic emissions for NL
```{r}
Data_TW_separated_NL <- data_micro_summed_over_polymers_NL |>
  mutate(TW_class = ifelse(Source == "Tyre wear", "Tyre wear", "Other sources")) |>
  group_by(TW_class, RUN, Scale) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

TW_Other_micro_plot_NL <- ggplot(Data_TW_separated_NL, aes(x = TW_class, y = Mass_Polymer_t, fill=TW_class)) +
  geom_violin() +
  scale_fill_discrete()+
  theme(legend.position="none")+
  scale_fill_manual(values = TW_other_colors) +
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "",
       y = "Microplastic emissies (ton)",
       title = "Microplastic emissies naar het milieu in 2019") +                   # Adjust labels
  coord_flip() +
  plot_theme 

TW_Other_micro_plot_NL

ggsave(paste0(Figure_folder, "TW_other_NL_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Figure comparing microplastic emissions tyre wear with other microplastic emissions for EU
```{r}
Data_TW_separated_EU <- data_micro_summed_over_polymers_EU |>
  mutate(TW_class = ifelse(Source == "Tyre wear", "Tyre wear", "Other sources")) |>
  group_by(TW_class, RUN, Scale) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t))

TW_Other_micro_plot_EU <- ggplot(Data_TW_separated_EU, aes(x = TW_class, y = Mass_Polymer_t, fill=TW_class)) +
  geom_violin() +
  scale_fill_discrete()+
  theme(legend.position="none")+
  scale_fill_manual(values = TW_other_colors) +
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "",
       y = "Microplastic emissies (ton)",
       title = "Microplastic emissies naar het milieu in 2019") +                   # Adjust labels
  coord_flip() +
  plot_theme 

TW_Other_micro_plot_EU

ggsave(paste0(Figure_folder, "TW_other_EU_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Plot comparing EU, NL and NL EU emissions
```{r}
TW_other <- bind_rows(Data_TW_separated, Data_TW_separated_EU, Data_TW_separated_NL) |>
    mutate(
    Scale = str_replace_all(Scale, "EU_with_NL", "EU + NL"),
    Scale = factor(Scale, levels = c("EU + NL", "EU", "NL")) # Reorder levels
  ) |>
  mutate(as.factor(Scale)) 

TW_other_scale_plot <- ggplot(TW_other, aes(x = Scale, y = Mass_Polymer_t, fill=TW_class)) +
    geom_violin() +
  theme(legend.position="bottom")+
  scale_fill_manual(values = TW_other_colors) +
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "",
       y = "Microplastic emissions (ton)")+
  coord_flip() +
  plot_theme 

TW_other_scale_plot

ggsave(paste0(Figure_folder, "TW_other_EU_NL_comparison_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

TW to environmental compartments
```{r}
TW_data <- data_micro_summed_over_polymers |>
  mutate(TW_class = ifelse(Source == "Tyre wear", "Tyre wear", "Other sources")) |>
  filter(Source == "Tyre wear") |>
  mutate(To_Compartment = str_remove(To_Compartment, fixed(" (micro)")))

TW_env_comp_plot <- ggplot(TW_data, aes(x = To_Compartment, y = Mass_Polymer_t, fill=To_Compartment)) +
  geom_violin() +
  theme(legend.position="none")+
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "", y = "Microplastic emissions (ton)") +                   # Adjust labels
  coord_flip() +
  plot_theme 

TW_env_comp_plot

ggsave(paste0(Figure_folder, "TW_to_env_comps_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Stacked barplot distribution over different sinks TWP vs. other sources
```{r}
barplot_data <- data_micro_summed_over_polymers |>
  mutate(TW_class = ifelse(Source == "Tyre wear", "Tyre wear", "Other sources")) |>
  ungroup() |> 
  mutate(Environmental_compartment = case_when(
    Environmental_compartment == "soil" ~ "Other soil",
    Environmental_compartment == "air" ~ "Air",
    .default = Environmental_compartment
  )) |>
  mutate(Environmental_compartment = case_when(
    To_Compartment == "Road side soil (micro)" ~ "Roadside soil",
    To_Compartment == "Surface water (micro)" ~ "Fresh water",
    To_Compartment == "Sea water (micro)" ~ "Sea water",
    .default = Environmental_compartment
  )) |>
  group_by(Scale, TW_class, To_Compartment, Environmental_compartment, Year, Type, Material_Type) |> 
  summarise(Mass_Polymer_t = mean(Mass_Polymer_t),
            n = n()) |> 
  ungroup() |> 
  group_by(Scale, TW_class, Environmental_compartment, Year, Type) |> 
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t),
            n= n())

# Make a stacked barplot for what percentage goes to air, soil and water
sink_dist_barplot <- ggplot(barplot_data, aes(fill = Environmental_compartment, x = TW_class, y = Mass_Polymer_t)) +
  geom_bar(position = "fill", stat="identity", color = "transparent") +
  scale_fill_manual(values = sink_colors) +
  scale_x_discrete(labels = wrap_format(10)) +                   # Wraps text longer than 10 characters
  scale_y_continuous(labels = scales::percent) +
  plot_theme +
  labs(y= "Distribution",
       x="") +
  labs(fill='Sink type') +
   theme( # Override specific elements
    panel.grid.major = element_blank(), # Major grid lines removed
    panel.grid.minor = element_blank()  # Minor grid lines removed
  ) 

sink_dist_barplot

ggsave(paste0(Figure_folder, "Env_comp_barplot_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Figure violin plot from each source to soil types

```{r}
soil_data <- data_Year |>
  filter(Environmental_compartment == "soil") |>
  filter(Material_Type == "micro")

source_soil_plot <- ggplot(soil_data, aes(x = IenW_Source, y = Mass_Polymer_t, fill = To_Compartment)) +
  geom_violin() +
  scale_fill_discrete()+
  theme(legend.position="right")+
  scale_y_log10(labels = scales::number_format())+          
  labs(x = "", y = "Microplastic emissions (ton)") +                   # Adjust labels
  coord_flip() +
  plot_theme 

source_soil_plot

ggsave(paste0(Figure_folder, "Sources_to_soil_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Plot TWP over time
```{r}
excluded_compartments <- c("Landfill", "Elimination", "Secondary material reuse")

TW_mean_data <- TW_emissions_over_time |>
  mutate(Mass_Polymer_t = Mass_Polymer_kt*1000) |>
  group_by(Source, To_Compartment, Year, Scale, Polymer) |>
  summarise(Mean = mean(Mass_Polymer_t)) |>
  mutate(Year = as.integer(Year)) |>
  filter(!To_Compartment %in% excluded_compartments) |>
  mutate(To_Compartment = str_replace_all(To_Compartment, " \\(micro\\)", ""))

TW_over_time <- ggplot(TW_mean_data, aes(x = Year, y = Mean, colour = To_Compartment, group = To_Compartment)) +
  geom_line(size=1) +
  plot_theme +          
  labs(x = "Year", y = "Mean microplastic emission (ton)", title = "Tyre wear emissions to environmental compartments") +
  scale_color_manual(values = to_comp_colors)

TW_over_time

ggsave(paste0(Figure_folder, "Tyre_wear_concentration_over_time_", MFAtype, "_", format(Sys.time(),'%Y%m%d'),".png"),
       width = 20, height = 10)
```

Tyre wear sankey for NL + EU
```{r}
palette_name <-"RdYlBu"

# Calculate mean flows for TW
TW_sankey_data <- TW_emissions_over_time |>
  filter(Year == 2019) |>
  group_by(Source, Polymer, To_Compartment) |>
  summarise(Mass_Polymer_kt = mean(Mass_Polymer_kt)) |>
  mutate(From_Compartment = "Tyre wear") |>
  mutate(Mass_Polymer_t = Mass_Polymer_kt*1000)

links <- TW_sankey_data |>
  group_by(Source, To_Compartment) |>
  summarise(Mass_Polymer_t = sum(Mass_Polymer_t)) |>
  filter(Mass_Polymer_t !=0)|>
  rename(source = Source, target = To_Compartment, value = Mass_Polymer_t)

# Prepare node labels
source_label_names <- links |>
  group_by(source) |>
  summarise(value = sum(value) / 1000, 2) |>
  mutate(value = format(value, scientific = TRUE, digits = 3))|>
  mutate(label = paste0(source, ': ', value, " kt"))

target_label_names <- links |>
  group_by(target) |>
  summarise(value = sum(value) / 1000, 2) |>
  mutate(value = format(value, scientific = TRUE, digits = 3))|>
  mutate(label = paste0(target, ': ', value, " kt"))

all_names <- c(source_label_names$source, target_label_names$target)
all_labels <- c(source_label_names$label, target_label_names$label)

# Make dataframe with node names
nodes <- data.frame(name = all_names, label = all_labels)

# Add columns to links (needed for sankey function)
links$source_id <- match(links$source, nodes$name) - 1
links$target_id <- match(links$target, nodes$name) - 1

# Create a color scheme
viridis_colors <- viridis(length(unique(links$target)), option = "D")

color_scale <- paste0(
  'd3.scaleOrdinal()',
  '.domain(["', paste(unique(links$target), collapse = '", "'), '"])',
  '.range(["', paste(viridis_colors, collapse = '", "'), '"])')

# Make the sankey diagram
sankey <- sankeyNetwork(Links = links, 
                        Nodes = nodes, 
                        Source = 'source_id',
                        Target = 'target_id',
                        Value = 'value', 
                        NodeID = 'label',
                        fontSize = 25, 
                        #colourScale=color_scale,
                        LinkGroup='target')

sankey 

filename<- "Sankey_tyre_wear_to_sinks_"

# Save the sankey diagram as an HTML file
html_file <- paste0(Figure_folder, filename, ".html")
saveNetwork(sankey, file = html_file, selfcontained = TRUE)
```
