---
title: "Testing Mass Balance"
author: "Valerie de Rijk"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) 
```

#Mass balance of particulates 
In this vignette, we assess the mass balance of the SB model. Due to the nature of the solver, small deviations from 0 are expected. However, we like to investigate the steady state computation for different particles to assess the stability of the solver.

## Initiation
First, we initiate the World with some dummy emissions to a chosen compartment. 


```{r solver}
library(dplyr)
library(readr)
library(tidyr)
library(stringr)


emissions <- data.frame(Abbr = "w1CS", 
                        Emis = 35)
emissions$Emis<- emissions$Emis * 1000 / (365.25 * 24 * 60 * 60)
print(emissions$Emis)# convert 1 t/y to si units: kg/s
```

## Solvers and substances 
Here, we solve for three different substances and two different solvers. We solve the equilibrium masses to later compute the mass balance. 

```{r list}
# Define the list of substances
substances <- c("nC60_10nm", "nAg_10nm", "nZnO")  # Add all the substances you want to process "1-aminoanthraquinone", # no class
# substances <- c(
#                           "1-HYDROXYANTHRAQUINONE", # acid
#                           "1-Hexadecanamine, N,N-dimethyl-", # base
#                           "1-Chloro-2-nitro-propane", 
#                            "Zn(II)", 
#                           "Be(II)", 
#                           "FUMARIC ACID", 
#                           "Closantel"
# ) # neutral
# Define the list of solvers
solvers <- c("SBsteady", "SB1Solve")

# Initialize an empty list to store results
results_list <- list()

# Loop through each substance and each solver
for (solver in solvers) {
  for (substance in substances) {
    # Print the current substance and solver (optional)
    cat("Processing substance:", substance, "with solver:", solver, "\n")
    
    # Initialize the world with the current substance
    source("baseScripts/initWorld_onlyParticulate.R")
    World$NewSolver(solver)
  
    # Solve for the steady solution
    SolutionSteady <- World$Solve(emissions)
    
    # Add a column for the substance and solver
    SolutionSteady$Substance <- substance
    SolutionSteady$Solver <- solver
    
    # Append the results to the list
    results_list[[paste(substance, solver, sep = "_")]] <- SolutionSteady
  }
}

# Combine all results into a single dataframe
SolutionSteady_all <- do.call(rbind, results_list)





```

## Compartment of interest
Here we define our compartment of interest to compute the mass balance for. 
```{r defining compartment of interest}
ScaleInterest <- "Continental"
SubCompartInterest <- "river"
SpeciesInterest <- "Solid"
```

## Mass balance check 
In this section we compute the mass balances for our compartment of interest and visualize it. We print the top rows and differences. 

```{r mass balance check}
# Initialize lists and data frames
mass_balance_results <- list()
total_abs_differences <- data.frame(Substance = character(), Solver = character(), TotalAbsDifference = numeric(), stringsAsFactors = FALSE)
top_rows_list <- list()
total_mass_difference_list <- list()
kaas_filtered_list <- list()
results_list <- list()
mass_balance_results <- list()
total_abs_differences <- data.frame(Substance = character(), Solver = character(), TotalAbsDifference = numeric(), stringsAsFactors = FALSE)
top_rows_list <- list()
total_mass_difference_list <- list()
kaas_filtered_list <- list()

# Loop over solvers and substances
for (solver in solvers) {
  for (substance in substances) {
    # Print the current substance and solver (optional)
    cat("Processing substance:", substance, "with solver:", solver, "\n")
    
    # Initialize the world with the current substance
    source("baseScripts/initWorld_onlyParticulate.R")
    World$NewSolver(solver)
  
    # Solve for the steady solution
    SolutionSteady <- World$Solve(emissions)
    
    # Add a column for the substance and solver
    SolutionSteady$Substance <- substance
    SolutionSteady$Solver <- solver
    
    # Append the results to the list
    results_list[[paste(substance, solver, sep = "_")]] <- SolutionSteady
    
    # Remove the SolutionSteady object to free up memory and ensure it's not reused
    rm(SolutionSteady)
    
    # Mass balance processing
    cat("Processing mass balance for substance:", substance, "\n")
    key <- paste(substance, solver, sep = "_")
    SolutionSteady <- results_list[[key]]
    
    # Create unique combinations
    unique_combinations <- SolutionSteady |>
      select(Scale, SubCompart, Species) |>
      distinct()
    
    # Create emission rows
    emission_rows <- unique_combinations |>
      mutate(
        Value = "emission",
        EqMass = if_else(Scale == "Continental" & SubCompart == "river" & Species == "Solid", emissions$Emis, 0)
      )
    
    # Adding "from" and "to" acronyms to the R K matrix
    kaas <- as_tibble(World$kaas)
    kaas <- kaas |>
      left_join(SolutionSteady, by = c("fromScale" = "Scale", "fromSubCompart" = "SubCompart", "fromSpecies" = "Species"))
    
    # Multiply k by the corresponding EqMass
    kaas <- kaas |>
      mutate(k = k * EqMass)
    kaas <- kaas |>
      mutate(k = if_else(fromSubCompart == toSubCompart &
                           fromScale == toScale &
                           fromSpecies == toSpecies,
                         -abs(k),
                         k))
    
    # Save kaas rows for certain subcompartments 
    kaas_filtered <- kaas |>
      filter((fromScale == ScaleInterest & fromSubCompart == SubCompartInterest & fromSpecies == SpeciesInterest) |
             (toScale ==  ScaleInterest & toSubCompart == SubCompartInterest & toSpecies == SpeciesInterest))
    
    # Aggregate k values based on toScale, toSubCompart, and toSpecies
    aggregated_to <- kaas |>
      group_by(toScale, toSubCompart, toSpecies) |>
      summarise(k_sum_to = sum(k)) |>
      ungroup()
    
    # Aggregate k values based on fromScale, fromSubCompart, and fromSpecies
    aggregated_from <- kaas |>
      filter(k >= 0) |>
      group_by(fromScale, fromSubCompart, fromSpecies) |>
      summarise(k_sum_from = sum(k)) |>
      ungroup()
    
    new_rows_from <- aggregated_from |>
      mutate(
        Value = "transportfrom",
        Scale = fromScale,
        SubCompart = fromSubCompart,
        Species = fromSpecies,
        EqMass = k_sum_from
      ) |>
      select(Value, Scale, SubCompart, Species, EqMass)
    
    new_rows_to <- aggregated_to |>
      mutate(
        Value = "transportto",
        Scale = toScale,
        SubCompart = toSubCompart,
        Species = toSpecies,
        EqMass = k_sum_to
      ) |>
      select(Value, Scale, SubCompart, Species, EqMass)
    
    # Combine SolutionSteady with new rows
    massbalancecheck <- bind_rows(new_rows_from, new_rows_to, emission_rows)
    massbalancecheck_sorted <- massbalancecheck |>
      arrange(Scale, SubCompart, Species)
    
    # Check differences
    massbalance_difference <- massbalancecheck_sorted |>
      pivot_wider(names_from = Value, values_from = EqMass) |>
      mutate(difference = transportto + emission - transportfrom) |>
      select(Scale, SubCompart, Species, difference)
    
    total_abs_difference <- sum(abs(massbalance_difference$difference), na.rm = TRUE)
    total_mass_difference <- sum(massbalance_difference$difference, na.rm = TRUE)
    
    # Add columns for substance and solver
    massbalance_difference$Substance <- substance
    massbalance_difference$Solver <- solver
    kaas_filtered$Substance <- substance
    kaas_filtered$Solver <- solver
    
    # Print largest differences
    top_5_rows <- massbalance_difference |>
      arrange(desc(abs(difference))) |>
      head(5)
    
    # Append the results to the lists
    mass_balance_results[[key]] <- massbalance_difference
    total_abs_differences <- rbind(total_abs_differences, data.frame(Substance = substance, Solver = solver, TotalAbsDifference = total_abs_difference))
    top_rows_list[[key]] <- top_5_rows
    total_mass_difference_list[[key]] <- total_mass_difference
    kaas_filtered_list[[key]] <- kaas_filtered
    rm(World)
  }
}

# Combine all mass balance results into a single dataframe
massbalance_difference_all <- do.call(rbind, mass_balance_results)
top_rows_all <- do.call(rbind, top_rows_list)
kaas_filtered_all <- do.call(rbind, kaas_filtered_list)

# Print the combined results (optional)
print(massbalance_difference_all)
print(top_rows_all)
print(kaas_filtered_all)


library(ggplot2)
ggplot(total_abs_differences, aes(x = Substance, y = TotalAbsDifference, fill = Solver)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Absolute Differences by Substance and Solver", x = "Substance", y = "Total Absolute Difference [kg]") +
  theme_minimal()

# Plotting the mass balance differences
ggplot(massbalance_difference_all, aes(x = interaction(Scale, SubCompart, Species), y = difference, fill = Substance)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Solver, scales = "free", ncol = 2) +
  labs(title = "Mass Balance Differences by Scale, SubCompart, and Species", x = "Combination (Scale, SubCompart, Species)", y = "Difference Mass Balance [kg]") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal()

top_rows_all$Substance <- factor(top_rows_all$Substance, levels = unique(top_rows_all$Substance))
top_rows_all$Solver <- factor(top_rows_all$Solver, levels = unique(top_rows_all$Solver))


# Plotting the top differences as grouped bars with facets for solvers
ggplot(top_rows_all, aes(x = interaction(SubCompart, Scale, Species), y = difference, fill = Substance)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.4) +
  labs(title = "Top Differences by Compartment, Scale, and Species", x = "Combination (Compartment, Scale, Species)", y = "Difference Mass Balance [kg]", fill = "Substance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ Solver, scales = "free", ncol = 2)
```

Here, we zoom in to what is happening at the individual compartments. 
```{r all plots}
library(dplyr)
library(ggplot2)
# Funcdata()# Function to process and plot data for a given substance

process_substance <- function(data, substance_name) {
  filtered_data <- data |>
    filter(Solver == "SBsteady", Substance == substance_name)
  
  
  filtered_data <- filtered_data |>
  mutate(
    condition_met = fromSubCompart == "agriculturalsoil" & fromScale == "Continental" ,
    k = if_else(condition_met, -k, k)
  )
  print(filtered_data)
  
  # Make the ones with the specified from columns negative
  # Aggregate k values for k_Advection
  k_Advection_sum <- sum(filtered_data$k[filtered_data$process == "k_Advection"])
  filtered_data <- filtered_data |>
    group_by(process) |>
    mutate(k = if_else(process == "k_Advection", k_Advection_sum, k)) |>
    filter(!duplicated(process))  # Remove duplicates after aggregation
  
  # Ensure k_Degradation remains negative
  filtered_data <- filtered_data |>
    mutate(k = if_else(process == "k_Degradation", -abs(k), k))
  
  # Calculate the sum of k_Degradation
  k_Degradation_sum <- sum(filtered_data$k[filtered_data$process == "k_Degradation"])
  print(paste("k_Degradation_sum for", substance_name, ":", k_Degradation_sum))
  
  # Calculate the sum of Desorption, Resuspension, and Sedimentation
  other_processes_sum <- sum(filtered_data$k[filtered_data$process %in% c("k_Desorption", "k_Resuspension", "k_Sedimentation")])
  
  # Calculate the difference
  difference <- other_processes_sum + k_Degradation_sum + k_Advection_sum
  print(paste("Difference for", substance_name, ":", difference))
  
  # Plot the data
  plot <- ggplot(filtered_data, aes(x = process, y = k, fill = process)) +
    geom_bar(stat = "identity") +
    scale_fill_brewer(palette = "Set1") +  # Choose a color palette (e.g., Set1)
    theme_minimal() +
    labs(
      title = paste("k values for", substance_name, "and SBsteady for", SubCompartInterest, "at", ScaleInterest),
      x = "net Process",
      y = "k [k*EqMass]"
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   geom_text(aes(label = paste("Difference:", round(difference, 6))), 
              x = 1, y = max(filtered_data$k) * 0.8, 
              hjust = 0, vjust = 0, color = "black", size = 3.5)
  
  
  print(plot)  # Explicitly print the plot
}

# Example usage
# Assuming kaas_filtered_all is your original data frame containing all relevant columns
unique_substances <- unique(kaas_filtered_all$Substance)

# Apply the function to all unique substances
for (substance in unique_substances) {
  process_substance(kaas_filtered_all, substance)
}


```

## Sensitivity Analysis 
To discover the influence of the degradation and dissolution on the mass balance of the particulates, we initiate a sensitivity analysis with different degradation rates and different dissolution rates. 
```{r Check with varying kdegradation}
Substance <- "nAg_10nm"
source("baseScripts/initWorld_onlyParticulate.R")
emissions <- data.frame(Abbr = "w1CS", 
                        Emis = 100/(3600*24*365))

ScaleInterest <- "Continental"
SubCompartInterest <- "river"
SpeciesInterest <- "Solid"
# Initialize lists and data frames
mass_balance_results <- list()
total_abs_differences <- data.frame(Degradation = numeric(), TotalAbsDifference = numeric(), stringsAsFactors = FALSE)
top_rows_list <- list()
total_mass_difference_list <- list()
kaas_filtered_list <- list()

# Generate a log scale sequence for degradation rates
log_seq <- seq(from = log10(1e-20), to = log10(1e-1), length.out = 10)

# Convert the log scale sequence back to the original scale
kdeg_range <- 10^log_seq
print(kdeg_range)

World$fetchData("Substance")
# Loop over degradation rates
for (deg in kdeg_range) {
  cat("Processing mass balance degradation:", deg, "\n")
  World$SetConst(Substance = Substance)
  print(World$fetchData("Substance"))
  kdeg <- data.frame(
  SubCompart = c("agriculturalsoil", "air", "cloudwater", "deepocean", 
                 "freshwatersediment", "lake", "lakesediment", 
                 "marinesediment", "naturalsoil", "othersoil", 
                 "river", "sea"),
  kdeg = rep(deg, 12)  
)

  ToPaste <- lapply(list(kdeg), function(x) {
    varName <- names(x)[!names(x) %in% The3D]
    stopifnot(length(varName)==1)
    # one line with 2 disadvantages of tidyverse..:
    as.data.frame(pivot_longer(data = x, cols = all_of(varName), names_to = "varName", values_to = "Waarde"))
  })
  
  dfs <- do.call(bind_rows, ToPaste)
  print(World$fetchData("Substance"))
  World$mutateVars(dfs)
  World$UpdateKaas(mergeExisting = FALSE)
  World$NewSolver("SBsteady")
  #World$Solve(emissions)
  SolutionSteady <-World$Solve(emissions)
  SolutionSteady2 <- as.data.frame(SolutionSteady)
  

  # Create unique combinations
  unique_combinations <- SolutionSteady2 |>
    select(Scale, SubCompart, Species) |>
    distinct()

  # Create emission rows
  emission_rows <- unique_combinations |>
    mutate(
      Value = "emission",
      EqMass = if_else(Scale == "Continental" & SubCompart == "river" & Species == "Solid", emissions$Emis, 0)
    )

  # Adding "from" and "to" acronyms to the R K matrix
  kaas <- as_tibble(World$kaas)
  kaas <- kaas |>
    left_join(SolutionSteady, by = c("fromScale" = "Scale", "fromSubCompart" = "SubCompart", "fromSpecies" = "Species"))

  # Multiply k by the corresponding EqMass
  kaas <- kaas |>
    mutate(k = k * EqMass)
  kaas <- kaas |>
    mutate(k = if_else(fromSubCompart == toSubCompart &
                         fromScale == toScale &
                         fromSpecies == toSpecies,
                       -abs(k),
                       k))

  # Save kaas rows for certain subcompartments 
  kaas_filtered <- kaas |>
    filter((fromScale == ScaleInterest & fromSubCompart == SubCompartInterest & fromSpecies == SpeciesInterest) |
           (toScale == ScaleInterest & toSubCompart == SubCompartInterest & toSpecies == SpeciesInterest))

  # Aggregate k values based on toScale, toSubCompart, and toSpecies
  aggregated_to <- kaas |>
    group_by(toScale, toSubCompart, toSpecies) |>
    summarise(k_sum_to = sum(k)) |>
    ungroup()

  # Aggregate k values based on fromScale, fromSubCompart, and fromSpecies
  aggregated_from <- kaas |>
    filter(k >= 0) |>
    group_by(fromScale, fromSubCompart, fromSpecies) |>
    summarise(k_sum_from = sum(k)) |>
    ungroup()

  new_rows_from <- aggregated_from |>
    mutate(
      Value = "transportfrom",
      Scale = fromScale,
      SubCompart = fromSubCompart,
      Species = fromSpecies,
      EqMass = k_sum_from
    ) |>
    select(Value, Scale, SubCompart, Species, EqMass)

  new_rows_to <- aggregated_to |>
    mutate(
      Value = "transportto",
      Scale = toScale,
      SubCompart = toSubCompart,
      Species = toSpecies,
      EqMass = k_sum_to
    ) |>
    select(Value, Scale, SubCompart, Species, EqMass)

  # Combine SolutionSteady with new rows
  massbalancecheck <- bind_rows(new_rows_from, new_rows_to, emission_rows)

  massbalancecheck_sorted <- massbalancecheck |>
    arrange(Scale, SubCompart, Species)

  # Check differences
  massbalance_difference <- massbalancecheck_sorted |>
    pivot_wider(names_from = Value, values_from = EqMass) |>
    mutate(difference = transportto + emission - transportfrom) |>
    select(Scale, SubCompart, Species, difference)

  total_abs_difference <- sum(abs(massbalance_difference$difference), na.rm = TRUE)
  total_mass_difference <- sum(massbalance_difference$difference, na.rm = TRUE)

  # Add columns for deg rate
  massbalance_difference$Degradation <- deg
  kaas_filtered$Degradation <- deg

  # Print largest differences
  top_5_rows <- massbalance_difference |>
    arrange(desc(abs(difference))) |>
    head(5)

  # Append the results to the lists
  key <- as.character(deg)
  mass_balance_results[[key]] <- massbalance_difference
  total_abs_differences <- rbind(total_abs_differences, data.frame(Degradation = deg, TotalAbsDifference = total_abs_difference))
  top_rows_list[[key]] <- top_5_rows
  total_mass_difference_list[[key]] <- total_mass_difference
  kaas_filtered_list[[key]] <- kaas_filtered
  rm(SolutionSteady)

}  
# Combine all mass balance results into a single dataframe
massbalance_difference_all <- do.call(rbind, mass_balance_results)
top_rows_all <- do.call(rbind, top_rows_list)
#kaas_filtered_all <- do.call(rbind, kaas_filtered_list)

# Print the combined results (optional)
print(massbalance_difference_all)
print(top_rows_all)

# Generate the plot
ggplot(total_abs_differences, aes(x = as.factor(Degradation), y = TotalAbsDifference, fill = as.factor(Degradation))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = paste("Total Absolute Differences by Degradation for", Substance), 
       x = "Degradation rate [s-1]", 
       y = "Total Absolute Difference [kg]") +
  scale_x_discrete(labels = function(x) format(as.numeric(x), scientific = TRUE, digits = 2)) +
  scale_fill_manual(values = scales::brewer_pal(palette = "Set3")(length(unique(total_abs_differences$Degradation)))) +
  theme_minimal() +
  theme(legend.position = "none") # Remove legend
```

```{r same as above but with kdis}
Substance <- "nC60_10nm"
source("baseScripts/initWorld_onlyParticulate.R")
emissions <- data.frame(Abbr = "w1CS", 
                        Emis = 1000/(3600*24*365))

ScaleInterest <- "Continental"
SubCompartInterest <- "river"
SpeciesInterest <- "Solid"

# Initialize lists and data frames
mass_balance_results <- list()
total_abs_differences <- data.frame(Dissolution = numeric(), TotalAbsDifference = numeric(), stringsAsFactors = FALSE)
top_rows_list <- list()
total_mass_difference_list <- list()
kaas_filtered_list <- list()

# Generate a log scale sequence for dissolution rates
log_seq <- seq(from = log10(1e-20), to = log10(1e-1), length.out = 10)

# Convert the log scale sequence back to the original scale
kdis_range <- 10^log_seq
print(kdis_range)

World$fetchData("Substance")
# Loop over dissolution rates
for (dis in kdis_range) {
  cat("Processing mass balance dissolution:", dis, "\n")
  World$SetConst(Substance = Substance)
  print(World$fetchData("Substance"))
  kdis <- data.frame(
    SubCompart = c("agriculturalsoil", "air", "cloudwater", "deepocean", 
                   "freshwatersediment", "lake", "lakesediment", 
                   "marinesediment", "naturalsoil", "othersoil", 
                   "river", "sea"),
    kdis = rep(dis, 12)
  )

  ToPaste <- lapply(list(kdis), function(x) {
    varName <- names(x)[!names(x) %in% The3D]
    stopifnot(length(varName) == 1)
    as.data.frame(pivot_longer(data = x, cols = all_of(varName), names_to = "varName", values_to = "Waarde"))
  })
  
  dfs <- do.call(bind_rows, ToPaste)
  print(World$fetchData("Substance"))
  World$mutateVars(dfs)
  World$UpdateKaas(mergeExisting = FALSE)
  World$NewSolver("SBsteady")
  SolutionSteady <- World$Solve(emissions)
  SolutionSteady2 <- as.data.frame(SolutionSteady)

  unique_combinations <- SolutionSteady2 |>
    select(Scale, SubCompart, Species) |>
    distinct()

  emission_rows <- unique_combinations |>
    mutate(
      Value = "emission",
      EqMass = if_else(Scale == "Continental" & SubCompart == "river" & Species == "Solid", emissions$Emis, 0)
    )

  kaas <- as_tibble(World$kaas)
  kaas <- kaas |>
    left_join(SolutionSteady, by = c("fromScale" = "Scale", "fromSubCompart" = "SubCompart", "fromSpecies" = "Species"))

  kaas <- kaas |>
    mutate(k = k * EqMass)
  kaas <- kaas |>
    mutate(k = if_else(fromSubCompart == toSubCompart &
                         fromScale == toScale &
                         fromSpecies == toSpecies,
                       -abs(k),
                       k))

  kaas_filtered <- kaas |>
    filter((fromScale == ScaleInterest & fromSubCompart == SubCompartInterest & fromSpecies == SpeciesInterest) |
           (toScale == ScaleInterest & toSubCompart == SubCompartInterest & toSpecies == SpeciesInterest))

  aggregated_to <- kaas |>
    group_by(toScale, toSubCompart, toSpecies) |>
    summarise(k_sum_to = sum(k)) |>
    ungroup()

  aggregated_from <- kaas |>
    filter(k >= 0) |>
    group_by(fromScale, fromSubCompart, fromSpecies) |>
    summarise(k_sum_from = sum(k)) |>
    ungroup()

  new_rows_from <- aggregated_from |>
    mutate(
      Value = "transportfrom",
      Scale = fromScale,
      SubCompart = fromSubCompart,
      Species = fromSpecies,
      EqMass = k_sum_from
    ) |>
    select(Value, Scale, SubCompart, Species, EqMass)

  new_rows_to <- aggregated_to |>
    mutate(
      Value = "transportto",
      Scale = toScale,
      SubCompart = toSubCompart,
      Species = toSpecies,
      EqMass = k_sum_to
    ) |>
    select(Value, Scale, SubCompart, Species, EqMass)

  massbalancecheck <- bind_rows(new_rows_from, new_rows_to, emission_rows)

  massbalancecheck_sorted <- massbalancecheck |>
    arrange(Scale, SubCompart, Species)

  massbalance_difference <- massbalancecheck_sorted |>
    pivot_wider(names_from = Value, values_from = EqMass) |>
    mutate(difference = transportto + emission - transportfrom) |>
    select(Scale, SubCompart, Species, difference)

  total_abs_difference <- sum(abs(massbalance_difference$difference), na.rm = TRUE)
  total_mass_difference <- sum(massbalance_difference$difference, na.rm = TRUE)

  massbalance_difference$Dissolution <- dis
  kaas_filtered$Dissolution <- dis

  top_5_rows <- massbalance_difference |>
    arrange(desc(abs(difference))) |>
    head(5)

  key <- as.character(dis)
  mass_balance_results[[key]] <- massbalance_difference
  total_abs_differences <- rbind(total_abs_differences, data.frame(Dissolution = dis, TotalAbsDifference = total_abs_difference))
  top_rows_list[[key]] <- top_5_rows
  total_mass_difference_list[[key]] <- total_mass_difference
  kaas_filtered_list[[key]] <- kaas_filtered
  rm(SolutionSteady)
}

massbalance_difference_all <- do.call(rbind, mass_balance_results)
top_rows_all <- do.call(rbind, top_rows_list)

print(massbalance_difference_all)
print(top_rows_all)

ggplot(total_abs_differences, aes(x = as.factor(Dissolution), y = TotalAbsDifference, fill = as.factor(Dissolution))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = paste("Total Absolute Differences by Dissolution for", Substance), 
       x = "Dissolution rate [s-1]", 
       y = "Total Absolute Difference [kg]") +
  scale_x_discrete(labels = function(x) format(as.numeric(x), scientific = TRUE, digits = 2)) +
  scale_fill_manual(values = scales::brewer_pal(palette = "Set3")(length(unique(total_abs_differences$Dissolution)))) +
  theme_minimal() +
  theme(legend.position = "none")
```