---
title: "Comparison between k's in Excel and in R"
author: "Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
projectRoot <- paste(getwd(), "..", "..", "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(echo = TRUE, fig.width = 14, fig.height = 8)
```

Install 2024.11.0 version of SimpleBox - the version used for verification of SimpleBox in Excel
```{R}
source("vignettes/CaseStudies/SB_in_R_paper/InstallSBoo.R")

setwd("SimpleBox/SBooScripts")
```

```{r Load packages}
library(openxlsx)
library(tidyverse)
library(ggplot2)
library(plotly)
```

Define output folder
```{R}
output_folder <- "../../vignettes/CaseStudies/SB_in_R_paper/Output/"
```

This vignette demonstrates the differences between SimpleBox in Excel and 
SimpleBox in R. The focus is calculating the relative differences for k's between
Excel and R, as the functions for some process and variable calculations have been
updated. To see which processes have changed, please see the verification 
vignettes for each substance type. 

# Calculate differences for engineered nanoparticles (example substance: nAg_10nm).

# Step 1. Compare the k's from SBExcel to SBR

# Create functions needed to compare Excel and R results

```{r comparison_1, message=FALSE, include=TRUE, echo=FALSE, warning=FALSE}
# Function to get the k's from an Excel file using the filename
GetExcelKaas <- function(SBExcelName){
  SBexcel.K <- openxlsx::read.xlsx(SBExcelName,
                    colNames = FALSE,
                    namedRegion = "K")
  SBexcel.Names <- openxlsx::read.xlsx(SBExcelName,
                    colNames = FALSE,
                    namedRegion = "box_names")
  
  colnames(SBexcel.K) <- SBexcel.Names
  
  SBexcel.K$to <-  as.character(SBexcel.Names)
  
  SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")
  
  if(World$fetchData("ChemClass") != "particle"){
    SBexcel.K <- SBexcel.K |>
      mutate(from = paste0(from, "U")) |>
      mutate(to = paste0(to, "U"))
  } else {
    SBexcel.K <- SBexcel.K
  }
}

# Function to get k's from the World object and attach the correct abbreviations
GetRKaas <- function(){
  #adding "from" and "to" acronyms to the R K matrix
  kaas <- as_tibble(World$kaas) 
  
  #R version does not us the acronyms of the excel version, set-up to convert them
  #Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
  accronym_map <- c("marinesediment" = "sd2",
                    "freshwatersediment" = "sd1",
                    "lakesediment" = "sd0", #SB Excel does not have this compartment. To do: can we turn this off (exclude this compartment) for testing?
                    "agriculturalsoil" = "s2",
                    "naturalsoil" = "s1",
                    "othersoil" = "s3",
                    "air" = "a",
                    "deepocean" = "w3",
                    "sea" = "w2",
                    "river" = "w1",
                    "lake" = "w0", 
                    "cloudwater" = "cw")
  
  accronym_map2 <- c("Arctic" = "A",
                     "Moderate" = "M",
                     "Tropic" = "T",
                     "Continental" = "C",
                     "Regional" = "R")
  
  accronym_map3 <- c("Dissolved" = "D", 
                     "Gas" = "G", 
                     "Large" = "P", 
                     "Small" = "A",
                     "Solid" = "S", 
                     "Unbound" = "U")
  
  
  kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                                        accronym_map2[fromScale], 
                                        accronym_map3[fromSpecies]),
                         to = paste0(accronym_map[toSubCompart], 
                                     accronym_map2[toScale], 
                                     accronym_map3[toSpecies]))
  
  kaas <-
    kaas |>
    mutate(from =
             ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") &
                      (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"),
                    str_replace_all(from, c("sd2"="sd","s1"="s")),
                    from)) |>
    mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") &
                         (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))
}

# Function to get the diagonal k's (removal processes) from Excel k's dataframe
GetDiagonalExcel <- function(SBexcel.K){
  diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] 
  
  #filter out dissolved and gas processes
  filtered_excel <- diagonal_excel[!endsWith(diagonal_excel$from, "D"), ]
  filtered_excel <- filtered_excel[!endsWith(filtered_excel$from, "G"), ]
}

# Function to get the R k's without the removal processes
GetFromToKaasR <- function(kaas){
  kaas2 <- kaas |>  
    filter(from != to) |> #filtering the diagonals out
    group_by(from, to) %>% summarize(k = sum(k)) |>
    mutate(fromto = paste(sep="_", from, to))
}

# Function to sum all k's in R that have the same from and to abbreviation
GetDiagonalR <- function(kaas){
  #filtered_R <-aggregate(k ~ from, data = kaas, FUN = sum) 
  filtered_R <- kaas |>
    group_by(from) |>
    summarise(k = sum(k))
}

#filtered_R <- filtered_R[!endsWith(filtered_R$from, "U"), ]

# Function to make one dataframe containing the k's for removal processes from Excel and R, and calculate the absolute and relative differences
MergeDiagonalExcelR <- function(filtered_R, filtered_excel){
  merged_diagonal <- merge(filtered_R, filtered_excel, by = "from", suffixes = c("_R", "_Excel")) 
  merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
  merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 
  sorted_diff <- merged_diagonal[order(abs(merged_diagonal$diff), decreasing = TRUE), ] |>
    mutate(reldif  = diff/k_R)
}

# Function to make one dataframe containing the k's for non-removal processes and calculate the absolute and relative differences
MergeFromToExcelR <- function(kaas2, SBexcel.K){
  SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
  SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)
  
  mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))
  
  mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K
  
  mergedkaas <- as_tibble(mergedkaas) |> mutate(reldif = diff/k_R) |> select(-fromto_R) |> select(-fromto_Excel)
}
```

# Calculate differences between Excel and R 

## Nanomaterials 
```{r Make relative difference table for nanomaterials (nAg_10nm)}
# Initialize the world for nAg_10nm
substance <- "nAg_10nm"

source("baseScripts/initWorld_onlyParticulate.R")

World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/20240815 SimpleBox4plastics_", substance, ".xlsx")

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_Nano <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "ENP") |>
  mutate(substance = Substance)
```

### Figures
```{R}
Table_Nano_SB_abbrs <- Table_Nano %>%
  mutate(
    from = case_when(
      str_sub(from, -2, -2) %in% c("A", "M", "T") & str_sub(from, 1, 2) == "sd" ~
        str_replace(from, "^sd", "sd1"),
      str_sub(from, -2, -2) %in% c("A", "M", "T") & str_sub(from, 1, 1) == "s" ~
        str_replace(from, "^s", "s1"),
      TRUE ~ from),
    to = case_when(
      str_sub(to, -2, -2) %in% c("A", "M", "T") & str_sub(to, 1, 2) == "sd" ~
        str_replace(to, "^sd", "sd1"),
      str_sub(to, -2, -2) %in% c("A", "M", "T") & str_sub(to, 1, 1) == "s" ~
        str_replace(to, "^s", "s1"),
      TRUE ~ to
    ) )

world_states_prefixed_from <- World$states$asDataFrame %>%
  rename_with(~paste0("from_", .), .cols = -Abbr)

world_states_prefixed_to <- World$states$asDataFrame %>%
  rename_with(~paste0("to_", .), .cols = -Abbr)

from_to <- Table_Nano_SB_abbrs |>
  filter(from != to) |>
  left_join(world_states_prefixed_from, by = c("from" = "Abbr")) |>
  left_join(world_states_prefixed_to, by = c("to" = "Abbr"))

diagonal <- Table_Nano_SB_abbrs |>
  filter(from == to)|>
  left_join(world_states_prefixed_from, by = c("from" = "Abbr")) |>
  left_join(world_states_prefixed_to, by = c("to" = "Abbr"))
```

```{R}
# Positive diff: R value higher than Excel value. 

# From-to
for(i in unique(from_to$from_Scale)){
  # Absolute
  plot_data <- from_to |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))
    p <- ggplot(plot_data, aes(x = tolabel, y = fromlabel, color = diff)) +
    geom_point(size = 4) +
    scale_color_gradient2(
      low = "blue", mid = "white", high = "red", midpoint = 0,
      name = "Difference"
    ) +
    labs(
      x = "To",
      y = "From",
      title = paste("Absolute difference", substance, ":", i)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  print(p)
  
  ggsave(paste0(output_folder, "Absolute difference from to ", substance, " ", i, ".png"))
  
  # Relative
    plot_data <- from_to |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))
    p <- ggplot(plot_data, aes(x = tolabel, y = fromlabel, color = reldif)) +
    geom_point(size = 4) +
    scale_color_gradient2(
      low = "blue", mid = "white", high = "red", midpoint = 0,
      name = "Relative difference"
    ) +
    labs(
      x = "To",
      y = "From",
      title = paste("Relative difference", substance, ":", i)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  print(p)
  
  ggsave(paste0(output_folder, "Relative difference from to ", substance, " ", i, ".png"))
}

# Diagonal
for(i in unique(diagonal$from_Scale)){
  # Absolute
  plot_data <- diagonal |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))

  p <- ggplot(plot_data, aes(x = fromlabel, y = diff, color = diff)) +
  geom_point(size = 4) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    x = "Compartiment",
    y = "Absolute difference",
    title = paste("Absolute difference", substance, ":", i)
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(paste0(output_folder, "Absolute difference removal ", substance, " ", i, ".png"))
  
  #Relative
    plot_data <- diagonal |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))

  p <- ggplot(plot_data, aes(x = fromlabel, y = diff, color = reldif)) +
  geom_point(size = 4) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    x = "Compartiment",
    y = "Relative difference",
    title = paste("Relative difference", substance, ":", i)
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(paste0(output_folder, "Relative difference removal ", substance, " ", i, ".png"))
}
```

## Microplastics

```{r Make relative difference table for microplastics}
# Initialize the world for microplastics
substance <- "microplastic"
source("baseScripts/initWorld_onlyPlastics.R")
World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/20240815 SimpleBox4plastics.xlsx") 

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_Microplastics <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "Microplastic")  |>
  mutate(substance = Substance)
```



## Acids

```{r Make relative difference table for acids}
substance <- "1-HYDROXYANTHRAQUINONE"
source("baseScripts/initWorld_onlyMolec.R")
World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.05_20240816_",substance,".xlsm") 

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_Acids <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "Acid")  |>
  mutate(substance = Substance)
```

## Bases

```{r Make relative difference table for bases}
substance <- "1-Hexadecanamine, N,N-dimethyl-"
source("baseScripts/initWorld_onlyMolec.R")
World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.05_20240816_",substance,".xlsm") 

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_Bases <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "Base")  |>
  mutate(substance = Substance)
```

## Metals

```{r Make relative difference table for metals}
substance <- "Sb(III)"
source("baseScripts/initWorld_onlyMolec.R")
World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.05_20240816_",substance,".xlsm") 

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_Metals <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "Metal")  |>
  mutate(substance = Substance)
```

## Without a class

```{r Make relative difference table for molecules without a class}
substance <- "1-aminoanthraquinone"
source("baseScripts/initWorld_onlyMolec.R")
World$substance <- substance

# Calculate the relative difference between k's in Excel and R
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.05_20240816_",substance,".xlsm") 

SBexcel.K <- GetExcelKaas(SBExcelName = SBExcelName)
kaas <- GetRKaas()
filtered_excel <- GetDiagonalExcel(SBexcel.K)
kaas2 <- GetFromToKaasR(kaas)
filtered_R <- GetDiagonalR(kaas)
sorted_diff <- MergeDiagonalExcelR(filtered_R, filtered_excel)
mergedkaas <- MergeFromToExcelR(kaas2, SBexcel.K) 

Table_NoClass <- bind_rows(mergedkaas, sorted_diff) |>
  select(from, to, diff, reldif) |>
  mutate(Substance = "No class")  |>
  mutate(substance = Substance)
```

# Combine the tables
```{R Combine the tables into one}
Combined_table <- bind_rows(Table_Acids, Table_Bases, Table_Metals, Table_NoClass, Table_Microplastics, Table_Nano) |>
  mutate(reldif = reldif) |>
  select(-diff)

Combined_table_wide <- Combined_table |>
  pivot_wider(names_from = "Substance", values_from = reldif)

Combined_large_diffs <- Combined_table |>
  filter(reldif > 0.1)

Large_diffs_wide <- Combined_large_diffs |>
    pivot_wider(names_from = "Substance", values_from = reldif) 

# Add states to dataframes
# Combined_table_wide <- Combined_table_wide |>
#   left_join(World$states$asDataFrame, by= "Abbr")
# 
# states <- World$states$asDataFrame

# Download the table with all relative differces
write.xlsx(Combined_table_wide, "vignettes/Development/Quality control/Publication/Combined_table_all.xlsx")

# Download the table with the largest relative differences
write.xlsx(Combined_large_diffs, "vignettes/Development/Quality control/Publication/Combined_table_large_diffs.xlsx")

# Make and save table for molecules only 
Table_molec <- Combined_table_wide |>
  select(-c(Microplastic, ENP)) |>
  drop_na()

write.xlsx(Table_molec, "vignettes/Development/Quality control/Publication/Combined_table_molec.xlsx")

# Make and save table for particles only
Table_particulate <- Combined_table_wide |>
  select(-c(Acid, Base, Metal, `No class`)) |>
  drop_na()

write.xlsx(Table_particulate, "vignettes/Development/Quality control/Publication/Combined_table_particulate.xlsx")
```

# Make figures
### Figures
```{R}
Table_Nano_SB_abbrs <- Table_Nano %>%
  mutate(
    from = case_when(
      str_sub(from, -2, -2) %in% c("A", "M", "T") & str_sub(from, 1, 2) == "sd" ~
        str_replace(from, "^sd", "sd1"),
      str_sub(from, -2, -2) %in% c("A", "M", "T") & str_sub(from, 1, 1) == "s" ~
        str_replace(from, "^s", "s1"),
      TRUE ~ from),
    to = case_when(
      str_sub(to, -2, -2) %in% c("A", "M", "T") & str_sub(to, 1, 2) == "sd" ~
        str_replace(to, "^sd", "sd1"),
      str_sub(to, -2, -2) %in% c("A", "M", "T") & str_sub(to, 1, 1) == "s" ~
        str_replace(to, "^s", "s1"),
      TRUE ~ to
    ) )

world_states_prefixed_from <- World$states$asDataFrame %>%
  rename_with(~paste0("from_", .), .cols = -Abbr)

world_states_prefixed_to <- World$states$asDataFrame %>%
  rename_with(~paste0("to_", .), .cols = -Abbr)

from_to <- Table_Nano_SB_abbrs |>
  filter(from != to) |>
  left_join(world_states_prefixed_from, by = c("from" = "Abbr")) |>
  left_join(world_states_prefixed_to, by = c("to" = "Abbr"))

diagonal <- Table_Nano_SB_abbrs |>
  filter(from == to)|>
  left_join(world_states_prefixed_from, by = c("from" = "Abbr")) |>
  left_join(world_states_prefixed_to, by = c("to" = "Abbr"))
```

```{R}
# Positive diff: R value higher than Excel value. 
for(substance in unique(Combined_table$Substance)){
  # From-to
for(i in unique(from_to$from_Scale)){
  # Absolute
  plot_data <- from_to |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))
    p <- ggplot(plot_data, aes(x = tolabel, y = fromlabel, color = diff)) +
    geom_point(size = 4) +
    scale_color_gradient2(
      low = "blue", mid = "white", high = "red", midpoint = 0,
      name = "Difference"
    ) +
    labs(
      x = "To",
      y = "From",
      title = paste("Absolute difference", substance, ":", i)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  print(p)
  
  ggsave(paste0(output_folder, "Absolute difference from to ", substance, " ", i, ".png"))
  
  # Relative
    plot_data <- from_to |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))
    p <- ggplot(plot_data, aes(x = tolabel, y = fromlabel, color = reldif)) +
    geom_point(size = 4) +
    scale_color_gradient2(
      low = "blue", mid = "white", high = "red", midpoint = 0,
      name = "Relative difference"
    ) +
    labs(
      x = "To",
      y = "From",
      title = paste("Relative difference", substance, ":", i)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  print(p)
  
  ggsave(paste0(output_folder, "Relative difference from to ", substance, " ", i, ".png"))
}

# Diagonal
for(i in unique(diagonal$from_Scale)){
  # Absolute
  plot_data <- diagonal |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))

  p <- ggplot(plot_data, aes(x = fromlabel, y = diff, color = diff)) +
  geom_point(size = 4) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    x = "Compartiment",
    y = "Absolute difference",
    title = paste("Absolute difference", substance, ":", i)
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(paste0(output_folder, "Absolute difference removal ", substance, " ", i, ".png"))
  
  #Relative
    plot_data <- diagonal |>
    filter(from_Scale == i) |>
    mutate(fromlabel = paste(from_Scale, from_SubCompart, from_Species),
           tolabel = paste(to_Scale, to_SubCompart, to_Species))

  p <- ggplot(plot_data, aes(x = fromlabel, y = diff, color = reldif)) +
  geom_point(size = 4) +
  scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    x = "Compartiment",
    y = "Relative difference",
    title = paste("Relative difference", substance, ":", i)
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)
  
  ggsave(paste0(output_folder, "Relative difference removal ", substance, " ", i, ".png"))
}
  
}

```
