---
title: "Model output"
author: "Joris Quik, Jaap Slootweg, Anne Hids"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
## impossible to generalise? projectRoot <- paste(getwd(), "..", sep = "/")
projectRoot <- paste(getwd(), "..", "..", sep = "/")
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

*SimpleBox 4.0 produces two sorts of output; a steady-state or "level
III" output, and the quasi-dynamic or "level IV" output. The
steady-state output is developed when the conditions (loadings and
environmental conditions) remain constant in time for a sufficiently
long period. The mass flows and concentrations are constant in time, as
the sum of the mass balance equation terms is equal to zero for all
boxes, and the n steady-state concentrations can be solved from the n
linear mass balance equations. This steady-state solution is obtained by
means of a matrix inversion routine. The mass flows and concentrations
that characterise this steady state are given in output tables. The
quasi-dynamic output is developed when, after a change in conditions
(loadings or environmental conditions), the mass flows and
concentrations develop toward a new steady state, according to the mass
balance equations. This computation is made by numerical integration of
the set of mass balance equations from time zero, with all
concentrations at zero, to infinite time with all concentrations
constant at steady state. The standard procedure is to compute the
steady-state solution first and carry out the dynamic response
computation afterwards, as an option.*

# *Initialize*

*This vignettes demonstrates the mechanism behind SBoo. The model first
needs to be initialized. Before initialisation a substances needs to be
selected, otherwise the "default stubstance" is used. initWorld.R does
just that, and calculates SBvars, flows and k's.*

```{r warning=FALSE}
#We need to initialize, by default a molecular substance is selected
Potential_substances <- c("1-aminoanthraquinone", # no class
                          "1-HYDROXYANTHRAQUINONE", # acid
                          "1-Hexadecanamine, N,N-dimethyl-", # base
                          "1-Chloro-2-nitro-propane", # neutral
                          "Sb(III)" # metal
                          ) 
              
substance <- Potential_substances[3]
source("baseScripts/initWorld_onlyMolec.R")

length(unique(World$kaas$process))

ProcessMolFunctions <- c("k_Adsorption", "k_Advection", "k_Burial",
                          "k_Degradation", "k_Deposition", "k_Desorption",
                          "k_Erosion", "k_Escape", 
                         "k_Leaching", "k_Resuspension", "k_Runoff", "k_Sedimentation", 
                         "k_Volatilisation")

ProcessMolFunctions[!ProcessMolFunctions %in% unique(World$kaas$process)]

unique(World$kaas$process)

```

```{r data xlsx and R}
#Comparing K's from R model to K's from excel model
library(openxlsx)
library(tidyverse)
#read in K matrix from excel
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.04_20240711_",substance,".xlsm") 

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names
SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

length(SBexcel.K$k)

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)
unique(kaas$fromScale)
unique(kaas$fromSubCompart)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", 
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale]))

# Issue that compartments sediment and soil at global scale in excel have sd and s as acronyms instead of sd2 and s1
kaas <- 
  kaas |> 
  mutate(from = 
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") & 
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"), 
                          str_replace_all(from, c("sd2"="sd","s1"="s")), 
                          from)) |> 
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") & 
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

  

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

diagonal_R <- 
  aggregate(k ~ from, data = kaas, FUN = sum) #R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

```

```{r diagonal}
#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R

ifelse( (length(diagonal_R$k)-2)==length(diagonal_excel$k) ,
       print("ok"), stop("too many compartment combo's")) 

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(diagonal_R, diagonal_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 

#Plot of differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = diff)) +
  geom_boxplot()

```

*Why are these compartments in SB-R instance: \>
diagonal*$from[!diagonal$*from %in% diagonal_excel\$from] [1] "s2A"
"s2M" "s2T" "s3A" "s3M" "s3T" "sd0A" "sd0C" "sd0M" "sd0R" "sd0T" "sd1A"
"sd1M" "sd1T" "w0A" "w0M" "w0T" "w1A" "w1M" "w1T" [21] "w3C" "w3R"*

*This is due to the degradation rate constant being calculated for all
combinations of subcompartments. There is currently no data sheet where
the exclusion based on combination of subcompartment and scale can be
set. So these types of exclusions are needed in the process function
itself.*

*So now the compartments are the correct amount based on fixing
exclusions in k_Degradation.*

```{r rate_constants}
#Dataframe of all other boxes

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

SBexcel.K$fromto[!SBexcel.K$fromto %in% kaas2$fromto] # No xlsx missing


kaas2$fromto[!kaas2$fromto %in% SBexcel.K$fromto] # more k's in SBoo than in SBxl

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)

#Column "diff" in df "mergedkaas" shows the difference between the R k output and the k's in Excel. 

#Visualize the diff per process in boxplots: which compartments (fromto boxes) have the largest differences between R and the excel model?

#Diff per "from" compartment
ggplot(mergedkaas, aes (x = from, y = diff)) +
  geom_boxplot()

#Diff per "to" compartment
ggplot(mergedkaas, aes (x = to, y = diff)) +
  geom_boxplot()

# "To" and "from" in one plot
ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") 


ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") 
#If you want to inspect per process instead of fromto box: look at the original kaas dataframe

```

```{r get differences larger than 1 percentile}

large_differences <- mergedkaas |>
  filter(relDif > 0.001) 

# print("Differences in k's between R and Excel larger thans 1 percentile")
# 
# knitr::kable(large_differences_before, format="markdown")
# 
# large_differences_before <- large_differences_before |>
#   select(from, to, k_R, k_Excel)

```

```{r s/w > a}
# Get processes involved 

toair <- kaas |>
  filter(toSubCompart == "air") |>
  filter(fromSubCompart != "air")

# process to look at is volatilisation

# Kaw differs between R and Excel

# The value for the variable R is more accurate in R than in excel (8.314462618 in R, 8.314 in Excel)
# Changing this with the test variable did not solve the issue

# Sol25 is the same
# MW is the same
# t25 is the same

# So pvap25 must be different, as it is the only remaining variable. 

# For the acid 1-HYDROXYANTHRAQUINONE, the pvap value from the substance csv was rounded to 2 decimals. In excel this number has much more decimals, causing the difference in kaw, kacompw and eventually volatilisation. 

# This should still be fixed in the substances csv.

```

```{r diff aR/aC > s2/s3 KEEP FOR FINAL VERSION!!!}

# Get the kaas from soil to air and see what processes are involved
airsoil <- kaas |>
  filter(fromScale == "Regional") |>
  filter(toSubCompart == "agriculturalsoil" | fromSubCompart == "othersoil") |>
  filter(fromSubCompart == "air")

# To calculate the Gasabs from air to soil, FRorig_spw for natural soil (and freshwater) was used also for other and agricultural soil (and seawater) 
# in excel. 
# In R however, the gasabs is calculated using the FRorig_spw for each specific soil type. By using the 'Test' variable, it was
# possible to temporarly change the FRorig_spw and FRorig in R to the value used in Excel. This fixed the large relative difference for this
# flux between excel and R. Conclusion: GASABS (used to calculate k_Adsorption) in R is calculated specifically for each 
# subcompartment, while in excel this variable is only calculated once for soil and once for water.  


```

```{r sedimentation to water KEEP FOR FINAL VERSION}
resus <- kaas |>
  filter(from == "sd2R") |>
  filter(to == "w2R") 

# The processes involved are resuspension and desorption.

# Get the vertical distance of marinesediment at continental scale
VD <- World$fetchData("VertDistance") |>
  filter(Scale == "Continental") |>
  filter(SubCompart == "marinesediment")

# Multiply the kaas by VD to be able to compare them to excel
resus <- resus |>
  mutate(mult_kaas = k*VD$VertDistance)

# Desorption is more or less the same in excel as in R, there is a larger difference in resuspension. 

# There was a mistake in Excel: Netsedrate for continental seawater was set to 0, while this should have been 2.74*10^-11


# The Kacompw differs slightly from the kacompw in excel
```

```{r Advection regional sea to continental sea}

#AdvAir$execute(debugAt = list(toScale = "Continental"))
World$moduleList[["CORRsoil"]]$execute(debugAt = list())
```

```{r water to sediment KEEP FOR FINAL VERSION}
# Get kaas from water to sediment

w2s <- kaas |>
  filter(from == "w1C") |>
  filter(to == "sd1C")

# Get the vertical distance of marinesediment at continental scale
VD <- World$fetchData("VertDistance") |>
  filter(Scale == "Continental") |>
  filter(SubCompart == "freshwatersediment")

# Multiply the kaas by VD to be able to compare them to excel
w2s <- w2s |>
  mutate(mult_kaas = k*VD$VertDistance)

# The reason that the k's from w to sd are not the same in R as in excel is that the settling velocities are different, 
# because of a difference in calculation. In excel, the settling velocity is calculated as 2.5/(24*3600). In R, the 
# settling velocity is calculated using a function. 

```

``` {r comparison of steady state emissions}

library(stringi)

World$NewSolver("SBsteady")

emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), Emis = c(10000, 10000, 10000) ) # convert 1 t/y to si units: kg/s

MW <- World$fetchData("MW")

emissions <- emissions |>
  mutate(Emis = Emis*1000/(MW*365.25*24*60*60))

SSsolve.R <- World$Solve(emissions)

SSsolve.excel <- read.xlsx(SBExcelName,
                             sheet=11,
                             colNames=TRUE,
                             rows=c(44,45))

SSsolve.excel <- SSsolve.excel |>
  select(-c(STEADY.STATE, X2)) |>
  pivot_longer(names_to = "Abbr", values_to = "EqMass", cols = everything()) 

SSsolve.R <- SSsolve.R |> mutate(Abbr =  paste0(accronym_map[SubCompart], 
                            accronym_map2[Scale]))

merged_SS_before <- merge(SSsolve.R, SSsolve.excel, by="Abbr", suffixes = c(".R", ".Excel")) |>
  mutate(absdiff = EqMass.R-EqMass.Excel) |>
  mutate(reldiff = absdiff/EqMass.R)

print("Difference in emissions between R and Excel")

knitr::kable(merged_SS, format="markdown")

```
