---
title: "Comparison of 2025.4.0 to 2026.3.1"
author: "Joris Quik"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

source("InstallSBoo.R")

```

# Setup for comparison

The below scripts are setup to compare a specific release to the current
version in use. E.g. loaded in R-Studio. (SBoo and SBooScripts). This
can be a release, e.g. the tag or specific branches in SBoo and
SBooScripts.

```{R Specify tag and load}
tag <- "2025.04.0"
# branch_sboo <- "FLux_work"
# branch_sbooscripts <- "Fluxwork"

InstallSBoo(Release = tag)
```

Select random substances for testing

```{R select substances, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
# excluded_substances <- c("Graphene Oxide", "Chitosan", "nTiO2_P25_CaLIBRAte_D6_3", "nAg_NanoFase", "nanoparticle", "GO-Chitosan")

substance_current <- read.csv("data/Substances.csv")
substances_other <-  read.csv("SimpleBox/SBooScripts/data/Substances.csv")

substances <- inner_join(substance_current,substances_other)

# Select 5 random substances for each class from substances csv
# and always select the default microplastic.

mp <- substances |>
  filter(Substance == "microplastic")

set.seed(492)

substances <- substances |>
  group_by(ChemClass) |>
  slice_sample(n = 5, replace = FALSE) |>
  ungroup()

substances <- rbind(substances, mp) |>
  distinct()

```

## Calculate for installed version

Calculate first order rate constants for the other version. Here we
apply the new initWorldOther from the current branch as there are
missinng links to the new install folder in the old release. This can be
changed in future comparisons.

```{R warning=FALSE}

# SBInstallFolder

# Calculate the kaas 
other_kaas <- data.frame()
  
for(substance in substances$Substance){
 
  source("baseScripts/initWorldOther.R")
    
  other_kaas <- rbind(other_kaas,  World$kaas |>
    mutate(Substance = substance,
           ChemClass = World$fetchData("ChemClass")))
}

```

## Calculate for currently loaded version

Calculate the first order rate constants for the current branch

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
```

```{R Run SB and get advection ks for FLux_work branch, echo=TRUE, message=TRUE, warning=FALSE, results="hide"}

current_kaas <- data.frame()
  
for(substance in substances$Substance){

  source("baseScripts/initWorld.R")
    
  current_kaas <- rbind(current_kaas,  World$kaas |>
                          mutate(Substance = substance,
                                 ChemClass = World$fetchData("ChemClass")))
}
```

# Compare rate constants, k's.

```{R Calculate the differences between ks, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
common_cols <- setdiff(intersect(colnames(current_kaas), colnames(other_kaas)), "k")

kaas_comparison <- merge(other_kaas, current_kaas, by=common_cols, suffixes = c("_Old", "_New"))

kaas_comparison <- kaas_comparison |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_New)

changed_kaas_all <- kaas_comparison |>
  filter(diff != 0) |>
  mutate(full_name = paste0("From ", fromSpecies, "_",fromSubCompart, "_", fromScale, 
                            " to ", toSpecies, "_", toSubCompart, "_", toScale))


head(changed_kaas_all)

unique(changed_kaas_all$process)

changed_kaas <- changed_kaas_all |>
  filter(abs(rel_diff) > 0.001) 

unique(changed_kaas$process)

```

The following processes seem to have changed k's:
`r unique(changed_kaas_all$process)`, but only
`r unique(changed_kaas$process)` changed more than 0.1%. We plot these
changes to check which k's are affected Below the relative difference in
rate constants is seen in figures:

```{R Plot the differences between ks, fig.height=15, fig.width=15}
custom_theme <- function() {
  theme(
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    plot.margin = margin(t = 10, r = 10, b = 30, l = 100) # ,
    # panel.background = element_rect(fill = "white", colour = "grey"),   # White background
    # panel.grid.major = element_line(colour = "grey")#,                   # Black major grid lines
    # panel.grid.minor = element_line(colour = "black")                    # Black minor grid lines
  )
}

  # p1 <- 
    ggplot(changed_kaas, aes(x = full_name, y = rel_diff)) +
    geom_point() +
    facet_wrap( ~ process, scales = "free") +
    #ggtitle(paste0("Relative difference in k's for ", unique(changed_kaas$process)[i]," between ", tag, " and current branch")) +
    # geom_hline(yintercept = 0.001, color = "red") +
    # geom_hline(yintercept = -0.001, color = "red") +
    custom_theme() +
    labs(x = "From to Compartment", y = "Relative diff in k")

LakeReduction <-   changed_kaas |> filter(process == "k_Advection" &
                            fromSubCompart == "lake") |> pull(rel_diff) |> min()

RiverIncrease <-   changed_kaas |> filter(process == "k_Advection") |> pull(rel_diff) |> max()


    
```

# Specific's on differences

## Advection

These results clearly show the changes in discharge calculation from
lake to river now being made dependent on the rain rate and the fraction
runoff reaching lake water. This reduces (in default scenario) the
advection rate constant by `r round(-LakeReduction, digits = 3)*100` %.
As expected this slightly increases the outflow from river to sea, given
the same rainrate. This also changes the mixing of regional and
continental sea's by the same amount. An increase of
`r round(RiverIncrease, digits = 3)*100` %.

## Heteroaglomeration

The large differences, more than 100 times lower k values are for two
specific substances:

```{r echo=FALSE}
SpecSubs <- changed_kaas |> filter(process == "k_HeteroAgglomeration.wsd") |> distinct(Substance) |> pull()
SpecSubs
```

This happens due to these two substances being the first to also
introduce a shape dimension (short, long, intermediate side's) for
calculation of the settling velocity in v 2025.4.0. As the current
version (soon to be 2026.3.1) already includes updated algorithms to
include shape in the calculation of the gravitational settling collision
frequency, this results in the large difference in
k_HeteroAgglomeration.wsd.

Further update of shape related work will be further reviewed in
upcoming release aimed at microplastics.
