---
title: "Comparison water advection update"
author: "Anne Hids, Joris Quik"
date: "2025-05-27"
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r init, message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse)
source("baseScripts/initWorld_onlyMolec.R")

f_MutateHelper <- function(varName = "landFRAC",
                           varData = c(0.6000, 0.0050, 0.2700, 0.1000, 0.0250, 0.6000, 0.0050, 0.2700, 0.1000, 0.0250)){
  varWorld <- World$fetchData(varName)
  varWorld[,varName] <- varData
  varWorld <-
    varWorld |> pivot_longer(cols = varName,
                             names_to = "varName",
                             values_to = "Waarde")
  World$mutateVars(varWorld)
  World$UpdateDirty(varName)
  return(World$fetchData(varName))
}



```

The advection flow calculations for the freshwater compartments are
revised using a different implementation of advective flows related to
the lake compartment and some other changes (from v2025.4.0 and the
Excel implementation). Below the changes are outlined, and finally a
comparison between the rate constants (k's) of the new and old version
is made.

## Update RainOnFreshwater function

In the previous version of SimpleBox, RainOnFreshwater was set to 0 for
lake compartments. This is resolved by using the same calculation for
lake as was already used for the river compartments.

This was the previous function:

```{R Old RainOnFreshwater function, include=TRUE}
# RainOnFreshwater
# World$fetchData("RainOnFreshwater") # new implementation provides

RainOnFreshwater <- function (RAINrate, Area, SubCompartName) {
  if (SubCompartName %in% c("river", "lake")) {
    #TODO resolve lake issues in waterflow; for now: old formulas
    if (SubCompartName == "lake"){
      return(0)
    } else {
      # RAINrateToSI is generated from units !
      return(RAINrate * Area)
    }     
  } else    return(NA)
}

CalcVariable <- World$NewCalcVariable("RainOnFreshwater") # load the old function above
# CalcVariable
# CalcVariable$needVars
# CalcVariable$execute()
old <- World$CalcVar("RainOnFreshwater")

```

The updated function now also adds rain on lake freshwater and replaces
the 0's for lake water at global scales with NA's which causes them to
be filtered out automatically.

```{R New RainOnFreshwater function, include=TRUE}
RainOnFreshwater <- function (RAINrate, Area, SubCompartName) {
  if (SubCompartName %in% c("river", "lake")) {
    # RAINrateToSI is generarted from units !
    return(RAINrate * Area)
  } else    return(NA)
}

CalcVariable <- World$NewCalcVariable("RainOnFreshwater") # load the old function above
# CalcVariable
# CalcVariable$needVars
# CalcVariable$execute()
World$CalcVar("RainOnFreshwater")

```

## Optimisation of FracROWatComp

This function calculates the area fraction of river and lake of the
total freshwater area at Regional and Continental scales. This is needed
to calculate how much runoff goes from soil to the different water
compartments.

The new function calculates the same as the old function with two
exceptions: -The function is shorter -There are NA's returned instead of
1's for scales which are not Regional or Continental and the SubCompart
is not river or lake.

```{R Old FracROWatComp function, include=TRUE}
FracROWatComp <- function(all.landFRAC, all.Matrix, Matrix, SubCompartName, ScaleName) {
  compFrac <- all.landFRAC$landFRAC[all.landFRAC$SubCompart == SubCompartName & all.landFRAC$Scale ==  ScaleName]
  all.landFrac <- as_tibble(all.landFRAC)
  all.Matrix <- as_tibble(all.Matrix)
  mergeddata <- left_join(
    x = all.landFRAC,
    y = all.Matrix,
    by = join_by(SubCompart))
  
  if ((Matrix == "water") & (ScaleName %in% c("Regional", "Continental"))) {
    # total landfrac of (fresh) water compartments
    waterFrac <- mergeddata |>
      filter(Matrix == "water" & Scale == ScaleName) |>
      summarise(waterFrac = sum(landFRAC, na.rm = TRUE)) |>
      pull(waterFrac)
    return(compFrac / waterFrac)
  } else {
    return(1)
  }
}


CalcVariable <- World$NewCalcVariable("FracROWatComp") # load the old function above
# CalcVariable
# CalcVariable$needVars
# CalcVariable$execute()
old <- World$CalcVar("FracROWatComp")

```

```{R New FracROWatComp function, include=TRUE}
FracROWatComp <- function(all.landFRAC, all.Matrix, Matrix, SubCompartName, ScaleName) {
  # browser()
  
  if ((Matrix == "water") & (ScaleName %in% c("Regional", "Continental"))) {
    
    compFrac <- all.landFRAC$landFRAC[all.landFRAC$SubCompart == SubCompartName & all.landFRAC$Scale == ScaleName]
    mergeddata <- merge(all.landFRAC, all.Matrix)
    waterFrac <- sum(mergeddata$landFRAC[mergeddata$Matrix == "water" & mergeddata$Scale == ScaleName])
    return(compFrac / waterFrac)
    
  } else if ((SubCompartName == "sea") & (ScaleName %in% c("Tropic", "Moderate", "Arctic"))){ 
    return(1)
  }  else
  {
    return(NA)
  }
}

CalcVariable <- World$NewCalcVariable("FracROWatComp") # load the old function above
# CalcVariable
# CalcVariable$needVars
# CalcVariable$execute()
World$CalcVar("FracROWatComp")

```

## Update ContRiver2Reg

For the 'river' SubCompart at 'Continental' Scale, the flow from
Continental to Regional river is now calculated as:

(sum of all runoff + the sum of all rain on freshwater) \*
dischargefraction between regional and continental scale.

The previously used function can be seen in the chunk below

```{R Old ContRiver2Reg, include=TRUE}

f_MutateHelper(varName = "dischargeFRAC",
               varData = c(0.1,0))

x_ContRiver2Reg <- function (ScaleName, SubCompartName, 
                             all.RunoffFlow, RainOnFreshwater, 
                             dischargeFRAC, LakeFracRiver){
  switch (ScaleName,
          "Continental" = {
            switch (SubCompartName,
                    "river" = { 
                      SumRainRunoff <- sum(all.RunoffFlow$RunoffFlow[all.RunoffFlow$Scale == "Continental"])
                      River2sea  <- RainOnFreshwater + SumRainRunoff * (1-dischargeFRAC)
                      Lake2River <- LakeFracRiver * River2sea
                      
                      return((RainOnFreshwater + SumRainRunoff + Lake2River) * dischargeFRAC)
                    },
                    return(NA)
            )
          },
          return(NA)
  )
}
# World$fetchData("RunoffFlow")
CalcVariable <- World$NewCalcVariable("x_ContRiver2Reg") # load the old function above
# World$fetchData("dischargeFRAC")

# World$fetchData("dischargeFRAC")

# World$fetchData("Runoff")

old <- World$CalcVar("x_ContRiver2Reg")



```

The updated function is simpler, more straightforward. the outflow
depends purely on the fraction of Rain and Runoff going to the regional
scale.

```{R New ContRiver2Reg, include=TRUE}
x_ContRiver2Reg <- function (ScaleName, SubCompartName, 
                             all.RunoffFlow, all.RainOnFreshwater, 
                             dischargeFRAC){
  
  switch (ScaleName,
          "Continental" = {
            switch (SubCompartName,
                    "river" = {
                      SumRainRunoff <- sum(all.RunoffFlow$RunoffFlow[all.RunoffFlow$Scale == ScaleName])+
                        sum(all.RainOnFreshwater$RainOnFreshwater[all.RainOnFreshwater$Scale == ScaleName])
                      # River2sea  <- RainOnFreshwater + SumRunoff * (1-dischargeFRAC)
                      # Lake2River <- LakeFracRiver * River2sea
                      
                      return((SumRainRunoff) * dischargeFRAC)
                    },
                    return(NA)
            )
          },
          return(NA)
  )
  
}
CalcVariable <- World$NewCalcVariable("x_ContRiver2Reg") # load the old function above
new <- World$CalcVar("x_ContRiver2Reg")

diff <- old$x_ContRiver2Reg - new$x_ContRiver2Reg
diff
# diff/old$x_ContRiver2Reg
# diff/new$x_ContRiver2Reg

```

The difference between the two functions is because in previous
itteration LakeFracRiver was used (World\$fetchData("LakeFracRiver")) to
calculate the flow from up stream fresh water (e.g. lake) (w0) to the
major downstream freshwater compartment, e.g. river (w1). This is now
done based on the calculated RunoffFlow going to w0 plus Rain on w0.
Details on the move away from using LakeFracRiver, see below.

## Update LakeOut function

The Lake compartment (w0) is an upstream compartment within a Scale that
has outflow to fresh water (w1). The outflow is dependent on the
fraction runoff going to lake (FracROWatComp) and the direct Rain on
lake fresh water (RainOnFreshwater). The flow from lake to river at
Regional and Continental scale is now calculated as:

RainOnFreshwater + FracROWatComp\*SumRunoff

```{R Old LakeOut function, include=TRUE}
x_LakeOutflow <- function(all.x_RiverDischarge, 
                          all.x_ContRiver2Reg, 
                          LakeFracRiver, 
                          ScaleName){
  x_RiverDischarge <- all.x_RiverDischarge$flow[all.x_RiverDischarge$fromScale==ScaleName]
  if(ScaleName == "Continental") {
    return(LakeFracRiver * x_RiverDischarge)
  } 
  if(ScaleName == "Regional"){
    x_ContRiver2Reg <- all.x_ContRiver2Reg$flow
    return(LakeFracRiver * (x_RiverDischarge + x_ContRiver2Reg))
  } 
  else NA
}


CalcVariable <- World$NewCalcVariable("x_LakeOutflow") # load the old function above
old <- World$CalcVar("x_LakeOutflow")

```

```{R New LakeOut function, include=TRUE}
x_LakeOutflow <- function(RainOnFreshwater,
                          all.RunoffFlow,
                          FracROWatComp,
                          SubCompartName,
                          ScaleName){
  switch(SubCompartName, # if this is coded with if statement it fails as SubCompartName for Arctic is NA
         "lake" = {
           SumRunoff <- sum(all.RunoffFlow$RunoffFlow[all.RunoffFlow$Scale == ScaleName])
           return(RainOnFreshwater + FracROWatComp*SumRunoff)
         },
         NA
  )
}

CalcVariable <- World$NewCalcVariable("x_LakeOutflow") # load the old function above
World$CalcVar("x_LakeOutflow")


# World$CalcVar("FracROWatComp")

```

The new implementation results in a lower LakeOut flow because it is not
fixed to a set fraction of the river discharge, but calculated based on
the fraction of Runoff going to lake which is based on the surface are
of the lake versus surface area of the river fresh water compartment.

# Compare new with previous implementation

The below scripts should be run to compare two versions of SBoo. This
uses a function for downloading and ordering the folders needed to
compare.

TO DO: fetch development at a certain date/commit hash instead of most
recent version in a branch. This would ensure consequent outcomes of
this comparison, no matter when the script is run or if a branch is
deleted.

```{R Specify directories, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
source("vignettes/Development/Quality control/ComparisonFunctions.R")
# Release = "2025.04.0"
folderpaths <- CompareFilesPrep(Release = "2025.04.0", # specify the tag to compare to
                                Test_SBoo = "FLux_work", # the branch name
                                Test_SBooScripts = "Fluxwork", # the branch name
                                Temp_Folder = "C:/Temp") # test folder




```

As advection processes do not depend on substance variables, we only
need to compare the advection flows for one substance, in this case
1-aminoanthraquinone.

```{R Specify substance, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
# substance_names <- "1-aminoanthraquinone" 
substances <- read.csv("data/Substances.csv")
# substances <- substances |>
#   filter(ChemClass != "particle")

substances <- substances |>
  filter(ChemClass == "")

substances <- substances[c(1,2),]
```

First remove all functions from the global environment to avoid errors

```{R Clean environment, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])
source(paste0(folderpaths[2], "/SBooScripts/vignettes/Development/Quality control/MassBalanceFunction.R"))

```

## Main version / development branch

```{R Run SB and get advection ks for development, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
# Save the original working directory for later
original_wd <- getwd()

# Change the wd to the wd of the main branches
setwd(paste0(folderpaths[1], "/SBooScripts"))

# Calculate the kaas 
main_kaas <- data.frame()
# main_solution <- "start"

for(i in 1:nrow(substances)){
  # Get the substance type
  subst_row <- substances[i, ]
  
  substance <- subst_row$Substance
  cc <- subst_row$ChemClass
  
  if(substance == 'microplastic' | substance == "TRWP"){
    source("baseScripts/initWorld_onlyPlastics.R")
  } else if(cc == "particle"){
    source("baseScripts/initWorld_onlyParticulate.R")
  } else{
    source("baseScripts/initWorld_onlyMolec.R")
  }
  
  kaas <- World$kaas |>
    # filter(process == "k_Advection") |>
    mutate(Substance = substance)
  
  main_kaas <- rbind(main_kaas, kaas)
  
  
  if(World$fetchData("ChemClass") == "particle"){
    emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), 
                            Emis = c(10000, 10000, 10000)) # convert 1 t/y to si units: kg/s
    emissions <- emissions |>
      mutate(Emis = Emis * 1000 / (365 * 24 * 60 * 60))
  } else {
    emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), 
                            Emis = c(10000, 10000, 10000)) # convert 1 t/y to si units: mol/s
    emissions <- emissions |>
      mutate(Emis = Emis * 1000 / (365 * 24 * 60 * 60))
  }
  
  World$NewSolver("SteadyStateSolver")
  World$Solve(emissions)
  
  solution <- World$Masses()
  names(solution)[names(solution) == "Mass_kg"] <- "main_Mass_kg"
  solution$Substance <- substances$Substance[i]
  
  if(i == 1) main_solution <- solution else {
    main_solution <- merge(main_solution,solution,all = TRUE)}
  
  solution_conc <- World$Concentration()
  names(solution_conc)[names(solution_conc) == "Concentration"] <- "main_Conc."
  solution_conc$Substance <- substances$Substance[i]
  if(i == 1) main_solution_conc <- solution_conc else {
    main_solution_conc <- merge(main_solution_conc,solution_conc,all = TRUE)}
  
  
  
  MB_main <- MB_function1(kaas = as_tibble(World$kaas),
                          Masses =  World$Masses())
  
  if (i == 1) main_MassBalance <- MB_main else {
    main_MassBalance <- merge(main_MassBalance, MB_main, all = TRUE) 
  }
  
  Kmatrix <- World$exportEngineR()
  
  # Strange that there would be a value here
  # Kmatrix["s1TU","w2TU"]
  
  #       to     from
  Kmatrix["w2TU","s1TU"]
  
  # But no value here:
  kaas |> filter(toScale == "Tropic" & toSubCompart == "sea")
  # test to see why below we get a flow from soil to sea
  
  
  Masses_Korder <- solution[match(colnames(Kmatrix),solution$Abbr),]
  CompFlows <-  Kmatrix %*% diag(Masses_Korder$main_Mass_kg)
  colnames(CompFlows) <- row.names(CompFlows)
  CompFlows <-  as.data.frame(CompFlows)
  CompFlows$To <- colnames(CompFlows)
  CompFlows <-
    CompFlows |> as_tibble() |> 
    pivot_longer(-To,
                 names_to = "From",
                 values_to = "Flow_kg/s") |> 
    mutate( Substance = substances$Substance[i] )
  
  if (i == 1) main_CompFlows <- CompFlows else {
    main_CompFlows <- merge(main_CompFlows, CompFlows, all = TRUE) 
  }
  
}

```


|To   |From |Substance                             |Flow_kg/s_main |Flow_kg/s_test |diff     |    reldiff|
|:----|:----|:-------------------------------------|:--------------|:--------------|:--------|----------:|
|s1AU |aAU  |(4-chloro-2-methylphenoxy)acetic acid |1.2e-01        |3.6e-01        |2.4e-01  |  0.4982982|
|s1AU |aAU  |1-aminoanthraquinone                  |3.9e-02        |1.1e+00        |1.1e+00  |  0.9334502|
|s1MU |aMU  |(4-chloro-2-methylphenoxy)acetic acid |1.2e+00        |2.7e+00        |1.5e+00  |  0.3777706|
|s1MU |aMU  |1-aminoanthraquinone                  |4.5e-01        |8.8e+00        |8.3e+00  |  0.9021398|
|s1TU |aTU  |(4-chloro-2-methylphenoxy)acetic acid |1.7e-02        |3.2e-02        |1.5e-02  |  0.3144034|
|s1TU |aTU  |1-aminoanthraquinone                  |5.2e-03        |8.0e-02        |7.4e-02  |  0.8775798|
|w2AU |s1AU |(4-chloro-2-methylphenoxy)acetic acid |1.3e-01        |0.0e+00        |-1.3e-01 | -1.0000000|
|w2AU |s1AU |1-aminoanthraquinone                  |4.9e-01        |0.0e+00        |-4.9e-01 | -1.0000000|
|w2MU |s1MU |(4-chloro-2-methylphenoxy)acetic acid |4.7e-01        |0.0e+00        |-4.7e-01 | -1.0000000|
|w2MU |s1MU |1-aminoanthraquinone                  |1.8e+00        |0.0e+00        |-1.8e+00 | -1.0000000|
|w2TU |s1TU |(4-chloro-2-methylphenoxy)acetic acid |9.5e-01        |0.0e+00        |-9.5e-01 | -1.0000000|
|w2TU |s1TU |1-aminoanthraquinone                  |3.6e+00        |0.0e+00        |-3.6e+00 | -1.0000000|

|To   |From |Substance                             |Flow_kg/s_main |Flow_kg/s_test |diff     |    reldiff|
|:----|:----|:-------------------------------------|:--------------|:--------------|:--------|----------:|
|aAU  |s1AU |(4-chloro-2-methylphenoxy)acetic acid |1.3e-10        |3.8e-10        |2.5e-10  |  0.4982982|
|aAU  |s1AU |1-aminoanthraquinone                  |2.9e-10        |8.4e-09        |8.1e-09  |  0.9334502|
|aMU  |s1MU |(4-chloro-2-methylphenoxy)acetic acid |1.9e-08        |4.1e-08        |2.3e-08  |  0.3777706|
|aMU  |s1MU |1-aminoanthraquinone                  |9.4e-08        |1.8e-06        |1.7e-06  |  0.9021398|
|aTU  |s1TU |(4-chloro-2-methylphenoxy)acetic acid |1.8e-09        |3.5e-09        |1.7e-09  |  0.3144034|
|aTU  |s1TU |1-aminoanthraquinone                  |1.1e-08        |1.7e-07        |1.6e-07  |  0.8775798|
|w2AU |s1AU |(4-chloro-2-methylphenoxy)acetic acid |2.7e-04        |0.0e+00        |-2.7e-04 | -1.0000000|
|w2AU |s1AU |1-aminoanthraquinone                  |2.5e-04        |0.0e+00        |-2.5e-04 | -1.0000000|
|w2MU |s1MU |(4-chloro-2-methylphenoxy)acetic acid |6.4e-03        |0.0e+00        |-6.4e-03 | -1.0000000|
|w2MU |s1MU |1-aminoanthraquinone                  |6.8e-03        |0.0e+00        |-6.8e-03 | -1.0000000|
|w2TU |s1TU |(4-chloro-2-methylphenoxy)acetic acid |2.6e-04        |0.0e+00        |-2.6e-04 | -1.0000000|
|w2TU |s1TU |1-aminoanthraquinone                  |2.4e-04        |0.0e+00        |-2.4e-04 | -1.0000000|

## New implementation.

```{R Clean environment again, include=FALSE}
rm(list = ls()[sapply(ls(), function(x) is.function(get(x)))])

source(paste0(folderpaths[2], "/SBooScripts/vignettes/Development/Quality control/MassBalanceFunction.R"))

```

```{R Run SB and get output data, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}
# Change the wd to the wd of the main branches
setwd(paste0(folderpaths[2], "/SBooScripts"))

test_kaas <- data.frame()
test_solution <- NA
test_MassBalance <- NA
test_CompFlows <- NA

for(i in 1:nrow(substances)){
  # Get the substance type
  subst_row <- substances[i, ]
  
  substance <- subst_row$Substance
  cc <- subst_row$ChemClass
  
  if(substance == 'microplastic' | substance == "TRWP"){
    source("baseScripts/initWorld_onlyPlastics.R")
  } else if(cc == "particle"){
    source("baseScripts/initWorld_onlyParticulate.R")
  } else{
    source("baseScripts/initWorld_onlyMolec.R")
  }
  
  #   mutLandFrac <- World$fetchData("landFRAC")
  # mutLandFrac$landFRAC <- c(0.6000, 0.0050, 0.2700, 0.1000, 0.0250, 0.6000, 0.0050, 0.2700, 0.1000, 0.0250)
  # mutLandFrac <-
  #   mutLandFrac |> pivot_longer(cols = "landFRAC",
  #                             names_to = "varName",
  #                             values_to = "Waarde")
  # World$mutateVars(mutLandFrac)
  # World$UpdateDirty("landFRAC")
  
  
  kaas <- World$kaas |>
    # filter(process == "k_Advection") |>
    mutate(Substance = substance)
  
  test_kaas <- rbind(test_kaas, kaas)
  
  if(World$fetchData("ChemClass") == "particle"){
    emissions <- data.frame(Abbr = c("aRS", "s2RS", "w1RS"), 
                            Emis = c(10000, 10000, 10000)) # convert 1 t/y to si units: kg/s
    emissions <- emissions |>
      mutate(Emis = Emis * 1000 / (365 * 24 * 60 * 60))
  } else {
    emissions <- data.frame(Abbr = c("aRU", "s2RU", "w1RU"), 
                            Emis = c(10000, 10000, 10000)) # convert 1 t/y to si units: mol/s
    emissions <- emissions |>
      mutate(Emis = Emis * 1000 / (365 * 24 * 60 * 60))
  }
  
  World$NewSolver("SteadyStateSolver")
  World$Solve(emissions)
  
  solution <- World$Masses()
  names(solution)[names(solution) == "Mass_kg"] <- "test_Mass_kg"
  solution$Substance <- substances$Substance[i]
  if(i == 1) test_solution <- solution else {
    test_solution <- merge(test_solution,solution,all = TRUE)}
  
  solution_conc <- World$Concentration()
  names(solution_conc)[names(solution_conc) == "Concentration"] <- "test_Conc."
  solution_conc$Substance <- substances$Substance[i]
  if(i == 1) test_solution_conc <- solution_conc else {
    test_solution_conc <- merge(test_solution_conc,solution_conc,all = TRUE)}
  
  MB_test <- MB_function1(kaas = as_tibble(World$kaas),
                          Masses =  World$Masses())
  
  if (i == 1) test_MassBalance <- MB_test else {
    test_MassBalance <- merge(test_MassBalance, MB_test, all = TRUE) 
  }
  
  Kmatrix <- World$exportEngineR()
  Masses_Korder <- solution[match(colnames(Kmatrix),solution$Abbr),]
  
  # Masses_Korder$Abbr == colnames(Kmatrix)
  # Masses_Korder$Abbr == rownames(Kmatrix)
  
  
  
  Kmatrix["s1TU","w2TU"]
  
  Kmatrix["w2TU","s1TU"]
  
  CompFlows <-  Kmatrix %*% diag(Masses_Korder$test_Mass_kg)
  colnames(CompFlows) <- row.names(CompFlows)
  CompFlows <-  as.data.frame(CompFlows)
  CompFlows$To <- colnames(CompFlows)
  CompFlows <-
    CompFlows |> as_tibble() |> 
    pivot_longer(-To,
                 names_to = "From",
                 values_to = "Flow_kg/s") |> 
    mutate( Substance = substances$Substance[i] )
  
  if (i == 1) test_CompFlows <- CompFlows else {
    test_CompFlows <- merge(test_CompFlows, CompFlows, all = TRUE) 
  }
  
  
}

setwd(original_wd)

```

## Compare the two versions
The comparison is done for different outcomes. Starting the the rate constrations, followed by the relative difference in mass in each compartment and the flows to and from compartments to further understand the differences in mass.

### rate constants

```{R Comparison of rate constants k's, results ="hide", echo=TRUE, message=FALSE, warning=FALSE}

kaas_comparison <- 
  merge(main_kaas |> 
          rename(
            k_Old = k
          ), 
        test_kaas |> 
          rename(
            k_New = k),
  all = TRUE) |>
  mutate(diff = k_New-k_Old) |> # If this number is positive, the New_k is higher than the Old_k (higher advection rate with new method)
  mutate(rel_diff = diff/k_Old) |> as_tibble()

changed_kaas <- 
  kaas_comparison |>
  filter(diff != 0) |>
  mutate(full_name = paste0("From ", fromSubCompart, "_", fromScale, " to ", toSubCompart, "_", toScale)) |> 
  mutate(fromname = paste0(fromSubCompart, "_", fromScale)) |>
  mutate(toname = paste0(toSubCompart, "_", toScale))

ggplot(changed_kaas, mapping = aes(x = toname, y = fromname, color = rel_diff)) + 
  geom_point(aes(shape = Substance)) + 
  labs(
    title = "Relative difference between old and new advection k's",
    x = "To",  
    y = "From",  
    color = "Relative difference"  
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.background = element_blank()  
  )

table_for_display <- changed_kaas |>
  select(fromScale, fromSubCompart, toScale, toSubCompart, Substance, k_Old, k_New, diff, rel_diff) |>
  mutate(diff = format(diff, scientific = TRUE, digits = 2)) |>
  mutate(rel_diff = format(rel_diff, scientific = TRUE, digits = 2))

knitr::kable(table_for_display)

```

### mass

```{R comparison of mass}

solution_comparison <- 
  merge(main_solution,test_solution) |> 
  mutate(diff = test_Mass_kg-main_Mass_kg) |> 
  mutate(rel_diff = diff/main_Mass_kg)

solution_comparison |>
  filter(rel_diff > 0.001) 


# ggplot(solution_comparison, mapping = aes(x = Abbr, y = diff, fill = Substance)) + 
#   geom_bar(stat = "identity", position = "dodge") +
#   labs(
#     title = "Difference between old and new steadystate mass",
#     x = "Compartment",  
#     y = "(Mass_new - Mass_old)"
#   ) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1), 
#     panel.grid.major = element_line(size = 0.2, color = "gray90"),
#     panel.background = element_blank()  
#   )

ggplot(solution_comparison, mapping = aes(x = Abbr, y = rel_diff, fill = Substance)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Relative difference between old and new steadystate mass",
    x = "Compartment",  
    y = "(Mass_new-Mass_old)/Mass_old"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.background = element_blank()  
  )



```

And concentration.

```{r comparison of concentration (solution_conc)}

con_solution_comparison <- 
  merge(main_solution_conc,test_solution_conc) |> 
  mutate(diff = test_Conc.-main_Conc.) |> 
  mutate(rel_diff = diff/main_Conc.)

ggplot(con_solution_comparison, mapping = aes(x = Abbr, y = rel_diff, fill = Substance)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Relative difference between old and new steadystate concentrations",
    x = "Compartment",  
    y = "(New-Previous)/Previous"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.background = element_blank()  
  )

```

### Flows
```{r}

test_MassBalance$version <- "new"
main_MassBalance$version <- "2025.04.0"

MB_test_main <- rbind(test_MassBalance,main_MassBalance)



ggplot(MB_test_main |> 
         mutate(Trans_from = -1*Trans_from,
                Removal_kg_s = -1*Removal_kg_s) |> 
         pivot_longer(!c(Compartment,Substance,version),
                      names_to = "Flow",
                      values_to = "Mass flow (kg/s)") , 
       aes(x = Compartment, y = `Mass flow (kg/s)`, fill = Flow)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~Substance*version) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    panel.grid.major.y = element_line(color = "grey80", size = 0.7),
    panel.grid.minor.y = element_line(color = "grey90", size = 0.5)
  ) +
  scale_y_continuous(breaks = scales::extended_breaks(20))

MB_test_main_wide <-
  MB_test_main |> pivot_wider(names_from = version, values_from =  c(Trans_to,   Trans_from, Emission_kg_s, Removal_kg_s,Diff_Flows)) |> 
  mutate(reldiff_from = (Trans_from_new-Trans_from_2025.04.0)/Trans_from_2025.04.0,
         reldiff_to = (Trans_to_new-Trans_to_2025.04.0)/Trans_to_2025.04.0,
         reldiff_removal = (Removal_kg_s_new-Removal_kg_s_2025.04.0)/Removal_kg_s_2025.04.0)

ggplot(MB_test_main_wide |> 
         pivot_longer(cols = c(reldiff_from, reldiff_to, reldiff_removal),
                      names_to = "Flow type",
                      values_to = "RelDiffFlow") |> 
         select(Compartment, Substance, `Flow type`, RelDiffFlow), 
       aes(x = Compartment, y = RelDiffFlow, fill = `Flow type`)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~Substance) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    panel.grid.major.y = element_line(color = "grey80", size = 0.7),
    panel.grid.minor.y = element_line(color = "grey90", size = 0.5)
  ) +
  scale_y_continuous(breaks = scales::extended_breaks(20))


# test_CompFlows$version <- "new"
# main_CompFlows$version <- "2025.04.0"
main_CompFlows <- main_CompFlows |> 
  rename(`Flow_kg/s_main` = `Flow_kg/s`)
test_CompFlows <- test_CompFlows |> 
  rename(`Flow_kg/s_test` = `Flow_kg/s`)

# sum(main_CompFlows$`Flow_kg/s_main`) - sum(test_CompFlows$`Flow_kg/s_test`)


CompFlowsDiff <- 
  merge(main_CompFlows, test_CompFlows) |> 
  mutate(diff = `Flow_kg/s_test`-`Flow_kg/s_main`,
         reldiff = (`Flow_kg/s_test`-`Flow_kg/s_main`)/abs(`Flow_kg/s_main`+`Flow_kg/s_test`)) |> 
  filter(diff != 0,
         `Flow_kg/s_main` > 0,
         abs(reldiff)>0.001)



ggplot(CompFlowsDiff, mapping = aes(x = To, y = From, color = reldiff)) + 
  geom_point(aes(shape = Substance)) + 
  labs(
    title = "Relative difference between old and new flow's",
    x = "To",  
    y = "From",  
    color = "Relative difference"  
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.grid.major = element_line(size = 0.2, color = "gray90"),
    panel.background = element_blank()  
  )

table_for_display <- CompFlowsDiff |>
  filter(abs(reldiff)>0.001) |>
  mutate(diff = format(diff, scientific = TRUE, digits = 2),
         # rel_diff = format(rel_diff, scientific = TRUE, digits = 2),
         `Flow_kg/s_main` = format(`Flow_kg/s_main`, scientific = TRUE, digits = 2),
         `Flow_kg/s_test` = format(`Flow_kg/s_test`, scientific = TRUE, digits = 2))

knitr::kable(table_for_display)


```



### Mass balance

```{r Mass balance check}

ggplot(rbind(main_MassBalance,test_MassBalance), 
       aes(x = Compartment, y = Diff_Flows, fill = Substance)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~version) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    panel.grid.major.y = element_line(color = "grey80", size = 0.7),
    panel.grid.minor.y = element_line(color = "grey90", size = 0.5)
  ) +
  scale_y_continuous(breaks = scales::extended_breaks(20))

```

From the mass balance test (flows into and out of a compartment should be in balance (sum = 0)) it can be seen that there are some much larger differences in the mass balance of the previous release "2025.04.0". This is mainly due to small tweaks in the solver by lowering stol variable in rootSolve::runsteady. In next release we might also change to rootSolve::steady. See SteadyStateSolver2.
