---
title: "10.1 Molecular verification clean"
author: "Anne Hids"
date: "2024-07-03"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::knit_meta()
knitr::opts_chunk$set(echo = TRUE)
## impossible to generalise? projectRoot <- paste(getwd(), "..", sep = "/")
#projectRoot <- paste(getwd(), "..", "..", sep = "/")
projectRoot <- "~/GitHub/SimpleBox/SBooScripts"
knitr::opts_knit$set(root.dir = projectRoot) #assuming vignette is in a direct subfolder of the project
```

*SimpleBox 4.0 produces two sorts of output; a steady-state or "level
III" output, and the quasi-dynamic or "level IV" output. The
steady-state output is developed when the conditions (loadings and
environmental conditions) remain constant in time for a sufficiently
long period. The mass flows and concentrations are constant in time, as
the sum of the mass balance equation terms is equal to zero for all
boxes, and the n steady-state concentrations can be solved from the n
linear mass balance equations. This steady-state solution is obtained by
means of a matrix inversion routine. The mass flows and concentrations
that characterise this steady state are given in output tables. The
quasi-dynamic output is developed when, after a change in conditions
(loadings or environmental conditions), the mass flows and
concentrations develop toward a new steady state, according to the mass
balance equations. This computation is made by numerical integration of
the set of mass balance equations from time zero, with all
concentrations at zero, to infinite time with all concentrations
constant at steady state. The standard procedure is to compute the
steady-state solution first and carry out the dynamic response
computation afterwards, as an option.*

# *Initialize*

*This vignettes demonstrates the mechanism behind SBoo. The model first
needs to be initialized. Before initialisation a substances needs to be
selected, otherwise the "default stubstance" is used. initWorld.R does
just that, and calculates SBvars, flows and k's.*

```{r warning=FALSE}
#We need to initialize, by default a molecular substance is selected
Potential_substances <- c("1-aminoanthraquinone", # no class
                          "1-HYDROXYANTHRAQUINONE", # acid
                          "1-Hexadecanamine, N,N-dimethyl-", # base
                          "1-Chloro-2-nitro-propane", # neutral
                          "Sb(III)", # metal
                          "nTiO2_10nm", # nanoparticle (class particle)
                          "microplastic") # microplastic (class particle)
substance <- Potential_substances[1]
source("baseScripts/initWorld_onlyMolec.R")
```

```{r data xlsx and R, include=FALSE, eval=FALSE}
#Comparing K's from R model to K's from excel model
library(openxlsx)
library(tidyverse)
#read in K matrix from excel
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.04_20240405_molec_test.xlsm") #molecular version for now

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

SBexcel.K$to <-  as.character(SBexcel.Names)
SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", 
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale]))

# Issue that compartments sediment and soil at global scale in excel have sd and s as acronyms instead of sd2 and s1
kaas <- 
  kaas |> 
  mutate(from = 
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") & 
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"), 
                          str_replace_all(from, c("sd2"="sd","s1"="s")), 
                          from)) |> 
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") & 
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

diagonal_R <- 
  aggregate(k ~ from, data = kaas, FUN = sum) #R model has k values per process, not per box. For the "diagonal" ("from = to") boxes, this is different than in the excel version. summing all the "froms" here to be able to compare them with excel matrix. This should result in one value for each compartment (scale-subcomp-species combo)

```

```{r diagonal, include=FALSE}
#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R

ifelse( (length(diagonal_R$k)-2)==length(diagonal_excel$k) ,
       print("ok"), stop("too many compartment combo's")) 

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(diagonal_R, diagonal_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 

#Plot of differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = diff)) +
  geom_boxplot()

```

*Why are these compartments in SB-R instance: \>
diagonal*$from[!diagonal$*from %in% diagonal_excel\$from] [1] "s2A"
"s2M" "s2T" "s3A" "s3M" "s3T" "sd0A" "sd0C" "sd0M" "sd0R" "sd0T" "sd1A"
"sd1M" "sd1T" "w0A" "w0M" "w0T" "w1A" "w1M" "w1T" [21] "w3C" "w3R"*

*This is due to the degradation rate constant being calculated for all
combinations of subcompartments. There is currently no data sheet where
the exclusion based on combination of subcompartment and scale can be
set. So these types of exclusions are needed in the process function
itself.*

*So now the compartments are the correct amount based on fixing
exclusions in k_Degradation.*

```{r rate_constants, include=FALSE}
#Dataframe of all other boxes

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

SBexcel.K$fromto[!SBexcel.K$fromto %in% kaas2$fromto] # No xlsx missing


kaas2$fromto[!kaas2$fromto %in% SBexcel.K$fromto] # more k's in SBoo than in SBxl

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)

#Column "diff" in df "mergedkaas" shows the difference between the R k output and the k's in Excel. 

#Visualize the diff per process in boxplots: which compartments (fromto boxes) have the largest differences between R and the excel model?

#Diff per "from" compartment
ggplot(mergedkaas, aes (x = from, y = diff)) +
  geom_boxplot()

#Diff per "to" compartment
ggplot(mergedkaas, aes (x = to, y = diff)) +
  geom_boxplot()

# "To" and "from" in one plot
ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") 


ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") 
#If you want to inspect per process instead of fromto box: look at the original kaas dataframe

```

## Difference in sedimentation rates

``` {r, include=FALSE}

large_differences <- mergedkaas |>
  filter(relDif > 0.001)

# Differences larger than 1 percentile all have to do with sedimentation/resuspension. 

difference <- kaas |> filter(from == "w1C")

################# Check difference in sedimentation rate #######################  

# Chosen k is from sea to marine sediment

sed_sea.R <- kaas |>
  filter(process == "k_Sedimentation") |>
  filter(fromScale == "Continental") |>
  filter(fromSubCompart == "sea") |>
  filter(toSubCompart == "marinesediment")

sed_sea.excel <- read.xlsx(SBExcelName,
                             sheet=9,
                             colNames=FALSE,
                             cols=c(7,12),
                             rows=154)

print(diff_sed <- sed_sea.excel$X2 - sed_sea.R$k)

# There is a difference between R and Excel of 1.47*10^-9

# Check if there is a difference between the settling velocities 
setvel.R <- World$fetchData("SettlVelocitywater") |>
  filter(Scale == "Continental")

setvel.excel <- read.xlsx(SBExcelName,
                             sheet=9,
                             colNames=FALSE,
                             cols=c(7,10),
                             rows=163)

print(diff_set <- setvel.excel$X2 - setvel.R$SettlVelocitywater) 

# There is a difference between the setting velocities of 3.5*10^-8 

# Check if there is a difference between the fractions in water
frinw.R <- World$fetchData("FRinw") |>
  filter(Scale == "Continental") |>
  filter(SubCompart == "sea")

frinw.excel <- read.xlsx(SBExcelName,
                             sheet=9,
                             colNames=FALSE,
                             cols=c(7,12),
                             rows=84)

print(diff_frinw <- frinw.excel$X2-frinw.R$FRinw)

# No difference in fraction in water between excel and R. 

# Adsorption comparison
ads.R <- kaas |>
  filter(process == "k_Adsorption") |>
  filter(fromScale == "Continental") |>
  filter(fromSubCompart == "sea")

ads.excel <- read.xlsx(SBExcelName,
                             sheet=9,
                             colNames=FALSE,
                             cols=c(7,12),
                             rows=164)

print(diff_ads <- ads.excel$X2-ads.R$k)

# Difference is 5.14*10^-11

kwsd.R <- World$fetchData("kwsd") |>
  filter(SubCompart == "sea" | SubCompart == "marinesediment")

```

## Perform the same test, but with the setvel in excel altered to the setvel in R

``` {r, include=FALSE}

source("baseScripts/initWorld_onlyMolec.R")

#read in K matrix from excel
SBExcelName <- paste0("vignettes/Development/Quality control/SBExcel/SimpleBox4.04_20240405_molec_test_sv.xlsm") #molecular version for now

SBexcel.K <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "K")
SBexcel.Names <- read.xlsx(SBExcelName,
                  colNames = FALSE,
                  namedRegion = "box_names")

colnames(SBexcel.K) <- SBexcel.Names

#long format version of Excel K matrix
# colnames(SBexcel.K) <- c("aR",	"w0R",	"w1R",	"w2R",	"sd1R", "sd2R",	"s1R",	"s2R",	"s3R",	"aC",	"w0C",	"w1C",	"w2C",	"sd1C",	"sd2C",	"s1C",	"s2C",	"s3C",	"aM",	"w2M",	"w3M",	"sdM",	"sM",	"aA",	"w2A",	"w3A",	"sdA",	"sA",	"aT",	"w2T",	"w3T",	"sdT",	"sT")
SBexcel.K$to <-  as.character(SBexcel.Names)

SBexcel.K <- pivot_longer(SBexcel.K, cols =  as.character(SBexcel.Names), values_to = "k", names_to = "from")

length(SBexcel.K$k)

#adding "from" and "to" acronyms to the R K matrix
kaas <- as_tibble(World$kaas)
unique(kaas$fromScale)
unique(kaas$fromSubCompart)

#R version does not us the acronyms of the excel version, set-up to convert them
#Note: this map creates the wrong acronym for soil and sediment at global scale, this is fixed afterwards
accronym_map <- c("marinesediment" = "sd2",
                "freshwatersediment" = "sd1",
                "lakesediment" = "sd0", 
                "agriculturalsoil" = "s2",
                "naturalsoil" = "s1",
                "othersoil" = "s3",
                "air" = "a",
                "deepocean" = "w3",
                "sea" = "w2",
                "river" = "w1",
                "lake" = "w0")

accronym_map2 <- c("Arctic" = "A",
                   "Moderate" = "M",
                   "Tropic" = "T",
                   "Continental" = "C",
                   "Regional" = "R")

kaas <- kaas |> mutate(from =  paste0(accronym_map[fromSubCompart], 
                            accronym_map2[fromScale]),
               to = paste0(accronym_map[toSubCompart], 
                           accronym_map2[toScale]))

# Issue that compartments sediment and soil at global scale in excel have sd and s as acronyms instead of sd2 and s1
kaas <- 
  kaas |> 
  mutate(from = 
                   ifelse((fromScale == "Tropic" | fromScale == "Arctic" | fromScale == "Moderate") & 
                            (fromSubCompart == "marinesediment" | fromSubCompart == "naturalsoil"), 
                          str_replace_all(from, c("sd2"="sd","s1"="s")), 
                          from)) |> 
  mutate(to = ifelse((toScale == "Tropic" | toScale == "Arctic" | toScale == "Moderate") & 
                       (toSubCompart == "marinesediment" | toSubCompart == "naturalsoil"), str_replace_all(to, c("sd2"="sd","s1"="s")), to))

  

diagonal_excel <- SBexcel.K[SBexcel.K$from == SBexcel.K$to,] #all the diagonals in excel are negative values -> sums of all the "froms" of that compartment

diagonal_R <- 
  aggregate(k ~ from, data = kaas, FUN = sum)

#Check for any differences in what compartments there are between versions
diagonal_R$from[!diagonal_R$from %in% diagonal_excel$from] # "sd0C" "sd0R" were added in sboo v1
diagonal_excel$from[!diagonal_excel$from %in% diagonal_R$from] #all excel diagonals are in those of R

ifelse( (length(diagonal_R$k)-2)==length(diagonal_excel$k) ,
       print("ok"), stop("too many compartment combo's")) 

#Single dataframe with both the R and excel diagonals
merged_diagonal <- merge(diagonal_R, diagonal_excel, by = "from", suffixes = c("_R", "_Excel")) 
merged_diagonal$k_Excel <- -merged_diagonal$k_Excel #Turning the "negative" values from the Excel matrix into positive ones
merged_diagonal$diff <- merged_diagonal$k_R - merged_diagonal$k_Excel 

#Plot of differences of the diagonals per compartment
ggplot(merged_diagonal, aes(x = from, y = diff)) +
  geom_boxplot()

kaas2 <- kaas |>  
  filter(from != to) |> #filtering the diagonals ou
  group_by(from, to) %>% summarize(k = sum(k)) #R version sometimes has multiple k's per fromto box; excel only has the summed k's per box

kaas2$fromto <- paste(sep = "_", kaas2$from, kaas2$to)

SBexcel.K <- filter(SBexcel.K, k != 0 & SBexcel.K$from != SBexcel.K$to)
SBexcel.K$fromto <- paste(sep = "_", SBexcel.K$from, SBexcel.K$to)

SBexcel.K$fromto[!SBexcel.K$fromto %in% kaas2$fromto] # No xlsx missing


kaas2$fromto[!kaas2$fromto %in% SBexcel.K$fromto] # more k's in SBoo than in SBxl

mergedkaas <- merge(kaas2, SBexcel.K, by = c("from", "to"), suffixes = c("_R", "_Excel"))

mergedkaas$diff <- mergedkaas$k_R - mergedkaas$k_Excel #compare R k to Excel K

mergedkaas <- as_tibble(mergedkaas) |> mutate(relDif = diff/k_R)

#Column "diff" in df "mergedkaas" shows the difference between the R k output and the k's in Excel. 

#Visualize the diff per process in boxplots: which compartments (fromto boxes) have the largest differences between R and the excel model?

#Diff per "from" compartment
ggplot(mergedkaas, aes (x = from, y = diff)) +
  geom_boxplot()

#Diff per "to" compartment
ggplot(mergedkaas, aes (x = to, y = diff)) +
  geom_boxplot()

# "To" and "from" in one plot
ggplot(mergedkaas, aes(x = to, y = from, color = abs(diff))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red", trans = "log10") 


ggplot(mergedkaas, aes(x = to, y = from, color = abs(relDif))) + 
  geom_point() +
  scale_color_gradient(low = "green", high = "red") 
#If you want to inspect per process instead of fromto box: look at the original kaas dataframe

```
