---
title: "R Notebook"
output:
  html_notebook
editor_options: 
  chunk_output_type: console
---

## testing SB dynamic output

```{r load data, include=FALSE}
# load("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/241203SB_Masses.RData")
# contains a dataframe with DeSolve::ode() output.
load("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/20241202_SB_Masses_v3.RData")
```

```{r basic info, echo=FALSE}
Combined_results <- rbind (Combined_results,Combined_results_TW)

sapply(Combined_results,class)
unique(Combined_results$time)
unique(Combined_results$Polymer)
unique(Combined_results$Source)
```

OK, but clear the the seconds output of time is not exactly the seconds in years "(24x3600x365.25)"

```{r completeness check, echo=FALSE}

Mass_all <-
  Combined_results |>
  mutate(time_year = time/(365.25*24*3600))

time_less <-
  Mass_all |> ungroup() |> 
  group_by(Source,Polymer,RUN) |> count(Polymer) |> filter(n<101)

time_2050 <-
  Mass_all |> ungroup() |>
  group_by(Source,Polymer,RUN) |> count(Polymer) |> filter(n==101)


```

There are `r round(100*length(time_less$RUN)/length(time_2050$RUN),2)` % RUN's (n = `r length(time_less$RUN)`) is incomplete meaning the solver output did not reach 2050.

No RUNs are missing for 1950

```{r}
Mass_all |> ungroup() |> mutate(time_year = as.integer(time_year)) |> 
  group_by(Source,Polymer,time_year) |> count(Polymer) |> filter(n<1000 & time_year == 1950)

# Mass_all |> ungroup() |> mutate(time_year = as.integer(time_year)) |> 
#   group_by(Source,Polymer,time_year) |> count(Polymer) |> filter(time_year == 1950)
# 
# 
# as_tibble(Mass_all) |> filter(Polymer=="NR") |> select(time_year,Polymer,Source,w1RS)

```

But for 2019, several polymers have runs that do not have a solution:

```{r}

Mass_all |> ungroup() |> mutate(time_year = as.integer(round(time_year,0))) |> 
  group_by(Polymer,time_year) |> count(Polymer)  |> filter(time_year == 2050) |> 
  pivot_wider(names_from = Polymer, values_from = n)



```

We continue the analysis with the Runs that have data for 2019.

```{r}

Mass_2019 <- Mass_all |> filter(time_year == 2019)

load(filepaths2[1])
States <- Output$SBoutput[[1]]$States
Mass_2019 <-
  Mass_2019 |> 
  pivot_longer(!c(time,time_year, RUN, Polymer, Source), names_to = "Abbr", values_to = "Mass") |>
  filter(!str_starts(Abbr, "emis")) |>
  left_join(States, by="Abbr") |>
  ungroup()

source("baseScripts/initWorld_onlyPlastics.R")

Conc_SAP_2019 <-
  Mass_2019 |> 
  # mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  # ungroup() |> 
  # sum over species
  group_by(time,time_year,RUN, Polymer, Source, Scale, SubCompart) |>
  summarise(Mass_SAP = sum(Mass)) |>
  # ungroup() |>
  left_join(World$fetchData("Volume"), 
            by=c("Scale", "SubCompart")) |>
  
  mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  ungroup() |>
  group_by(time,time_year,RUN, Polymer, Source, Scale, SubCompart) |>
  summarise(Mass_SAP = sum(Mass_SAP),
            Volume = sum(Volume)) |> 
  
  left_join(World$fetchData("FRACw"),
            by=c("Scale", "SubCompart")) |> 
  left_join(World$fetchData("FRACa"),
            by=c("Scale", "SubCompart")) |> 
  left_join(World$fetchData("rhoMatrix"),
            by=c("SubCompart")) |> 
  left_join(World$fetchData("Matrix"),
            by=c("SubCompart"))|> 
  mutate(conc_kg_m3 = Mass_SAP/Volume) |> 
  mutate(Unit = "kg/m3") |> 
  mutate(Concentration =
           case_match(Matrix,
                      "air" ~ conc_kg_m3*1000000000,
                      "water" ~ conc_kg_m3*1000000,
                      "soil" ~ conc_kg_m3  / ((1 - FRACw - FRACa) * rhoMatrix)*1000, # RhoWater needed (we need to define subcompartment variables for rhoWater, rhoSolid and rhoAir)
                      "sediment" ~ conc_kg_m3  / ((1 - FRACw - FRACa) * rhoMatrix)*1000,
                      .default = conc_kg_m3),
         Unit =
           case_match(Matrix,
                      "air" ~ "ug/m3",
                      "water" ~ "ug/L",
                      "soil" ~ "g/kg dw",
                      "sediment" ~ "g/kg dw",
                      .default = Unit)) |>
  mutate(SubCompartName = paste0(SubCompart, " (", Unit, ")")) |>
  select(-c(conc_kg_m3, FRACw, FRACa, rhoMatrix, Matrix, Volume))

plot_data <-
  Mass_2019 |> 
  # mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  # ungroup() |> 
  # sum over species
  group_by(time,time_year,RUN, Polymer, Source, Scale, SubCompart) |>
  summarise(Mass_SAP = sum(Mass)) |>
  # ungroup() |>
  left_join(World$fetchData("Volume"), 
            by=c("Scale", "SubCompart")) |>
  
  mutate(SubCompart = ifelse(SubCompart == "cloudwater", "air", SubCompart)) |>
  ungroup() |>
  group_by(time,time_year,RUN, Polymer, Source, Scale, SubCompart) |>
  summarise(Mass_SAP = sum(Mass_SAP),
            Volume = sum(Volume)) |> 
  filter(Scale == "Regional") |> ungroup() |> 
  group_by(RUN,Source,Scale,SubCompart,time_year) |> 
  summarise(Mass_SAP  = sum(Mass_SAP),
            Volume = sum(Volume)) |> 
  left_join(World$fetchData("FRACw"),
            by=c("Scale", "SubCompart")) |> 
  left_join(World$fetchData("FRACa"),
            by=c("Scale", "SubCompart")) |> 
  left_join(World$fetchData("rhoMatrix"),
            by=c("SubCompart")) |> 
  left_join(World$fetchData("Matrix"),
            by=c("SubCompart"))|> 
  mutate(conc_kg_m3 = Mass_SAP/Volume) |> 
  mutate(Unit = "kg/m3") |> 
  mutate(Concentration =
           case_match(Matrix,
                      "air" ~ conc_kg_m3*1000000000,
                      "water" ~ conc_kg_m3*1000000,
                      "soil" ~ conc_kg_m3  / ((1 - FRACw - FRACa) * rhoMatrix)*1000, # RhoWater needed (we need to define subcompartment variables for rhoWater, rhoSolid and rhoAir)
                      "sediment" ~ conc_kg_m3  / ((1 - FRACw - FRACa) * rhoMatrix)*1000,
                      .default = conc_kg_m3),
         Unit =
           case_match(Matrix,
                      "air" ~ "ug/m3",
                      "water" ~ "ug/L",
                      "soil" ~ "g/kg dw",
                      "sediment" ~ "g/kg dw",
                      .default = Unit)) |>
  mutate(SubCompartName = paste0(SubCompart, " (", Unit, ")")) |>
  select(-c(conc_kg_m3, FRACw, FRACa, rhoMatrix, Matrix, Volume))



ggplot(plot_data, mapping = aes(x = SubCompartName, y = Concentration, fill = Source)) +  
  geom_violin() + 
  labs(x = "Compartment",
       y = "Concentration") +
  plot_theme +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides="l",  
                      short = unit(0.07, "cm"),
                      mid = unit(0.07, "cm"),
                      long = unit(0.1, "cm"),
                      size = 0.25) +
  scale_fill_manual(values = Source_colors) +
  theme(legend.position = "bottom") +   
  guides(fill = guide_legend(title = NULL))



```

```{r}

#Try GSA
load("/rivm/r/E121554 LEON-T/03 - uitvoering WP3/Deliverable 3.5/SB_Material_parameters.RData")


Par_Probs <-
  Material_Parameters_long |> 
  mutate(par_name = paste0(VarName,Polymer,SubCompart,Species)) |> 
  distinct(RUN,value,VarName,.keep_all = TRUE) |>
  # filter(Polymer == "NR") |> 
  select(-c(VarName,Source,Scale,SubCompart,Species,Distribution,
            Polymer,Unit)) |> pivot_wider(names_from = par_name,
                                          values_from = value) |> 
  drop_na()

EmissionProB <- DPMFA_SBoutput$DPMFA_sink_micro |> unnest(Emis) |> filter(Year %in% c(1989,1999,2009,2019)) |> ungroup() |> 
  group_by(Abbr,Year,RUN) |> 
  summarise(Mass_kg_s = sum(Mass_kg_s)) |> 
  mutate(emis_name = paste0(Year,Abbr)) |> ungroup() |> 
  select(-c(Year,Abbr)) |> 
  pivot_wider(names_from = emis_name,
              values_from = Mass_kg_s) |> left_join(DPMFA_SBoutput$NR_SBR_fractions)

ProbS <- 
  plot_data |> filter(SubCompart == "river" & Source == "Tyre wear") |> 
  ungroup() |> 
  select(RUN,Concentration) |> 
  left_join(EmissionProB) |> 
  left_join(Par_Probs) |> 
  drop_na() |> 
  # mutate_all(~if_else(. == 0, 1e-20, .)) |> # make 0's very small numbers 1e-20
  mutate_all(log) |> # log transform
  drop_na() |> # drop any rows with NA's
  select(-where(~ var(.) == 0))# remove columns with 0 variance (are constant)

probX <- ProbS |> select(-c(RUN, Concentration))
probY <- ProbS |> pull(Concentration)
library(sensitivity)
# library(readxl)
library(ggplot2)
library(ks) ### ks needed for sensiFdiv function


m <- sensiFdiv(model = NULL, X=probX, fdiv = "TV", nboot = 0, conf = 0.95,   scale = TRUE)
tell(m, y=probY,S)

# ggplot(m, ylim = c(0, 1))
#prepare output for ggplot
borg_d_temp <- tibble(Variable = colnames(probX),
                      delta= m$S$original)

# TCdelta_part1 <- borg_d_temp 
# TCsens_data <-
#   TCsens_data |> add_row(TCdelta_part1)


mydf <-borg_d_temp |> arrange(desc(delta))

# Sensplot <- ggplot(borg_d_temp, aes(x = delta,reorder(TC,delta))) + geom_bar(stat="identity")+
#   theme_light()+
#   theme(axis.text.y = element_text(size = 10),
#         axis.title.y = element_blank(),
#         # plot.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())
# 
# 
# pdata <-
#   borg_d_temp |> drop_na()
#   # filter(Source == iSource)
# p <- ggplot(pdata, aes(x=Polymer, y=TC, fill= delta)) + 
#   geom_tile(stat="identity")+
#   scale_fill_viridis_c(alpha = 0.5, option = "A", direction = -1)+
#   theme_light()+
#   theme(axis.text.y = element_text(size = 8),
#         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
#         # plot.background = element_blank(),
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())


```
